### 2013-03-08 EconGrowth-R-Functions.Rnw
Now including graphs for both US and ZA in the presentation.

<<setup_parent, echo=FALSE, message=FALSE, eval=TRUE>>=
require(knitr)
require(lattice)
require(latticeExtra)
require(ggplot2)
require(car)
require(mosaic)
require(xtable)
require(nlmrt)
require(micEconCES)
# tikz allows use of LaTeX formatting and font in graphs. Allows for a consistent look across the paper.
# See http://r-forge.r-project.org/R/?group_id=440 for instructions on installing tikzDevice.
require(tikzDevice) 
trellis.par.set(theme=theme.mosaic())
options(width=75)
options(show.signif.stars=TRUE)
opts_chunk$set(
  # ********************************************************************************************************
  eval=TRUE,
  # ********************************************************************************************************
  dev='tikz',    #Allows LaTeX code in graphical output. E.g., "$y$" for a variable name in the legend for a graph.
  tidy=FALSE,
  comment=NA,
  #Tells whether to cache output from chunks are saved. Cacheing saves time. 
  #However, references to LaTeX tables DO NOT WORK if cacheing is turned on (i.e., cache=TRUE).
  cache=FALSE,
  warning=FALSE,  #Tells whether to show warnings in the output.
  message=FALSE, #Turns off messages for all chunks. Set TRUE on an individual chunk to see it.
  echo=FALSE     #Tells whether to echo code for all chunks. Set TRUE on an individual chunk to see its code.
)

# Statistical significance levels. We'll work with 95% CIs
ciLevel <- 0.95
ciHalfLevel <- ciLevel + (1.0-ciLevel)/2.0
# List of countries
countryAbbrevs <- c(US="US", UK="UK", JP="JP", CN="CN", ZA="ZA", SA="SA", IR="IR", TZ="TZ", ZM="ZM")
countryAbbrevsAlph <- c(CN="CN", IR="IR", JP="JP", SA="SA", TZ="TZ", UK="UK", US="US", ZA="ZA", ZM="ZM")
countryAbbrevsU <- c(US="US", UK="UK", JP="JP") #Only these countries have useful work data
countryNamesAlph <- c(CN="China", IR="Iran", JP="Japan", SA="Saudi Arabia", TZ="Tanzania", UK="United Kingdom", US="USA", ZA="South Africa", ZM="Zambia") #In alphabetical order.
yLimitsForGDPGraphs <- list(c(1,10), c(1,4), c(1,4), c(1,4), c(1,4), c(1,4), c(1,4), c(1,4), c(1,4)) # Alph order
energyTypes <- c(Q="Q", X="X", U="U") # List of energy types
factors <- c(K="K", L="L", Q="Q", X="X", U="U") # List of factors of production
########### Several global parameters for graphs. Set here and use below to ensure consistent appearance of graphs.
# Set the order for presenting countries in 3x3 lattice graphs. Default is alphabetical. 
# "1" means first alphabetically.
countryOrderForGraphs <- c(7,6,3,1,8,5,4,2,9) # Sets the order as US, UK, JP, CN, ZA, TZ, SA, IR, ZM.
timeTics <- c(1980, 1990, 2000, 2010)
yTicsForIndexedGraphs <- c(1,2,3,4,5,6,7,8,9,10) #y tic mark locations.

maxWidth <- 6.5 #Inches
# Full page lattice plot sizes
ninePanelLayoutSpec <- c(3,3) #indicates a 3x3 arrangement of panels.
onePanelLayoutSpec <- c(1,1) #indicates a 1x1 arrangement of panels.
ninePanelGraphWidth <- maxWidth     
ninePanelGraphHeight <- 7.5  #Inches
presentationGraph1RowHeight <- 2.8 #Inches
presentationGraph2ColWidth  <- 2 #Inches
keyTextSize <- 0.85 #85% of normal size
keyColumns <- 1 #Want only 1 column in the key for lattice graphs
defaultKeyXLoc <- 0.01 #x position of the key. This default is good for 9-panel graphs.
defaultKeyYLoc <- 0.95 #y position of the key. This default is good for 9-panel graphs.
# Parameter graphs are tricky to ensure same size. I've found that with height set to 5 in for the 
# cd graphs and 2.93 in for the sf graphs, I get a nearly-exact match with graph size. From these settings I 
# can set up 2 equations:
#    5 in = 2*hgraph + haxis
#    2.93 in = hgraph + haxis
# to estimate that h_axis = 0.86 in and hgraph = 2.05 in.
h_param_graph <- 2.07 # Inches for a single panel
h_param_axis <- 0.86 # Inches (also includes whatever margin is applied)
sfParameterGraphHeight <- h_param_graph + h_param_axis
cdParameterGraphHeight <- 2*h_param_graph + h_param_axis
cesParameterGraphHeight <- 3*h_param_graph + h_param_axis
linexParameterGraphHeight <- h_param_graph + h_param_axis
sfParameterGraphWidth <- maxWidth
cdParameterGraphWidth <- maxWidth
cesParameterGraphWidth <- maxWidth
linexParameterGraphWidth <- maxWidth
# Other graph parameters that apply to all graphs
scaleTextSize <- 0.8  #Multiple of normal size
scaleTickSize <- -0.5 #50% of normal size and pointing INWARD!
@

<<Data Functions, eval=TRUE>>=
#################################
# This function loads data given a country abbreviation.
# The file name from which data will be loaded is assumed to be of the form 
# "<countryAbbrev>Data.txt". The file is assumed to exist in a subfolder of this project called "data".
#
# countryAbbrev is a string containing the 2-letter abbreviation for the country, e.g. "US" or "CN"
#
# returns a data.frame with the data that has been loaded
##
loadData <- function(countryAbbrev){
  # Read the data file as a table with a header.  
  fileName <- paste("data/", countryAbbrev, "Data.txt", sep="")
  dataTable <- read.table(fileName, header = TRUE)
  return(dataTable)
}

###############################
# Tells whether we have data for this combination of country and energy type for a single-factor curve fit.
#
# returns logical value (TRUE or FALSE)
##
haveDataSF <- function(countryAbbrev, factor){
  # The design here is to go through several cases where data is unavailable and
  # return FALSE. If we make it through all of those cases, we'll return TRUE.
  if (!(countryAbbrev %in% countryAbbrevs)){
    # If we don't know the country, we don't have data
    return(FALSE)
  }
  if (missing(factor) || is.na(factor)){
    # If we don't have a factor, we don't have data.
    return(FALSE)
  }
  if (!(factor %in% factors)){
    # If we don't know the factor, we don't have data
    return(FALSE)
  }
  if (factor == "U" && (!(countryAbbrev %in% countryAbbrevsU))){
    # We have U data for only a few countries.
    return(FALSE)
  }
  return(TRUE)
}

###############################
# Tells whether we have data for this combination of country and energy type for a Cobb-Douglas curve fit.
# a missing energyType or energyType == NA means that we want to develop a model without energy, if possible.
#
# returns logical value (TRUE or FALSE)
##
haveDataCD <- function(countryAbbrev, energyType){
  # The design here is to go through several cases where data is unavailable and
  # return FALSE. If we make it through all of those cases, we'll return TRUE.
  if (!(countryAbbrev %in% countryAbbrevs)){
    # If we don't know the country, we don't have data
    return(FALSE)
  }
  if (missing(energyType) || is.na(energyType)){
    # If we don't want energy included, we have data for all known countries. We need to check this case
    # first, BEFORE doing various checks on the energyType.
    return(TRUE)
  }
  if (!(energyType %in% energyTypes)){
    # If we want to include energy, we know only a few types. So, return false if we don't know the energy type.
    return(FALSE)
  }
  if (energyType == "U" && (!(countryAbbrev %in% countryAbbrevsU))){
    # We have U data for only a few countries.
    return(FALSE)
  }
  return(TRUE)
}

#####################
# This function adds NA rows to the end of a data.frame to ensure that the df is 
# the same length as the country data set
# returns a modified version of df that includes the padded rows filled with "NA".
##
padRows <- function(countryAbbrev, df){
  nRowsData <- nrow(loadData(countryAbbrev)) #Gets number of rows in data set
  nRowsDF <- nrow(df)
  nRowsToAdd <- nRowsData - nRowsDF
  nCol <- ncol(df)
  dfToAppend <- as.data.frame(matrix(NA, ncol=nCol, nrow=nRowsToAdd))
  colnames(dfToAppend) <- colnames(df)
  return(rbind(df, dfToAppend))
}

##############################
# Returns an integer representing the column index for some data
# dataTable the data.frame in which you want to change column names
# factor should be a string and one of Year, Y, K, L, Q, X, or U
##
columnIndex <- function(dataTable, factor){
  if (factor == "Year"){
    colName <- "iYear"
  } else if (factor == "Y"){
    colName <- "iGDP"
  } else if (factor == "K"){
    colName <- "iCapStk"
  } else if (factor == "L"){
    colName <- "iLabor"
  } else if (factor == "Q"){
    colName <- "iQ"
  } else if (factor == "X"){
    colName <- "iX"
  } else if (factor == "U"){
    colName <- "iU"
  } else {
    print(paste("Unknown factor:", factor, "in colIndex. Terminating execution."))
    quit()
  }
  # Get the desired column index.
  colIndex <- which(names(dataTable) %in% colName) #Find index of desired column
  return(colIndex)  
}

##############################
# Replaces a column name with the given string
# dataTable the data.frame that you're working with
# factor should be a string and one of Year, Y, K, L, Q, X, or U
# newName should be a string and the desired new name of the column
# returns dataTable with a new name for one of its factor column.
##
replaceColName <- function(dataTable, factor, newName){
  colIndex <- columnIndex(dataTable, factor)
  colnames(dataTable)[colIndex] <- newName #Change desired column name to newName
  return(dataTable)
}

##############################
# Returns an xtable of covariance information for the specified country.
##
covarianceTable <- function(countryAbbrev){
  data <- loadData(countryAbbrev)
  # Grab only the data we want.
  dataForCovar <- cbind(data$iGDP, data$iCapStk, data$iLabor, data$iQ, data$iX)
  names <- c("$y$", "$k$", "$l$", "$q$", "$x$")
  if (countryAbbrev %in% countryAbbrevsU){
    # Add U data only if we have it.
    dataForCovar <- cbind(dataForCovar, data$iU)
    names <- c("$y$", "$k$", "$l$", "$q$", "$x$", "$u$")
  }
  covarResults <- cor(dataForCovar)
  colnames(covarResults) <- names
  rownames(covarResults) <- names
  covarXtable <- xtable(covarResults, 
                      caption=paste(countryAbbrev, "covariance table."), 
                      label=paste("tab:Covariance_", countryAbbrev, sep=""))
  return(covarXtable)
}

###############################
# Prints a covariance table for countryAbbrev
##
printCovarTable <- function(countryAbbrev){
  print(covarianceTable(countryAbbrev), 
                      caption.placement="top", 
                      sanitize.rownames.function = identity,
                      sanitize.colnames.function = identity,
                      size="\\tiny",
                      table.placement="H")
}

####################################
# Creates a graph that displays all of the factors of production for all countries (if you leave off countryAbbrev)
# or a specific country (if you supply a 2-letter abbreviation for a country that we know).
# Provide a value for textScaling to scale the size of the text on the graph. This is expecially useful
# for converting between graphs in a paper and graphs in a beamer presentation.
# textScaling = 1.0 is good for a paper. textScaling = 0.6 is good for a beamer presentation.
# keyXLoc and keyYLoc are locations for the graph's legend.
##
createHistoricalLatticeGraph <- function(countryAbbrev, textScaling = 1.0, keyXLoc = defaultKeyXLoc, keyYLoc = defaultKeyYLoc){
  # Code for all graphs, regardless of whether we want to focus on a specific country
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1, 5, 2, 4, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 2, 2, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "black", "black", "red", "blue", "darkorange") #line and symbol colors
  symbols <- c(1, NA, NA, NA, NA, NA)  #NA gives no symbol.
  # Code that deals with items that are specific to whether we want all countries or a specific country.
  if (missing(countryAbbrev)){
    # We want a graph with panels for all countries
    dataTable <- loadData("All")
    factorLevels <- countryNamesAlph # We want all countries shown
    indexCond <- list(countryOrderForGraphs) # We want all countries in this order
    layout <- ninePanelLayoutSpec # Show all countries
    yLimits <- yLimitsForGDPGraphs
  } else {
    # We want only a specific country
    dataTable <- loadData(countryAbbrev)
    # Select the correct y limits
    index <- which(countryAbbrevsAlph %in% countryAbbrev)
    # The following lines use [index:index] as a convenient way of subsetting.
    # This has the added benefit of maintaining the correct classes for things.
    yLimits <- yLimitsForGDPGraphs[index:index]       # Pick limits for the country we want.
    factorLevels <- countryNamesAlph[index:index]     # Only show the country we have chosen.
    indexCond <- list(c(1))                           # We want only one country.
    layout <- onePanelLayoutSpec                      # We want only one panel in the graph.
  }
  graph <- xyplot(iGDP+iCapStk+iLabor+iQ+iX+iU ~ Year | Country, data=dataTable,
                  type = graphType,
                  index.cond = indexCond, #orders the panels.
                  layout = layout, 
                  # Sets strip parameters
                  strip = strip.custom(factor.levels=factorLevels, # Sets text for factor levels
                                       bg="transparent", # Sets background transparent to match the graph itself.
                                       par.strip.text=list(cex=textScaling) # Scales text in the strip.
                                       ),
                  as.table = TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("$y$", "$k$", "$l$", "$q$", "$x$", "$u$")),
                           type=graphType,
                           cex=keyTextSize * textScaling, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize * textScaling, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              x=list(at=timeTics),
                              y=list(relation="free",  #allows each axis to be different
                                     at=yTicsForIndexedGraphs)), #specifies location for tics
                  ylim=yLimits, #y axis limits
                  #axis labels and scaling
                  xlab=list(label="", cex=textScaling), 
                  ylab=list(label="Indexed (1980=1 or 1991=1)", cex=textScaling)
                  ) 
  return(graph)
}
@

<<single-factor functions, eval=TRUE>>=
####################
# Returns an nls single-factor model for the country and factor specified.
# factor should be one of "K", "L", "Q", "X", or "U".
##
singleFactorModel <- function(countryAbbrev, factor){
  dataTable <- loadData(countryAbbrev) #Load the data that we need.
  # We'll change the name of the desired column to "f"
  dataTable <- replaceColName(dataTable, factor, "f")
  # Now do the fit.
  lambdaGuess <- 0.0 #guessing lambda = 0 means there is no technological progress.
  # Some economies need different guesses
  if (countryAbbrev == "IR" && factor == "X"){
    mGuess <- 0.7
  } else {
    # This works for every other combination of country and factor.
    mGuess <- 0.5
  }
  start <- list(lambda=lambdaGuess, m=mGuess)
  # Runs a non-linear least squares fit to the data. We've replaced beta with 1-alpha for simplicity.
  model <- iGDP ~ exp(lambda*iYear) * f^m
  modelSF <- nls(formula=model, data=dataTable, start = start)
#   model <- iGDP ~ exp(lambda*iYear) * f^m
#   modelSF <- wrapnls(formula=model, data=dataTable, start=list(lambda=lambdaGuess, m=mGuess))
  return(modelSF)
}

#########################
# Takes the single-factor fitted models and creates per-country predictions for them.
# Returns a data.frame with the predictions.
##
singleFactorPredictions <- function(countryAbbrev, factor){
  # Can't make predictions for any of CN, ZA, SA, IR, TZ, or ZM if we're interested in U
  if (!(haveDataSF(countryAbbrev, factor))){
    #If we don't have data for this combination of countryAbbrev and energyType, 
    # return a column of NAs if the above conditions have been met.
    nRows <- 21 # All of these countries need 21 rows.
    df <- as.data.frame(matrix(NA, ncol = 1, nrow = nRows))
    colnames(df) <- "pred"
    return(df)
  }
  model <- singleFactorModel(countryAbbrev, factor)
  pred <- predict(model) #See http://stackoverflow.com/questions/9918807/how-get-plot-from-nls-in-r
  df <- data.frame(pred)
  # Pad with rows as necessary
  df <- padRows(countryAbbrev, df)
  return(df)
}

#########################
# Takes the single-factor fitted models and creates a single column of predicted GDP values
# that corresponds, row for row, with the AllData.txt file.
##
singleFactorPredictionsColumn <- function(factor){
  out <- do.call("rbind", lapply(countryAbbrevs, singleFactorPredictions, factor=factor))
  colnames(out) <- c(paste("predGDP", factor, sep=""))
  return(out)
}

##############################
# Creates a graph that plots predicted GDP as lines, one for each single factor, and historical GDP 
# data as open circles.
##
createSFLatticeGraph <- function(countryAbbrev, textScaling = 1.0, keyXLoc = defaultKeyXLoc, keyYLoc = defaultKeyYLoc){
  dataTable <- loadData("All") #Grab the raw data
  predictionsK <- singleFactorPredictionsColumn("K") #Predictions from SF with K
  predictionsL <- singleFactorPredictionsColumn("L") #Predictions from SF with L
  predictionsQ <- singleFactorPredictionsColumn("Q") #Predictions from SF with Q
  predictionsX <- singleFactorPredictionsColumn("X") #Predictions from SF with X
  predictionsU <- singleFactorPredictionsColumn("U") #Predictions from SF with U
  #Now add the predictions columns to the data.
  dataTable <- cbind(dataTable, predictionsK, predictionsL, predictionsQ, predictionsX, predictionsU)
  # Code for all graphs, regardless of whether we want to focus on a specific country
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1, 5, 2, 4, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 2, 2, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "black", "black", "red", "blue", "darkorange") #line and symbol colors
  symbols <- c(1, NA, NA, NA, NA, NA)  #NA gives no symbol.
  # Code that deals with items that are specific to whether we want all countries or a specific country.  
  if (missing (countryAbbrev)){
    # We want a graph with panels for all countries
    yLimits <- yLimitsForGDPGraphs
    factorLevels <- countryNamesAlph # We want all countries shown
    indexCond <- list(countryOrderForGraphs) # We want all countries in this order
    layout <- ninePanelLayoutSpec # Show all countries
  } else {
    # We want only a specific country
    dataTable <- subset(dataTable, Country == countryAbbrev)    
    # Select the correct y limits
    index <- which(countryAbbrevsAlph %in% countryAbbrev)
    # The following lines use [index:index] as a convenient way of subsetting.
    # This has the added benefit of maintaining the correct classes for things.
    yLimits <- yLimitsForGDPGraphs[index:index]       # Pick limits for the country we want.
    factorLevels <- countryNamesAlph[index:index]     # Only show the country we have chosen.
    indexCond <- list(c(1))                           # We want only one country.
    layout <- onePanelLayoutSpec                      # We want only one panel in the graph.
  }  
  graph <- xyplot(iGDP+predGDPK+predGDPL+predGDPQ+predGDPX+predGDPU ~ Year | Country, data=dataTable,
                  type=graphType,
                  index.cond=indexCond, #orders the panels.
                  layout=layout, #indicates a 3x3 arrangement of panels.
                  strip = strip.custom(factor.levels=factorLevels, 
                                       bg="transparent", # Sets background transparent to match the graph itself.
                                       par.strip.text=list(cex=textScaling) # Scales text in the strip.
                                       ),
                  as.table=TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("Actual", "With $k$", "With $l$", "With $q$", "With $x$", "With $u$")),
                           type=graphType,
                           cex=keyTextSize * textScaling, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize * textScaling, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              x=list(at=timeTics),
                              y=list(relation="free",  #allows each axis to be different
                                     at=yTicsForIndexedGraphs)), #specifies location for tics
                  ylim=yLimits, #y axis limits
                  #axis labels and scaling
                  xlab=list(label="", cex=textScaling), 
                  ylab=list(label="Indexed (1980=1 or 1991=1)", cex=textScaling)
                  )
  return(graph)
}

#################################################
# Calculates parameter estimates and confidence intervals
# for the single factor production function for a given a country.
#
# countryAbbrev is a string containing the 2-letter abbreviation for the country, e.g. "US" or "CN"
# factor is a string, one of "K", "L", "Q", "X", or "U"
#
# returns a data.frame of data for the Cobb-Douglas model. 
# First row is the +95% CI on all parameters
# Second row contains the parameter estimates
# Third row is the -95% CI on all parameters
# Each column has names: lambda and m corresponding to the parameters in the model.
##
singleFactorData <- function(countryAbbrev, factor){
  #First, check to see if we want useful work (U) AND one of the countries for which we don't have data.
  if (!haveDataSF(countryAbbrev, factor)){
    #Return a column of NAs if we don't have data for this factor
    nRows <- 3 # +95% CI, SF, and -95% CI.
    nCols <- 2 # lambda, m
    df <- as.data.frame(matrix(NA, ncol = nCols, nrow = nRows))
    colnames(df) <- c("lambda", "m")
    rownames(df) <- c("+95% CI", "SF", "-95% CI")
    return(df)
  }
  # We have a combination of country and factor for which we have data.
  modelSF <- singleFactorModel(countryAbbrev, factor)
  summarySF <- summary(modelSF) # Gives the nls summary table.
  ciSF <- confint(modelSF, level = ciLevel)  # Calculates confidence intervals for the SF model.
  dofSF <- summarySF$df[2] # Gives the degrees of freedom for the model.
  tvalSF <- qt(ciHalfLevel, df = dofSF)
  #Single factor model
  lambda <- as.numeric(coef(modelSF)["lambda"])
  m <- as.numeric(coef(modelSF)["m"])
  # Combine all estimates and their confidence intervals into data frames with intelligent row names
  estSF <- data.frame(lambda = lambda, m = m)
  row.names(estSF) <- "SF"
  # The [1] subscripts pick off the lower confidence interval
  lowerSF <- data.frame(lambda=ciSF["lambda","2.5%"], 
                        m=ciSF["m", "2.5%"])
  row.names(lowerSF) <- "-95% CI"
  # The [2] subscripts pick off the upper confidence interval
  upperSF <- data.frame(lambda=ciSF["lambda","97.5%"], 
                        m=ciSF["m", "97.5%"])
  row.names(upperSF) <- "+95% CI"
  # Now create the data for a table and return it
  dataSF <- rbind(upperSF, estSF, lowerSF)
  return(dataSF)
}

############
# Creates a row for the single factor parameters table for the given country (2-letter code) and factor.
##
singleFactorCountryRow <- function(countryAbbrev, factor){
 dataSF <- singleFactorData(countryAbbrev, factor)
 out <- cbind(dataSF["-95% CI", "lambda"], dataSF["SF", "lambda"], dataSF["+95% CI", "lambda"],
              dataSF["-95% CI", "m"],  dataSF["SF", "m"],  dataSF["+95% CI", "m"])
 return(out)
}

############################
# Aggregates the single-factor results into a big data table for the given factor.
##
singleFactorParamsTable <- function(factor){
  #Do rbind on the results of creating a row in the table for every country abbreviation that we know.
  dataSF <- do.call("rbind", lapply(countryAbbrevs, singleFactorCountryRow, factor=factor))
  if (factor == "K"){
    factorString <- "$\\alpha$"
  } else if (factor == "L"){
    factorString <- "$\\beta$"
  } else {
    factorString <- "$\\gamma$"
  }
  colnames(dataSF) <- c(" ", "$\\lambda$", " ", 
                        " ", factorString,  " ")
  rownames(dataSF) <- countryAbbrevs
  tableSF <- xtable(dataSF, 
                    caption=paste("Single-factor model (with $", tolower(factor), "$) parameters for 1980-2011 (US, UK, JP), 1991-2010 (CN and ZA), and 1991-2011 (SA, IR, TZ, and ZM). (Parameter estimates beneath symbol. 95\\% confidence interval bounds to left and right.)", sep=""), 
                    label=paste("tab:SF_Parameters_With_", factor, sep=""),
                    digits = c(1, 4,4,4, 2,2,2),
                    align = "r|ccc|ccc") #Sets alignment of the numbers in the columns
  return(tableSF)
}

###########################################
# Creates a number of rows in a data.frame that contain information 
# about the coefficients of a single factor model for countryAbbrev.
# Each parameter has its own row with confidence intervals.
# The country name is in a column. Which parameter is involved is
# also in a column.
# 
# The return type is a data.frame.
##
singleFactorCountryRowsForParamsGraph <- function(countryAbbrev, factor){
  #Create three rows, one for each parameter. Each row is a data.frame so that it is plottable!
  dataSF <- singleFactorData(countryAbbrev, factor)
  valueRow <- "SF"
  # Create rows for each parameter
  lambdaRow <- data.frame(country = countryAbbrev, 
                          parameter = "lambda", 
                          lowerCI = dataSF["-95% CI", "lambda"], 
                          value = dataSF[valueRow, "lambda"], 
                          upperCI = dataSF["+95% CI", "lambda"])
  mRow <- data.frame(country = countryAbbrev, 
                     parameter = "m",
                     lowerCI = dataSF["-95% CI", "m"],
                     value = dataSF[valueRow, "m"],  
                     upperCI = dataSF["+95% CI", "m"])
  table <- rbind(lambdaRow, mRow)
  return(table)
}

###############################
# Prints a covariance table for the given factor of production
##
printSFParamsTable <- function(factor){
  print(singleFactorParamsTable(factor),
      caption.placement="top", 
      sanitize.colnames.function = identity, 
      size="\\tiny", 
      table.placement="H")
}

#############################
# Creates a graph with confidene intervals for the single-factor model for the given factor
##
createSFParamsGraph <- function(factor){
  # Set the name of the 2nd graph (m)
  factorLabel <- paste("$", tolower(factor), "$", sep="")
  # Create a data table with the following columns:
  # country abbrev, parameter (lambda or m), -95% CI, value, +95% CI
  dataTable <- do.call("rbind", lapply(countryAbbrevs, singleFactorCountryRowsForParamsGraph, factor=factor))
  graph <- segplot(country ~ upperCI + lowerCI | parameter, 
                   data = dataTable,
                   centers = value, #identifies where the dots should be placed
                   draw.bands = FALSE, #provides nicer error bars
                   horizontal = FALSE, #makes error bars vertical and puts the countries in the x axis
                   layout = c(2,1), #2 column, 1 row
                   col = "black", #Sets line color to black
                   lwd = 1, #Sets line width to 1.0
                   strip = strip.custom(bg="white", factor.levels=c("$\\lambda$", factorLabel)),
                   ylim = list(c(-0.05, 0.1), c(-0.5, 1.5)), #y axis limits
                   scales = list(cex=scaleTextSize, #controls text size on scales.
                                 tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                                 x=list(cex=0.75), #reduces text size so that country abbrevs are legible
                                 y=list(relation="free", #allow each axis to be different
                                        at=list(c(0.0, 0.05, 0.1),   #y tick mark for lambda
                                                c(0.0, 0.5, 1.0)
                                                )
                                        )
                                 )
                   )

  return(graph)
}
@

<<cobb-douglas functions, eval=TRUE>>=
####################
# Returns an nls Cobb-Douglas model (without energy) for the country specified. 
# No energy is included in the function to be fitted.
##
cdModel <- function(countryAbbrev){
  dataTable <- loadData(countryAbbrev) #Load the data that we need.
  # Run the non-linear least squares fit to the data. No energy term desired in the Cobb-Douglas equation.
  # Establish guess values for alpha and lambda.
  lambdaGuess <- 0.0 #guessing lambda = 0 means there is no technological progress.
  alphaGuess <- 0.7 #0.7 gives good results for all countries.  
  start <- list(lambda=lambdaGuess, alpha=alphaGuess)
  # Runs a non-linear least squares fit to the data. We've replaced beta with 1-alpha for simplicity.
  model <- iGDP ~ exp(lambda*iYear) * iCapStk^alpha * iLabor^(1 - alpha)
  modelCD <- nls(formula=model, data=dataTable, start = start,
                 #Include the next 3 lines to fit with constraints.
                 #algorithm = "port",
                 #lower = list(lambda=-Inf, alpha=0),
                 #upper = list(lambda=Inf, alpha=1)
                 )
# print("str(modelCD)")
# print(str(modelCD))
# print("")
# print(modelCD$)
#   # We estimated alpha with the above code, but we have beta = 1.0 - alpha. So, get 
#   # an estimate for beta and add it to the model.
#   summaryCD <- summary(modelCD)
# print("summaryCD")
# print(summaryCD)
#   dofCD <- summaryCD$df[2] # Gives the degrees of freedom for the model.
#   tvalCD <- qt(ciHalfLevel, df = dofCD)
#   alpha <- coef(modelCD)["alpha"]
#   betaEst <- deltaMethod(modelCD, "1 - alpha") # Estimates beta and its standard error (SE).
# print("betaEst")
# print(betaEst)
#   # Calculate beta and its confidence interval and report it.
#   # Now calculate a confidence interval on beta
#   betaCICD <- with(betaEst, Estimate + c(-1.0, 1.0) * tvalCD * SE); betaCICD # Gives the confidence interval on beta.  
# print("betaCICD")
# print(betaCICD)

  return(modelCD)
}

####################
# Returns an nls Cobb-Douglas model for the country specified
# energyType is one of "Q", "X", or "U".
##
cdeModel <- function(countryAbbrev, energyType){
  dataTable <- loadData(countryAbbrev) #Load the data that we need.
  # We need to do the Cobb-Douglas fit with the desired energy data.
  # To achieve the correct fit, we'll change the name of the desired column
  # to "iEToFit" and use "iEToFit" in the nls function. 
  dataTable <- replaceColName(dataTable, energyType, "iEToFit")
  # Establish guess values for lambda, alpha, and beta.
  lambdaGuess <- 0.0 #guessing lambda = 0 means there is no technological progress.
  alphaGuess <- 0.9
  betaGuess <- 1.0 - alphaGuess
  aGuess <- alphaGuess
  bGuess <- alphaGuess + betaGuess
  start <- list(lambda=lambdaGuess, a=aGuess, b=bGuess)
  # Now actually do the fit, using the column name "iEToFit".
  # gamma is a free parameter.
  # Reparameterize to ensure that we meet the constraints:
  # * alpha + beta + gamma = 1.0.
  # * alpha, beta, and gamma are all between 0.0 and 1.0.
  # To do this, we reparameterize as
  # * 0 < a < 1
  # * 0 < b < 1
  # * alpha = min(a, b)
  # * beta = b - a
  # * gamma = 1 - max(a, b)
  model <- iGDP ~ exp(lambda*iYear) * iCapStk^min(a,b) * iLabor^abs(b-a) * iEToFit^(1.0 - max(a,b))
  modelCDe <- nls(formula=model, data = dataTable, start = start,
                  control = nls.control(maxiter = 200, 
                                        tol = 1e-05, 
                                        minFactor = 1/1024, 
                                        printEval=FALSE, #Tells whether to print details of curve fit process.
                                        warnOnly=TRUE),
                  #Include the next 3 lines to fit with constraints.
                  #algorithm = "port",
                  #lower = list(lambda=-Inf, a=0, b=0),
                  #upper = list(lambda= Inf, a=1, b=1)
                  )
  return(modelCDe)
}

##############################
# Returns a Cobb-Douglas model where gamma (the exponent on the energy term) has been fixed
##
cdeFixedGammaModel <- function(countryAbbrev, energyType, gamma){
  dataTable <- loadData(countryAbbrev) #Load the data that we need.
  # We need to do the Cobb-Douglas fit with the desired energy data.
  # To achieve the correct fit, we'll change the name of the desired column
  # to "iEToFit" and use "iEToFit" in the nls function. 
  dataTable <- replaceColName(dataTable, energyType, "iEToFit")
  # We're fixing the value of gamma.
  # Calculate a new column iEGamma = iEToFit^gamma
  iEGamma <- with(dataTable, iEToFit^gamma)
  dataTable <- cbind(dataTable, iEGamma)
  # We already know the value of gamma. We don't want to estimate it.
  # Establish guess values for lambda and alpha.
  lambdaGuess <- 0.0 #guessing lambda = 0 means there is no technological progress.
  alphaGuess <- 0.7*(1-gamma) #ensure a consistent starting point.
  start <- list(lambda=lambdaGuess, alpha=alphaGuess)
  model <- iGDP ~ exp(lambda*iYear) * iCapStk^alpha * iLabor^((1-gamma) - alpha) * iEGamma
  modelCDe <- nls(formula=model, data = dataTable, start = start,
                  control = nls.control(maxiter = 200, 
                                        tol = 1e-05, 
                                        minFactor = 1/1024, 
                                        printEval=FALSE, #Tells whether to print details of curve fit process.
                                        warnOnly=TRUE),
                  #Include the next 3 lines to fit with constraints.
                  algorithm = "port",
                  lower = list(lambda=-Inf, alpha=0),
                  upper = list(lambda=Inf, alpha=1.0-gamma)
                  )
  return(modelCDe)
}

####################
# Returns an nls Cobb-Douglas model for the country specified
# Give an energyType ("Q", "X", or "U") if you want to include an energy term
# If you supply a value for the gamma argument, a fit with fixed gamma will be provided.
# This function dispatches to cdModel, cdeModel, or cdFixedGammaModel based on which arguments are specified.
##
cobbDouglasModel <- function(countryAbbrev, energyType, gamma){
  if (missing(energyType)){
    # Fit the Cobb-Douglas model without energy.
    return(cdModel(countryAbbrev))
  }
  if (!missing(gamma)){
    # Fit the Cobb-Douglas model with fixed value of gamma
    return(cdeFixedGammaModel(countryAbbrev, energyType, gamma))
  }
  # Fit the Cobb-Douglas model with gamma as a free parameter
  return(cdeModel(countryAbbrev, energyType))
}

###############################
# Does a grid search over values of gamma for a CD production function
##
cdeGridData <- function(countryAbbrev, energyType, gammaGrid){
  gamma <- gammaGrid
  models <- lapply(gamma, cobbDouglasModel, energyType=energyType, countryAbbrev=countryAbbrev)
  return(models)
}

#############################
# This function creates a series of graphs for Cobb-Douglas with energy functions
# showing various parameters as a function of gamma.
##
cdeGridGraphs <- function(countryAbbrev, energyType){
  gammaGrid <- seq(0.0, 0.95, by=0.05)
  #print(paste(countryAbbrev, energyType))
  models <- cdeGridData(countryAbbrev, energyType, gammaGrid)
  SSE <- lapply(models, function(model) summary(model)$sigma)
  lambda <- lapply(models, function(model) coef(model)["lambda"])
  alpha <- lapply(models, function(model) coef(model)["alpha"])
  data <- cbind(lambda, alpha, gammaGrid, SSE)
  plot(gammaGrid, SSE)
  plot(gammaGrid, lambda)
  plot(gammaGrid, alpha)
}

#########################
# Takes the Cobb-Douglas fitted models and creates per-country predictions for them.
# Returns a data.frame with the predictions.
##
cobbDouglasPredictions <- function(countryAbbrev, energyType){
  # Can't make predictions for any of CN, ZA, SA, IR, TZ, or ZM if we're interested in U
  if (!missing(energyType)){
    # Consider this replacement only if energyType has been specified.
    if (!(haveDataCD(countryAbbrev, energyType))){
      # If we don't have data for this combination of countryAbbrev and energyType, 
      # return a column of NAs when the above conditions have been met.
      nRows <- 21 # All of these countries need 21 rows.
      df <- as.data.frame(matrix(NA, ncol = 1, nrow = nRows))
      colnames(df) <- "pred"
      return(df)
    }
  }
  model <- cobbDouglasModel(countryAbbrev, energyType)
  pred <- predict(model) #See http://stackoverflow.com/questions/9918807/how-get-plot-from-nls-in-r
  df <- data.frame(pred)
  # Pad with rows as necessary
  df <- padRows(countryAbbrev, df)
  return(df)
}

#########################
# Takes the Cobb-Douglas fitted models and creates a single column of predicted GDP values
# that corresponds, row for row, with the AllData.txt file.
##
cobbDouglasPredictionsColumn <- function(energyType){
  if (missing(energyType)){
    out <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasPredictions)) #rbinds all the predictions together
    colnames(out) <- c("predGDP")
  } else {
    out <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasPredictions, energyType=energyType))
    colnames(out) <- c(paste("predGDP", energyType, sep=""))
  }
  return(out)
}

#################################################
# Calculates parameter estimates and confidence intervals
# for the Cobb-Douglas production function given a country.
#
# countryAbbrev is a string containing the 2-letter abbreviation for the country, e.g. "US" or "CN"
# energyType is a string, one of "Q", "X", or "U"
#
# returns a data.frame of data for the Cobb-Douglas model. 
# First row is the +95% CI on all parameters
# Second row contains the parameter estimates
# Third row is the -95% CI on all parameters
# Each column has names: lambda, alpha, beta, gamma, corresponding to the parameters in the model.
##
cobbDouglasData <- function(countryAbbrev, energyType){
  #First, check to see if we want useful work (U) AND one of the countries for which we don't have data.
  if (!haveDataCD(countryAbbrev, energyType)){
    #Return a column of NAs if the above conditions have been met.
    nRows <- 3 # +95% CI, CDe, and -95% CI.
    nCols <- 4 # lambda, alpha, beta, and gamma
    df <- as.data.frame(matrix(NA, ncol = nCols, nrow = nRows))
    colnames(df) <- c("lambda", "alpha", "beta", "gamma")
    rownames(df) <- c("+95% CI", "CDe", "-95% CI")
    return(df)
  }
  # We have a combination of country and energy type for which we have data.
  modelCD <- cobbDouglasModel(countryAbbrev, energyType)
  summaryCD <- summary(modelCD) # Gives the nls summary table.
  ciCD <- confint(modelCD, level = ciLevel)  # Calculates confidence intervals for the CD model.
  dofCD <- summaryCD$df[2] # Gives the degrees of freedom for the model.
  tvalCD <- qt(ciHalfLevel, df = dofCD)
  if (missing(energyType)){
    #Cobb-Douglas without energy
    alpha <- as.numeric(coef(modelCD)["alpha"])
    beta <- 1.0 - alpha
    beta.est <- deltaMethod(modelCD, "1 - alpha") # Estimates beta and its standard error (SE).
    # Calculate beta and its confidence interval and report it.
    # Now calculate a confidence interval on beta
    betaCICD <- with(beta.est, Estimate + c(-1.0, 1.0) * tvalCD * SE); betaCICD # Gives the confidence interval on beta.  
    # Combine all estimates and their confidence intervals into data frames with intelligent row names
    estCD <- data.frame(lambda=coef(modelCD)["lambda"], alpha=coef(modelCD)["alpha"], beta=beta, gamma=0)
    row.names(estCD) <- "CD"
    # The [1] subscripts pick off the lower confidence interval
    lowerCD <- data.frame(lambda=ciCD["lambda","2.5%"], alpha=ciCD["alpha", "2.5%"], beta=betaCICD[1], gamma=NA) 
    row.names(lowerCD) <- "-95% CI"
    # The [2] subscripts pick off the lower confidence interval
    upperCD <- data.frame(lambda=ciCD["lambda","97.5%"], alpha=ciCD["alpha", "97.5%"], beta=betaCICD[2], gamma=NA)
    row.names(upperCD) <- "+95% CI"
  } else {
    #Cobb-Douglas with energy
    a <- as.numeric(coef(modelCD)["a"])
    b <- as.numeric(coef(modelCD)["b"])
    lambda <- as.numeric(coef(modelCD)["lambda"])
    alpha <- a
    beta <- b - a
    gamma <- 1.0 - b
    # Report results with SE
    beta.est <- deltaMethod(modelCD, "b-a") # Reports results for beta, because beta = b - a.
    gamma.est <- deltaMethod(modelCD, "1-b") # Reports results for gamma, because gamma = 1 - b.
    # Now calculate confidence intervals.
    betaCICD <- with(beta.est, Estimate + c(-1.0, 1.0) * tvalCD * SE) # Gives the confidence interval on beta.
    gammaCICD <- with(gamma.est, Estimate + c(-1.0, 1.0) * tvalCD * SE) # Gives the confidence interval on gamma.
    # Combine all estimates and their confidence intervals into data frames with intelligent row names
    estCD <- data.frame(lambda = lambda, alpha = alpha, beta = beta, gamma = gamma)
    row.names(estCD) <- "CDe"
    # The [1] subscripts pick off the lower confidence interval
    lowerCD <- data.frame(lambda=ciCD["lambda","2.5%"], 
                          alpha=ciCD["a", "2.5%"], 
                          beta=betaCICD[1], 
                          gamma=gammaCICD[1])
    row.names(lowerCD) <- "-95% CI"
    # The [2] subscripts pick off the upper confidence interval
    upperCD <- data.frame(lambda=ciCD["lambda","97.5%"], 
                          alpha=ciCD["a", "97.5%"], 
                          beta=betaCICD[2], 
                          gamma=gammaCICD[2])
    row.names(upperCD) <- "+95% CI"
  }
  # Now create the data for a table and return it
  dataCD <- rbind(upperCD, estCD, lowerCD)
  return(dataCD)
}

############
# Creates a row for the Cobb Douglas parameters table for the given country (2-letter code) and energyType (Q, X, or U)
##
cobbDouglasCountryRow <- function(countryAbbrev, energyType){
  dataCD <- cobbDouglasData(countryAbbrev, energyType)
  if (missing(energyType)){
    out <- cbind(dataCD["-95% CI", "lambda"], dataCD["CD", "lambda"], dataCD["+95% CI", "lambda"],
                 dataCD["-95% CI", "alpha"],  dataCD["CD", "alpha"],  dataCD["+95% CI", "alpha"],
                 dataCD["-95% CI", "beta"],   dataCD["CD", "beta"],   dataCD["+95% CI", "beta"])
  } else {
    out <- cbind(dataCD["-95% CI", "lambda"], dataCD["CDe", "lambda"], dataCD["+95% CI", "lambda"],
                 dataCD["-95% CI", "alpha"],  dataCD["CDe", "alpha"],  dataCD["+95% CI", "alpha"],
                 dataCD["-95% CI", "beta"],   dataCD["CDe", "beta"],   dataCD["+95% CI", "beta"],
                 dataCD["-95% CI", "gamma"],  dataCD["CDe", "gamma"],  dataCD["+95% CI", "gamma"])
  }
  return(out)
}

###########################################
# Creates a number of rows in a data.frame that contain information 
# about the coefficients of a Cobb-Douglas model for countryName.
# Each parameter has its own row with confidence intervals.
# The country name is in a column. Which parameter is involved is
# also in a column.
# 
# The return type is a data.frame.
##
cobbDouglasCountryRowsForParamsGraph <- function(countryName, energyType){
  #Create three rows, one for each parameter. Each row is a data.frame so that it is plottable!
  if (missing(energyType)){
    valueRow <- "CD"
    dataCD <- cobbDouglasData(countryName)
  } else {
    valueRow <- "CDe"
    dataCD <- cobbDouglasData(countryName, energyType)
  }
  # These rows are common to the "with" and "without" energy cases.
  lambdaRow <- data.frame(country = countryName, 
                          parameter = "lambda", 
                          lowerCI = dataCD["-95% CI", "lambda"], 
                          value = dataCD[valueRow, "lambda"], 
                          upperCI = dataCD["+95% CI", "lambda"])
  alphaRow <- data.frame(country = countryName, 
                         parameter = "alpha",
                         lowerCI = dataCD["-95% CI", "alpha"],
                         value = dataCD[valueRow, "alpha"],  
                         upperCI = dataCD["+95% CI", "alpha"])
  betaRow <- data.frame(country = countryName, 
                        parameter = "beta",
                        lowerCI = dataCD["-95% CI", "beta"],
                        value = dataCD[valueRow, "beta"], 
                        upperCI = dataCD["+95% CI", "beta"])
  if (missing(energyType)){
    table <- rbind(lambdaRow, alphaRow, betaRow)
  } else {
    gammaRow <- data.frame(country = countryName, 
                           parameter = "gamma",
                           lowerCI = dataCD["-95% CI", "gamma"],
                           value = dataCD[valueRow, "gamma"], 
                           upperCI = dataCD["+95% CI", "gamma"])
    table <- rbind(lambdaRow, alphaRow, betaRow, gammaRow)
  }
  return(table)
}

############################
# Aggregates the Cobb-Douglas results into a big data table. No energy.
##
cobbDouglasParamsTableNoEnergy <- function(){
  #Do rbind on the results of creating a row in the table for every country abbreviation that we know.
  dataCD <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasCountryRow))
  colnames(dataCD) <- c(" ", "$\\lambda$", " ", 
                        " ", "$\\alpha$",  " ",
                        " ", "$\\beta$",   " ")
  rownames(dataCD) <- countryAbbrevs
  tableCD <- xtable(dataCD, 
                    caption="Cobb-Douglas model (without energy) parameters for 1980-2011 (US, UK, JP), 1991-2010 (CN and ZA), and 1991-2011 (SA, IR, TZ, and ZM). (Parameter estimates beneath symbol. 95\\% confidence interval bounds to left and right.)", 
                    label="tab:CD_Parameters_No_Energy",
                    digits = c(1, 4,4,4, 2,2,2, 2,2,2),
                    align = "r|ccc|ccc|ccc") #Sets alignment of the numbers in the columns
  return(tableCD)
}

############################
# Aggregates the Cobb-Douglas results into a big data table for the given energyType.
##
cobbDouglasParamsTableWithEnergy <- function(energyType){
  #Do rbind on the results of creating a row in the table for every country abbreviation that we know.
  dataCD <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasCountryRow, energyType=energyType))
  colnames(dataCD) <- c(" ", "$\\lambda$", " ", 
                        " ", "$\\alpha$",  " ",
                        " ", "$\\beta$",   " ",
                        " ", "$\\gamma$",  " ")
  rownames(dataCD) <- countryAbbrevs
  tableCD <- xtable(dataCD, 
                    caption=paste("Cobb-Douglas model (with $", tolower(energyType), "$) parameters for 1980-2011 (US, UK, JP), 1991-2010 (CN and ZA), and 1991-2011 (SA, IR, TZ, and ZM). (Parameter estimates beneath symbol. 95\\% confidence interval bounds to left and right.)", sep=""), 
                    label=paste("tab:CD_Parameters_With_", energyType, sep=""),
                    digits = c(1, 4,4,4, 2,2,2, 2,2,2, 2,2,2),
                    align = "r|ccc|ccc|ccc|ccc") #Sets alignment of the numbers in the columns
  return(tableCD)
}

############################
# Prints a table with parameters from a Cobb-Douglas model for the given energyType. 
# Leave energyType blank to print a table for Cobb-Douglas without energy.
##
printCDParamsTable <- function(energyType){
  if (missing(energyType)){
    print(cobbDouglasParamsTableNoEnergy(), 
        caption.placement="top", 
        sanitize.colnames.function = identity, 
        size="\\tiny", 
        table.placement="H")
  } else {
    print(cobbDouglasParamsTableWithEnergy(energyType), 
      caption.placement="top", 
      sanitize.colnames.function = identity, 
      size="\\tiny",
      table.placement="H")
  }
}

#############################
# Creates a graph with confidence intervals for the Cobb-Douglas model for the given energyType
##
createCDParamsGraph <- function(energyType){
  if (missing(energyType)){
    # Create a data table with the following columns:
    # country abbrev, parameter (lambda, alpha, or beta), -95% CI, value, +95% CI
    dataTable <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasCountryRowsForParamsGraph))
    graph <- segplot(country ~ upperCI + lowerCI | parameter, 
                     data = dataTable, 
                     centers = value, #identifies where the dots should be placed
                     draw.bands = FALSE, #provides nicer error bars
                     horizontal = FALSE, #makes error bars vertical and puts the countries in the x axis
                     layout = c(2,2), #2 column, 2 row
                     index.cond = list(c(2,3,1)), #orders the panels as alpha, beta, lambda, gamma
                     strip = strip.custom(factor.levels=c("$\\lambda$", "$\\alpha$", "$\\beta$"), #sets labels in strip 
                                          bg="white"), #sets background color in strip
                     col = "black", #Sets line color to black
                     lwd = 1, #Sets line width to 1.0
                     ylim = list(c(-0.05, 0.1), c(-0.5, 1.5), c(-0.5, 1.5)), #y axis limits
                     scales = list(cex=scaleTextSize, #controls text size on scales.
                                   tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                                   x=list(cex=0.75), #reduces text size so that country abbrevs are legible
                                   y=list(relation="free", #allow each axis to be different
                                          at=list(c(0.0, 0.05, 0.1),   #y tick mark for lambda
                                                  c(0.0, 0.5, 1.0), 
                                                  c(0.0, 0.5, 1.0)
                                                  )
                                          )
                                   )
                     )
   } else {
    # Create a data table with the following columns:
    # country abbrev, parameter (lambda, alpha, or beta), -95% CI, value, +95% CI
    dataTable <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasCountryRowsForParamsGraph, energyType=energyType))
    graph <- segplot(country ~ upperCI + lowerCI | parameter, 
                     data = dataTable, 
                     centers = value, #identifies where the dots should be placed
                     draw.bands = FALSE, #provides nicer error bars
                     horizontal = FALSE, #makes error bars vertical and puts the countries in the x axis
                     layout = c(2,2), #2 column, 2 row
                     index.cond = list(c(2,3,1,4)), #orders the panels as alpha, beta, lambda, gamma
                     #set labels and bg color in strip
                     strip = strip.custom(factor.levels=c("$\\lambda$", "$\\alpha$", "$\\beta$", "$\\gamma$"),  
                                          bg="white"),
                     col = "black", #Sets line color to black
                     lwd = 1, #Sets line width to 1.0
                     ylim = list(c(-0.05, 0.1), c(-0.5, 1.5), c(-0.5, 1.5), c(-0.5, 1.5)), #y axis limits
                     scales = list(cex=scaleTextSize, #controls text size on scales.
                                   tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                                   x=list(cex=0.75), #reduces text size so that country abbrevs are legible
                                   y=list(relation="free", #allow each axis to be different
                                          at=list(c(0.0, 0.05, 0.1),   #y tick mark for lambda
                                                  c(0.0, 0.5, 1.0), 
                                                  c(0.0, 0.5, 1.0),
                                                  c(0.0, 0.5, 1.0)
                                                  )
                                          )
                                   )
                     )
   }
   return(graph)
}

##############################
# Creates a graph that plots predicted GDP as lines and GDP data as open circles.
##
createCDLatticeGraph <- function(countryAbbrev, textScaling = 1.0, keyXLoc = defaultKeyXLoc, keyYLoc = defaultKeyYLoc){
  dataTable <- loadData("All") #Grab the raw data
  predictions <- cobbDouglasPredictionsColumn()     #Predictions from CD without energy
  predictionsQ <- cobbDouglasPredictionsColumn("Q") #Predictions from CD with Q
  predictionsX <- cobbDouglasPredictionsColumn("X") #Predictions from CD with X
  predictionsU <- cobbDouglasPredictionsColumn("U") #Predictions from CD with U
  #Now add the predictions columns to the data.
  dataTable <- cbind(dataTable, predictions, predictionsQ, predictionsX, predictionsU) 
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1, 2, 4, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 2, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "black", "red", "blue", "darkorange") #line and symbol colors
  symbols <- c(1, NA, NA, NA, NA)  #NA gives no symbol.
  # Code that deals with items that are specific to whether we want all countries or a specific country.  
  if (missing (countryAbbrev)){
    # We want a graph with panels for all countries
    yLimits <- yLimitsForGDPGraphs
    factorLevels <- countryNamesAlph # We want all countries shown
    indexCond <- list(countryOrderForGraphs) # We want all countries in this order
    layout <- ninePanelLayoutSpec # Show all countries
  } else {
    # We want only a specific country
    dataTable <- subset(dataTable, Country == countryAbbrev)    
    # Select the correct y limits
    index <- which(countryAbbrevsAlph %in% countryAbbrev)
    # The following lines use [index:index] as a convenient way of subsetting.
    # This has the added benefit of maintaining the correct classes for things.
    yLimits <- yLimitsForGDPGraphs[index:index]       # Pick limits for the country we want.
    factorLevels <- countryNamesAlph[index:index]     # Only show the country we have chosen.
    indexCond <- list(c(1))                           # We want only one country.
    layout <- onePanelLayoutSpec                      # We want only one panel in the graph.
  }  
  graph <- xyplot(iGDP+predGDP+predGDPQ+predGDPX+predGDPU ~ Year | Country, data=dataTable,
                  type=graphType,
                  index.cond=indexCond, #orders the panels.
                  layout=layout, #indicates a 3x3 arrangement of panels.
                  strip = strip.custom(factor.levels=factorLevels, 
                                       bg="transparent", # Sets background transparent to match the graph itself.
                                       par.strip.text=list(cex=textScaling) # Scales text in the strip.
                                       ),
                  as.table=TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("Actual", "No energy", "With $q$", "With $x$", "With $u$")),
                           type=graphType,
                           cex=keyTextSize * textScaling, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize * textScaling, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              x=list(at=timeTics),
                              y=list(relation="free",  #allows each axis to be different
                                     at=yTicsForIndexedGraphs)), #specifies location for tics
                  ylim=yLimits, #y axis limits
                  #axis labels and scaling
                  xlab=list(label="", cex=textScaling), 
                  ylab=list(label="Indexed (1980=1 or 1991=1)", cex=textScaling)
                  )
  return(graph)
}
@

<<ces functions, eval=TRUE>>=
# Set some common names for CES variables.
xNames = c("iCapStk", "iLabor", "iEToFit")
tName = "iYear"
yName = "iGDP"

######################################
# Loads data and trims rows where necessary.
# Also, replaces the name of the desired energy column with "iEToFit".
# Returns a data.frame with the loaded data.
##
loadAndPrepDataForCES <- function(countryAbbrev, energyType){
  dataTable <- loadData(countryAbbrev)
  if (energyType == "U"){
    # Trim the dataset to include only those years for which U is available.
    dataTable <- subset(dataTable, !is.na(iU))
  }
  # Trim data from China and South Africa that is missing labor information for 2011.
  dataTable <- subset(dataTable, !is.na(iLabor))
  # We need to do the CES fit with the desired energyType.
  # To achieve the correct fit, we'll change the name of the desired column
  # to "iEToFit" and use "iEToFit" in the nls function.
  dataTable <- replaceColName(dataTable, energyType, "iEToFit")
  return(dataTable)
}

####################
# Returns a cesEst model for the country and energyType specified.
# energyType should be one of Q", "X", or "U".
##
cesModel <- function(countryAbbrev, energyType){
  # Load the data that we need.
  dataTable <- loadAndPrepDataForCES(countryAbbrev, energyType)
  control=nls.lm.control(maxiter=1000, maxfev=2000)
  if (countryAbbrev == "US") {
    # With Q, The unconstrained optimization leads to rho1 = -0.72 and sigma1 = 3.57, which is not 
    # economically meaningful.  So, a grid search was performed, and it confirmed that low values of 
    # rho1 yield a better fit. In fact, the lower the value of rho1, the better. So, we'll set it 
    # here at rho1=0.0, which makes the elasticity of substitution between k and l sigma1 = 1.0, 
    # giving the Cobb-Douglas form for the k l portion of the CES function. Similar results are found
    # for X and U.
    rho1 = 0.0
    # Unconstrained optimization leads to rho = 25.7 and sigma = 0.0374, indicating that (kl) and (e) are 
    # complimentary.  We'll let the estCES function solve for the best value.
    modelCES <- cesEst(data=dataTable, yName=yName, xNames=xNames, tName=tName, control=control,
                       method="LM",
                       rho1=rho1
                       )
  } else if (countryAbbrev == "UK") {
    # The unconstrained optimization leads to rho1 = -2.4 and sigma1 = -0.71, which is not economically meaningful.
    # So, a grid search was performed, and it confirmed that low values of rho1 yield a better fit. In fact, the 
    # lower the value of rho1, the better. So, we'll set it here at rho1=0.0, which makes the elasticity of 
    # substitution between k and l sigma1 = 1.0, giving the Cobb-Douglas form for the k l portion of the CES function.
    # rho1 = c(0.00, 0.01, 0.02)
    rho1 = 0.0
    modelCES <- cesEst(data=dataTable, yName=yName, xNames=xNames, tName=tName, control=control,
                       method="LM",
                       rho1=rho1
                       )
  } else if (countryAbbrev == "JP") {
    # The unconstrained optimization leads to rho1 = 7.99 and sigma1 = 0.11, indicating complementarity between k & l,
    # but different from the results for the US and UK (where k & l were found to be substitutes).  However, the
    # p value for rho1 is very high (0.3), indicating that rho1 is not making a significant contribution to the model. 
    # The unconstrained fit gives R^ = 0.996 and sigma = 0.103, indicating complementarity between (kl) and (e).
    # 
    # In fact, we could force rho1 = 0.0 and achieve rho = 280 and sigma = 0.00356 with R^2 = 0.988, a very small
    # decrease, indeed.  Note that both the constrained and unconstrained fits give small sigma, indicating
    # that (kl) and (e) are complimentary.
    modelCES <- cesEst(data=dataTable, yName=yName, xNames=xNames, tName=tName, control=control,
                       method="LM"
                       )
  } else if (countryAbbrev == "CN"){
    # Unconstrained optimization: rho1 < -0.2290, rho = 14.1651, R^2 = 0.9984. This corresponds to sigma1 = 1.297
    # and sigma = 0.0659, indicating that (kl) and (e) are complements and k & l are substitutes.  However, none of the
    # rho or sigma values are statistically significant.
    # Now, sigma1 > 1 is not economically meaningful, so we'll set rho1 = 0 to force sigma1 = 1.0. When we do that, 
    # we obtain sigma1 = 1.0 (so perfect substitution between k and l) and sigma = 7.19 (again, not economically
    # meaningful) with R^2 = 0.9983.  Again, none of the rho or sigma values are statistically significant.
    # Furthermore, the only significant varluable is gamma = 1.0731.
    # So, for the final fit, I'll put rho1 = 0.0 and rho = 0.0. The final R^2 = 0.9983.
    rho1 = 0.0
    rho = 0.0
    modelCES <- cesEst(data=dataTable, yName=yName, xNames=xNames, tName=tName, control=control,
                       method="LM",
                       rho1=rho1,
                       rho=rho                     
                       )
  } else if (countryAbbrev == "ZA"){
    # An unconstrained fit works for South Africa.  Results are sigma1 = 0.05, signalling complementarity
    # between k and l.  sigma = 0.03, indicating complementatity between (kl) and (e).
    modelCES <- cesEst(data=dataTable, yName=yName, xNames=xNames, tName=tName, control=control,
                       method="LM",
                       )
  } else if (countryAbbrev == "SA"){
    # An unconstrained fit gives rho = -6.33, rho1 = 9.36 (sigma1 = 0.0965) and R^2 = 0.9892.  
    # This result is not economically meaningful.
    # Constraining rho = 0 gives rho1 = -582, again not economically meaningful.
    # So, do a grid search over meaningful regions to see what is best.
    # The result (to with +/- 0.1) is rho = 0.0 and rho1 = 5.7, with R^2 = 0.9885.
    # However, neither the the rho or nor the sigma values are statistically significant.
    rho1 = 5.7
    rho = 0.0
    modelCES <- cesEst(data=dataTable, yName=yName, xNames=xNames, tName=tName, control=control,
                       method="LM",
                       rho1 = rho1,
                       rho = rho
                       )
  } else if (countryAbbrev == "IR"){
    # Iran works with an unconstrained fit.
    modelCES <- cesEst(data=dataTable, yName=yName, xNames=xNames, tName=tName, control=control,
                       method="LM",
                       )
  } else if (countryAbbrev == "TZ"){
    # Using the PORT algorithm yields convergence and R^2 = 0.9974 with economically menaingful results.
    # However, none of the rho or sigma values are statistically significant.
    modelCES <- cesEst(data=dataTable, yName=yName, xNames=xNames, tName=tName, control=control,
                       method="PORT",
                       )
  } else if (countryAbbrev == "ZM"){
    # ZM works with the PORT algorithm, but not the LM algorithm.
    modelCES <- cesEst(data=dataTable, yName=yName, xNames=xNames, tName=tName, control=control,
                       method="PORT",
                       rho=0.0
                       )
  }
  return(modelCES)
}

#########################
# Takes the CES fitted models and creates per-country predictions for them.
# Returns a data.frame with the predictions.
##
cesPredictions <- function(countryAbbrev, energyType){
  dataTable <- loadAndPrepDataForCES(countryAbbrev, energyType)
  # Can't make predictions for any of CN, ZA, SA, IR, TZ, or ZM if we're interested in U
  if (!(haveDataSF(countryAbbrev, energyType))){
    # If we don't have data for this combination of countryAbbrev and energyType, 
    # return a column of NAs if the above conditions have been met.
    nRows <- 21 # All of these countries need 21 rows.
    df <- as.data.frame(matrix(NA, ncol = 1, nrow = nRows))
    colnames(df) <- "pred"
    return(df)
  }
  model <- cesModel(countryAbbrev, energyType)
  pred <- fitted(model)
  df <- data.frame(pred)
  # Pad with rows as necessary
  df <- padRows(countryAbbrev, df)
  return(df)
}

#########################
# Takes the CES fitted models and creates a single column of predicted GDP values
# that corresponds, row for row, with the AllData.txt file.
##
cesPredictionsColumn <- function(energyType){
  out <- do.call("rbind", lapply(countryAbbrevs, cesPredictions, energyType=energyType))
  colnames(out) <- c(paste("predGDP", energyType, sep=""))
  return(out)
}

##############################
# Creates a graph that plots predicted GDP as lines, one for each single factor, and historical GDP 
# data as open circles.
##
createCESLatticeGraph <- function(countryAbbrev, textScaling = 1.0, keyXLoc = defaultKeyXLoc, keyYLoc = defaultKeyYLoc){
  dataTable <- loadData("All") #Grab the raw data
  predictionsQ <- cesPredictionsColumn("Q") #Predictions from CES with Q
  predictionsX <- cesPredictionsColumn("X") #Predictions from CES with X
  predictionsU <- cesPredictionsColumn("U") #Predictions from CES with U
  #Now add the predictions columns to the data.
  dataTable <- cbind(dataTable, predictionsQ, predictionsX, predictionsU) 
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 2, 4, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "red", "blue", "darkorange") #line and symbol colors
  symbols <- c(1, NA, NA, NA)  #NA gives no symbol.
  # Code that deals with items that are specific to whether we want all countries or a specific country.  
  if (missing(countryAbbrev)){
    # We want a graph with panels for all countries
    yLimits <- yLimitsForGDPGraphs
    factorLevels <- countryNamesAlph # We want all countries shown
    indexCond <- list(countryOrderForGraphs) # We want all countries in this order
    layout <- ninePanelLayoutSpec # Show all countries
  } else {
    # We want only a specific country
    dataTable <- subset(dataTable, Country == countryAbbrev)    
    # Select the correct y limits
    index <- which(countryAbbrevsAlph %in% countryAbbrev)
    # The following lines use [index:index] as a convenient way of subsetting.
    # This has the added benefit of maintaining the correct classes for things.
    yLimits <- yLimitsForGDPGraphs[index:index]       # Pick limits for the country we want.
    factorLevels <- countryNamesAlph[index:index]     # Only show the country we have chosen.
    indexCond <- list(c(1))                           # We want only one country.
    layout <- onePanelLayoutSpec                      # We want only one panel in the graph.
  }
  graph <- xyplot(iGDP+predGDPQ+predGDPX+predGDPU ~ Year | Country, data=dataTable,
                  type=graphType,
                  index.cond=indexCond, #orders the panels.
                  layout=layout,
                  strip = strip.custom(factor.levels=factorLevels, 
                                       bg="transparent", # Sets background transparent to match the graph itself.
                                       par.strip.text=list(cex=textScaling) # Scales text in the strip.
                                       ),
                  as.table=TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("Actual", "With $q$", "With $x$", "With $u$")),
                           type=graphType,
                           cex=keyTextSize * textScaling, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize * textScaling, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              x=list(at=timeTics),
                              y=list(relation="free",  #allows each axis to be different
                                     at=yTicsForIndexedGraphs)), #specifies location for tics
                  ylim=yLimits, #y axis limits
                  #axis labels and scaling
                  xlab=list(label="", cex=textScaling), 
                  ylab=list(label="Indexed (1980=1 or 1991=1)", cex=textScaling)
                  )
  return(graph)
}

#################################################
# Calculates parameter estimates and confidence intervals
# for the CES production function given a country and an energyType.
#
# countryAbbrev is a string containing the 2-letter abbreviation for the country, e.g. "US" or "CN"
# energyType is a string, one of "Q", "X", or "U"
#
# returns a data.frame of data for the LINEX model.
# First row is the +95% CI on all parameters
# Second row contains the parameter estimates
# Third row is the -95% CI on all parameters
# Each column has names: gamma, lambda, delta_1, delta, rho_1, rho
##
cesData <- function(countryAbbrev, energyType){
  #First, check to see if we want useful work (U) AND one of the countries for which we don't have data.
  if (!haveDataCD(countryAbbrev, energyType)){
    #Return a column of NAs if the above conditions have been met.
    nRows <- 3 # +95% CI, CDe, and -95% CI.
    nCols <- 6 # gamma, lambda, delta_1, delta, sigma_1, and sigma
    df <- as.data.frame(matrix(NA, ncol = nCols, nrow = nRows))
    colnames(df) <- c("gamma", "lambda", "delta_1", "delta", "sigma_1", "sigma")
    rownames(df) <- c("+95% CI", "CES", "-95% CI")
    return(df)
  }
  # We have a combination of country and energy type for which we have data.
  modelCES <- cesModel(countryAbbrev, energyType)
  summaryCES <- summary(modelCES) # Gives a summary table.
#print(summaryCES)
  # Calculate the degrees of freedom as N_obs - N_params
  dofCES <- nrow(loadAndPrepDataForCES(countryAbbrev, energyType)) - 6 # Gives the degrees of freedom for the model.
#print(dofCES)
  tvalCES <- qt(ciHalfLevel, df = dofCES)
#print(tvalCES)
  # Calculate confidence intervals for the parameters of the model
  ciCES <- confint(modelCES, level = ciLevel)  # Calculates confidence intervals for the CES model.
#print(ciCES)
  # Get confidence intervals for the elasticities of substitution
  sigma_1.est <- deltaMethod(modelCES, "1.0 / (1.0 + rho_1)") # Estimates sigma_1 and its standard error (SE).
#print(sigma_1.est)
  sigma_1CI <- with(sigma_1.est, Estimate + c(-1.0, 1.0) * tvalCES * SE)
  # There are some cases where the confidence intervals for sigma_1 are huge. Set them to -Inf or Inf as appropriate
  if (is.nan(sigma_1CI[1]) || (sigma_1CI[1] < -1.0e10)){
    sigma_1CI[1] <- -Inf 
  }
  if (is.nan(sigma_1CI[2]) || (sigma_1CI[2] > 1.0e10)){
    sigma_1CI[2] <- Inf
  }
#print("sigma_1CI:")
#print(sigma_1CI)
  sigma.est <- deltaMethod(modelCES, "1.0 / (1.0 + rho)") # Estimates sigma and its standard error (SE)
#print(sigma.est)
  sigmaCI <- with(sigma.est, Estimate + c(-1.0, 1.0) * tvalCES * SE)
#print("sigmaCI:")
#print(sigmaCI)
  # Report results with SE
  gamma <- coef(modelCES)["gamma"] #as.numeric(coef(modelCES)["gamma"])
  lambda <- coef(modelCES)["lambda"] #as.numeric(coef(modelCES)["lambda"])
  delta_1 <- coef(modelCES)["delta_1"]
  delta <- coef(modelCES)["delta"]
  sigma_1 <- sigma_1.est$Estimate
  sigma <- sigma.est$Estimate
  # Combine all estimates and their confidence intervals into data frames with intelligent row names
  estCES <- data.frame(gamma=gamma, lambda=lambda, delta_1=delta_1, delta=delta, sigma_1=sigma_1, sigma=sigma)
  row.names(estCES) <- "CES"
#print(estCES)
  # The [1] subscripts pick off the lower confidence interval
#print(ciCES["gamma","2.5 %"])
  lowerCES <- data.frame(gamma=ciCES["gamma","2.5 %"], lambda=ciCES["lambda", "2.5 %"],
                         delta_1=ciCES["delta_1", "2.5 %"], delta=ciCES["delta", "2.5 %"],
                         sigma_1=sigma_1CI[1], sigma=sigmaCI[1])
  row.names(lowerCES) <- "-95% CI"
#print(lowerCES)
  # The [2] subscripts pick off the upper confidence interval
  upperCES <- data.frame(gamma=ciCES["gamma","97.5 %"], lambda=ciCES["lambda", "97.5 %"],
                         delta_1=ciCES["delta_1", "97.5 %"], delta=ciCES["delta", "97.5 %"],
                         sigma_1=sigma_1CI[2], sigma=sigmaCI[2])
  row.names(upperCES) <- "+95% CI"
#print(upperCES)
  # Now create the data for a table and return it
  dataCES <- rbind(upperCES, estCES, lowerCES)
#print("dataCES")
#print(dataCES)
  return(dataCES)
}

############
# Creates a row for the CES parameters table for the given country (2-letter code) and energyType (Q, X, or U)
##
cesCountryRow <- function(countryAbbrev, energyType){
  dataCES <- cesData(countryAbbrev, energyType)
  out <- cbind(dataCES["-95% CI", "gamma"], dataCES["CES", "gamma"], dataCES["+95% CI", "gamma"],
               dataCES["-95% CI", "lambda"], dataCES["CES", "lambda"], dataCES["+95% CI", "lambda"],
               dataCES["-95% CI", "delta_1"], dataCES["CES", "delta_1"], dataCES["+95% CI", "delta_1"],
               dataCES["-95% CI", "delta"], dataCES["CES", "delta"], dataCES["+95% CI", "delta"],
               dataCES["-95% CI", "sigma_1"], dataCES["CES", "sigma_1"], dataCES["+95% CI", "sigma_1"],
               dataCES["-95% CI", "sigma"], dataCES["CES", "sigma"], dataCES["+95% CI", "sigma"]
               )
  return(out)
}

###########################################
# Creates a number of rows in a data.frame that contain information 
# about the coefficients of a CES model for countryName and energyType
# Each parameter has its own row with confidence intervals.
# The country name is in a column. Which parameter is involved is
# also in a column.
# 
# The return type is a data.frame.
##
cesCountryRowsForParamsGraph <- function(countryName, energyType){
  #Create six rows, one for each parameter. Each row is a data.frame so that it is plottable!
  dataCES <- cesData(countryName, energyType)
  gammaRow <- data.frame(country = countryName, 
                          parameter = "gamma", 
                          lowerCI = dataCES["-95% CI", "gamma"], 
                          value = dataCES["CES", "gamma"], 
                          upperCI = dataCES["+95% CI", "gamma"])
  lambdaRow <- data.frame(country = countryName, 
                         parameter = "lambda",
                         lowerCI = dataCES["-95% CI", "lambda"],
                         value = dataCES["CES", "lambda"],  
                         upperCI = dataCES["+95% CI", "lambda"])
  delta_1Row <- data.frame(country = countryName, 
                         parameter = "delta_1",
                         lowerCI = dataCES["-95% CI", "delta_1"],
                         value = dataCES["CES", "delta_1"],  
                         upperCI = dataCES["+95% CI", "delta_1"])
  deltaRow <- data.frame(country = countryName, 
                         parameter = "delta",
                         lowerCI = dataCES["-95% CI", "delta"],
                         value = dataCES["CES", "delta"],  
                         upperCI = dataCES["+95% CI", "delta"])
  sigma_1Row <- data.frame(country = countryName, 
                         parameter = "sigma_1",
                         lowerCI = dataCES["-95% CI", "sigma_1"],
                         value = dataCES["CES", "sigma_1"],  
                         upperCI = dataCES["+95% CI", "sigma_1"])
  sigmaRow <- data.frame(country = countryName, 
                         parameter = "sigma",
                         lowerCI = dataCES["-95% CI", "sigma"],
                         value = dataCES["CES", "sigma"],  
                         upperCI = dataCES["+95% CI", "sigma"])
  table <- rbind(gammaRow, lambdaRow, delta_1Row, deltaRow, sigma_1Row, sigmaRow)
  return(table)
}

############################
# Aggregates the CES results for gamma, delta_1, and delta into a table for the given energyType.
##
cesParamsTableA <- function(energyType){
  dataCES <- do.call("rbind", lapply(countryAbbrevs, cesCountryRow, energyType=energyType))
  colnames(dataCES) <- c(" ", "$\\gamma$",    " ", 
                         " ", "$\\lambda$",   " ",
                         " ", "$\\delta_1$",  " ",
                         " ", "$\\delta$",    " ",
                         " ", "$\\sigma_1$",  " ",
                         " ", "$\\sigma$",    " ")
  rownames(dataCES) <- countryAbbrevs
  tableCESa <- xtable(dataCES[,c(1,2,3, 7,8,9, 10,11,12)], #Picks up gamma, delta_1, and delta
                      caption=paste("CES model (with $", tolower(energyType), "$). $\\gamma$, $\\delta_1$, and $\\delta$ parameters for 1980-2011 (US, UK, JP), 1991-2010 (CN and ZA), and 1991-2011 (SA, IR, TZ, and ZM). (Parameter estimates beneath symbol. 95\\% confidence interval bounds to left and right.)", sep=""), 
                      label=paste("tab:CES_Parameters_With_", energyType, sep=""),
                      digits = c(1, 2,2,2, 2,2,2, 2,2,2),
                      align = "r|ccc|ccc|ccc" #Sets alignment of the numbers in the columns
                      ) 
  return(tableCESa)
}

############################
# Aggregates the CES results for lambda, sigma_1, and sigma into a table for the given energyType.
##
cesParamsTableB <- function(energyType){
  dataCES <- do.call("rbind", lapply(countryAbbrevs, cesCountryRow, energyType=energyType))
  colnames(dataCES) <- c(" ", "$\\gamma$",    " ", 
                         " ", "$\\lambda$",   " ",
                         " ", "$\\delta_1$",  " ",
                         " ", "$\\delta$",    " ",
                         " ", "$\\sigma_1$",  " ",
                         " ", "$\\sigma$",    " ")
  rownames(dataCES) <- countryAbbrevs
  tableCESb <- xtable(dataCES[,c(4,5,6, 13,14,15, 16,17,18)], #Picks up lambda, sigma_1, and sigma
                      caption=paste("CES model (with $", tolower(energyType), "$). $\\lambda$, $\\sigma_1$, and $\\sigma$ parameters for 1980-2011 (US, UK, JP), 1991-2010 (CN and ZA), and 1991-2011 (SA, IR, TZ, and ZM). (Parameter estimates beneath symbol. 95\\% confidence interval bounds to left and right.)", sep=""), 
                      label=paste("tab:CES_Parameters_With_", energyType, sep=""),
                      digits = c(1, 4,4,4, 2,2,2, 2,2,2),
                      align = "r|ccc|ccc|ccc"  #Sets alignment of the numbers in the columns
                      )
  return(tableCESb)
}

#############################
# Creates a graph with confidence intervals for the CES model for the given energyType
##
createCESParamsGraph <- function(energyType){
  # Create a data table with the following columns:
  # country abbrev, parameter (gamma, lambda, delta_1, delta, sigma_1, sigma), -95% CI, value, +95% CI
  dataTable <- do.call("rbind", lapply(countryAbbrevs, cesCountryRowsForParamsGraph, energyType=energyType))
  graph <- segplot(country ~ upperCI + lowerCI | parameter, 
                   data = dataTable, 
                   centers = value, #identifies where the dots should be placed
                   as.table = TRUE, #indexing of panels starts in upper left and goes across rows.
                   draw.bands = FALSE, #provides nicer error bars
                   horizontal = FALSE, #makes error bars vertical and puts the countries in the x axis
                   layout = c(2,3), #2 column, 3 row
                   # Orders the panels as gamma, lambda, delta_1, delta, sigma_1, sigma
                   index.cond = list(c(1,2,3,4,5,6)), 
                   #set labels and bg color in strip
                   strip = strip.custom(factor.levels=c("$\\gamma$", "$\\lambda$", "$\\delta_1$",
                                                        "$\\delta$", "$\\sigma_1$", "$\\sigma$"), bg="white"),
                   col = "black", #Sets line color to black
                   lwd = 1, #Sets line width to 1.0
                   ylim = list(c(0.9, 1.1), c(-0.05, 0.05), c(0.0, 1.0),
                               c(0.0, 1.0), c(0.0, 1.0), c(0.0, 1.0)), #y axis limits
                   scales = list(cex=scaleTextSize, #controls text size on scales.
                                 tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                                 x=list(cex=0.75), #reduces text size so that country abbrevs are legible
                                 y=list(relation="free", #allow each axis to be different
                                        at=list(c(0.9, 1.0, 1.1), c(-0.05, 0.0, 0.05),
                                                c(0.0, 0.5, 1.0), c(0.0, 0.5, 1.0), 
                                                c(0.0, 0.5, 1.0), c(0.0, 0.5, 1.0))   #y tick marks
                                       )
                   )
  )
  return(graph)
}

############################
# Prints a table with gamma, delta_1, and delta parameters from a CES model for the given energyType. 
##
printCESParamsTableA <- function(energyType){
  print(cesParamsTableA(energyType), 
    caption.placement="top", 
    sanitize.colnames.function = identity, 
    size="\\tiny",
    table.placement="H")
}

############################
# Prints a table with lambda, sigma_1, and sigma parameters from a CES model for the given energyType. 
##
printCESParamsTableB <- function(energyType){
  print(cesParamsTableB(energyType), 
    caption.placement="top", 
    sanitize.colnames.function = identity, 
    size="\\tiny",
    table.placement="H")
}
@

<<LINEX functions, eval=TRUE>>=
####################
# Returns an nls linex model for the country and energyType specified.
# energyType should be one of Q", "X", or "U".
##
linexModel <- function(countryAbbrev, energyType){
  # Load the data that we need.
  dataTable <- loadData(countryAbbrev)
  # We need to do the CES fit with the desired energyType.
  # To achieve the correct fit, we'll change the name of the desired column
  # to "iEToFit" and use "iEToFit" in the nls function.
  dataTable <- replaceColName(dataTable, energyType, "iEToFit")
  a_0Guess <- 0.5
  c_tGuess <- 1.0
  if (countryAbbrev == "SA"){
    # Need adjusted guess values, becasue k and l are above GDP for SA.
    a_0Guess <- -1
    c_tGuess <- -6

  }
  # Runs a non-linear least squares fit to the data with constraints
  modelLINEX <- nls(iGDP ~ iEToFit * exp(a_0*(2.0 - (iLabor+iEToFit)/iCapStk) + a_0 * c_t *(iLabor/iEToFit - 1.0)), 
                  data=dataTable,
                  start=list(a_0=a_0Guess, c_t=c_tGuess),
                  control=nls.control(maxiter=50, tol=1e-06, minFactor=1/1024, printEval=FALSE, warnOnly=TRUE),
#                   algorithm = "port",
#                   lower = list(a_0=-INF, c_t=-INF),
#                   upper = list(a_0= INF, c_t= INF)
                  )
  return(modelLINEX)
}

#########################
# Takes the LINEX fitted models and creates per-country predictions for them.
# Returns a data.frame with the predictions.
##
linexPredictions <- function(countryAbbrev, energyType){
  # Can't make predictions for any of CN, ZA, SA, IR, TZ, or ZM if we're interested in U
  if (!(haveDataSF(countryAbbrev, energyType))){
    # If we don't have data for this combination of countryAbbrev and energyType, 
    # return a column of NAs if the above conditions have been met.
    nRows <- 21 # All of these countries need 21 rows.
    df <- as.data.frame(matrix(NA, ncol = 1, nrow = nRows))
    colnames(df) <- "pred"
    return(df)
  }
  model <- linexModel(countryAbbrev, energyType)
  pred <- predict(model) #See http://stackoverflow.com/questions/9918807/how-get-plot-from-nls-in-r
  df <- data.frame(pred)
  # Pad with rows as necessary
  df <- padRows(countryAbbrev, df)
  return(df)
}

#########################
# Takes the LINEX fitted models and creates a single column of predicted GDP values
# that corresponds, row for row, with the AllData.txt file.
##
linexPredictionsColumn <- function(energyType){
  out <- do.call("rbind", lapply(countryAbbrevs, linexPredictions, energyType=energyType))
  colnames(out) <- c(paste("predGDP", energyType, sep=""))
  return(out)
}

##############################
# Creates a graph that plots predicted GDP as lines, one for each single factor, and historical GDP 
# data as open circles.
##
createLINEXLatticeGraph <- function(countryAbbrev, textScaling = 1.0, keyXLoc = defaultKeyXLoc, keyYLoc = defaultKeyYLoc){
  dataTable <- loadData("All") #Grab the raw data
  predictionsQ <- linexPredictionsColumn("Q") #Predictions from LINEX with Q
  predictionsX <- linexPredictionsColumn("X") #Predictions from LINEX with X
  predictionsU <- linexPredictionsColumn("U") #Predictions from LINEX with U
  #Now add the predictions columns to the data.
  dataTable <- cbind(dataTable, predictionsQ, predictionsX, predictionsU) 
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 2, 4, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "red", "blue", "darkorange") #line and symbol colors
  symbols <- c(1, NA, NA, NA)  #NA gives no symbol.
  # Code that deals with items that are specific to whether we want all countries or a specific country.  
  if (missing (countryAbbrev)){
    # We want a graph with panels for all countries
    yLimits <- yLimitsForGDPGraphs
    factorLevels <- countryNamesAlph # We want all countries shown
    indexCond <- list(countryOrderForGraphs) # We want all countries in this order
    layout <- ninePanelLayoutSpec # Show all countries
  } else {
    # We want only a specific country
    dataTable <- subset(dataTable, Country == countryAbbrev)    
    # Select the correct y limits
    index <- which(countryAbbrevsAlph %in% countryAbbrev)
    # The following lines use [index:index] as a convenient way of subsetting.
    # This has the added benefit of maintaining the correct classes for things.
    yLimits <- yLimitsForGDPGraphs[index:index]       # Pick limits for the country we want.
    factorLevels <- countryNamesAlph[index:index]     # Only show the country we have chosen.
    indexCond <- list(c(1))                           # We want only one country.
    layout <- onePanelLayoutSpec                      # We want only one panel in the graph.
  }  
  graph <- xyplot(iGDP+predGDPQ+predGDPX+predGDPU ~ Year | Country, data=dataTable,
                  type=graphType,
                  index.cond=indexCond, #orders the panels.
                  layout=layout,
                  strip = strip.custom(factor.levels=factorLevels, 
                                       bg="transparent", # Sets background transparent to match the graph itself.
                                       par.strip.text=list(cex=textScaling) # Scales text in the strip.
                                       ),
                  as.table=TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("Actual", "With $q$", "With $x$", "With $u$")),
                           type=graphType,
                           cex=keyTextSize * textScaling, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize * textScaling, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              x=list(at=timeTics),
                              y=list(relation="free",  #allows each axis to be different
                                     at=yTicsForIndexedGraphs)), #specifies location for tics
                  ylim=yLimits, #y axis limits
                  #axis labels and scaling
                  xlab=list(label="", cex=textScaling), 
                  ylab=list(label="Indexed (1980=1 or 1991=1)", cex=textScaling)
                  )
  return(graph)
}

#################################################
# Calculates parameter estimates and confidence intervals
# for the LINEX production function given a country and an energyType.
#
# countryAbbrev is a string containing the 2-letter abbreviation for the country, e.g. "US" or "CN"
# energyType is a string, one of "Q", "X", or "U"
#
# returns a data.frame of data for the LINEX model.
# First row is the +95% CI on all parameters
# Second row contains the parameter estimates
# Third row is the -95% CI on all parameters
# Each column has names: a_0 and c_t, corresponding to the parameters in the model.
##
linexData <- function(countryAbbrev, energyType){
  #First, check to see if we want useful work (U) AND one of the countries for which we don't have data.
  if (!haveDataCD(countryAbbrev, energyType)){
    #Return a column of NAs if the above conditions have been met.
    nRows <- 3 # +95% CI, CDe, and -95% CI.
    nCols <- 2 # a_0 and c_t
    df <- as.data.frame(matrix(NA, ncol = nCols, nrow = nRows))
    colnames(df) <- c("a_0", "c_t")
    rownames(df) <- c("+95% CI", "CDe", "-95% CI")
    return(df)
  }
  # We have a combination of country and energy type for which we have data.
  modelLINEX <- linexModel(countryAbbrev, energyType)
  summaryLINEX <- summary(modelLINEX) # Gives the nls summary table.
  dofLINEX <- summaryLINEX$df[2] # Gives the degrees of freedom for the model.
  tvalLINEX <- qt(ciHalfLevel, df = dofLINEX)
  ciLINEX <- confint(modelLINEX, level = ciLevel)  # Calculates confidence intervals for the LINEX model.
  # Report results with SE
  a_0 <- as.numeric(coef(modelLINEX)["a_0"])
  c_t <- as.numeric(coef(modelLINEX)["c_t"])
  # Combine all estimates and their confidence intervals into data frames with intelligent row names
  estLINEX <- data.frame(a_0 = a_0, c_t = c_t)
  row.names(estLINEX) <- "LINEX"
  # The [1] subscripts pick off the lower confidence interval
  lowerLINEX <- data.frame(a_0=ciLINEX["a_0","2.5%"], c_t=ciLINEX["c_t", "2.5%"])
  row.names(lowerLINEX) <- "-95% CI"
  # The [2] subscripts pick off the upper confidence interval
  upperLINEX <- data.frame(a_0=ciLINEX["a_0","97.5%"], c_t=ciLINEX["c_t", "97.5%"])
  row.names(upperLINEX) <- "+95% CI"
  # Now create the data for a table and return it
  dataLINEX <- rbind(upperLINEX, estLINEX, lowerLINEX)
  return(dataLINEX)
}

############
# Creates a row for the LINEX parameters table for the given country (2-letter code) and energyType (Q, X, or U)
##
linexCountryRow <- function(countryAbbrev, energyType){
  dataLINEX <- linexData(countryAbbrev, energyType)
  out <- cbind(dataLINEX["-95% CI", "a_0"], dataLINEX["LINEX", "a_0"], dataLINEX["+95% CI", "a_0"],
               dataLINEX["-95% CI", "c_t"], dataLINEX["LINEX", "c_t"], dataLINEX["+95% CI", "c_t"])
  return(out)
}

###########################################
# Creates a number of rows in a data.frame that contain information 
# about the coefficients of a LINEX model for countryName and energyType
# Each parameter has its own row with confidence intervals.
# The country name is in a column. Which parameter is involved is
# also in a column.
# 
# The return type is a data.frame.
##
linexCountryRowsForParamsGraph <- function(countryName, energyType){
  #Create three rows, one for each parameter. Each row is a data.frame so that it is plottable!
  dataLINEX <- linexData(countryName, energyType)
  a_0Row <- data.frame(country = countryName, 
                          parameter = "a_0", 
                          lowerCI = dataLINEX["-95% CI", "a_0"], 
                          value = dataLINEX["LINEX", "a_0"], 
                          upperCI = dataLINEX["+95% CI", "a_0"])
  c_tRow <- data.frame(country = countryName, 
                         parameter = "c_t",
                         lowerCI = dataLINEX["-95% CI", "c_t"],
                         value = dataLINEX["LINEX", "c_t"],  
                         upperCI = dataLINEX["+95% CI", "c_t"])
  table <- rbind(a_0Row, c_tRow)
  return(table)
}

#############################
# Creates a graph with confidence intervals for the LINEX model for the given energyType
##
createLINEXParamsGraph <- function(energyType){
  # Create a data table with the following columns:
  # country abbrev, parameter (a_0, c_t), -95% CI, value, +95% CI
  dataTable <- do.call("rbind", lapply(countryAbbrevs, linexCountryRowsForParamsGraph, energyType=energyType))
  graph <- segplot(country ~ upperCI + lowerCI | parameter, 
                   data = dataTable, 
                   centers = value, #identifies where the dots should be placed
                   draw.bands = FALSE, #provides nicer error bars
                   horizontal = FALSE, #makes error bars vertical and puts the countries in the x axis
                   layout = c(2,1), #2 column, 2 row
                   index.cond = list(c(1,2)), #orders the panels as alpha, beta, lambda, gamma
                   #set labels and bg color in strip
                   strip = strip.custom(factor.levels=c("$a_0$", "$c_t$"), bg="white"),
                   col = "black", #Sets line color to black
                   lwd = 1, #Sets line width to 1.0
                   ylim = list(c(-0.05, 1.0), c(-0.5, 5)), #y axis limits
                   scales = list(cex=scaleTextSize, #controls text size on scales.
                                 tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                                 x=list(cex=0.75), #reduces text size so that country abbrevs are legible
                                 y=list(relation="free", #allow each axis to be different
                                        at=list(c(0.0, 0.5, 1.0), c(0, 1, 2, 3, 4, 5))   #y tick marks
                                 )
                   )
  )
  return(graph)
}

############################
# Aggregates the LINEX results into a big data table for the given energyType.
##
linexParamsTable <- function(energyType){
  #Do rbind on the results of creating a row in the table for every country abbreviation that we know.
  dataLINEX <- do.call("rbind", lapply(countryAbbrevs, linexCountryRow, energyType=energyType))
  colnames(dataLINEX) <- c(" ", "$a_0$", " ", " ", "$c_t$",  " ")
  rownames(dataLINEX) <- countryAbbrevs
  tableLINEX <- xtable(dataLINEX, 
                    caption=paste("LINEX model (with $", tolower(energyType), "$) parameters for 1980-2011 (US, UK, JP), 1991-2010 (CN and ZA), and 1991-2011 (SA, IR, TZ, and ZM). (Parameter estimates beneath symbol. 95\\% confidence interval bounds to left and right.)", sep=""), 
                    label=paste("tab:LINEX_Parameters_With_", energyType, sep=""),
                    digits = c(1, 2,2,2, 2,2,2),
                    align = "r|ccc|ccc") #Sets alignment of the numbers in the columns
  return(tableLINEX)
}

############################
# Prints a table with parameters from a LINEX model for the given energyType. 
##
printLINEXParamsTable <- function(energyType){
  print(linexParamsTable(energyType), 
    caption.placement="top", 
    sanitize.colnames.function = identity, 
    size="\\tiny",
    table.placement="H")
}
@

<<AIC_Table_Functions, eval=TRUE>>=
###############################
# Creates an xtable object that holds the AIC values for each parameter estimation that we include.
##
createAICTable <- function(){
  ######################
  # Single-factor models
  ######################
  # Single-factor with K
  sfKModels <- lapply(countryAbbrevs, singleFactorModel, factor="K")
  aicSFk <- data.frame(lapply(sfKModels, AIC))
  rownames(aicSFk) <- "SF$k$"
  # Single-factor with L
  sfLModels <- lapply(countryAbbrevs, singleFactorModel, factor="L")
  aicSFl <- data.frame(lapply(sfLModels, AIC))
  rownames(aicSFl) <- "SF$l$"
  # Single-factor with Q
  sfQModels <- lapply(countryAbbrevs, singleFactorModel, factor="Q")
  aicSFq <- data.frame(lapply(sfQModels, AIC))
  rownames(aicSFq) <- "SF$q$"
  # Single-factor with X
  sfXModels <- lapply(countryAbbrevs, singleFactorModel, factor="X")
  aicSFx <- data.frame(lapply(sfXModels, AIC))
  rownames(aicSFx) <- "SF$x$"
  # Single-factor with U
  aicSFu <- cbind(US=AIC(singleFactorModel("US", "U")), 
                  UK=AIC(singleFactorModel("UK", "U")), 
                  JP=AIC(singleFactorModel("JP", "U")),
                  CN=NA, ZA=NA, SA=NA, IR=NA, TZ=NA, ZM=NA) #No U data for these countries.
  rownames(aicSFu) <- "SF$u$"
  ######################
  # Cobb-Douglas models
  ######################
  # Cobb-Douglas without energy
  cdModels <- lapply(countryAbbrevs, cobbDouglasModel)
  aicCD <- data.frame(lapply(cdModels, AIC))
  rownames(aicCD) <- "CD"
  # Cobb-Douglas with Q
  cdQModels <- lapply(countryAbbrevs, cobbDouglasModel, energyType="Q")
  aicCDq <- data.frame(lapply(cdQModels, AIC))
  rownames(aicCDq) <- "CD$q$"
  # Cobb-Douglas with X
  cdXModels <- lapply(countryAbbrevs, cobbDouglasModel, energyType="X")
  aicCDx <- data.frame(lapply(cdXModels, AIC))
  rownames(aicCDx) <- "CD$x$"
  # Cobb-Douglas with U
  aicCDu <- cbind(US=AIC(cobbDouglasModel("US", "U")), 
                  UK=AIC(cobbDouglasModel("UK", "U")), 
                  JP=AIC(cobbDouglasModel("JP", "U")),
                  CN=NA, ZA=NA, SA=NA, IR=NA, TZ=NA, ZM=NA) #No U data for these countries.
  rownames(aicCDu) <- "CD$u$"
  ######################
  # CES models
  ######################

  ######################
  # LINEX models
  ######################
  # LINEX with Q
  linexQModels <- lapply(countryAbbrevs, linexModel, energyType="Q")
  aicLINEXq <- data.frame(lapply(linexQModels, AIC))
  rownames(aicLINEXq) <- "LINEX$q$"
  # LINEX with X
  linexXModels <- lapply(countryAbbrevs, linexModel, energyType="X")
  aicLINEXx <- data.frame(lapply(linexXModels, AIC))
  rownames(aicLINEXx) <- "LINEX$x$"  
  # LINEX with U
  aicLINEXu <- cbind(US=AIC(linexModel("US", "U")), 
                  UK=AIC(linexModel("UK", "U")), 
                  JP=AIC(linexModel("JP", "U")),
                  CN=NA, ZA=NA, SA=NA, IR=NA, TZ=NA, ZM=NA) #No U data for these countries.
  rownames(aicLINEXu) <- "LINEX$u$"

  # Create the data.frame table
  out <- rbind(aicSFk, aicSFl, aicSFq, aicSFx, aicSFu, aicCD, aicCDq, aicCDx, aicCDu, aicLINEXq, aicLINEXx, aicLINEXu) 
  # Create the xtable with caption and label
  out <- xtable(out, caption="AIC values for all models.", label="tab:AICTable", digits=1)
  return(out)
}
@