### 2013-07-30 CES.Rnw
bug fixes in running multiple CES models

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{language} % Obtain language.sty from https://github.com/rpruim/latex
\usepackage{underscore} % Should be part of any LaTeX distribution.
\usepackage{float} % For precise figure positioning.
\usepackage{multirow} % To create multirow tables
\usepackage{amssymb} % To provide check marks

\title{Parameter Estimation for the CES Model}
\author{Matthew Kuperus Heun}
\begin{document}

<<setup, include=FALSE>>=
require(knitr)
require(plyr)
opts_chunk$set(
  tidy=FALSE,
  eval=TRUE
)
source("Econ-Growth-Functions2.R")
@

<<>>=
extractRSSdata <- function(object, ...) {
  res <- expand.grid( object$rho1Values, object$rhoValues )
  names(res) <- c('rho1', 'rho')
  res <- transform(res, RSS = as.vector(object$rssArray))
  res
}
@

\maketitle

% Add an argument if you want a label. Example: "\label{eq:CES}"
\newcommand{\ceskl}[1]{\begin{equation} #1
  y = \gamma \: A \: \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{-1/\rho_1}; 
      A = \mathrm{e}^{\lambda (t-t_0)}
\end{equation}}

\newcommand{\ceskle}[1]{\begin{equation} #1
  y = \gamma \: A \: \left\{\delta \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) e^{-\rho} \right\}^{-1/\rho}; 
      A = \mathrm{e}^{\lambda (t-t_0)}
\end{equation}
}

\newcommand{\ceslek}[1]{\begin{equation} #1
  y = \gamma \: A \: \left\{\delta \left[\delta_1 l^{-\rho_1} 
      + (1-\delta_1) e^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) k^{-\rho} \right\}^{-1/\rho}; 
      A = \mathrm{e}^{\lambda (t-t_0)}
\end{equation}
}

\newcommand{\cesekl}[1]{\begin{equation} #1
  y = \gamma \: A \: \left\{\delta \left[\delta_1 e^{-\rho_1} 
      + (1-\delta_1) k^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) l^{-\rho} \right\}^{-1/\rho}; 
      A = \mathrm{e}^{\lambda (t-t_0)}
\end{equation}
}

%++++++++++++++++++++++++++++++
\section{CES Production Function} 
\label{sec:CES}
%++++++++++++++++++++++++++++++

The Constant Elasticity of Substitution (CES) model without energy is

\ceskl{\label{eq:CES}}

\begin{itemize}
\item
Equation~\ref{eq:CES} is a CES production function with 
capital stock ($k$) and labor ($l$) factors of production. 
\item
$\gamma$ is a fitting parameter that represents productivity. 
\item
$\delta_1$ describes the importance of ($k$) relative to ($l$).
\item
The fitting parameter $\rho_1$ indicates elasticity
of substitution ($\sigma_1$) between capital ($k$) 
and labor ($l$). $\sigma_1 = \frac{1}{1+\rho_1}$.
\item
As $\rho_1 \rightarrow 0$, $\sigma_1 \rightarrow 1$,
and the embedded CES production function for $k$ and $l$ degenerates 
to the Cobb-Douglas production function.
\item
As $\sigma_1 \rightarrow 0$ ($\rho_1 \rightarrow \infty$),
$k$ and $l$ are perfect complements.
\item
As $\sigma_1 \rightarrow \infty$ ($\rho_1 \rightarrow -1$),
$k$ and $l$ are perfect substitutes.
\item
Constraints on the fitting parameters include 
  \begin{itemize}
  \item $\delta_1 \in (0,1)$
  \item $\rho_1 \in (-1,0) \cup (0,\infty)$
  \item $\sigma_1 \in (0,1) \cup (1,\infty)$
  \end{itemize}

\end{itemize}

The CES model with energy is

\ceskle{\label{eq:CESe}}
% \begin{equation} \label{eq:CESe}
%   y = \gamma \: A \: \left\{\delta \left[\delta_1 k^{-\rho_1} 
%       + (1-\delta_1)l^{-\rho_1} \right]^{\rho/\rho_1} 
%       + (1-\delta) e^{-\rho} \right\}^{-1/\rho}; 
%       A = \mathrm{e}^{\lambda (t-t_0)}
% \end{equation}

\begin{itemize}
\item
Equation~\ref{eq:CESe} augments Equation~\ref{eq:CES} with energy 
using a ($kl$)$e$ nesting structure, as is typical in the literature. 
\item
$\delta$ describes the importance of the capital/labor pair ($kl$) relative to energy ($e$).
\item
As $\delta \rightarrow 1$, the energy term ($e$) 
becomes increasingly unimportant for the model.
\item
The fitting parameter $\rho$ indicates elasticity
of substitution ($\sigma$) between the capital/labor pair ($kl$) 
and energy ($e$). $\sigma = \frac{1}{1+\rho}$.
\item
As $\rho \rightarrow 0$, $\sigma \rightarrow 1$,
and the CES production function for $(kl)$ and $(e)$ degenerates 
to the Cobb-Douglas production function.
\item
As $\sigma \rightarrow \infty$ ($\rho \rightarrow -1$), 
($kl$) and $e$ are perfect substitutes. 
\item
As $\sigma \rightarrow 0$ ($\rho \rightarrow \infty$), 
($kl$) and $e$ are perfect complements: 
no substitution is possible. 
\item 
Constraints on the fitting parameters include 
  \begin{itemize}
  \item $\delta \in (0,1)$,
  \item $\rho \in (-1,0) \cup (0,\infty)$.
  \item $\sigma \in (0,1) \cup (1,\infty)$
  \end{itemize}
\item
Equation~\ref{eq:CES} is a degenerate form of Equation~\ref{eq:CESe} 
where $\delta \rightarrow 1$ and $\rho$ is undefined in 
the sense that energy ($e$) is unimportant regardless of the 
value of $\rho$.
\end{itemize}

%++++++++++++++++++++++++++++++
\section{Fitting Without Constraints} 
\label{sec:CES-without-constraints}
%++++++++++++++++++++++++++++++

%------------------------------
\subsection{Standard Energy Nesting ($kl$)$e$}
\label{sec:kle-nesting}
%------------------------------

The \texttt{R} package \texttt{micEcon} provide function 
\texttt{cesEst} for estimating the parameters of the CES model.
If we use \texttt{cesEst} ``out of the box,'' which uses the
Levenberg-Marquardt (\texttt{LM}) optimization method, 
we often obtain parameter
estimates that violate the constraints above. 
The default starting values for the fitting parameters are 
\begin{itemize}
\item $\lambda = 0.015$
\item $\gamma = $ a value that makes the mean of the residuals is zero under certain assumptions
\item $\delta = 0.5$
\item $\delta_1 = 0.5$
\item $\rho = 0.25$ (implying that $\sigma = 0.8$) (if not otherwise specified)
\item $\rho_1 = 0.25$ (implying that $\sigma_1 = 0.8$) (if not otherwise specified)
\end{itemize}

We can perform unconstrained parameter estimation with the following code.

<<>>=
dataUK <- loadData(countryAbbrev="UK")
modelUKQ_LM <- cesEstPlus(data=dataUK, 
                      yName="iGDP", 
                      xNames=c("iCapStk", "iLabor", "iQ"), 
                      tName="iYear")
summary(modelUKQ_LM)
@

In the above example, we obtain convergence of the curve-fitting routine. 
However, there are two problems.

\begin{enumerate}
\item The estimate of $\rho_1 = $ \Sexpr{coef(modelUKQ_LM)["rho_1"]}
violates the constraint $\rho_1 \in (-1,0) \cup (0,\infty)$. 
This violation leads to $\sigma_1$ (\texttt{E_1_2 (HM)}) being undefined (\texttt{NA})
in the model summary.
\item $\delta \rightarrow 1$, meaning that the energy term
is unnecessary. The Cobb-Douglas model indicated that the 
capital stock term was unnecessary ($\alpha = 0$). So, the 
Cobb-Douglas and CES models seem to be in conflict.
\end{enumerate}

However, we should note that we are not respecting the constraints here. 
To assess whether the models are actually telling us 
something different, we need to be sure to respect constraints
with the CES model.

%------------------------------
\subsection{Alternative Energy Nesting ($le$)$k$}
\label{sec:lek-nesting}
%------------------------------

A second option is to use a different nesting structure
for the CES model, namely ($le$)($k$), as shown in Equation~\ref{eq:ceslek}.

\ceskle{}

\ceslek{\label{eq:ceslek}}

% \begin{equation} \label{eq:CESe2}
%   y = \gamma \: A \: \left\{\delta \left[\delta_1 l^{-\rho_1} 
%       + (1-\delta_1) e^{-\rho_1} \right]^{\rho/\rho_1} 
%       + (1-\delta) k^{-\rho} \right\}^{-1/\rho}; 
%       A = \mathrm{e}^{\lambda (t-t_0)}
% \end{equation}

In Equation~\ref{eq:ceslek}, $\delta \rightarrow 1$ indicates that capital stock ($k$)
is unnecessary.

<<>>=
modelUKQ_LM2 <- cesEstPlus(data=dataUK, 
                      yName="iGDP", 
                      xNames=c("iLabor", "iQ", "iCapStk"), 
                      tName="iYear")
summary(modelUKQ_LM2)
@

With ($le$)($k$) nesting, the fitting routine converges, and we find that 
$\delta = $ \Sexpr{coef(modelUKQ_LM2)["delta"]}, 
which violates the constraint $\delta \in (0,1)$. This result might indicate 
that capital stock ($k$) is unnecessary, thereby agreeing with 
results from the Cobb-Douglas model.

So, the indications about relative importance of different factors 
of production may be a function of nesting method in the CES model.
However, we need to be careful about convergence and respecting 
constraints.

%++++++++++++++++++++++++++++++
\section{Fitting With Constraints} 
\label{sec:CES-with-constraints}
%++++++++++++++++++++++++++++++

To fit while respecting the constraints, we can use the \texttt{L-BFGS-B} optimization
method.

<<>>=
modelUKQ_LBFGSB <- cesEstPlus(data=dataUK, 
                          yName="iGDP", 
                          xNames=c("iCapStk", "iLabor", "iQ"), 
                          tName="iYear",
                          method="L-BFGS-B")
summary(modelUKQ_LBFGSB)
@

\noindent The above approach correctly respects the constraints, 
but the \texttt{LM} algorithm does not converge.

To assess what is happening, we can use the built-in grid search
capability of \texttt{cesEstPlus}. Grid search is accessed by specifying
\texttt{rho} and/or \texttt{rho1} arguments and a range of values. 
Note that specifying a range of values for $\rho$ or $\rho_1$ that 
contains values less than $-1$ results in an error. 
The model with lowest $SSE$ will be returned from the grid search.

<<>>=
modelUKQ_grid <- cesEstPlus(data=dataUK, 
                        yName="iGDP", 
                        xNames=c("iCapStk", "iLabor", "iQ"), 
                        tName="iYear",
                        method="L-BFGS-B",
                        rho=seq(from=5, to=30, by=5), 
                        rho1=seq(from=-0.5, to=1.5, by=0.5))
summary(modelUKQ_grid)
@

The negative of the $SSE$ values from the grid search can be plotted as a function of 
$\rho$ and $\rho_1$ quite easily, as shown in Figure~\ref{fig:UKQ-grid-coarse}.
The \texttt{plot} function puts the negative of $SSE$ on the $z$ axis, 
so better curve fits appear higher on the $z$ axis of this plot.

<<UKQ-grid-coarse, fig.pos="H", fig.cap="$-SSE$ vs. $\\rho$ and $\\rho_1$. Coarse grid.">>=
plot(modelUKQ_grid)
print(class(modelUKQ_grid))
@

The best fit occurs when $\rho_1 \rightarrow -1$ (its boundary) and $\rho$ is large, 
but not so large that you fall over the cliff. In Figure~\ref{fig:UKQ-grid-coarse},
that is at $\rho = 15$ and $\rho_1 = -0.5$ (the lowest value of $\rho_1$ in the grid search).
Although, it should be noted that $\frac{\partial^2 SSE}{\partial \rho_1^2}$
is very small.

To explore the details at high values of $\rho$ and values of 
$\rho_1 \rightarrow -1$, we can do a detailed analysis with 
a finer grid.

<<>>=
modelUKQ_grid_detail <- cesEstPlus(data=dataUK, 
                        yName="iGDP", 
                        xNames=c("iCapStk", "iLabor", "iQ"), 
                        tName="iYear",
                        method="L-BFGS-B",
                        rho=seq(from=100, to=120, length.out=20), 
                        rho1=seq(from=1, to=2, length.out=20))
summary(modelUKQ_grid_detail)
levelplot( RSS ~ rho1 + rho, data=extractRSSdata(modelUKQ_grid_detail), 
           region=TRUE, 
           col.regions=terrain.colors)
@

We can build a graph as shown in Figure~\ref{fig:UKQ-grid-fine}.

<<UKQ-grid-fine, fig.pos="H", fig.cap="$-SSE$ vs. $\\rho$ and $\\rho_1$. Fine grid.">>=
plot(modelUKQ_grid_detail)
@

The best fit occurs when $\rho_1 \rightarrow -1$ (at its boundary
and indicating perfect substitutibility between $k$ and $l$)
and $\delta \rightarrow 1$ (indicating that energy is unnecessary
in this production function). Interestingly, the Cobb-Douglas
model with $q$ for the UK indicates that energy is important
($\gamma \approx 0.5$). Are the Cobb-Douglas and CES models
indicating different things?

The sudden dropoff in $-SSE$ at larger values of $\rho$ 
should be investigated.

%++++++++++++++++++++++++++++++
\section{Fitting Without Energy} 
\label{sec:CES-without-energy}
%++++++++++++++++++++++++++++++

Based on the fact that the best fit constrained model gives
$\delta \rightarrow 1$, we can try a no-energy CES model 
with the following code. 
(Note that the value of \texttt{rho} below 
corresponds to the parameter $\rho_1$ in Equation~\ref{eq:CES}.)

<<>>=
modelUK <- cesEstPlus(data=dataUK, 
                        yName="iGDP", 
                        xNames=c("iCapStk", "iLabor"), 
                        tName="iYear",
                        method="L-BFGS-B")
summary(modelUK)
@

With the no-energy model, we have achieved reasonable values for
all parameters. And, the parameter estimation algorithm converges.
However, we may be in a situation where Cobb-Douglas and CES
are in conflict regarding the role of energy in economic growth.

%++++++++++++++++++++++++++++++
\section{Effect of Nesting} 
\label{sec:effect-of-nesting}
%++++++++++++++++++++++++++++++
In Section~\ref{sec:lek-nesting}, we found that different results were
obtained with standard nesting ($kl$)$e$ vs. a differnet nesting, namely
($le$)$k$. To explore that further, we fit all 9 countries with 
all three nesting options:
\begin{itemize}
\item{($kl$)$e$} 
\end{itemize}
\ceskle{}
\begin{itemize}
\item{($le$)$k$}
\end{itemize}
\ceslek{}
\begin{itemize}
\item{($ek$)$l$}
\end{itemize}
\cesekl{\label{eq:cesekl}}

We use the \texttt{L-BFGS-B} method (as shown in Section~\ref{sec:CES-with-constraints}),
which respects constraints on the model parameters, and thermal energy ($q$) is 
used for the energy quantification.

<<>>=
nestingOptions <- c(kle="(kl)e", lek="(le)k", ekl="(ek)l")
fitCES <- function(countryAbbrev, nest, energyType="Q", method=c("PORT","L-BFGS-B")){
  data <- loadData(countryAbbrev=countryAbbrev)
  data <- replaceColName(data, energyType, "iEToFit")
  if (nest == "(kl)e"){
    xNames <- c("iCapStk", "iLabor", "iEToFit")
  } else if (nest == "(le)k"){
    xNames <- c("iLabor", "iEToFit", "iCapStk")
  } else if (nest == "(ek)l"){
    xNames <- c("iEToFit", "iCapStk", "iLabor")
  } else {
    stop(paste("Unknown nesting option", nest, "in fitCES."))
  }
  model <- cesModel(data=data, energyType=energyType, 
                  # yName="iGDP", 
                  xNames=xNames, 
                  # tName="iYear",
                  method=method)
  res <- naturalCoef( model ) 
  res <- transform( res, nest=nest, country=countryAbbrev, nest=nest, converge=model$convergence ) 
  return(res)
}

# Make all the models
kleModels <- ldply(countryAbbrevs, fitCES, nest="(kl)e")
lekModels <- ldply(countryAbbrevs, fitCES, nest="(le)k")
eklModels <- ldply(countryAbbrevs, fitCES, nest="(ek)l")
allModels <- rbind(kleModels, lekModels, eklModels)
# Ignore the leading .id column
allModels <- subset( allModels, select= -.id )
allModels <- transform( allModels, max.sse = pmax(sse.PORT, sse.L.BFGS.B))
# Sort rows based on country
allModels_ordered <- allModels[order(allModels$country) , ] 
# Reorder the columns to put the meta information first
allModels_ordered <- allModels_ordered[c("country", "nest", "converge", "gamma", "lambda", 
                                         "delta_1", "delta", "rho_1", "rho", "method", "sse","max.sse")]
# Create the LaTeX table
CESTable <- xtable(allModels_ordered, caption="CES coefficients for three nesting options.", 
                   label="tab:CES-coeffs-table", digits=3,
                   table.placement="H")
@

<<CESResultsTable, results='asis'>>=
print(CESTable,
      caption.placement="top", 
      table.placement="H",
      include.rownames=FALSE,
      hline.after=c(-1,0,3,6,9,12,15,18,21,24,nrow(CESTable)),
      digits=3
      )
@

One way to assess the data in Table~\ref{tab:CES-coeffs-table} is to compare indications 
from the Cobb-Douglas and CES models. 
Table~\ref{tab:cd-ces-comparison} does so.

\begin{table}[H] \caption{Comparison of indications from Cobb-Douglas and CES models with energy.} \label{tab:cd-ces-comparison} 
  \begin{center}
    \begin{tabular}{c|l|l|c}
Country & CD                                     & CES                                            & Agree? \\
\hline
US      & $\gamma = 0 \rightarrow e$ unnecessary & ($kl$)$e$, ($le$)$k \rightarrow e$ unnecessary & \checkmark \\
\hline
UK      & $\alpha = 0 \rightarrow k$ unnecessary & ($kl$)$e$, ($le$)$k \rightarrow k$ unnecessary & \checkmark \\
\hline
JP      & $\gamma = 0 \rightarrow e$ unnecessary & ($ek$)$l \rightarrow e$ unnecessary            & \checkmark $\:$ but no CES convergence\\
\hline
CN      & $\gamma = 0 \rightarrow e$ unnecessary & No indications                                 & Note: no CES convergence\\
\hline
ZA      & $\beta  = 0 \rightarrow l$ unnecessary & No indications                                 &      \\
\hline
SA      & No indications                         & No indications                                 & \checkmark \\
\hline
IR      & $\gamma = 0 \rightarrow e$ unnecessary & All nestings $\rightarrow e$ unnecessary       & \checkmark \\
\hline
TZ      & $\beta  = 0 \rightarrow l$ unnecessary & No indications                                 &      \\
\hline
\multirow{2}{*}{ZM} & $\beta  = 0 \rightarrow l$ unnecessary & \multirow{2}{*}{All nestings $\rightarrow l$ unnecessary} & \multirow{2}{*}{-- (partial)}\\
        & $\gamma = 0 \rightarrow e$ unnecessary &                                                &        \\

    \end{tabular}
  \end{center}
\end{table}

Table~\ref{tab:cd-ces-comparison} looks better than I was expecting. 
Note that in Section~\ref{sec:CES-without-constraints},
I was not assessing the indications from the CES model 
\emph{while also respecting the constraints}. 
When the constraints are respected, both Cobb-Douglas and CES
give similar indications for most economies.

I suspect that clearing up the non-convergence issues 
for the UK, JP, and CN will provide additional insights.
Martin may be able to provide additional insight on the South African (ZA)
and Tanzanian (TZ) situations, as ZA and TZ are the only countries with both 
(a) dissimilar indications from Cobb-Douglas and CES and 
(b) convergence of the CES models for all nestings.

%++++++++++++++++++++++++++++++
\section{Possible Ways Forward} 
\label{sec:CES-way-forward}
%++++++++++++++++++++++++++++++

Here are some brainstorm-y ideas on how to proceed, in no particular order.

\begin{enumerate}
\item I think we need to understand the cause of the 
steep dropoff in Figures~\ref{fig:UKQ-grid-coarse} and
\ref{fig:UKQ-grid-fine}.

\item We need to understand why the nesting structure appears
to affect the indications of which parameters are important.
This may involve a little bit of literature searching by Martin.

\item We may want to consider an algorithm to detect when
$\delta \rightarrow 1$ and switch to a no-energy
model as shown above.

\item Perhaps there is a degenerate form of the model that is 
easier to fit if we detect that $\rho_1 \rightarrow -1$.

\item It is possible to set $\rho_1$ to a specific number 
(in this case, we could choose $\rho_1 = -0.99$; $-1.0$ would 
blow up)
and have \texttt{cesEst} fit the remaining parameters.
(This is accomplished by supplying \texttt{rho} and/or
\texttt{rho1} parameters as single values, not 
ranges.)
A possible algorithm using this mechanism would be
similar to our algorithm for respecting the constraints
on the Cobb-Douglas model.
  \begin{enumerate}
  \item{Fit the unconstrained model}
  \item{If the unconstrained fit violates any constraints, 
  set the offending parameter at the constraint}
  \item{Re-fit the remaining the parameters}
  \end{enumerate}
\end{enumerate}

\end{document}