### 2013-07-30 EconGrowth2.Rnw
New file for the actual paper.

% This article has been prepared for publication in Energy Economics in RStudio with knitr.
% According to http://www.elsevier.com/author-schemas/the-elsarticle-latex-document-class, we should be using the
% elsarticle.cls file.
% According to http://cdn.elsevier.com/assets/pdf_file/0006/109392/journal_refstyles.pdf, we should be using
% elsarticle-template-2-harv.tex as the template for the text.
% Furthermore, we should be using model2-names.bst for the bibliographic references.
% The approach here is to load the frontmatter and backmatter from elsarticle-template-2-harv.tex
% both ahead of and behind the text for our paper.
% -- Matthew Kuperus Heun, 2013-01-18

%% This is file `elsarticle-template-2-harv.tex',
%%
%% Copyright 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle  Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for Elsevier's document class `elsarticle'
%% with harvard style bibliographic references
%%
%% $Id: elsarticle-template-2-harv.tex 155 2009-10-08 05:35:05Z rishi $
%% $URL: http://lenova.river-valley.com/svn/elsbst/trunk/elsarticle-template-2-harv.tex $
%%
\documentclass[preprint,authoryear,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,authoryear,1p,times]{elsarticle}
%% \documentclass[final,authoryear,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,authoryear,3p,times]{elsarticle}
%% \documentclass[final,authoryear,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,authoryear,5p,times]{elsarticle}
%% \documentclass[final,authoryear,5p,times,twocolumn]{elsarticle}

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%% or use the graphicx package for more complicated commands
%% \usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

\usepackage{soul} % Provides strikethrough text
\usepackage{float} % Allows precise positioning of tables and figures within the text
\usepackage{amsmath} % Allows \begin{equation*} \end{equation*} for unnumbered equations
\usepackage{multirow} % To create multirow tables
\usepackage{hyperref} % To create hyperlinks in the paper

% From http://economics.utoronto.ca/osborne/latex/BIBTEX.HTM
\newcommand{\citeapos}[1]{\citeauthor{#1}'s (\citeyear{#1})} % Posessive citations.

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon (default)
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   authoryear - selects author-year citations (default)
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%   longnamesfirst  -  makes first citation full author list
%%
%% \biboptions{longnamesfirst,comma}

% \biboptions{}

\journal{Energy Economics}

\begin{document}

<<setup, echo=FALSE, message=FALSE, eval=TRUE>>=
require(lattice)
require(latticeExtra)
require(ggplot2)
require(car)
require(mosaic)
require(xtable)
require(nlmrt)
require(micEconCES)
# Provides access to data melting. 
# See http://cran.r-project.org/web/packages/reshape2/reshape2.pdf
require(reshape2) 
# tikz allows use of LaTeX formatting and font in graphs. 
# Allows for a consistent look across the paper.
# See http://r-forge.r-project.org/R/?group_id=440 for instructions 
# on installing tikzDevice.
# require(tikzDevice) 
trellis.par.set(theme=theme.mosaic())
options(width=75)
options(show.signif.stars=TRUE)
opts_chunk$set(
  # ********************************************************
  eval=TRUE,
  # ********************************************************
  # tikz allows LaTeX code in graphical output. 
  # E.g., "$y$" for a variable name in the legend for a graph.  
  # dev='tikz',
  tidy=FALSE,
  comment=NA,
  # Tells whether to cache output from chunks are saved. Cacheing saves time. 
  # However, references to LaTeX tables DO NOT WORK if cacheing is turned on.
  cache=FALSE,
  warning=FALSE,  #Tells whether to show warnings in the output.
  # Turns off messages for all chunks. 
  # Set TRUE on an individual chunk to see it.
  message=FALSE, 
  # Tells whether to echo code for all chunks. 
  # Set TRUE on an individual chunk to see its code.
  echo=FALSE     
)
source("Econ-Growth-Functions2.R")
source("Graphics.R")
@

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{An Empirical Analysis of the Role of Energy in Economic Growth}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

\author[CalvinEngr]{Caleb Reese}
\author[CalvinEngr]{Lucas Timmer}
\author[CalvinEngr]{Matthew Kuperus Heun\corref{cor1}}
\ead{mkh2@calvin.edu}
\author[Stellenbosch]{Martin de Wit}
\author[CalvinMath]{Randall J. Pruim}

\cortext[cor1]{Corresponding author}
\address[CalvinEngr]{Engineering Department, Calvin College, Grand Rapids, MI 49546, USA}
\address[Stellenbosch]{Faculty of Economics and Management Sciences, School of Public Leadership, Stellenbosch University, P.O. Box 610, Bellville 7535, South Africa}
\address[CalvinMath]{Mathematics \& Statistics Department, Calvin College, Grand Rapids, MI 49546, USA}

\begin{abstract}
%% Text of abstract
**** Add abstract ****
\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
economic growth \sep energy \sep Cobb-Douglas \sep CES \sep LINEX
%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
\end{keyword}

\end{frontmatter}

% \linenumbers
%% main text

\section*{Notes from Martin}

**** Additional references requested by Martin. Incorporate into paper? ****

\citet{Lindenberger:2011dx}

\citet{Stresing:2008cq}

The remainder of this paper is organized as follows. Section~\ref{sec:Coordinates} 
discusses the coordinates of analysis described above. 
Section~\ref{sec:Sources_of_Data} details sources of data and calculations to 
generate time series for economic output and factors of production. The procedures 
for and results from parameter estimation for the various production functions 
are discussed in Section~\ref{sec:Parameter_Estimation}. 
Section~\ref{sec:Implications} discusses implications of the parameter estimation 
results, and Section~\ref{sec:Conclusion} concludes. An appendix provides details 
of all statistical results.

**** Notes from Martin ****

To be included somewhere.

Big picture
Why do we do this? Want to link energy and economics on national level 
and in a modeled approach,informing energy policy during energy transitions.
Usually macro energy economic models rely on PF with energy included. 
Key papers using CES in energy economics incl high impact such as 
Stern, Gerlagh, Tol etc. 
Cite in introduction met energy economists using PF approaches, specially CES.
But, which form to choose and using which parameters? 
Literature is inconclusive, very little actual empirical testing of such functions. 
Mostly CES in CGE models used by Wbank and others. 
CD adviced to be used by CBO. 
How do we select a functional form and what are the parameters to be used? On what basis do we select?

We aim. to search for empirically robust solution, one that stands up to statistical testing.

What we find is counterintuitive for many cases: dropping energy from the PF has little impact, 
but some notable exceptions. 
On other hand, only including energy in a PF of LINEX form also yields good results and stats tests. 
Is not a simple case for dropping energy, but acknowledging its key role. 
Do we support the call for better bottom-up understanding? 
What do we say on the limitations of macroeconomic modelling approaches altogether? 
(Note such limitations are well known in ME literature, for example capital controversy Robinson).
We speculate on implications for choosing PF in energy economic models that inform growth policy.

Points on paper so far
1. Does the relative small differences in fit between different production functions warrant 
such a large debate on validity of one vs other? Clearly not. 
CD been used by economists for ease of use and extreme flexibility (Miller).  
PF generally used on instrumentalist grounds if give reasonable good description of data. 
NK econ theory provide poor causal explanation of observed data, 
but yet CD shows very good fit (miller, p17). 
CES is  more general, but is a liability in forecasting because difficult 
to get estimate of EoS that is consistent across studies (Miller, p26). 
Aggregation is problem of both CD and CES functional forms, 
so being careful when making strong inferences that cannot be justified.

2. We think we can contribute by emphasising the need 
for proper stats testing or inconsistent reporting on use of preferred functions. 
Problems with empirical testing presented by Shen, Serrenho.
What are our issues with multivariate OLS (Serrenho)?

3. Are we joining the forced choice between production functions or 
acknowledge the similar problems with aggregate prod functions in general? 
Do we see this as sufficient evidence to doubt the top-down approach to econ growth altogether?

4. R2 is a very crude measure and only very minor improvements 
are achieved when others factors such as energy is added in a CD. 
We concur. Puzzling is conclusion that such minor improvements 
does prove the need to include energy in PF (Serrenho).

5. We do think it is a choice between CD and CES (Miller) as 
both have serious deficiencies after stats tests.

6. One central issue within PF is elasticity of subs (EoS). 
That is the \%ch in factor proportions ( k,l, etc) 
because of one unit ch in MRTS. 
MRTS is rate at which one factor can be substituted for another holding output constant. 
EoS is low in Leontief cases, L shaped isoquants.

7. Miller makes the point that even if parameters in model are stats artifacts, 
they are still useful for forecasting. Do we agree?

8. Is Ockham's razor relevant to what we are doing?

9. Assumptions of CD are strongly rejected by empirical work, Miller still makes case for using CD.

11. What about energy augmented

10. TZ case
CDno e: almost 80\% l
CD with e: 25\% energy, rest c,importance of l drops to 0
CES no e: 20\% c (sigma), 80\% l, but large range
CES with e: fig 35 e may be NB
LINEX: one of lowest ct parameters (low relative importance of labour eff compared to cap eff)
Does this mean that relative small increases in cap eff can really lift the economy?

11. ZA case
CD no e: roughly half c, l
CD with e: roughly half, c, e, importance of l drops to 0
CES no e: 10% c, 25% l ??
CES with e: fig 35 e may be NB
LINEX: one of highest ct parameters (high relative importance of labour eff compared to cap eff)
Does this mean that relative small increases in labour eff can really lift the economy?

12. Capital starved Zambia. Very high importance. Understand CES param better.

Discussion: What is Next?

a. A comparative analysis between diff PF reveal interesting 
enough information to warrant a more diffused approach to countries. 
Thus, we have a case for doing this for  all countries in the world 
(or at least those representative of countries in different groups 80/20)

b. If c is relatively more important than labour or energy 
for growth and if capital stock is low, a situation one would expect in LDCs, 
a relative small addition of capital stock can give relatively large returns in terms of growth.

c. Now it becomes interesting, in a 'stages of growth' approach, 
using LT econ data it may be the case that the importance 
of factors are different across growth stages. 
It would need a composite type of analysis - both stock and importance.

d. Is it possible to have evidence of diff stocks and diff importances across phases of growth? 
Maybe test this for a data rich nation that has worked through these stages, 
we try to identify them and do partial analysis on time periods coinciding with these stages.

Should we eliminate $u$ data? Only 3 countries. Only 20 years. Doesn't add much insight. 
Too bad, but perhaps necessary to simplify paper.

Discussions:

Counterfactuals: energy, cap, labour what ifs for future.

interpretation of type of economies, what does the data presentation say 
on relative change in each input and change in GDP?

Forcing const elast of subs in CES thru nested structures violates 
observation that more than two factor will not be const. elasticities Uzawa 1962.

R programming dealt with nested CES see paper discussion on Kemfert.

Next steps:

Stages of growth crowd implicitly assumes countries at similar stage of growth will
have similar values for $\alpha$, $\beta$, $\gamma$. However, our results show
that not to be true. Needs further exploration to cover more (all?) countries.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} 
\label{sec:Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

1. Set scene through energy economic modelling papers. 
Point out that PF approaches are alive and well (Stern, Tol, Gerlagh, World Bank, IMF), 
short lit review as part of a longer introduction.  
Mention competing PF approaches and parameter choice. 
On what basis selection and how estimate parameters?

2. Point out the need for adequate statistical testing and economic interpretation, 
citing some earlier critical papers on PF limitations. Take issue with

3. Spell out the contribution we will be making in this paper, 
we are taking the approach that all models are wrong but some are useful. 
We are after generating insights even after deeper assumptions of econ growth theory are very limiting.

4. Introduce coordinates of analysis.

The remainder of this paper is organized as follows: **** complete this sentence ****.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Coordinates of Analysis} 
\label{sec:Coordinates}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This section describes the three coordinates of analysis and briefly reviews 
literature related to each.

%++++++++++++++++++++++++++++++
\subsection{Mathematical Forms of the Energy-augmented Production Function} 
\label{sec:Prod_Func_Forms}
%++++++++++++++++++++++++++++++

In this paper, we assess three prominent energy-augmented production functions 
appear in the literature: Cobb-Douglas (CDe), Constant Elasticity of 
Substitution (CESe), and LINear EXponential (LINEX). The following subsections 
describe each.

%------------------------------
\subsubsection{Energy-augmented Cobb-Douglas Production Function (CDe)} 
\label{sec:CDe}
%------------------------------

Similar to Equation~\ref{eq:CD}, the energy-augmented Cobb-Douglas 
production function is

\begin{equation} \label{eq:CDe}
  y = A k^\alpha l^\beta e^\gamma; A = \mathrm{e}^{\lambda(t-t_0)}
\end{equation}

\noindent where $e \equiv E/E_0$, and $E$ is in units of energy per time, 
typically TJ/year. The energy-augmented Cobb-Douglas production function 
is often assumed to have constant returns to scale for the three factors 
of production: $\alpha + \beta + \gamma = 1$.

\citet{Ayres:2008cl} uses Equation~\ref{eq:CDe} with exergy ($x$) and 
useful work ($u$) quantifications for $e$ to describe historic Japanese 
and American economies from 1900-2000. \citet{Froling:2009vo} uses an 
energy-augmented Cobb-Douglas production function with thermal energy 
quantification ($q$) to describe output for the global economy from 
1800--1970. Fr\"{o}ling's model endogenizes labor with respect to fertility 
and mortality rates and separates thermal energy into two types: biomass 
energy and coal energy. \citet{Serrenho:2010vq} use the log transform of 
the energy-augmented Cobb-Douglas function with thermal energy 
quantification ($q$) to describe output for the EU-15 countries. 
Serrenho divides thermal energy into renewables, non-renewables, 
wastes, and imports. The main argument of these papers is that 
energy is an important driver of growth and needs to be included
as an essential factor of production.

**** Martin: add more here about the conclusions of each paper regarding 
the role of energy in economic growth? Add additional literature? 
Or, adjust as you see fit. Especially, are there any confidence intervals 
or statistical evaluation in each of these studies? ****

%------------------------------
\subsubsection{Energy-augmented CES Production Function (CESe)} 
\label{sec:CES}
%------------------------------

Other energy economists use an energy-augmented 
Constant Elasticity of Substitution (CES$e$) production function to 
describe economic growth. The \texttt{R} \citep{R} package 
\texttt{micEconCES} \citep{Henningsen:2011td} estimates CES production 
functions of the following forms (among others):

\begin{equation} \label{eq:CES}
  y = \gamma \: A \: \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{-1/\rho_1}; 
      A = \mathrm{e}^{\lambda (t-t_0)}
\end{equation}

\begin{equation} \label{eq:CESe}
  y = \gamma \: A \: \left\{\delta \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) e^{-\rho} \right\}^{-1/\rho}; 
      A = \mathrm{e}^{\lambda (t-t_0)}
\end{equation}

\noindent Equation~\ref{eq:CES} is a CES production function with 
capital stock ($k$) and labor ($l$) factors of production. 
Equation~\ref{eq:CESe} augments Equation~\ref{eq:CES} with energy 
using a ($kl$)($e$) nesting structure, as is typical in the literature. 
Equation~\ref{eq:CES} is a degenerate form of Equation~\ref{eq:CESe} 
where $\delta \rightarrow 1$ and $\rho \rightarrow -1$.

$\gamma$ is a fitting parameter that represents productivity. 
The fitting parameters $\rho_1$ and $\rho$ indicate elasticities 
of substitution ($\sigma_1$ and $\sigma$).
The elasticity of substitution 
between capital ($k$) and labor ($l$) is given by 
$\sigma_1 = \frac{1}{1+\rho_1}$, and
the elasticity of substitution between ($kl$) and ($e$) is given by 
$\sigma = \frac{1}{1+\rho}$. 
As $\rho_1 \rightarrow 0$, $\sigma_1 \rightarrow 1$,
and the embedded CES production function for $k$ and $l$ degenerates 
to the Cobb-Douglas production function.
Similarly, as $\rho \rightarrow 0$, $\sigma \rightarrow 1$,
and the CES production function for $(kl)$ and $(e)$ degenerates 
to the Cobb-Douglas production function.
As $\sigma \rightarrow \infty$ ($\rho \rightarrow -1$), 
($kl$) and ($e$) are perfect substitutes. 
As $\sigma \rightarrow 0$ ($\rho \rightarrow \infty$), 
($kl$) and ($e$) are perfect complements: 
no substitution is possible. Similarly, 
as $\sigma_1 \rightarrow 0$ ($\rho_1 \rightarrow \infty$),
$k$ and $l$ are perfect complements.
$\delta_1$ describes the relative importance of capital ($k$)
and labor ($l$), and
$\delta$ describes the importance of ($kl$) relative to ($e$).

Constraints on the fitting parameters include 
$\delta_1 \in (0,1)$,
$\delta \in (0,1)$,
$\rho_1 \in (-1,0) \cup (0,\infty)$, and
$\rho \in (-1,0) \cup (0,\infty)$. **** Describe here how 
we implemented the constraints. ****

\citet{VanderWerf:2008tc} evalutes various nested CES production 
functions for 12 OECD countries with respect to climate change policy. 
\citeauthor{VanderWerf:2008tc} concludes that ($kl$)($e$) nesting structure 
best fits historical data. \citet{Lecca:2011kr} also evaluate a variety of 
CES production functions using a KLEM structure to determine how the various 
options for nesting the energy factor of production affect predictive ability. 
\citet{Stern:2012ey} also use a CES production function of the form ($kl$)($e$). 
\citeauthor{Stern:2012ey} apply their CES function to the Swedish economy 
from 1800-2000 and demonstrate the importance of the transition from wood to 
fossil fuels for economic growth. \citet{Shen:2013vp} included energy 
in nested aggregated CES production functions 
for China and found that the best nesting structure is ($le$)($k$).

**** Martin: add more here about the conclusions of each paper regarding 
the role of energy in economic growth? Add additional literature? 
Or, adjust as you see fit. Especially, are there any confidence 
intervals or statistical evaluation in each of these studies? ****

%------------------------------
\subsubsection{LINEX Production Function} 
\label{sec:LINEX}
%------------------------------

A third production function, the LINear EXponential (LINEX) function 
\citep{Kummel:1980wx,Kummel:1982vy,Kummel:1985vz,Kummel:1989tf,Kummel:2002tx,Kummel:2010vz}, 
was derived from thermodynamic considerations:

\begin{equation} \label{eq:LINEX}
  y = Ae; 
  A = \mathrm{e}^{a_0\left[2 \left(1 - \frac{1}{\rho_k} \right) 
    + c_t \left(\vphantom{\frac{1}{\rho_k}}\rho_l - 1 \right)\right]},
\end{equation}

\noindent where $e$ is indexed energy and e is the base of 
the natural logarithm. The LINEX model assumes energy ($e$) 
is the only factor of production. 
The parameters $\rho_k$ and $\rho_l$ are defined as

\begin{equation} \label{eq:rho_k}
  \rho_k \equiv \frac{k}{(1/2)(l+e)}
\end{equation}

\noindent and

\begin{equation} \label{eq:rho_l}
  \rho_l \equiv \frac{l}{e}
\end{equation}

\noindent and represent 
(a) capital stock increase relative to the average of labor and energy and 
(b) labor increase relative to energy, respectively. 
Energy ($e$), as the only factor of production, is enhanced by $A$ if either

\begin{itemize}
\item{capital stock ($k$) has increased more than the average 
      of labor ($l$) and energy ($e$), i.e. if $\rho_k >$ 1, or}
\item{labor ($l$) has increased more than energy ($e$), i.e. if $\rho_l >$ 1.}
\end{itemize}

\noindent According to the LINEX model, an economy that increases 
capital stock ($k$) without a commensurate increase in labor ($l$) 
and energy ($e$) will experience an increase in output ($y$) in excess 
of its increase of energy consumption ($e$), because $A$ will be 
greater than 1. Similarly, an economy benefits by increasing labor ($l$) 
without a commensurate increase in energy ($e$). Thus, $\rho_k$ is an 
efficiency of using additional labor and energy to make 
additional capital stock available to the economy, 
and $\rho_l$ is an efficiency of using additional energy to make 
additional labor available to the economy. 
$a_0$ represents the overall importance of these efficiencies, 
and $c_t$ represents the relative importance 
of the labor efficiency term compared to the capital stock efficiency term. 
We treat $a_0$ and $c_t$ as fitting parameters; 
both are assumed constant with respect to time.

In contrast to both the Cobb-Douglas model (Equations~\ref{eq:CD} and 
\ref{eq:CDe}) and the CES model (Equations~\ref{eq:CES} and 
\ref{eq:CESe}), the LINEX model (Equation~\ref{eq:LINEX}) accounts 
for efficiencies among capital, labor, and energy directly and, 
therefore, does not employ a time-dependent 
technological progress term ($A$). \citet{Kummel:2002tx} use the 
LINEX production function to describe economic output 
for the economies of the U.S., Japan, and Germany. 
\citet{Warr:2010ga} compare the LINEX production function with 
the Cobb-Douglas production function to assess economic growth 
for Austria, the United Kingdom, the United States, and Japan. 
\citet{Ayres:2005kn} argued that introducing energy inputs 
in traditional production functions does not improve its explanatory power, 
but including useful work in a LINEX function does.

**** Martin: add more here about the conclusions of each paper 
regarding the role of energy in economic growth? Add additional literature? 
Or, adjust as you see fit. Especially, are there any confidence intervals 
or statistical evaluation in each of these studies? ****

%++++++++++++++++++++++++++++++
\subsection{Quantification of Energy} 
\label{sec:Quantification_of_Energy}
%++++++++++++++++++++++++++++++

The literature shows three quantifications of energy ($E$) used in 
energy-augmented production functions: thermal energy ($Q$), 
exergy ($X$), and useful work ($U$). Thermal energy ($Q$) 
is the heat content of an energy carrier. Exergy ($X$) is 
the maximum useful work that could be extracted from an energy carrier 
by a process that brings the system into equilibrium with 
a heat reservoir, typically the surrounding environment 
\citep[p.~78]{Ayres:2010ug}. Like \citeauthor{Ayres:2010ug}, 
we use chemical exergy (as opposed to thermal exergy or mechanical exergy) 
as our quantification for exergy ($x$). Useful work ($U$) 
includes the direct energy that goes into an economy, including work 
from machines, human muscles, and animals. Firms pay for thermal energy ($Q$), 
but useful work ($U$) alters the size and shape of materials, 
keeps the lights on, and transports goods to consumers. 
Not all forms of energy are created equal in terms of their ability 
to generate economic activity \citep{Cleveland:2000vl, Warr:2010ht}. 
Both exergy ($X$), and to a greater extent, useful work ($U$) 
quantifications of energy are intended to account for the usefulness 
of energy for generating economic activity, thereby reflecting 
the ``quality'' of each energy type and the efficiency of machines 
and devices typically used in energy conversion processes.

Most energy economists use thermal energy ($q$) 
for the energy factor of production ($e$) in energy-augmented 
production functions, including \citet{Stern:2012ey}, 
\citet{Froling:2009vo}, \citet{Serrenho:2010vq}, and 
\citet{Kummel:1980wx,Kummel:1982vy,Kummel:1985vz,Kummel:1989tf,Kummel:2002tx,Kummel:2010vz}. 
\citet{Ayres:2003ec} introduced the idea that exergy ($x$) and 
useful work ($u$) were important factors of production.

**** Martin: add more here about the conclusions of each paper 
regarding the role of energy in economic growth? 
Add additional literature? Or, adjust as you see fit. 
Especially, are there any confidence intervals or 
statistical evaluation in each of these studies? ****

We know of no statistically-rigorous comparisons of the 
predictive ability of each energy quantification in the 
various types of production functions discussed in 
Section~\ref{sec:Prod_Func_Forms}. We analyze all 
three quantifications of energy ($Q$, $X$, and $U$) 
for developed economies and thermal energy ($Q$) and 
exergy ($X$) for all other economies.

%++++++++++++++++++++++++++++++
\subsection{Economy Types} 
\label{sec:Economy_Types}
%++++++++++++++++++++++++++++++

In the energy-economic growth literature, the focus is 
almost exclusively on developed economies from North America 
or Europe. Examples include \citet{Kummel:1980wx} applying 
LINEX to describe the West German economy, \citet{Serrenho:2010vq} 
applying Cobb-Douglas variants to EU-15 countries, 
\citet{VanderWerf:2008tc} applying CES nesting variants 
to the 12 OECD countries, \citet{Lecca:2011kr} applying 
CES nesting variants to the Scottish economy, \citet{Stern:2012ey} 
applying CES to describe the Swedish economy, and \citet{Warr:2012cg} 
using LINEX and Cobb-Douglas to describe the Austrian, 
American, British, and Japanese economies.

But, the near-exclusive focus on developed economies closes off 
the possibility of discovering any role that energy plays 
in the economic development process. In this paper we analyze at least 
two economies in each of the following four categories: 
Developed, 
NICs,\footnote{Newly Industrialized Countries}
OPEC, and 
LDCs.\footnote{Lesser Developed Countries} 
(See Table~\ref{tab:Economy_Examples}.)

\begin{table} \caption{Example economies and years of evaluation.} \label{tab:Economy_Examples} 
  \begin{center}
    \begin{tabular}{r c c l l} 
      \multirow{3}{*}{Developed} & \includegraphics[height=0.11in]{flag/flag_US.png} & US & USA & \multirow{3}{*}{1980--2011}\\
                                 & \includegraphics[height=0.11in]{flag/flag_UK.png} & UK & United Kingdom & \\
                                 & \includegraphics[height=0.11in]{flag/flag_JP.png} & JP & Japan          & \\
      \hline
      \multirow{2}{*}{NICs}      & \includegraphics[height=0.11in]{flag/flag_CN.png} & CN & China & \multirow{2}{*}{1991--2011}        \\
                                 & \includegraphics[height=0.11in]{flag/flag_ZA.png} & ZA & South Africa & \\
      \hline
      \multirow{2}{*}{OPEC}      & \includegraphics[height=0.11in]{flag/flag_SA.png} & SA & Saudi Arabia & \multirow{2}{*}{1991--2011} \\
                                 & \includegraphics[height=0.11in]{flag/flag_IR.png} & IR & Iran         &  \\
      \hline
      \multirow{2}{*}{LDCs}      & \includegraphics[height=0.11in]{flag/flag_TZ.png} & TZ & Tanzania & \multirow{2}{*}{1991--2011}     \\
                                 & \includegraphics[height=0.11in]{flag/flag_ZM.png} & ZM & Zambia      & \\
    \end{tabular}
  \end{center}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Sources of Data}
\label{sec:Sources_of_Data}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

To assess energy-augmented models of economic growth along the 
coordinates of analysis outlined in Section~\ref{sec:Coordinates}, 
historical data for economic output ($Y$), capital stock ($K$), 
labor ($L$), thermal energy ($Q$), exergy ($X$), and useful work ($U$) 
are required. Subsections below describe sources for and calculations 
to obtain historical data. A final subsection presents the historical 
data in graphical form.

%++++++++++++++++++++++++++++++
\subsection{Economic Output} 
\label{sec:Economic_Output_Data}
%++++++++++++++++++++++++++++++

We measure economic output ($Y$) by gross domestic product (GDP). 
GDP data in real 2005 U.S. dollars for developed countries (1980--2011) 
and all other countries (1991--2011) was obtained from 
the World Bank's Global Economic Prospects (GEP) database 
\citep{WorldBankGEP:2013a}.

%++++++++++++++++++++++++++++++
\subsection{Capital Stock} 
\label{sec:Capital_Stock_Data}
%++++++++++++++++++++++++++++++

Gross capital stock in real 2005 U.S. dollars 
for developed countries (1980--2011) and 
all other countries (1991--2011) was obtained from 
the World Bank's Global Economic Prospects (GEP) database 
\citep{WorldBankGEP:2013a}.

%++++++++++++++++++++++++++++++
\subsection{Labor and Employment}
\label{sec:Labor_Data}
%++++++++++++++++++++++++++++++

Total employment and average hours worked per worker per year 
for the developed nations (United States, United Kingdom, and Japan) 
for 1980--2011 were obtained from the United States 
\citet[Tables 6 and 7]{BLS:2013a}. Total hours worked per year was 
calculated by multiplying total employment by average hours worked 
per worker per year.

The World Bank's World Development Indicators database 
\citep{WorldBankWDI:2013a} provides the employment to population 
ratio for people fifteen and older, total population, and a percentage 
of people under fifteen for other economies (South Africa, China, Iran, 
Saudi Arabia, Tanzania, and Zambia) for 1991--2011. The population 
of people under fifteen years of age was subtracted from total population 
to calculate the population of people fifteen and older. 
Total employment was calculated by multiplying the employment to 
population ratio of people fifteen and older by the population 
of people fifteen and older.

%++++++++++++++++++++++++++++++
\subsection{Thermal Energy, Exergy, and Useful Work} 
\label{sec:Energy_Data}
%++++++++++++++++++++++++++++++

To account for all energy input to an economy, we summed consumption data 
for many energy sources, including coal, oil and petroleum products, 
dry natural gas and natural gas plant liquids, nuclear power generation, 
and renewable energy sources (solar, wind, geothermal, biomass/fuel wood, 
and hydro). In the style of work by Warr and Ayres 
\citep{Ayres:2003ec, Ayres:2005kn, Warr:2010ht, Warr:2010ga, Warr:2012cg}, 
energy from human and animal muscle power was also included. 
Similar to the economic data discussed in 
Sections~\ref{sec:Economic_Output_Data}--\ref{sec:Labor_Data}, 
we generated energy data for developed countries from 1980--2011, 
with the exception of useful work data (1980--2000). For all other countries, 
we generated thermal energy and exergy data for 1991--2011. 
Details on each energy quantifications are given in subsections below.

%------------------------------
\subsubsection{Thermal Energy} 
\label{sec:Q_Data}
%------------------------------

Thermal energy ($Q$) for fuels (in units of TJ/year) was calculated by

\begin{equation} \label{eq:thermal_energy}
  Q_{fuel} = C_{fuel}h_{fuel}
\end{equation}

\noindent where $C_{fuel}$ is the consumption rate of the fuel 
(in units of short tons/year for coal, barrels/year for oil and petroleum, 
ft$^3$/year for dry natural gas and natural gas plant liquids, 
and m$^3$/year for biomass/fuel wood) and $h_{{fuel}}$ is 
the heat content of the fuel (in kJ/short ton for coal, kJ/barrel for oil, 
kJ/ft$^3$ for dry natural gas and natural gas plant liquids, 
and kJ/kg for biomass/fuel wood). Fuel wood density of 300~kg/m$^3$ 
was taken from \citet{Ragland:2013wp} at the low end of the range 
of density of dry wood and bark.

Yearly consumption rates ($C_{fuel}$) of coal, oil and petroleum, 
dry natural gas and natural gas plant liquids, and biofuels 
(ethanol and biodiesel) were obtained from the \citet{EIA:2013a} 
for all economies. Yearly average heat content data for coal, oil, 
and natural gas and natural gas plant liquids was obtained from the 
EIA for all economies. We used lower (net) heating values of 
26.8~MJ/kg \citep[Table~A-27]{Cengel:2011ww} and 37.8~MJ/kg 
\citep[Table~7.3]{Gupta:2010vg} for the heat content of ethanol 
and corn-based biodiesel, respectively. Yearly consumption and 
heat content data for biomass/fuel wood was obtained from the 
``Fuelwood'' time series of the FAO Statistics Database 
\citep{FAO:2013a}.

For renewable and nuclear energy sources, yearly electricity 
generation (in BTU/year) was available directly from 
the \citet{EIA:2013a} for all economies, so 
Equation~\ref{eq:thermal_energy} was unnecessary.

Animal and human muscle work was estimated using an approach 
similar to \citet{Ayres:2010ug}, using some data from Warr's 
REXS database \citep{Warr:2012a} and results from 
\citet{Wirsenius:2000vi}. For animal work, our analysis was 
restricted to draught animals (horses and mules). Muscle work 
from animals ($Q_{animals}$) was calculated using

\begin{equation} \label{eq:animal_muscle}
  Q_{animals} = Q_{m,horse}N_{horses} + Q_{m,mule}N_{mules}
\end{equation}

\noindent where $Q_m$ is metabolizable energy (in kcal/day-animal) 
and $N$ is the number of animals. Metabolizable energy values 
($Q_{m,horses}$ = 19,110 kcal/day-horse and $Q_{m,mules}$ = 
7,215 kcal/day-mule) were taken from \citet{Warr:2012a}. 
$Q_m$ represents the amount of energy in the feed consumed by 
animals for a day of work, and they are assumed constant for 
all countries and years. The number of horses and mules for 
each economy was taken from the ``Live Animals'' section of the 
FAO Statistics online database \citep{FAO:2013a}.

Human muscle power was calculated in a similar manner:

\begin{equation} \label{eq:human_muscle}
  Q_{workers} = Q_{m,workers} N_{workers}
\end{equation}

\noindent where the value of $Q_{m,workers}$ is the ``the amount 
of food available for human consumption'' (i.e., at the wholesale level) 
taken from the 
\citet[``\texttt{Dietary~Energy~Cons.~Countries}'' tab]{FAO:2013b}. 
The value for $N_{workers}$ was obtained from the labor time series 
for each economy as discussed in Section~\ref{sec:Labor_Data}.

%------------------------------
\subsubsection{Exergy} 
\label{sec:X_Data}
%------------------------------

Exergy ($X$), expressed in TJ/year, was calculated by multiplying 
the thermal energy ($Q$) values discussed in Section~\ref{sec:Q_Data} 
by a unitless exergy-to-thermal-energy allocation factor 
($f_{exergy-allocation}$) shown in Table~\ref{tab:Exergy_Allocation}.

\begin{equation} \label{eq:exergy}
  X_{fuel} = f_{exergy-allocation} Q_{fuel}
\end{equation}

\begin{table} \caption{Exergy allocation factors.} \label{tab:Exergy_Allocation} 
  \begin{center}
    \begin{tabular}{r l} 
      Fuel                                    & $f_{exergy-allocation}$ \\
      \hline
      Coal                                    & 1.088    \\
      Oil and petroleum                       & 1.073   \\      
      Dry natural gas and                     & \multirow{2}{*}{1.04}    \\
      natural gas liquids                     &         \\
      Corn-based biodiesel                    & 1.07  \\
      Biomass/fuel wood                       & 1.15   \\
      Ethanol biofuel                         & 1.099 \\
      Nuclear, wind,                          & \multirow{2}{*}{1.0}   \\
      solar PV, and hydro                     &                        \\
    \end{tabular}
  \end{center}
\end{table}

\noindent For carbon-based energy sources, $f_{exergy-allocation}$ 
is the ratio of chemical exergy to lower (net) heating value. 
The exergy allocation factors for coal, oil, natural gas, 
biodiesel, and biomass/fuel wood were obtained from 
\citet[Table~A-2]{Ayres:2003ec}. The exergy allocation ratio for 
liquid ethanol was calculated from the standard molar chemical exergy 
of 1357.7~kJ/mol \citep[Table~II]{Szargut:1988wv} and its lower 
(net) heating value of 26.8~MJ/kg \citep[Table~A-27]{Cengel:2011ww}. 
$f_{exergy-allocation}$ for each fuel was assumed to be constant 
through time across all economies. For nuclear, wind, 
solar photovoltaic, and hydro energy sources, the electricity generated 
is already pure exergy, so $f_{exergy-allocation}$ is unity.

To calculate the exergy of animal and human muscle power, 
the food-to-work conversion efficiency ($\eta_{food \rightarrow work}$) 
and the ratio of food intake to food available at wholesale 
($f_{intake/wholesale}$) were included.

\begin{equation} \label{eq:animal_muscle_exergy}
  X_{animals} = \eta_{food \rightarrow work,animals} Q_{animals}
\end{equation}

\begin{equation} \label{eq:human_muscle_exergy}
  X_{workers} = \eta_{food \rightarrow work,workers} f_{intake/wholesale} Q_{workers}
\end{equation}

\noindent $\eta_{food \rightarrow work, animals}$ and 
$\eta_{food \rightarrow work, workers}$ are animal and worker efficiency 
of converting food into energy, respectively. The animal feed-to-work 
conversion efficiency ($\eta_{feed \rightarrow work, animals}$ = 0.4) 
was taken from \citet{Warr:2012a} and assumed equal across all economies. 
The food-to-work efficiency for workers 
($\eta_{food \rightarrow work, workers}$ = 0.2) was taken from 
\citet[Table~B.1]{Ayres:2010ug}. Following \citet[p.~315]{Ayres:2010ug} 
and methodology from \citet{Warr:2012a}, an additional factor 
($f_{intake/wholesale}$) was included to account for ``wastage and 
losses of food in the household, e.g. during storage, in preparation 
and cooking, as plate-waste, or quantities fed to domestic animals 
and pets'' \citep[\texttt{notes} tab]{FAO:2013b}. 
Values of $f_{intake/wholesale}$ were calculated from 
\citet[Table~3.3 for wholesale and Table~3.4 for intake]{Wirsenius:2000vi}. 
$f_{intake/wholesale}$ varies by country as shown in 
Table~\ref{tab:f_intake_wholesale}. A similar intake-to-wholesale ratio 
could also be applied to animal work, but data to calculate the ratio was 
unavailable from \citet{Wirsenius:2000vi}, and $f_{intake/wholesale,animals}$ 
is assumed to be unity.\footnote{The error in total exergy input to 
any economy caused by assuming $f_{intake/wholesale,animals} = 1.0$ 
is very small, because animal muscle exergy is a small fraction of the 
total exergy supply to any economy, with a maximum of 0.41\% for China in 1992. 
It could also be argued that $f_{intake/wholesale, animals}$ is likely 
to be closer to unity than $f_{intake/wholesale, workers}$.}

\begin{table} \caption{Food intake-to-wholesale ratios 
                        ($f_{intake/wholesale}$) for all countries.} 
                        \label{tab:f_intake_wholesale}
  \begin{center}
    \begin{tabular}{r c c c c c c c c c}
      Country & US & UK & JP & CN & ZA & SA & IR & TZ & ZM \\
      \hline
      $f_{intake/wholesale}$ & 0.62 & 0.65 & 0.76 & 0.76 & 0.87 & 0.72 & 0.72 & 0.72 & 0.72\\
    \end{tabular}
  \end{center}
\end{table}

%------------------------------
\subsubsection{Useful Work} 
\label{sec:U_Data}
%------------------------------

The conversion of a raw energy carrier (e.g., coal) into 
productive energy for an economy (e.g., electricity) necessarily 
involves inefficiencies. In the case of coal consumption, 
only about one-third of the coal exergy is converted to useful electricity: 
the rest is waste heat. The concept of ``useful work,'' pioneered by 
Ayres and Warr \citep{Ayres:2003ec, Ayres:2005kn}, builds upon 
exergy analysis to consider efficiencies of both energy production and enduse. 
To calculate useful work ($U$), the total exergy input to an 
economy ($X$) is decomposed into several end uses: 
high temperature industrial heat (e.g., steel production), 
medium temperature heat (e.g., steam production), 
low temperature space heat (for residential heating), 
mechanical drive (energy used for prime mover engines of all types except electric motors), 
electricity (excluding lighting), 
lighting (from both electricity and fuels), and 
muscle work (from animals and workers) as shown in the following equation:

\begin{equation} \label{eq:useful_work}
  U = \sum_i{\varepsilon_i \phi_i X}
\end{equation}

\noindent where the subscript $i$ indicates the energy enduse, 
$\varepsilon_i$ is the efficiency of converting exergy into 
useful energy for each enduse, and $\phi_i$ gives 
the fraction of exergy input to an economy that goes toward 
the $i^\mathrm{th}$ enduse.

Values of $\varepsilon_i$ and $\phi_i$ were calculated from 
the REXS database supplied by \citet{Warr:2012a}. 
Values of $X$ were calculated by methods discussed in 
Section~\ref{sec:X_Data}. For most countries, values of 
$\phi_i$ are not constant with time, as economic sectors with 
different distributions of raw exergy inputs grow at different rates. 
\citeapos{Warr:2012a} REXS database contains information for 
developed economies (United States, United Kingdom, and Japan) 
from 1900--2000, and we used data from the 1980--2000 timeframe. 
Extending the REXS database to the BRICS, petrol, and LDC economies 
is beyond the scope of this work, because longitudinal data for 
the efficiency of different end uses ($\varepsilon_i$) and 
the distribution of exergy sources to end uses 
($\phi_i$) are extremely difficult to obtain.

%++++++++++++++++++++++++++++++
\subsection{Historical Data} 
\label{sec:Historical_Data}
%++++++++++++++++++++++++++++++

Figure~\ref{fig:Historical_Lattice_Graph} presents historical data, 
including indexed values for 
output ($y$), 
capital stock ($k$), 
labor ($l$), and 
the various energy quantifications ($q$, $x$, and $u$) 
for 1980--2011 (developed economies) and 1991--2011 (other economies). 
Factors of production track each other closely, 
and high covariances among these variables can be seen in 
Tables~\ref{tab:Covariance_US}--\ref{tab:Covariance_ZM}. 
**** Discuss the covariances in more detail after we 
finalize the covariances for U. ****

<<Historical_Lattice_Graph, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="Historical data: indexed GDP ($y$), capital stock ($k$), labor ($l$), thermal energy ($q$), exergy ($x$), and useful work ($u$).">>=
createHistoricalLatticeGraph()
@

Figure~\ref{fig:Historical_Lattice_Graph} indicates the importance 
of energy quantification for economic growth modeling. 
Thermal energy ($q$) and exergy ($x$) grow at similar rates, 
unless there is a significant raw energy carrier substitution in the economy 
(such as China's rapid increase of coal and natural gas consumption 
relative to petroleum since 2000). The growth rate of $u$ is 
significantly different from $q$ and $x$ for most countries, however. 
\citet{Warr:2012cg} show that from 1950-1990, useful work ($u$) 
grew much faster than $q$, $x$, $k$, or $l$ due to efficiency gains 
in the U.S. economy. The results of \citeauthor{Warr:2012cg} suggest 
that the time-dependent technological progress term ($A$) 
in the Cobb-Douglas and CES production functions 
(Equations~\ref{eq:CD}--\ref{eq:CESe}) may be a proxy for 
increasing energy efficiency in the U.S. economy.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Parameter Estimation}
\label{sec:ParameterEstimation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This section should cover both technique and results. 
Organize along coordinates of analysis.

1. Discuss parameter estimation for each model here.

2. Show scatter plot results.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion of Results}
\label{sec:DiscussionOfResults}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Drive towards new insights from our stats tested comparative approach

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:Conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Future Work}
\label{sec:FutureWork}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}
\label{sec:Acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Funding for Caleb Reese and Lucas Timmer was provided 
by the Jack and Lois Kuipers Applied Mathematics Endowment and 
the Calvin College Alumni Association. 
Partial support for Matthew Kuperus Heun was 
provided by a Calvin College Research Followship. 
Partial funding for Martin de Wit was provided 
by the South African National Research Foundation (NRF).\footnote{Any opinion, 
findings, and conclusions or recommendations expressed herein are those 
of the authors, and, therefore, NRF does not accept any liability in regard thereto.} 
The authors are grateful to Benjamin S. Warr (INSEAD) for providing 
the REXS database \citep{Warr:2012a}. 
The authors thank Loren L. Heun (Western Michigan University) 
for insightful critique of the statistical analyses presented herein.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Reproducible Research}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In the spirit of Reproducible Research \citep{Gandrud:2013vx}, 
all data, spreadsheets, \texttt{R} code \citep{R}, and other materials 
associated with this paper can be found at\\*
\protect\url{https://github.com/MatthewHeun/Econ-Growth-R-Analysis}.

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix
%% \section{}
%% \label{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Covariance Tables}
\label{sec:Covariance_Tables}
\setcounter{table}{0} % Restarts table number counter at 0 in this section.
\setcounter{figure}{0} % Restarts figure number counter at 0 in this section.
\setcounter{equation}{0} % Restarts equation counter to 0 in this section.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% References
%%
%% Following citation commands can be used in the body text:
%%
%%  \citet{key}  ==>>  Jones et al. (1990)
%%  \citep{key}  ==>>  (Jones et al., 1990)
%%
%% Multiple citations as normal:
%% \citep{key1,key2}         ==>> (Jones et al., 1990; Smith, 1989)
%%                            or  (Jones et al., 1990, 1991)
%%                            or  (Jones et al., 1990a,b)
%% \cite{key} is the equivalent of \citet{key} in author-year mode
%%
%% Full author lists may be forced with \citet* or \citep*, e.g.
%%   \citep*{key}            ==>> (Jones, Baker, and Williams, 1990)
%%
%% Optional notes as:
%%   \citep[chap. 2]{key}    ==>> (Jones et al., 1990, chap. 2)
%%   \citep[e.g.,][]{key}    ==>> (e.g., Jones et al., 1990)
%%   \citep[see][pg. 34]{key}==>> (see Jones et al., 1990, pg. 34)
%%  (Note: in standard LaTeX, only one note is allowed, after the ref.
%%   Here, one note is like the standard, two make pre- and post-notes.)
%%
%%   \citealt{key}          ==>> Jones et al. 1990
%%   \citealt*{key}         ==>> Jones, Baker, and Williams 1990
%%   \citealp{key}          ==>> Jones et al., 1990
%%   \citealp*{key}         ==>> Jones, Baker, and Williams, 1990
%%
%% Additional citation possibilities
%%   \citeauthor{key}       ==>> Jones et al.
%%   \citeauthor*{key}      ==>> Jones, Baker, and Williams
%%   \citeyear{key}         ==>> 1990
%%   \citeyearpar{key}      ==>> (1990)
%%   \citetext{priv. comm.} ==>> (priv. comm.)
%%   \citenum{key}          ==>> 11 [non-superscripted]
%% Note: full author lists depends on whether the bib style supports them;
%%       if not, the abbreviated list is printed even when full requested.
%%
%% For names like della Robbia at the start of a sentence, use
%%   \Citet{dRob98}         ==>> Della Robbia (1998)
%%   \Citep{dRob98}         ==>> (Della Robbia, 1998)
%%   \Citeauthor{dRob98}    ==>> Della Robbia

%% References with bibTeX database:

\bibliographystyle{model2-names}
%%\bibliography{<your-bib-database>}
\bibliography{EconGrowth.bib}

%% Authors are advised to submit their bibtex database files. They are
%% requested to list a bibtex style file in the manuscript if they do
%% not want to use model2-names.bst.

%% References without bibTeX database:

% \begin{thebibliography}{00}

%% \bibitem must have one of the following forms:
%%   \bibitem[Jones et al.(1990)]{key}...
%%   \bibitem[Jones et al.(1990)Jones, Baker, and Williams]{key}...
%%   \bibitem[Jones et al., 1990]{key}...
%%   \bibitem[\protect\citeauthoryear{Jones, Baker, and Williams}{Jones
%%       et al.}{1990}]{key}...
%%   \bibitem[\protect\citeauthoryear{Jones et al.}{1990}]{key}...
%%   \bibitem[\protect\astroncite{Jones et al.}{1990}]{key}...
%%   \bibitem[\protect\citename{Jones et al., }1990]{key}...
%%   \harvarditem[Jones et al.]{Jones, Baker, and Williams}{1990}{key}...
%%

% \bibitem[ ()]{}

% \end{thebibliography}

\end{document}

%%
%% End of file `elsarticle-template-2-harv.tex'.