### 2013-08-01 EconGrowth2.Rnw
Added resample scatter plots for (le)k and (ek)l nestings for CES.

% This article has been prepared for publication in Energy Economics in RStudio with knitr.
% According to http://www.elsevier.com/author-schemas/the-elsarticle-latex-document-class, we should be using the
% elsarticle.cls file.
% According to http://cdn.elsevier.com/assets/pdf_file/0006/109392/journal_refstyles.pdf, we should be using
% elsarticle-template-2-harv.tex as the template for the text.
% Furthermore, we should be using model2-names.bst for the bibliographic references.
% The approach here is to load the frontmatter and backmatter from elsarticle-template-2-harv.tex
% both ahead of and behind the text for our paper.
% -- Matthew Kuperus Heun, 2013-01-18

%% This is file `elsarticle-template-2-harv.tex',
%%
%% Copyright 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle  Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for Elsevier's document class `elsarticle'
%% with harvard style bibliographic references
%%
%% $Id: elsarticle-template-2-harv.tex 155 2009-10-08 05:35:05Z rishi $
%% $URL: http://lenova.river-valley.com/svn/elsbst/trunk/elsarticle-template-2-harv.tex $
%%
\documentclass[preprint,authoryear,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,authoryear,1p,times]{elsarticle}
%% \documentclass[final,authoryear,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,authoryear,3p,times]{elsarticle}
%% \documentclass[final,authoryear,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,authoryear,5p,times]{elsarticle}
%% \documentclass[final,authoryear,5p,times,twocolumn]{elsarticle}

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%% or use the graphicx package for more complicated commands
%% \usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

\usepackage{soul} % Provides strikethrough text
\usepackage{float} % Allows precise positioning of tables and figures within the text
\usepackage{amsmath} % Allows \begin{equation*} \end{equation*} for unnumbered equations
\usepackage{multirow} % To create multirow tables
\usepackage{hyperref} % To create hyperlinks in the paper

% From http://economics.utoronto.ca/osborne/latex/BIBTEX.HTM
\newcommand{\citeapos}[1]{\citeauthor{#1}'s (\citeyear{#1})} % Posessive citations.

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon (default)
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   authoryear - selects author-year citations (default)
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%   longnamesfirst  -  makes first citation full author list
%%
%% \biboptions{longnamesfirst,comma}

% \biboptions{}

\journal{Energy Economics}

\begin{document}

<<setup, echo=FALSE, message=FALSE, eval=TRUE>>=
require(lattice)
require(latticeExtra)
require(ggplot2)
require(car)
require(mosaic)
require(xtable)
require(nlmrt)
require(micEconCES)
# Provides access to data melting. 
# See http://cran.r-project.org/web/packages/reshape2/reshape2.pdf
require(reshape2) 
# tikz allows use of LaTeX formatting and font in graphs. 
# Allows for a consistent look across the paper.
# See http://r-forge.r-project.org/R/?group_id=440 for instructions 
# on installing tikzDevice.
# require(tikzDevice) 
trellis.par.set(theme=theme.mosaic())
options(width=75)
options(show.signif.stars=TRUE)
opts_chunk$set(
  # ********************************************************
  eval=TRUE,
  # ********************************************************
  # tikz allows LaTeX code in graphical output. 
  # E.g., "$y$" for a variable name in the legend for a graph.  
  # dev='tikz',
  tidy=FALSE,
  comment=NA,
  # Tells whether to cache output from chunks are saved. Cacheing saves time. 
  # However, references to LaTeX tables DO NOT WORK if cacheing is turned on.
  cache=FALSE,
  warning=FALSE,  #Tells whether to show warnings in the output.
  # Turns off messages for all chunks. 
  # Set TRUE on an individual chunk to see it.
  message=FALSE, 
  # Tells whether to echo code for all chunks. 
  # Set TRUE on an individual chunk to see its code.
  echo=FALSE     
)
source("Econ-Growth-Functions2.R")
source("Graphics.R")
@

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{An Empirical Analysis of the Role of Energy in Economic Growth}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

\author[CalvinEngr]{Caleb Reese}
\author[CalvinEngr]{Lucas Timmer}
\author[CalvinEngr]{Matthew Kuperus Heun\corref{cor1}}
\ead{mkh2@calvin.edu}
\author[Stellenbosch]{Martin de Wit}
\author[CalvinMath]{Randall J. Pruim}

\cortext[cor1]{Corresponding author}
\address[CalvinEngr]{Engineering Department, Calvin College, Grand Rapids, MI 49546, USA}
\address[Stellenbosch]{Faculty of Economics and Management Sciences, School of Public Leadership, Stellenbosch University, P.O. Box 610, Bellville 7535, South Africa}
\address[CalvinMath]{Mathematics \& Statistics Department, Calvin College, Grand Rapids, MI 49546, USA}

\begin{abstract}
%% Text of abstract
**** Add abstract ****
\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
economic growth \sep energy \sep Cobb-Douglas \sep CES \sep LINEX
%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
\end{keyword}

\end{frontmatter}

% \linenumbers
%% main text

\section*{Notes from Martin}

**** Additional references requested by Martin. Incorporate into paper? ****

\citet{Lindenberger:2011dx}

\citet{Stresing:2008cq}

**** Notes from Martin ****

To be included somewhere.

Big picture
Why do we do this? Want to link energy and economics on national level 
and in a modeled approach,informing energy policy during energy transitions.
Usually macro energy economic models rely on PF with energy included. 
Key papers using CES in energy economics incl high impact such as 
Stern, Gerlagh, Tol etc. 
Cite in introduction met energy economists using PF approaches, specially CES.
But, which form to choose and using which parameters? 
Literature is inconclusive, very little actual empirical testing of such functions. 
Mostly CES in CGE models used by Wbank and others. 
CD adviced to be used by CBO. 
How do we select a functional form and what are the parameters to be used? On what basis do we select?

We aim. to search for empirically robust solution, one that stands up to statistical testing.

What we find is counterintuitive for many cases: dropping energy from the PF has little impact, 
but some notable exceptions. 
On other hand, only including energy in a PF of LINEX form also yields good results and stats tests. 
Is not a simple case for dropping energy, but acknowledging its key role. 
Do we support the call for better bottom-up understanding? 
What do we say on the limitations of macroeconomic modelling approaches altogether? 
(Note such limitations are well known in ME literature, for example capital controversy Robinson).
We speculate on implications for choosing PF in energy economic models that inform growth policy.

Points on paper so far
1. Does the relative small differences in fit between different production functions warrant 
such a large debate on validity of one vs other? Clearly not. 
CD been used by economists for ease of use and extreme flexibility (Miller).  
PF generally used on instrumentalist grounds if give reasonable good description of data. 
NK econ theory provide poor causal explanation of observed data, 
but yet CD shows very good fit (miller, p17). 
CES is  more general, but is a liability in forecasting because difficult 
to get estimate of EoS that is consistent across studies (Miller, p26). 
Aggregation is problem of both CD and CES functional forms, 
so being careful when making strong inferences that cannot be justified.

2. We think we can contribute by emphasising the need 
for proper stats testing or inconsistent reporting on use of preferred functions. 
Problems with empirical testing presented by Shen, Serrenho.
What are our issues with multivariate OLS (Serrenho)?

3. Are we joining the forced choice between production functions or 
acknowledge the similar problems with aggregate prod functions in general? 
Do we see this as sufficient evidence to doubt the top-down approach to econ growth altogether?

4. R2 is a very crude measure and only very minor improvements 
are achieved when others factors such as energy is added in a CD. 
We concur. Puzzling is conclusion that such minor improvements 
does prove the need to include energy in PF (Serrenho).

5. We do think it is a choice between CD and CES (Miller) as 
both have serious deficiencies after stats tests.

6. One central issue within PF is elasticity of subs (EoS). 
That is the \%ch in factor proportions ( k,l, etc) 
because of one unit ch in MRTS. 
MRTS is rate at which one factor can be substituted for another holding output constant. 
EoS is low in Leontief cases, L shaped isoquants.

7. Miller makes the point that even if parameters in model are stats artifacts, 
they are still useful for forecasting. Do we agree?

8. Is Ockham's razor relevant to what we are doing?

9. Assumptions of CD are strongly rejected by empirical work, Miller still makes case for using CD.

11. What about energy augmented

10. TZ case
CDno e: almost 80\% l
CD with e: 25\% energy, rest c,importance of l drops to 0
CES no e: 20\% c (sigma), 80\% l, but large range
CES with e: fig 35 e may be NB
LINEX: one of lowest ct parameters (low relative importance of labour eff compared to cap eff)
Does this mean that relative small increases in cap eff can really lift the economy?

11. ZA case
CD no e: roughly half c, l
CD with e: roughly half, c, e, importance of l drops to 0
CES no e: 10\% c, 25\% l ??
CES with e: fig 35 e may be NB
LINEX: one of highest ct parameters (high relative importance of labour eff compared to cap eff)
Does this mean that relative small increases in labour eff can really lift the economy?

12. Capital starved Zambia. Very high importance. Understand CES param better.

Discussion: What is Next?

a. A comparative analysis between diff PF reveal interesting 
enough information to warrant a more diffused approach to countries. 
Thus, we have a case for doing this for  all countries in the world 
(or at least those representative of countries in different groups 80/20)

b. If c is relatively more important than labour or energy 
for growth and if capital stock is low, a situation one would expect in LDCs, 
a relative small addition of capital stock can give relatively large returns in terms of growth.

c. Now it becomes interesting, in a 'stages of growth' approach, 
using LT econ data it may be the case that the importance 
of factors are different across growth stages. 
It would need a composite type of analysis - both stock and importance.

d. Is it possible to have evidence of diff stocks and diff importances across phases of growth? 
Maybe test this for a data rich nation that has worked through these stages, 
we try to identify them and do partial analysis on time periods coinciding with these stages.

Should we eliminate $u$ data? Only 3 countries. Only 20 years. Doesn't add much insight. 
Too bad, but perhaps necessary to simplify paper.

Discussions:

Counterfactuals: energy, cap, labour what ifs for future.

interpretation of type of economies, what does the data presentation say 
on relative change in each input and change in GDP?

Forcing const elast of subs in CES thru nested structures violates 
observation that more than two factor will not be const. elasticities Uzawa 1962.

R programming dealt with nested CES see paper discussion on Kemfert.

Next steps:

Stages of growth crowd implicitly assumes countries at similar stage of growth will
have similar values for $\alpha$, $\beta$, $\gamma$. However, our results show
that not to be true. Needs further exploration to cover more (all?) countries.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} 
\label{sec:Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Thoughts on literature review and contextualisation of our work

Imagine ourselves walking into a conversation on the role of energy in the economy.

1. Listen to what others say

Identify key issues and important voices

\begin{itemize}

\item{energy has become an important strategic issue for economic growth, shocks in 1970s, spikes in energy prices generated renewed interest (IEA, IMF), supported by analysis on more structural depletion and technology race over time (Hamilton), cite reviews on energy and economics (Stern,...), cite articles that show close correlations energy use and economic growth (Cleveland etc)}

\item{economic theoretical frameworks adjusted: energy theory of value; Energy fundamental in biophysical economics (Cleveland, Hall), energy as one cost in production set; energy plays a role in production (extended Solow-Swan or CES (who introduced CES first) framework)}

\item{how to account for energy on a macro economic modeling? taken up in economic growth theory by (1) including energy in PF (stern, froling, tahvonen, Gerlagh, Tol, Kemfert, FEEM crowd, Cooper, Nel) or (2) new thermodynamically rooted Function (kümmel, lindenberger, ayres, warr). Mention energy in endogenous growth theory models (smulders, Mulder, vd Bergh) in SD modelling}

\item{countries are going through different stages in their development, and the role of energy in such phases is now started to be explored (frohling, tahvonen, dalgaard)}

\item{some have started to do work on empirical/statistical testing of PF with energy (Shen, walley, Serrenho,) and some have started doing comparative analysis of PF CD vs CES (miller)}

\end{itemize}

2. Find our own voice

What is our own argument?

We argue that macroeconomic modelling is a reality (being used for policy analysis), 
but needs to be used in a very careful way. 
Throughout decades, there are enough signals that 
macroeconomic growth theory has serious limitations. 
Not revisiting but refer to reviews (mishra, others).

We argue that the basis on which choice version of energy-augmented macroeconomic model 
and parameters are made, needs more detailed work, 
specifically detailed statistical testing and/or data analysis, including
the quantification of energy ($q$, $x$, or $u$).

We also argue that different types of countries may need different PF parameters 
to be used in modelling approaches, 
leave possibility open for different explanations of growth 
in different countries and at different stages of their development.

We present a comparative analysis among countries and of commonly used approaches in PF, 
focussing on the insights one may drawn on a detailed comparative analysis, 
rather then run with one selected PF.

What is general tone of conversation? Controversial, established? Cross-disciplinary?
We provide insights from cross-disciplinary work among statistics, engineering,
and environmental economics. This is a new approach to issues that have been 
discussed in recent years in the literature.

We are taking issue with untested (statistically) applications of functional forms 
of energy-augmented PF across different countries, with different energy quantifications. 
Also taking issue with idea that importance of energy in growth 
also means it is necessary to include it as a factor in PF (unnecessary vs. unimportant). 
Opens up a argument for the need of richer than just aggregate modelling approaches.

Who makes arguments we want to amplify?

Amplify voices who claim energy is important in understanding growth in economies (Stern, Tol, Gerlagh)

Amplify those who call for caution in applying aggregate PF by just adding energy to it. (who?)

Amplify those who call for testing (statistically) the benefit of energy-augmentation. (shen/whalley)

Amplify comparative analysis of role of energy between countries indifferent stages of development. (who?)

What established sources we build on? Question? Contradict? In our writing.

Build on articles using R to test econ growth PF, 
question what is reported when doing these tests (shen/whalley and others who have used micEconCES). 
Critique literature that calls for uncritical energy-augmentation of PF, 
without testing whether that make sense at all.

Does analysis differ from others in the conversation?

Yes, differs as follows:
\begin{itemize}
\item{if one do these test, report on all $\delta$, $\sigma$, $\gamma$, and $\lambda$ 
values as well as sse. 
Good fits not evidence of best values, must report uncertainties as well.}
\item{What motivates nesting of CES? Not well behaved $\rho$ results, driving $\sigma$s.
Different type of CES with parsimonious factors, sigmas can be generated but not that easy (idea Randy).
If you run all 3 nesting options, you can identify elasticities of all three factors in relation to each 
other. Furthermore, you can determine which two factors are driving overall behavior in the aggregate function.}
\item{necessary of different factors in PF to explain growth for countries in diff. development status}
\end{itemize}

What do we add to the conversation?

\begin{itemize}

\item{We are able to gain rich insights 
on what really drives growth from detailed analysis (including uncertainties) 
by testing all functional forms across many different 
economy types.}

\item{We were unable to gain new insights from the various energy quantifications that we tried.}

\item{We find, surprisingly, that energy is not necessary 
to be included alongside $k$ and $l$ in all PF. Models with energy and models without energy
all perform well on an R$^2$ basis.}

\item{We also find that

(1) covariance of energy($e$) with both $k$ and $l$ is very high, with few exceptions (UK?).

(2) with few exceptions, $\sigma_{(kl)e}$, $\sigma_{ke}$, and $\sigma_{le}$ 
are all very low (indicating complementarity) (verify!!!!), and

(3) energy-only models (SF$e$ and LINEX) achieve very good fits, just like CD$e$ and CES$e$.
(should we make this point or not?)}

\item{What does that mean? 
It is a fallacy to think that saying energy is unnecessary 
to include in an equation means energy is unimportant for economic growth. How to support this argument?

If you find that energy is unnecessary to include alongside capital and labor 
to explain growth in one economy at the macro, aggregated level, 
you should not assume that energy is unimportant as a driver for economic growth for all economies.
For developed economies, energy is captial intensive: you need to have large capital to make use 
of energy. In developing economies, you don't need capital to make use of energy. 
Rather, you need labor.

We need to verify the above by comparison with our results.}

\item{Also signals that different types of countries likely 
to react differently to energy constraints. 
Interpret $\sigma$ here, substitution possibilities.}

\end{itemize}

More work?

**** Notes on the introduction from Martin ****

1. Set scene through energy economic modelling papers. 
Point out that PF approaches are alive and well (Stern, Tol, Gerlagh, World Bank, IMF), 
short lit review as part of a longer introduction.  
Mention competing PF approaches and parameter choice. 
On what basis selection and how estimate parameters?

2. Point out the need for adequate statistical testing and economic interpretation, 
citing some earlier critical papers on PF limitations. Take issue with

3. Spell out the contribution we will be making in this paper, 
we are taking the approach that all models are wrong but some are useful. 
We are after generating insights even after deeper assumptions of econ growth theory are very limiting.

4. Introduce coordinates of analysis.

**** Orignial introduction that MKH wrote ****

Neoclassical economic growth theory attributes output ($Y$) to factors of production:
capital stock ($K$) and labor ($L$). The Cobb-Douglas production function is the 
canonical representation of indexed economic output ($y$) as a function of indexed 
labor ($l$), indexed capital stock ($k$), and an exogenous, time-dependent 
technological progress term ($A$):

\begin{equation} \label{eq:CD}
  y = A k^\alpha l^\beta; A = \mathrm{e}^{\lambda(t-t_0)}
\end{equation}

\noindent where $y \equiv Y/Y_\mathrm{0}$, $k \equiv K/K_\mathrm{0}$, and 
$l \equiv L/L_\mathrm{0}$; with $Y$ represented by GDP, $K$ expressed in 
currency units, $L$ expressed in workers or work-hours/year, time ($t$) measured in years, 
with e being the base of the natural logarithm, and subscripts ``0'' indicating values 
at an initial year. $A$ is a unitless technological progress term, with $\lambda$ 
indicating the pace of technological progress (in units of 
year$^{-1}$).\footnote{Dimensionless, indexed quantities are represented by 
lower-case symbols 
($y$, $k$, $l$, $e$, $q$, $x$, and $u$), and dimensional 
quantities are represented by upper-case symbols 
($Y$, $K$, $L$, $E$, $Q$, $X$, and $U$). One exception is the technolgical 
progress term $A$, which is dimensionless but written upper-case by convention.} 
Neoclassical economic growth theory as expressed in the Cobb-Douglas production 
function makes no attempt to describe the mechanisms by which technological 
progress ($A$) affects economic growth, nor does it attempt to describe how 
technological progress ($A$) might vary from one economy to the next. Mathematically, 
$A$ is a fitting parameter that aligns the predictions of Equation~\ref{eq:CD} 
with historical data. Typically, the constraint $\alpha + \beta = 1$ is assumed, 
indicating constant returns to scale.

As shown in Equation~\ref{eq:CD} above, neoclassical economic theory does not 
include an explicit term for the role of energy in economic growth. However, 
\citet{Bashmakov:2007ek}, \citet{Hamilton:2009ui}, \citet{Murphy:2011jh}, 
\citet{Benes:2012tm}, \citet{Kumhof:2012ua}, and others indicate that the 
importance of energy for economic growth may far outweigh its cost share in the 
economy. That neoclassical economic growth theory ignores energy is, therefore, 
cause for concern, and inquiry into the proper approach for including energy 
as a factor of production in economic growth modeling is both relevant and important.

Some energy economists augment economic growth models by including energy as an 
additional (or the only) factor of production. However, there is little agreement 
on how this should be done. Neither the mathematical form of the energy-augmented 
production function nor the quantification of energy are consistent in the 
literature. There has been little empirical work assessing the predictive ability 
of various energy-augmented production functions for different types of economies. 
And, very few studies of the importance of energy in economic growth include 
rigorous statistical assessments of energy-related fitting parameters. This paper 
attempts to contribute to the literature by applying statistically rigorous 
methods along three coordinates of analysis: form of the energy augmented 
production function, energy quantification, and type of economy.

The remainder of this paper is organized as follows. Section~\ref{sec:Coordinates} 
discusses the coordinates of analysis described above. 
Section~\ref{sec:Sources_of_Data} details sources of data and calculations to 
generate time series for economic output and factors of production. The procedures 
for and results from parameter estimation for the various production functions 
are discussed in Section~\ref{sec:Parameter_Estimation}. 
Section~\ref{sec:Discussion_of_Results} discusses implications of the parameter estimation 
results, and Section~\ref{sec:Conclusion} concludes. An appendix provides details 
of all statistical results.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Coordinates of Analysis} 
\label{sec:Coordinates}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This section describes the three coordinates of analysis and briefly reviews 
literature related to each.

%++++++++++++++++++++++++++++++
\subsection{Mathematical Forms of the Energy-augmented Production Function} 
\label{sec:Prod_Func_Forms}
%++++++++++++++++++++++++++++++

In this paper, we assess three prominent energy-augmented production functions 
appear in the literature: Cobb-Douglas (CDe), Constant Elasticity of 
Substitution (CESe), and LINear EXponential (LINEX). The following subsections 
describe each.

%------------------------------
\subsubsection{Energy-augmented Cobb-Douglas Production Function (CDe)} 
\label{sec:CDe}
%------------------------------

Similar to Equation~\ref{eq:CD}, the energy-augmented Cobb-Douglas 
production function is

\begin{equation} \label{eq:CDe}
  y = A k^\alpha l^\beta e^\gamma; A = \mathrm{e}^{\lambda(t-t_0)}
\end{equation}

\noindent where $e \equiv E/E_0$, and $E$ is in units of energy per time, 
typically TJ/year. The energy-augmented Cobb-Douglas production function 
is often assumed to have constant returns to scale for the three factors 
of production: $\alpha + \beta + \gamma = 1$.

\citet{Ayres:2008cl} uses Equation~\ref{eq:CDe} with exergy ($x$) and 
useful work ($u$) quantifications for $e$ to describe historic Japanese 
and American economies from 1900-2000. \citet{Froling:2009vo} uses an 
energy-augmented Cobb-Douglas production function with thermal energy 
quantification ($q$) to describe output for the global economy from 
1800--1970. Fr\"{o}ling's model endogenizes labor with respect to fertility 
and mortality rates and separates thermal energy into two types: biomass 
energy and coal energy. \citet{Serrenho:2010vq} use the log transform of 
the energy-augmented Cobb-Douglas function with thermal energy 
quantification ($q$) to describe output for the EU-15 countries. 
Serrenho divides thermal energy into renewables, non-renewables, 
wastes, and imports. The main argument of these papers is that 
energy is an important driver of growth and needs to be included
as an essential factor of production.

**** Martin: add more here about the conclusions of each paper regarding 
the role of energy in economic growth? Add additional literature? 
Or, adjust as you see fit. Especially, are there any confidence intervals 
or statistical evaluation in each of these studies? ****

%------------------------------
\subsubsection{Energy-augmented CES Production Function (CESe)} 
\label{sec:CES}
%------------------------------

Other energy economists use an energy-augmented 
Constant Elasticity of Substitution (CES$e$) production function to 
describe economic growth. The \texttt{R} \citep{R} package 
\texttt{micEconCES} \citep{Henningsen:2011td} estimates CES production 
functions of the following forms (among others):

\begin{equation} \label{eq:CES}
  y = \gamma \: A \: \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{-1/\rho_1}; 
      A = \mathrm{e}^{\lambda (t-t_0)}
\end{equation}

\begin{equation} \label{eq:CESe}
  y = \gamma \: A \: \left\{\delta \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) e^{-\rho} \right\}^{-1/\rho}; 
      A = \mathrm{e}^{\lambda (t-t_0)}
\end{equation}

\noindent Equation~\ref{eq:CES} is a CES production function with 
capital stock ($k$) and labor ($l$) factors of production. 
Equation~\ref{eq:CESe} augments Equation~\ref{eq:CES} with energy 
using a ($kl$)($e$) nesting structure, as is typical in the literature. 
Equation~\ref{eq:CES} is a degenerate form of Equation~\ref{eq:CESe} 
where $\delta \rightarrow 1$ and $\rho \rightarrow -1$.

$\gamma$ is a fitting parameter that represents productivity. 
The fitting parameters $\rho_1$ and $\rho$ indicate elasticities 
of substitution ($\sigma_1$ and $\sigma$).
The elasticity of substitution 
between capital ($k$) and labor ($l$) is given by 
$\sigma_1 = \frac{1}{1+\rho_1}$, and
the elasticity of substitution between ($kl$) and ($e$) is given by 
$\sigma = \frac{1}{1+\rho}$. 
As $\rho_1 \rightarrow 0$, $\sigma_1 \rightarrow 1$,
and the embedded CES production function for $k$ and $l$ degenerates 
to the Cobb-Douglas production function.
Similarly, as $\rho \rightarrow 0$, $\sigma \rightarrow 1$,
and the CES production function for $(kl)$ and $(e)$ degenerates 
to the Cobb-Douglas production function.
As $\sigma \rightarrow \infty$ ($\rho \rightarrow -1$), 
($kl$) and ($e$) are perfect substitutes. 
As $\sigma \rightarrow 0$ ($\rho \rightarrow \infty$), 
($kl$) and ($e$) are perfect complements: 
no substitution is possible. Similarly, 
as $\sigma_1 \rightarrow 0$ ($\rho_1 \rightarrow \infty$),
$k$ and $l$ are perfect complements.
$\delta_1$ describes the relative importance of capital ($k$)
and labor ($l$), and
$\delta$ describes the importance of ($kl$) relative to ($e$).

Constraints on the fitting parameters include 
$\delta_1 \in (0,1)$,
$\delta \in (0,1)$,
$\rho_1 \in (-1,0) \cup (0,\infty)$, and
$\rho \in (-1,0) \cup (0,\infty)$. **** Describe here how 
we implemented the constraints. ****

\citet{VanderWerf:2008tc} evalutes various nested CES production 
functions for 12 OECD countries with respect to climate change policy. 
\citeauthor{VanderWerf:2008tc} concludes that ($kl$)($e$) nesting structure 
best fits historical data. \citet{Lecca:2011kr} also evaluate a variety of 
CES production functions using a KLEM structure to determine how the various 
options for nesting the energy factor of production affect predictive ability. 
\citet{Stern:2012ey} also use a CES production function of the form ($kl$)($e$). 
\citeauthor{Stern:2012ey} apply their CES function to the Swedish economy 
from 1800-2000 and demonstrate the importance of the transition from wood to 
fossil fuels for economic growth. \citet{Shen:2013vp} included energy 
in nested aggregated CES production functions 
for China and found that the best nesting structure is ($le$)($k$).

**** Martin: add more here about the conclusions of each paper regarding 
the role of energy in economic growth? Add additional literature? 
Or, adjust as you see fit. Especially, are there any confidence 
intervals or statistical evaluation in each of these studies? ****

%------------------------------
\subsubsection{LINEX Production Function} 
\label{sec:LINEX}
%------------------------------

A third production function, the LINear EXponential (LINEX) function 
\citep{Kummel:1980wx,Kummel:1982vy,Kummel:1985vz,Kummel:1989tf,Kummel:2002tx,Kummel:2010vz}, 
was derived from thermodynamic considerations:

\begin{equation} \label{eq:LINEX}
  y = Ae; 
  A = \mathrm{e}^{a_0\left[2 \left(1 - \frac{1}{\rho_k} \right) 
    + c_t \left(\vphantom{\frac{1}{\rho_k}}\rho_l - 1 \right)\right]},
\end{equation}

\noindent where $e$ is indexed energy and e is the base of 
the natural logarithm. The LINEX model assumes energy ($e$) 
is the only factor of production. 
The parameters $\rho_k$ and $\rho_l$ are defined as

\begin{equation} \label{eq:rho_k}
  \rho_k \equiv \frac{k}{(1/2)(l+e)}
\end{equation}

\noindent and

\begin{equation} \label{eq:rho_l}
  \rho_l \equiv \frac{l}{e}
\end{equation}

\noindent and represent 
(a) capital stock increase relative to the average of labor and energy and 
(b) labor increase relative to energy, respectively. 
Energy ($e$), as the only factor of production, is enhanced by $A$ if either

\begin{itemize}
\item{capital stock ($k$) has increased more than the average 
      of labor ($l$) and energy ($e$), i.e. if $\rho_k >$ 1, or}
\item{labor ($l$) has increased more than energy ($e$), i.e. if $\rho_l >$ 1.}
\end{itemize}

\noindent According to the LINEX model, an economy that increases 
capital stock ($k$) without a commensurate increase in labor ($l$) 
and energy ($e$) will experience an increase in output ($y$) in excess 
of its increase of energy consumption ($e$), because $A$ will be 
greater than 1. Similarly, an economy benefits by increasing labor ($l$) 
without a commensurate increase in energy ($e$). Thus, $\rho_k$ is an 
efficiency of using additional labor and energy to make 
additional capital stock available to the economy, 
and $\rho_l$ is an efficiency of using additional energy to make 
additional labor available to the economy. 
$a_0$ represents the overall importance of these efficiencies, 
and $c_t$ represents the relative importance 
of the labor efficiency term compared to the capital stock efficiency term. 
We treat $a_0$ and $c_t$ as fitting parameters; 
both are assumed constant with respect to time.

In contrast to both the Cobb-Douglas model (Equations~\ref{eq:CD} and 
\ref{eq:CDe}) and the CES model (Equations~\ref{eq:CES} and 
\ref{eq:CESe}), the LINEX model (Equation~\ref{eq:LINEX}) accounts 
for efficiencies among capital, labor, and energy directly and, 
therefore, does not employ a time-dependent 
technological progress term ($A$). \citet{Kummel:2002tx} use the 
LINEX production function to describe economic output 
for the economies of the U.S., Japan, and Germany. 
\citet{Warr:2010ga} compare the LINEX production function with 
the Cobb-Douglas production function to assess economic growth 
for Austria, the United Kingdom, the United States, and Japan. 
\citet{Ayres:2005kn} argued that introducing energy inputs 
in traditional production functions does not improve its explanatory power, 
but including useful work in a LINEX function does.

**** Martin: add more here about the conclusions of each paper 
regarding the role of energy in economic growth? Add additional literature? 
Or, adjust as you see fit. Especially, are there any confidence intervals 
or statistical evaluation in each of these studies? ****

%++++++++++++++++++++++++++++++
\subsection{Quantification of Energy} 
\label{sec:Quantification_of_Energy}
%++++++++++++++++++++++++++++++

The literature shows three quantifications of energy ($E$) used in 
energy-augmented production functions: thermal energy ($Q$), 
exergy ($X$), and useful work ($U$). Thermal energy ($Q$) 
is the heat content of an energy carrier. Exergy ($X$) is 
the maximum useful work that could be extracted from an energy carrier 
by a process that brings the system into equilibrium with 
a heat reservoir, typically the surrounding environment 
\citep[p.~78]{Ayres:2010ug}. Like \citeauthor{Ayres:2010ug}, 
we use chemical exergy (as opposed to thermal exergy or mechanical exergy) 
as our quantification for exergy ($x$). Useful work ($U$) 
includes the direct energy that goes into an economy, including work 
from machines, human muscles, and animals. Firms pay for thermal energy ($Q$), 
but useful work ($U$) alters the size and shape of materials, 
keeps the lights on, and transports goods to consumers. 
Not all forms of energy are created equal in terms of their ability 
to generate economic activity \citep{Cleveland:2000vl, Warr:2010ht}. 
Both exergy ($X$), and to a greater extent, useful work ($U$) 
quantifications of energy are intended to account for the usefulness 
of energy for generating economic activity, thereby reflecting 
the ``quality'' of each energy type and the efficiency of machines 
and devices typically used in energy conversion processes.

Most energy economists use thermal energy ($q$) 
for the energy factor of production ($e$) in energy-augmented 
production functions, including \citet{Stern:2012ey}, 
\citet{Froling:2009vo}, \citet{Serrenho:2010vq}, and 
\citet{Kummel:1980wx,Kummel:1982vy,Kummel:1985vz,Kummel:1989tf,Kummel:2002tx,Kummel:2010vz}. 
\citet{Ayres:2003ec} introduced the idea that exergy ($x$) and 
useful work ($u$) were important factors of production.

**** Martin: add more here about the conclusions of each paper 
regarding the role of energy in economic growth? 
Add additional literature? Or, adjust as you see fit. 
Especially, are there any confidence intervals or 
statistical evaluation in each of these studies? ****

We know of no statistically-rigorous comparisons of the 
predictive ability of each energy quantification in the 
various types of production functions discussed in 
Section~\ref{sec:Prod_Func_Forms}. We analyze all 
three quantifications of energy ($Q$, $X$, and $U$) 
for developed economies and thermal energy ($Q$) and 
exergy ($X$) for all other economies.

%++++++++++++++++++++++++++++++
\subsection{Economy Types} 
\label{sec:Economy_Types}
%++++++++++++++++++++++++++++++

In the energy-economic growth literature, the focus is 
almost exclusively on developed economies from North America 
or Europe. Examples include \citet{Kummel:1980wx} applying 
LINEX to describe the West German economy, \citet{Serrenho:2010vq} 
applying Cobb-Douglas variants to EU-15 countries, 
\citet{VanderWerf:2008tc} applying CES nesting variants 
to the 12 OECD countries, \citet{Lecca:2011kr} applying 
CES nesting variants to the Scottish economy, \citet{Stern:2012ey} 
applying CES to describe the Swedish economy, and \citet{Warr:2012cg} 
using LINEX and Cobb-Douglas to describe the Austrian, 
American, British, and Japanese economies.

But, the near-exclusive focus on developed economies closes off 
the possibility of discovering any role that energy plays 
in the economic development process. In this paper we analyze at least 
two economies in each of the following four categories: 
Developed, 
NICs,\footnote{Newly Industrialized Countries}
OPEC, and 
LDCs.\footnote{Lesser Developed Countries} 
(See Table~\ref{tab:Economy_Examples}.)

\begin{table} \caption{Example economies and years of evaluation.} \label{tab:Economy_Examples} 
  \begin{center}
    \begin{tabular}{r c c l l} 
      \multirow{3}{*}{Developed} & \includegraphics[height=0.11in]{flag/flag_US.png} & US & USA & \multirow{3}{*}{1980--2011}\\
                                 & \includegraphics[height=0.11in]{flag/flag_UK.png} & UK & United Kingdom & \\
                                 & \includegraphics[height=0.11in]{flag/flag_JP.png} & JP & Japan          & \\
      \hline
      \multirow{2}{*}{NICs}      & \includegraphics[height=0.11in]{flag/flag_CN.png} & CN & China & \multirow{2}{*}{1991--2011}        \\
                                 & \includegraphics[height=0.11in]{flag/flag_ZA.png} & ZA & South Africa & \\
      \hline
      \multirow{2}{*}{OPEC}      & \includegraphics[height=0.11in]{flag/flag_SA.png} & SA & Saudi Arabia & \multirow{2}{*}{1991--2011} \\
                                 & \includegraphics[height=0.11in]{flag/flag_IR.png} & IR & Iran         &  \\
      \hline
      \multirow{2}{*}{LDCs}      & \includegraphics[height=0.11in]{flag/flag_TZ.png} & TZ & Tanzania & \multirow{2}{*}{1991--2011}     \\
                                 & \includegraphics[height=0.11in]{flag/flag_ZM.png} & ZM & Zambia      & \\
    \end{tabular}
  \end{center}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Sources of Data}
\label{sec:Sources_of_Data}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

To assess energy-augmented models of economic growth along the 
coordinates of analysis outlined in Section~\ref{sec:Coordinates}, 
historical data for economic output ($Y$), capital stock ($K$), 
labor ($L$), thermal energy ($Q$), exergy ($X$), and useful work ($U$) 
are required. Subsections below describe sources for and calculations 
to obtain historical data. A final subsection presents the historical 
data in graphical form.

%++++++++++++++++++++++++++++++
\subsection{Economic Output} 
\label{sec:Economic_Output_Data}
%++++++++++++++++++++++++++++++

We measure economic output ($Y$) by gross domestic product (GDP). 
GDP data in real 2005 U.S. dollars for developed countries (1980--2011) 
and all other countries (1991--2011) was obtained from 
the World Bank's Global Economic Prospects (GEP) database 
\citep{WorldBankGEP:2013a}.

%++++++++++++++++++++++++++++++
\subsection{Capital Stock} 
\label{sec:Capital_Stock_Data}
%++++++++++++++++++++++++++++++

Gross capital stock in real 2005 U.S. dollars 
for developed countries (1980--2011) and 
all other countries (1991--2011) was obtained from 
the World Bank's Global Economic Prospects (GEP) database 
\citep{WorldBankGEP:2013a}.

%++++++++++++++++++++++++++++++
\subsection{Labor and Employment}
\label{sec:Labor_Data}
%++++++++++++++++++++++++++++++

Total employment and average hours worked per worker per year 
for the developed nations (United States, United Kingdom, and Japan) 
for 1980--2011 were obtained from the United States 
\citet[Tables 6 and 7]{BLS:2013a}. Total hours worked per year was 
calculated by multiplying total employment by average hours worked 
per worker per year.

The World Bank's World Development Indicators database 
\citep{WorldBankWDI:2013a} provides the employment to population 
ratio for people fifteen and older, total population, and a percentage 
of people under fifteen for other economies (South Africa, China, Iran, 
Saudi Arabia, Tanzania, and Zambia) for 1991--2011. The population 
of people under fifteen years of age was subtracted from total population 
to calculate the population of people fifteen and older. 
Total employment was calculated by multiplying the employment to 
population ratio of people fifteen and older by the population 
of people fifteen and older.

%++++++++++++++++++++++++++++++
\subsection{Thermal Energy, Exergy, and Useful Work} 
\label{sec:Energy_Data}
%++++++++++++++++++++++++++++++

To account for all energy input to an economy, we summed consumption data 
for many energy sources, including coal, oil and petroleum products, 
dry natural gas and natural gas plant liquids, nuclear power generation, 
and renewable energy sources (solar, wind, geothermal, biomass/fuel wood, 
and hydro). In the style of work by Warr and Ayres 
\citep{Ayres:2003ec, Ayres:2005kn, Warr:2010ht, Warr:2010ga, Warr:2012cg}, 
energy from human and animal muscle power was also included. 
Similar to the economic data discussed in 
Sections~\ref{sec:Economic_Output_Data}--\ref{sec:Labor_Data}, 
we generated energy data for developed countries from 1980--2011, 
with the exception of useful work data (1980--2000). For all other countries, 
we generated thermal energy and exergy data for 1991--2011. 
Details on each energy quantifications are given in subsections below.

%------------------------------
\subsubsection{Thermal Energy} 
\label{sec:Q_Data}
%------------------------------

Thermal energy ($Q$) for fuels (in units of TJ/year) was calculated by

\begin{equation} \label{eq:thermal_energy}
  Q_{fuel} = C_{fuel}h_{fuel}
\end{equation}

\noindent where $C_{fuel}$ is the consumption rate of the fuel 
(in units of short tons/year for coal, barrels/year for oil and petroleum, 
ft$^3$/year for dry natural gas and natural gas plant liquids, 
and m$^3$/year for biomass/fuel wood) and $h_{{fuel}}$ is 
the heat content of the fuel (in kJ/short ton for coal, kJ/barrel for oil, 
kJ/ft$^3$ for dry natural gas and natural gas plant liquids, 
and kJ/kg for biomass/fuel wood). Fuel wood density of 300~kg/m$^3$ 
was taken from \citet{Ragland:2013wp} at the low end of the range 
of density of dry wood and bark.

Yearly consumption rates ($C_{fuel}$) of coal, oil and petroleum, 
dry natural gas and natural gas plant liquids, and biofuels 
(ethanol and biodiesel) were obtained from the \citet{EIA:2013a} 
for all economies. Yearly average heat content data for coal, oil, 
and natural gas and natural gas plant liquids was obtained from the 
EIA for all economies. We used lower (net) heating values of 
26.8~MJ/kg \citep[Table~A-27]{Cengel:2011ww} and 37.8~MJ/kg 
\citep[Table~7.3]{Gupta:2010vg} for the heat content of ethanol 
and corn-based biodiesel, respectively. Yearly consumption and 
heat content data for biomass/fuel wood was obtained from the 
``Fuelwood'' time series of the FAO Statistics Database 
\citep{FAO:2013a}.

For renewable and nuclear energy sources, yearly electricity 
generation (in BTU/year) was available directly from 
the \citet{EIA:2013a} for all economies, so 
Equation~\ref{eq:thermal_energy} was unnecessary.

Animal and human muscle work was estimated using an approach 
similar to \citet{Ayres:2010ug}, using some data from Warr's 
REXS database \citep{Warr:2012a} and results from 
\citet{Wirsenius:2000vi}. For animal work, our analysis was 
restricted to draught animals (horses and mules). Muscle work 
from animals ($Q_{animals}$) was calculated using

\begin{equation} \label{eq:animal_muscle}
  Q_{animals} = Q_{m,horse}N_{horses} + Q_{m,mule}N_{mules}
\end{equation}

\noindent where $Q_m$ is metabolizable energy (in kcal/day-animal) 
and $N$ is the number of animals. Metabolizable energy values 
($Q_{m,horses}$ = 19,110 kcal/day-horse and $Q_{m,mules}$ = 
7,215 kcal/day-mule) were taken from \citet{Warr:2012a}. 
$Q_m$ represents the amount of energy in the feed consumed by 
animals for a day of work, and they are assumed constant for 
all countries and years. The number of horses and mules for 
each economy was taken from the ``Live Animals'' section of the 
FAO Statistics online database \citep{FAO:2013a}.

Human muscle power was calculated in a similar manner:

\begin{equation} \label{eq:human_muscle}
  Q_{workers} = Q_{m,workers} N_{workers}
\end{equation}

\noindent where the value of $Q_{m,workers}$ is the ``the amount 
of food available for human consumption'' (i.e., at the wholesale level) 
taken from the 
\citet[``\texttt{Dietary~Energy~Cons.~Countries}'' tab]{FAO:2013b}. 
The value for $N_{workers}$ was obtained from the labor time series 
for each economy as discussed in Section~\ref{sec:Labor_Data}.

%------------------------------
\subsubsection{Exergy} 
\label{sec:X_Data}
%------------------------------

Exergy ($X$), expressed in TJ/year, was calculated by multiplying 
the thermal energy ($Q$) values discussed in Section~\ref{sec:Q_Data} 
by a unitless exergy-to-thermal-energy allocation factor 
($f_{exergy-allocation}$) shown in Table~\ref{tab:Exergy_Allocation}.

\begin{equation} \label{eq:exergy}
  X_{fuel} = f_{exergy-allocation} Q_{fuel}
\end{equation}

\begin{table} \caption{Exergy allocation factors.} \label{tab:Exergy_Allocation} 
  \begin{center}
    \begin{tabular}{r l} 
      Fuel                                    & $f_{exergy-allocation}$ \\
      \hline
      Coal                                    & 1.088    \\
      Oil and petroleum                       & 1.073   \\      
      Dry natural gas and                     & \multirow{2}{*}{1.04}    \\
      natural gas liquids                     &         \\
      Corn-based biodiesel                    & 1.07  \\
      Biomass/fuel wood                       & 1.15   \\
      Ethanol biofuel                         & 1.099 \\
      Nuclear, wind,                          & \multirow{2}{*}{1.0}   \\
      solar PV, and hydro                     &                        \\
    \end{tabular}
  \end{center}
\end{table}

\noindent For carbon-based energy sources, $f_{exergy-allocation}$ 
is the ratio of chemical exergy to lower (net) heating value. 
The exergy allocation factors for coal, oil, natural gas, 
biodiesel, and biomass/fuel wood were obtained from 
\citet[Table~A-2]{Ayres:2003ec}. The exergy allocation ratio for 
liquid ethanol was calculated from the standard molar chemical exergy 
of 1357.7~kJ/mol \citep[Table~II]{Szargut:1988wv} and its lower 
(net) heating value of 26.8~MJ/kg \citep[Table~A-27]{Cengel:2011ww}. 
$f_{exergy-allocation}$ for each fuel was assumed to be constant 
through time across all economies. For nuclear, wind, 
solar photovoltaic, and hydro energy sources, the electricity generated 
is already pure exergy, so $f_{exergy-allocation}$ is unity.

To calculate the exergy of animal and human muscle power, 
the food-to-work conversion efficiency ($\eta_{food \rightarrow work}$) 
and the ratio of food intake to food available at wholesale 
($f_{intake/wholesale}$) were included.

\begin{equation} \label{eq:animal_muscle_exergy}
  X_{animals} = \eta_{food \rightarrow work,animals} Q_{animals}
\end{equation}

\begin{equation} \label{eq:human_muscle_exergy}
  X_{workers} = \eta_{food \rightarrow work,workers} f_{intake/wholesale} Q_{workers}
\end{equation}

\noindent $\eta_{food \rightarrow work, animals}$ and 
$\eta_{food \rightarrow work, workers}$ are animal and worker efficiency 
of converting food into energy, respectively. The animal feed-to-work 
conversion efficiency ($\eta_{feed \rightarrow work, animals}$ = 0.4) 
was taken from \citet{Warr:2012a} and assumed equal across all economies. 
The food-to-work efficiency for workers 
($\eta_{food \rightarrow work, workers}$ = 0.2) was taken from 
\citet[Table~B.1]{Ayres:2010ug}. Following \citet[p.~315]{Ayres:2010ug} 
and methodology from \citet{Warr:2012a}, an additional factor 
($f_{intake/wholesale}$) was included to account for ``wastage and 
losses of food in the household, e.g. during storage, in preparation 
and cooking, as plate-waste, or quantities fed to domestic animals 
and pets'' \citep[\texttt{notes} tab]{FAO:2013b}. 
Values of $f_{intake/wholesale}$ were calculated from 
\citet[Table~3.3 for wholesale and Table~3.4 for intake]{Wirsenius:2000vi}. 
$f_{intake/wholesale}$ varies by country as shown in 
Table~\ref{tab:f_intake_wholesale}. A similar intake-to-wholesale ratio 
could also be applied to animal work, but data to calculate the ratio was 
unavailable from \citet{Wirsenius:2000vi}, and $f_{intake/wholesale,animals}$ 
is assumed to be unity.\footnote{The error in total exergy input to 
any economy caused by assuming $f_{intake/wholesale,animals} = 1.0$ 
is very small, because animal muscle exergy is a small fraction of the 
total exergy supply to any economy, with a maximum of 0.41\% for China in 1992. 
It could also be argued that $f_{intake/wholesale, animals}$ is likely 
to be closer to unity than $f_{intake/wholesale, workers}$.}

\begin{table} \caption{Food intake-to-wholesale ratios 
                        ($f_{intake/wholesale}$) for all countries.} 
                        \label{tab:f_intake_wholesale}
  \begin{center}
    \begin{tabular}{r c c c c c c c c c}
      Country & US & UK & JP & CN & ZA & SA & IR & TZ & ZM \\
      \hline
      $f_{intake/wholesale}$ & 0.62 & 0.65 & 0.76 & 0.76 & 0.87 & 0.72 & 0.72 & 0.72 & 0.72\\
    \end{tabular}
  \end{center}
\end{table}

%------------------------------
\subsubsection{Useful Work} 
\label{sec:U_Data}
%------------------------------

The conversion of a raw energy carrier (e.g., coal) into 
productive energy for an economy (e.g., electricity) necessarily 
involves inefficiencies. In the case of coal consumption, 
only about one-third of the coal exergy is converted to useful electricity: 
the rest is waste heat. The concept of ``useful work,'' pioneered by 
Ayres and Warr \citep{Ayres:2003ec, Ayres:2005kn}, builds upon 
exergy analysis to consider efficiencies of both energy production and enduse. 
To calculate useful work ($U$), the total exergy input to an 
economy ($X$) is decomposed into several end uses: 
high temperature industrial heat (e.g., steel production), 
medium temperature heat (e.g., steam production), 
low temperature space heat (for residential heating), 
mechanical drive (energy used for prime mover engines of all types except electric motors), 
electricity (excluding lighting), 
lighting (from both electricity and fuels), and 
muscle work (from animals and workers) as shown in the following equation:

\begin{equation} \label{eq:useful_work}
  U = \sum_i{\varepsilon_i \phi_i X}
\end{equation}

\noindent where the subscript $i$ indicates the energy enduse, 
$\varepsilon_i$ is the efficiency of converting exergy into 
useful energy for each enduse, and $\phi_i$ gives 
the fraction of exergy input to an economy that goes toward 
the $i^\mathrm{th}$ enduse.

Values of $\varepsilon_i$ and $\phi_i$ were calculated from 
the REXS database supplied by \citet{Warr:2012a}. 
Values of $X$ were calculated by methods discussed in 
Section~\ref{sec:X_Data}. For most countries, values of 
$\phi_i$ are not constant with time, as economic sectors with 
different distributions of raw exergy inputs grow at different rates. 
\citeapos{Warr:2012a} REXS database contains information for 
developed economies (United States, United Kingdom, and Japan) 
from 1900--2000, and we used data from the 1980--2000 timeframe. 
Extending the REXS database to the BRICS, petrol, and LDC economies 
is beyond the scope of this work, because longitudinal data for 
the efficiency of different end uses ($\varepsilon_i$) and 
the distribution of exergy sources to end uses 
($\phi_i$) are extremely difficult to obtain.

%++++++++++++++++++++++++++++++
\subsection{Historical Data} 
\label{sec:Historical_Data}
%++++++++++++++++++++++++++++++

Figure~\ref{fig:Historical_Lattice_Graph} presents historical data, 
including indexed values for 
output ($y$), 
capital stock ($k$), 
labor ($l$), and 
the various energy quantifications ($q$, $x$, and $u$) 
for 1980--2011 (developed economies) and 1991--2011 (other economies). 
Factors of production track each other closely, 
and high covariances among these variables can be seen in 
Tables~\ref{tab:Covariance_US}--\ref{tab:Covariance_ZM}. 
**** Do we want to discuss the covariances? ****

<<Historical_Lattice_Graph, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="Historical data: indexed GDP ($y$), capital stock ($k$), labor ($l$), thermal energy ($q$), exergy ($x$), and useful work ($u$).", eval=FALSE>>=
createHistoricalLatticeGraph()
@

Figure~\ref{fig:Historical_Lattice_Graph} indicates the importance 
of energy quantification for economic growth modeling. 
Thermal energy ($q$) and exergy ($x$) grow at similar rates, 
unless there is a significant raw energy carrier substitution in the economy 
(such as China's rapid increase of coal and natural gas consumption 
relative to petroleum since 2000). The growth rate of $u$ is 
significantly different from $q$ and $x$ for most countries, however. 
\citet{Warr:2012cg} show that from 1950-1990, useful work ($u$) 
grew much faster than $q$, $x$, $k$, or $l$ due to efficiency gains 
in the U.S. economy. The results of \citeauthor{Warr:2012cg} suggest 
that the time-dependent technological progress term ($A$) 
in the Cobb-Douglas and CES production functions 
(Equations~\ref{eq:CD}--\ref{eq:CESe}) may be a proxy for 
increasing energy efficiency in the U.S. economy.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Parameter Estimation}
\label{sec:Parameter_Estimation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\texttt{R} software, makes detailed testing possible. R$^2$ not a good measure, 
limited insights t and p values, especially on PFs that have parameters with constraints

We found that we are able to maintain good fits while respecting theoretical foundations
of the models (i.e., respect parameter constraints). No need to relax parameter constraints
and, thereby, contradict micro foundations.

Grid search etc. (Matt, Randy to populate)

This section should cover both technique and results. 
Organize along coordinates of analysis.

1. Discuss parameter estimation for each model here.

CES Model: It is possible for L-BFGS-B to report convergence and PORT to report non-convergence 
but nevertheless reach a better solution (smaller sse).

2. Show scatter plot results.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion of Results}
\label{sec:Discussion_of_Results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Drive towards new insights from our stats tested comparative approach..

**** Notes for interpretation of results ****

CD assumes $\sigma_1 = 1$. CES doesn't assume a value for $\sigma_1$. 
The $\sigma_1$ vs. $\delta_1$ graphs for CES without energy often have 
two frequent values for $\sigma_1$.  One of those frequent values is
nearer to $\sigma_1 = 1$. When CD and CES without energy do not agree on 
share values (CD: $\alpha$ compared to $\beta$, CES: $\delta_1$) for the 
base curve fit (red dot), they do agree for the data that clusters near 
$\sigma_1 = 1$ for the CES fits.

Country interpretation

Comparison Cobb Douglas and CES results.

%++++++++++++++++++++++++++++++
\subsection{United States} 
\label{sec:InterpUS}
%++++++++++++++++++++++++++++++

CD: Energy is unnecessary

CES with $q$:

($kl$): $\delta_1 = 0.3, \sigma_1 = 0.8$. $k$ is less necessary. Complements, but close to substitutes.
signal of substitutability in the spread.

($kl$)$e$: energy is unnec ($\delta = 1$). $\delta_1 = 0.45$: $k$ and $l$ are about equally necessary.
$\sigma_1 = 0.8$: $k$ and $l$ are slight complements. Spread on $0.25 < \sigma_1 < \infty$, signalling 
possibilities of substitution.

($el$)$k$: energy is unnec. ($\delta_1$ is 0);

($le$)$k$ no indication
Relative consistent signal that energy is unnec.

CES with $u$:

($kl$)$e$: $\delta = 1$, signalling $e$ unnecessary. $\sigma = 0.8$, indicating clost to substitutability,
but still complementary. 
$\delta_1 = 0.25$, with large spread. not strong for $k$.
$\sigma_1 \rightarrow \infty$, but with large spread from 0.2 upward. Signals complementarity
between $k$ and $l$.

%++++++++++++++++++++++++++++++
\subsection{United Kingdom} 
\label{sec:InterpUK}
%++++++++++++++++++++++++++++++

CD: Capital is unnec.

CES with $q$:

($kl$): $\delta_1 = 0.7, \sigma_1 = 0.35$. $kl$ complementary, $k$ is more important, because $\delta_1 > 0.5$. Large spread on $\delta_1$. Complementary signal is consistent, however.

($kl$)$e$: energy is unnec ($\delta = 1$). $\delta_1 = 0.25$, $k$ is less necessary. 
$\sigma_1 \rightarrow \infty$, giving strong $kl$ substitutability signal.

($le$)$k$ capital is unnec. ($\delta = 1$);

($ek$)$l$ ($\delta_1 = 0$) energy is unnec.

Thus, inconsistent results, possibly look at different time series 
as massive changes in economic policy in Uk economic history

CES with $u$:

($kl$)$e$: $\delta = 1$, $u$ is unnecessary. $\sigma = 0.8$, approache complementarity between $kl$ and $e$.
$\delta_1 = 0.1$. $\sigma_1 \rightarrow \infty$, indicating substitutability between $k$ and $l$.

%++++++++++++++++++++++++++++++
\subsection{Japan} 
\label{sec:InterpJP}
%++++++++++++++++++++++++++++++

CD: energy is unnec.

CES with $q$:

($kl$): $\delta_1 = 0.7, \sigma_1 = 0.25$. Very little scatter. Both are well-known. 
$l$ is close to unnecessary. $k$ and $l$ are complements.

($kl$)$e$: no indications, all are important. $\delta = 0.6$, both $kl$ and $e$ are necessary.
$0.5 < \delta < 1$.
$\sigma=0.1$ signalling complementarity betwen $kl$ and $e$.
$\delta_1 = 0.8$ signalling $l$ is almost unnecessary. $\sigma_1 = 0.1$, signalling complementarity 
between $k$ and $l$.

($le$)$k$: no indications, all are important.

($ek$)$l$: no indications, all are important.

weak signal that labour is less necessary 
(($kl$)$e$ $delta_1 = 0.86$ and ($ek$)$l$ $\delta = 0.89$)

Thus, inconsistent results.

CES with $u$:

($kl$)$e$: $\delta = 0.6, \sigma = 0.4$. Complementarity between $kl$ and $e$.
$\delta_1 = 1$, indicating $l$ unnecessary. $\sigma_1 = 0.1$, indicating complementarity
between $k$ and $l$. Need only one of $k$ or $l$ in your equation, but they are
complemetary! You need one to have the other.

%++++++++++++++++++++++++++++++
\subsection{China} 
\label{sec:InterpCN}
%++++++++++++++++++++++++++++++

CD: energy not necessary, capital less necessary

CES with $q$:

($kl$): $\delta_1 = 0.3$, so $k$ is getting close to unnecessary. $\sigma_1 = 1.25$.
$k$ and $l$ are substitutes. Large spread in $\delta_1$ from 0.1 to 1.0. Large spread in
$\sigma_1$, but most points have $\sigma_1 > 1$.

($kl$)$e$: $\delta = 0.8$. however, strong signal at $0.8 < \delta < 1$, 
signalling that energy is less necessary. $0.5 < \delta < 1$. $\sigma = 0.5$, 
signalling complementarity between $kl$ and $e$. 
$\delta_1 = 1$, signalling that $l$ is unnecessary. $\sigma_1 = 0.1$, signalling
that $k$ and $l$ are complementary. Both $\delta_1$ and $\sigma_1$ have large spreads, 
so our confidence in these results is low.

($le$)$k$: labour is unnnec ($\delta_1 = 0$);

($ek$)$l$ energy is unnec. ($\delta_1 \rightarrow 0$)

Strong signal that energy is unnec, but the labour and capital signals are conflicting

%++++++++++++++++++++++++++++++
\subsection{South Africa} 
\label{sec:InterpZA}
%++++++++++++++++++++++++++++++

CD: Labour is unnec.

CES with $q$:

($kl$): $\delta_1 = 0.35, \sigma_1 = 0.8$. $k$ is relatively less necessary. $k$ and $l$ 
are complements but relatively close to being substitutes. Spread is tight between 
$0.25 < \delta_1 < 0.5$ and $\sigma_1$ is consistently complementary 
between $0.1 < \sigma_1 < 0.8$.

($kl$)$e$: labour is unnec. ($\delta_1 = 1$). $\delta = 0.25$, signalling complemtarity 
between $kl$ and $e$, but large spread from 0 to 1. $\sigma$ and $\sigma_1$ are both 0, 
indicating perfect complements betwen $kl$ and $e$ and between $k$ and $l$.

($le$)$k$ labour is unnec. ($\delta_1 = 0$);

($ek$)$l$ labour is unnec. ($\delta = 1$)

Consistent signal that labour is unnec.

%++++++++++++++++++++++++++++++
\subsection{Saudi Arabia} 
\label{sec:InterpSA}
%++++++++++++++++++++++++++++++

CD: no indications

CES with $q$:

($kl$): $\delta_1 = 0.85$. $l$ is unnecessary. $\sigma_1 = 0.1$ indicating strong complementarity between 
$k$ and $l$. Spread of $0.15 < \delta_1 < 1$. $0 < \sigma_1 < 0.8$, signalling complementarity.

($kl$)$e$: $\delta = 0.6$, $e$ is less necessary (weak signal). Big spread on $\delta$ between 0 and 0.8.
$\sigma = 0.1$ indicating complementarity between $kl$ and $e$. $\delta_1 = 0.8$ 
(with large spread between 0.3 and 1), labor is less necessary.
$\sigma_1 = 0.1$: very complementary between $k$ and $l$.

($le$)$k$:

($ek$)$l$: weak signal that labour is unnec ($\delta = 0.93$)

Almost consistent signal that no indocations.

%++++++++++++++++++++++++++++++
\subsection{Iran} 
\label{sec:InterpIR}
%++++++++++++++++++++++++++++++

CD: energy is unnec.

CES with $q$:

($kl$): $\delta_1 = 0.3$. Spread $0 < \delta_1 < 0.7$. $0 < \sigma_1 < 0.8$. $\sigma_1 = 0$. 
Strong complementarity. $k$ is less necessary compared to $l$.

($kl$)$e$: energy is unnec ($\delta=1$). $\sigma = 0.75$ indicating complementarity between $kl$ and $e$, approaching substitutibility.
$\delta_1 = 0.3, \sigma_1 = 0$ indicating strong perfect complementarity between $k$ and $l$.

($le$)$k$: energy is unnec. ($\delta_1=1$);

($ek$)$l$: labour is unnec. ($\delta=1$)

%++++++++++++++++++++++++++++++
\subsection{Tanzania} 
\label{sec:InterpTZ}
%++++++++++++++++++++++++++++++

CD: labour in unnec.

CES with $q$:

($kl$): $\delta_1 = 0.8$ $0.3 < \delta_1 < 1$ signalling labor is less necessary.
$\sigma_1 = 0.1$ $0 < \sigma_1 < 0.8$, signalling complementary $k$ and $l$.

($kl$)$e$: labour is unnec ($\delta_1=0.97$). $\delta = 0.45$, so $kl$ and $e$ both necessary, but 
$\sigma \rightarrow \infty$, so $kl$ and $e$ has perfect substitutibility.
$\sigma_1 = 0.1$, so $k$ and $l$ are perfect complements. Spread on $\delta$ is large (0.2 to 1).
So, we are unsure about the value of $\delta$.

($le$)$k$: weak signal that energy is unnec. ($\delta_1=0.92$); 
very weal signal capital is unnec ($\delta=0.84$);

($ek$)$l$: labour is unnec. $\delta=0.96$

Consistent signal that labour is unnec, with exception of one nesting.

%++++++++++++++++++++++++++++++
\subsection{Zambia} 
\label{sec:InterpZM}
%++++++++++++++++++++++++++++++

CD: labour is unnec; energy is unnec.

CES with $q$:

($kl$): $\delta_1 = 1$: $l$ is unnecessary. $0.5 < \delta_1 < 1$. $\sigma_1 = 0.8$ without any 
noticeable spread indicating complemetary.

($kl$)$e$: labour is unnec. ($\delta_1=1$). $\delta = 0.6$, $kl$ and $e$ are both necessary. 
$\sigma \rightarrow \infty$, signalling substitutability between $kl$ and $e$. $\sigma_1 = 0.8$ with very
large spread between 0 and 2. Approaching substitutability between $k$ and $l$ and perfect substitutibility
between $kl$ and $e$. Very little certainty here.

($le$)$k$: ($\delta_1=0$);

($ek$)$l$: labour is unnec. ($\delta=1$)

Consistent signal that labour is unnec.

Discussion of results:

How to get better consistent signals?

\begin{itemize}
\item{Use CD as starting points and do CES convergence on that basis.}
\item{Do scatterplots to get range of uncertainty inputs, then re-interpret. delta against sigma plots, as well as $\delta_1$ and $\sigma_1$ plots.}
\end{itemize}

Thus, to do the fitting well we need to understand:
\begin{itemize}
\item{nesting}
\item{algorithm}
\item{starting values}
\end{itemize}
esp. for CES.

(Maybe illustrate with one case study such as China what implications are).

How do we compare with LINEX results?

CES is a broader specification of CD, LINEX is not. In LINEX no delta's nor sigmas are generated. Only efficiency terms are generated in LINEX, thus in other words economic growth is only possible by increased efficiency. 
The efficiency of using energy is driven by capital stock. The objective/growth engine is to increase capital at a rate that is faster then growth in labour and energy.  Increasing labour relative to energy gives a muted benefit. 
LINEX derived for Germany: only really applicable when energy and capital are complements.

Discussion in relation to other papers. 
Problems with other papers:
\begin{itemize}
\item{If not giving you all parameters}
\item{Not explain nesting approach}
\item{No CI on data}
\end{itemize}

Economic interpretation (revisit after scatterplot information)
Developed economies: Capital and energy is complementary (US: labour and energy substitutable)
NICs: Capital and energy complementary, but labour and energy more substitutable
OPEC: Labour and energy complementary (mixed results on capital and energy)
LDCs: Labour and energy complementary, Capital and energy perfect substitutes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:Conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Future Work}
\label{sec:FutureWork}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}
\label{sec:Acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Funding for Caleb Reese and Lucas Timmer was provided 
by the Jack and Lois Kuipers Applied Mathematics Endowment and 
the Calvin College Alumni Association. 
Partial support for Matthew Kuperus Heun was 
provided by a Calvin College Research Followship. 
Partial funding for Martin de Wit was provided 
by the South African National Research Foundation (NRF).\footnote{Any opinion, 
findings, and conclusions or recommendations expressed herein are those 
of the authors, and, therefore, NRF does not accept any liability in regard thereto.} 
The authors are grateful to Benjamin S. Warr (INSEAD) for providing 
the REXS database \citep{Warr:2012a}. 
The authors thank Loren L. Heun (Western Michigan University) 
for insightful critique of the statistical analyses presented herein.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Reproducible Research}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In the spirit of Reproducible Research \citep{Gandrud:2013vx}, 
all data, spreadsheets, \texttt{R} code \citep{R}, and other materials 
associated with this paper can be found at\\*
\protect\url{https://github.com/MatthewHeun/Econ-Growth-R-Analysis}.

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix
%% \section{}
%% \label{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Covariance Tables}
\label{sec:Covariance_Tables}
\setcounter{table}{0} % Restarts table number counter at 0 in this section.
\setcounter{figure}{0} % Restarts figure number counter at 0 in this section.
\setcounter{equation}{0} % Restarts equation counter to 0 in this section.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% References
%%
%% Following citation commands can be used in the body text:
%%
%%  \citet{key}  ==>>  Jones et al. (1990)
%%  \citep{key}  ==>>  (Jones et al., 1990)
%%
%% Multiple citations as normal:
%% \citep{key1,key2}         ==>> (Jones et al., 1990; Smith, 1989)
%%                            or  (Jones et al., 1990, 1991)
%%                            or  (Jones et al., 1990a,b)
%% \cite{key} is the equivalent of \citet{key} in author-year mode
%%
%% Full author lists may be forced with \citet* or \citep*, e.g.
%%   \citep*{key}            ==>> (Jones, Baker, and Williams, 1990)
%%
%% Optional notes as:
%%   \citep[chap. 2]{key}    ==>> (Jones et al., 1990, chap. 2)
%%   \citep[e.g.,][]{key}    ==>> (e.g., Jones et al., 1990)
%%   \citep[see][pg. 34]{key}==>> (see Jones et al., 1990, pg. 34)
%%  (Note: in standard LaTeX, only one note is allowed, after the ref.
%%   Here, one note is like the standard, two make pre- and post-notes.)
%%
%%   \citealt{key}          ==>> Jones et al. 1990
%%   \citealt*{key}         ==>> Jones, Baker, and Williams 1990
%%   \citealp{key}          ==>> Jones et al., 1990
%%   \citealp*{key}         ==>> Jones, Baker, and Williams, 1990
%%
%% Additional citation possibilities
%%   \citeauthor{key}       ==>> Jones et al.
%%   \citeauthor*{key}      ==>> Jones, Baker, and Williams
%%   \citeyear{key}         ==>> 1990
%%   \citeyearpar{key}      ==>> (1990)
%%   \citetext{priv. comm.} ==>> (priv. comm.)
%%   \citenum{key}          ==>> 11 [non-superscripted]
%% Note: full author lists depends on whether the bib style supports them;
%%       if not, the abbreviated list is printed even when full requested.
%%
%% For names like della Robbia at the start of a sentence, use
%%   \Citet{dRob98}         ==>> Della Robbia (1998)
%%   \Citep{dRob98}         ==>> (Della Robbia, 1998)
%%   \Citeauthor{dRob98}    ==>> Della Robbia

%% References with bibTeX database:

\bibliographystyle{model2-names}
%%\bibliography{<your-bib-database>}
\bibliography{EconGrowth.bib}

%% Authors are advised to submit their bibtex database files. They are
%% requested to list a bibtex style file in the manuscript if they do
%% not want to use model2-names.bst.

%% References without bibTeX database:

% \begin{thebibliography}{00}

%% \bibitem must have one of the following forms:
%%   \bibitem[Jones et al.(1990)]{key}...
%%   \bibitem[Jones et al.(1990)Jones, Baker, and Williams]{key}...
%%   \bibitem[Jones et al., 1990]{key}...
%%   \bibitem[\protect\citeauthoryear{Jones, Baker, and Williams}{Jones
%%       et al.}{1990}]{key}...
%%   \bibitem[\protect\citeauthoryear{Jones et al.}{1990}]{key}...
%%   \bibitem[\protect\astroncite{Jones et al.}{1990}]{key}...
%%   \bibitem[\protect\citename{Jones et al., }1990]{key}...
%%   \harvarditem[Jones et al.]{Jones, Baker, and Williams}{1990}{key}...
%%

% \bibitem[ ()]{}

% \end{thebibliography}

\end{document}

%%
%% End of file `elsarticle-template-2-harv.tex'.