### 2014-04-22 Testing/ExperimentingWithPlots.Rnw
Now using iK, iL, iY for data columns. batchEcon2.R now working.

\documentclass{article}

\begin{document}

<<message=FALSE>>=
require(knitr)
opts_chunk$set(
  dev="pdf",
  tidy=FALSE
)
require(ggplot2)
require(plyr)
source('Econ-Growth-Functions2.R')
@

<<plot1>>=
for (nest in c("(kl)e", "(le)k", "(ek)l") ) {
  nestData <- loadCESSpaghettiGraphData("Q", nest= nest, 
                archive="data_archive/data_resample_2013-11-16_Best_Algorithm_Saving_Models_n=50.zip")
  print( 
    qplot( Year, iY, group=ResampleNumber, 
           data=nestData, # subset(foo, Country=="US"), 
           facets = ~Country, geom="line", alpha=I(0.15) ) + 
      geom_line(data=subset(nestData, Type=="actual"), 
                colour="red") +
      labs(title=nest) +
      theme_minimal()
  )
}
@

\end{document}