### 2014-05-16 Article/EconGrowth2-GraphsOnly.Rnw
Edits to the Graphs-Only version of the article.

% This article has been prepared for publication in Energy Economics in RStudio with knitr.
% According to http://www.elsevier.com/author-schemas/the-elsarticle-latex-document-class, we should be using the
% elsarticle.cls file.
% According to http://cdn.elsevier.com/assets/pdf_file/0006/109392/journal_refstyles.pdf, we should be using
% elsarticle-template-2-harv.tex as the template for the text.
% Furthermore, we should be using model2-names.bst for the bibliographic references.
% The approach here is to load the frontmatter and backmatter from elsarticle-template-2-harv.tex
% both ahead of and behind the text for our paper.
% -- Matthew Kuperus Heun, 2013-01-18

%% This is file `elsarticle-template-2-harv.tex',
%%
%% Copyright 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle  Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for Elsevier's document class `elsarticle'
%% with harvard style bibliographic references
%%
%% $Id: elsarticle-template-2-harv.tex 155 2009-10-08 05:35:05Z rishi $
%% $URL: http://lenova.river-valley.com/svn/elsbst/trunk/elsarticle-template-2-harv.tex $
%%
\documentclass[preprint,authoryear,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,authoryear,1p,times]{elsarticle}
%% \documentclass[final,authoryear,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,authoryear,3p,times]{elsarticle}
%% \documentclass[final,authoryear,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,authoryear,5p,times]{elsarticle}
%% \documentclass[final,authoryear,5p,times,twocolumn]{elsarticle}

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%% or use the graphicx package for more complicated commands
%% \usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

\usepackage{wrapfig}    % Allows wrapping of text around figures
\usepackage{soul}       % Provides strikethrough text
\usepackage{float}      % Allows precise positioning of tables and figures within the text
\usepackage{amsmath}    % Allows \begin{equation*} \end{equation*} for unnumbered equations
\usepackage{multirow}   % To create multirow tables
\usepackage{hyperref}   % To create hyperlinks in the paper
\usepackage{microtype}  % produces hanging punctuation and beautiful type

% From http://economics.utoronto.ca/osborne/latex/BIBTEX.HTM
\newcommand{\citeapos}[1]{\citeauthor{#1}'s (\citeyear{#1})} % Posessive citations.

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon (default)
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   authoryear - selects author-year citations (default)
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%   longnamesfirst  -  makes first citation full author list
%%
%% \biboptions{longnamesfirst,comma}

% \biboptions{}

\usepackage[shownotes]{authNote}

\journal{Energy Economics}

\begin{document}

<<setup, echo=FALSE, message=FALSE, eval=TRUE>>=
require(EconModels)
require(EconData)
require(lattice)
require(latticeExtra)
require(ggplot2)
require(car)
require(mosaic)
require(xtable)
require(nlmrt)
# require(micEconCES)
require(plyr)
require(dplyr)
# Provides access to data melting. 
# See http://cran.r-project.org/web/packages/reshape2/reshape2.pdf
require(reshape2) 
# tikz allows use of LaTeX formatting and font in graphs. 
# Allows for a consistent look across the paper.
# See http://r-forge.r-project.org/R/?group_id=440 for instructions 
# on installing tikzDevice.
# require(tikzDevice) 
trellis.par.set(theme=theme.mosaic())
options(width=75)
options(show.signif.stars=TRUE)
require(knitr)

# Load historical data
# In Rstudio, this paper is generated from it's directory, which is "Article".
# We need to define paths relative to the location of the paper. 
# Relative path to historical data from Article
historicalData <- Econ2011 # From the EconData package

# Load coefficients from all models in one large data frame
# coeffs_df <- tbl_df( readRDS("../AllCoef.rda") )
coeffs_df <- tbl_df(AllCoef)
# Rearrange the country factors to obtain our preferred ordering.
coeffs_df$country <- as.factor(coeffs_df$country)
for (lev in rev(countryAbbrevs)) { coeffs_df$country <- relevel(coeffs_df$country, ref=lev) }
# Change the names of the energy factors to the names for the graph
coeffs_df$energy <- replace(coeffs_df$energy, which(is.na(coeffs_df$energy)), "none")
coeffs_df$energy <- replace(coeffs_df$energy, which(coeffs_df$energy=="iQ"), "q")
coeffs_df$energy <- replace(coeffs_df$energy, which(coeffs_df$energy=="iX"), "x")
coeffs_df$energy <- replace(coeffs_df$energy, which(coeffs_df$energy=="iU"), "u")
# Rearrange the energy factors to obtain our preferred ordering.
coeffs_df$energy <- as.factor(coeffs_df$energy)
energyLevels <- c(none="none", q="q", x="x", u="u")
for (lev in rev(energyLevels)) { coeffs_df$energy <- relevel(coeffs_df$energy, ref=lev) }

# Load fitted data from all models in one large data frame
# fitted_df <- tbl_df( readRDS("../AllFitted.rda") )
fitted_df <- tbl_df(AllFitted)
# Rearrange the energy factors to obtain our preferred ordering.
fitted_df$energy <- replace(fitted_df$energy, which(is.na(fitted_df$energy)), "none")
fitted_df$energy <- replace(fitted_df$energy, which(fitted_df$energy=="iQ"), "q")
fitted_df$energy <- replace(fitted_df$energy, which(fitted_df$energy=="iX"), "x")
fitted_df$energy <- replace(fitted_df$energy, which(fitted_df$energy=="iU"), "u")
fitted_df$energy <- as.factor(fitted_df$energy)
for (lev in rev(energyLevels)) { fitted_df$energy <- relevel(fitted_df$energy, ref=lev) }

# Load the original models.
# oModels <- readRDS(file=file.path("..", "oModels.Rdata"))
oModels <- OrigModels

opts_chunk$set(
  # ********************************************************
  eval=TRUE,
  # ********************************************************
  # tikz allows LaTeX code in graphical output. 
  # E.g., "$y$" for a variable name in the legend for a graph.  
  # dev='tikz',
  tidy=FALSE,
  comment=NA,
  # Tells whether to cache output from chunks are saved. Cacheing saves time. 
  # However, references to LaTeX tables DO NOT WORK if cacheing is turned on.
  cache=FALSE,
  warning=FALSE,  #Tells whether to show warnings in the output.
  # Turns off messages for all chunks. 
  # Set TRUE on an individual chunk to see it.
  message=FALSE, 
  # Tells whether to echo code for all chunks. 
  # Set TRUE on an individual chunk to see its code.
  echo=FALSE     
)
@

<<constants, include=FALSE, eval=TRUE>>=
# Graph Heights
maxHeight <- 8.5 # Inches
ninePanelScatterGraphHeight <- maxHeight # Inches
# ninePanelTernaryGraphHeight <- 7 # Inches
# ninePanelSpaghettiGraphHeight <- 7 # Inches

# Graph Widths
maxWidth <- 6.5 # Inches
fourPanelScatterGraphWidth <- maxWidth # Inches
# threePanelScatterGraphWidth <- maxWidth*0.75 # Inches
# onePanelScatterGraphWidth <- maxWidth*0.25
# fourPanelTernaryGraphWidth <- maxWidth
# twoPanelTernaryGraphWidth <- maxWidth*0.45
# fourPanelSpaghettiGraphWidth <- maxWidth
# twoPanelSpaghettiGraphWidth <- maxWidth*0.55
# onePanelSpaghettiGraphWidth <- maxWidth*0.4

presentationGraph1RowHeight <- 2.8 # Inches
presentationGraph2ColWidth <- 2 # Inches
presentationGraph1ColWidth <- 2*presentationGraph2ColWidth
# keyTextSize <- 0.85 # 85% of normal size
# keyColumns <- 1 # Want only 1 column in the key for lattice graphs
# defaultKeyXLoc <- 0.01 # x position of the key. This default is good for 9-panel graphs.
# defaultKeyYLoc <- 0.95 # y position of the key. This default is good for 9-panel graphs.
# Parameter graphs are tricky to ensure same size. I've found that with height set to 5 in for the 
# cd graphs and 2.93 in for the sf graphs, I get a nearly-exact match with graph size. From these settings I 
# can set up 2 equations:
#    5 in = 2*hgraph + haxis
#    2.93 in = hgraph + haxis
# to estimate that h_axis = 0.86 in and hgraph = 2.05 in.
h_param_graph <- 2.07 # Inches for a single panel
h_param_axis <- 0.86 # Inches (also includes whatever margin is applied)
# sfParameterGraphHeight <- h_param_graph + h_param_axis
cdParameterGraphHeight <- 2*h_param_graph + h_param_axis
cesParameterGraphHeight <- 3*h_param_graph + h_param_axis
linexParameterGraphHeight <- h_param_graph + h_param_axis
sfParameterGraphWidth <- maxWidth
cdParameterGraphWidth <- maxWidth
cesParameterGraphWidth <- maxWidth
linexParameterGraphWidth <- maxWidth
# Other graph parameters that apply to all graphs
scaleTextSize <- 0.8  # Multiple of normal size
scaleTickSize <- -0.5 # 50% of normal size and pointing INWARD!

spaghettiGraphLevel <- 0.95 # Spaghetti graph bands will encompass 95% of the data.
@

% These equations are used in several locations in the paper.
% By defining commands for them, we ensure consistent appearance
% of the equations when they are used.

% Add an argument if you want a label. Example: "\label{eq:CES}"
\newcommand{\ceskl}[1]{\begin{equation} #1
  y = \gamma \: A \: \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{-1/\rho_1}; 
      A = \mathrm{e}^{\lambda (t-t_0)}
\end{equation}}

\newcommand{\ceskle}[1]{\begin{equation} #1
  y = \gamma \: A \: \left\{\delta \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) e^{-\rho} \right\}^{-1/\rho}; 
      A = \mathrm{e}^{\lambda (t-t_0)}
\end{equation}
}

\newcommand{\ceslek}[1]{\begin{equation} #1
  y = \gamma \: A \: \left\{\delta \left[\delta_1 l^{-\rho_1} 
      + (1-\delta_1) e^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) k^{-\rho} \right\}^{-1/\rho}; 
      A = \mathrm{e}^{\lambda (t-t_0)}
\end{equation}
}

\newcommand{\cesekl}[1]{\begin{equation} #1
  y = \gamma \: A \: \left\{\delta \left[\delta_1 e^{-\rho_1} 
      + (1-\delta_1) k^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) l^{-\rho} \right\}^{-1/\rho}; 
      A = \mathrm{e}^{\lambda (t-t_0)}
\end{equation}
}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{An Empirical Analysis of the Role of Energy in Economic Growth}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

\author[CalvinEngr]{Matthew Kuperus Heun\corref{cor1}}
\ead{mkh2@calvin.edu}
\author[Stellenbosch]{Martin de Wit}
\author[CalvinMath]{Randall J. Pruim}
\author[CalvinEngr]{Caleb Reese}
\author[CalvinEngr]{Lucas Timmer}

\cortext[cor1]{Corresponding author}
\address[CalvinEngr]{Engineering Department, Calvin College, Grand Rapids, MI 49546, USA}
\address[Stellenbosch]{Faculty of Economics and Management Sciences, School of Public Leadership, Stellenbosch University, P.O. Box 610, Bellville 7535, South Africa}
\address[CalvinMath]{Mathematics \& Statistics Department, Calvin College, Grand Rapids, MI 49546, USA}

\begin{abstract}
%% Text of abstract
**** Add abstract ****
\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
economic growth \sep energy \sep Cobb-Douglas \sep CES \sep LINEX
%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
\end{keyword}

\end{frontmatter}

% \linenumbers
%% main text

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} 
\label{sec:Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{equation} \label{eq:CD}
  y = \theta \mathrm{e}^{\lambda(t-t_0)} k^\alpha l^\beta \;,
\end{equation}
where 
\begin{itemize}
	\item $y \equiv Y/Y_{0}$,  % why was 0 typeset as mathrm??
	\item $k \equiv K/K_{0}$, 
	\item $l \equiv L/L_{0}$, 
	\item $Y$ represented by GDP, 
	\item $K$ expressed in currency units, 
	\item $L$ expressed in workers or work-hours/year, 
	\item $t$ (time) is measured in years, 
	\item e is the base of the natural logarithm, and 
	\item 0 subscripts indicate values at an initial year.%
\footnote{Dimensionless, indexed quantities are represented by 
lower-case symbols 
($y$, $k$, $l$, $e$, $q$, $x$, and $u$), and dimensional 
quantities are represented by upper-case symbols 
($Y$, $K$, $L$, $E$, $Q$, $X$, and $U$). 
Model parameters are represented by Greek letters
($\alpha$, $\beta$, $\lambda$, $\theta$).} 
\end{itemize}

\begin{equation} \label{eq:CDe}
  y = \theta \mathrm{e}^{\lambda(t-t_0)}  k^\alpha l^\beta e^\gamma \;,
\end{equation}
where $e \equiv E/E_0$, and $E$ is in units of energy per time, 
typically TJ/year. The energy-augmented Cobb-Douglas production function 
is often assumed to have constant returns to scale for the three factors 
of production: $\alpha + \beta + \gamma = 1$.

\ceskl{\label{eq:CES}}

\ceskle{\label{eq:CESkle}}

Constraints on the fitting parameters include 
$\delta_1 \in [0,1]$,
$\delta \in [0,1]$,
$\rho_1 \in [-1,0) \cup (0,\infty)$, and
$\rho \in [-1,0) \cup (0,\infty)$.

\ceslek{\label{eq:CESlek}}

\cesekl{\label{eq:CESekl}}

A third production function, the LINear EXponential (LINEX) function
\begin{equation} \label{eq:LINEX}
  y = \theta \; \mathrm{e}^{a_0\left[2 \left(1 - \frac{1}{\rho_k} \right) 
    + c_t \left(\vphantom{\frac{1}{\rho_k}}\rho_l - 1 \right)\right]} \; e \;,
\end{equation}

\begin{equation} \label{eq:rho_k}
  \rho_k \equiv \frac{k}{(1/2)(l+e)}
\end{equation}
and
\begin{equation} \label{eq:rho_l}
  \rho_l \equiv \frac{l}{e}
\end{equation}
and represent 
(\ref{eq:rho_k}) capital stock increase relative to the average of labor and energy and 
(\ref{eq:rho_l}) labor increase relative to energy.
The effect of energy ($e$) on production is enhanced if either

\begin{itemize}
\item{capital stock ($k$) has increased more than the average 
      of labor ($l$) and energy ($e$), i.e. if $\rho_k >$ 1, or}
\item{labor ($l$) has increased more than energy ($e$), i.e. if $\rho_l >$ 1.}
\end{itemize}
%
An economy that increases 
capital stock ($k$) and labor ($l$) at the same rate as energy will have
capital and labor efficiency ratios ($\rho_k$ and $\rho_l$, respectively) of unity.
In that case, economic output ($y$) increases at the same rate as energy consumption ($e$).
According to the LINEX model, an economy that increases 
capital stock ($k$) without a commensurate increase in labor ($l$) 
and energy ($e$) will experience an increase in output ($y$) in excess 
of its increase of energy consumption ($e$), because
\(
\mathrm{e}^{a_0\left[2 \left(1 - \frac{1}{\rho_k} \right) 
    + c_t \left(\vphantom{\frac{1}{\rho_k}}\rho_l - 1 \right)\right]}
	> 1.
\) 
Similarly, an economy benefits by increasing labor ($l$) 
without a commensurate increase in energy ($e$). 
Thus, $\rho_k$ is an 
efficiency of using additional labor and energy to make 
additional capital stock available to the economy, 
and $\rho_l$ is an efficiency of using additional energy to make 
additional labor available to the economy. 
Increases in $\rho_k$ and $\rho_l$ provide upward pressure on economic output ($y$),
as the only factor of production ($e$) is used more efficiently.

The parameter $a_0$ represents the overall importance of efficiencies $\rho_k$ and $\rho_l$, 
and the parameter $c_t$ represents the relative importance 
of the labor efficiency term ($\rho_l$) compared to the capital stock efficiency term ($\rho_k$). 
Both $a_0$ and $c_t$ are assumed constant with respect to time.

In this paper we analyze at least 
two economies in each of the following four categories: 
Developed, 
NICs,\footnote{Newly Industrialized Countries}
OPEC, and 
LDCs.\footnote{Lesser Developed Countries} 
(See Table~\ref{tab:Economy_Examples}.)

\begin{table} \caption{Example economies and years of evaluation.} \label{tab:Economy_Examples} 
  \begin{center}
    \begin{tabular}{r c c l l} 
      \multirow{3}{*}{Developed} & \includegraphics[height=0.11in]{flag/flag_US.png} & US & USA & \multirow{3}{*}{1980--2011}\\
                                 & \includegraphics[height=0.11in]{flag/flag_UK.png} & UK & United Kingdom & \\
                                 & \includegraphics[height=0.11in]{flag/flag_JP.png} & JP & Japan          & \\
      \hline
      \multirow{2}{*}{NICs}      & \includegraphics[height=0.11in]{flag/flag_CN.png} & CN & China & \multirow{2}{*}{1991--2011}        \\
                                 & \includegraphics[height=0.11in]{flag/flag_ZA.png} & ZA & South Africa & \\
      \hline
      \multirow{2}{*}{OPEC}      & \includegraphics[height=0.11in]{flag/flag_SA.png} & SA & Saudi Arabia & \multirow{2}{*}{1991--2011} \\
                                 & \includegraphics[height=0.11in]{flag/flag_IR.png} & IR & Iran         &  \\
      \hline
      \multirow{2}{*}{LDCs}      & \includegraphics[height=0.11in]{flag/flag_TZ.png} & TZ & Tanzania & \multirow{2}{*}{1991--2011}     \\
                                 & \includegraphics[height=0.11in]{flag/flag_ZM.png} & ZM & Zambia      & \\
    \end{tabular}
  \end{center}
\end{table}

\clearpage

<<Historical_Graph, fig.pos="H", fig.align='center', fig.width=3, fig.height=7.5, fig.cap="Historical data: indexed GDP ($y$), capital stock ($k$), labor ($l$), and thermal energy ($q$).">>=
historicalPlot(historicalData)
@

<<CD_Ternary_Scatter_Graph, fig.pos="H", fig.align='center', fig.width=5.8, fig.height=7.7, fig.cap="Cobb-Douglas resample model parameters.">>=
cdCoeffsData <- filter(coeffs_df, model=="cd", nestStr %in% c("iK+iL", "iK+iL+iQ", "iK+iL+iX", "iK+iL+iU"))
standardTriPlot(data=cdCoeffsData, facet_formula=country~energy)
@

<<CD_GDP_Spaghetti_Graph, fig.pos="H", fig.align='center', fig.width=6, fig.height=8.2, fig.cap="Cobb-Douglas fitted models.">>=
cdFittedData <- filter(fitted_df, model=="cd", nest %in% c("iK+iL", "iK+iL+iQ", "iK+iL+iX", "iK+iL+iU"))
spaghettiPlot(data=cdFittedData, facet_formula=Country~energy, level=spaghettiGraphLevel) +
    labs(y="Indexed GDP (1980=1 or 1991=1)")
@

% <<CES_lambda_gamma_graph, fig.pos="H", fig.align='center', fig.width=fourPanelScatterGraphWidth, fig.height=ninePanelScatterGraphHeight, fig.cap="$\\lambda$ and $\\gamma$ values for CES resample models.">>=
% cesCoeffsData <- filter(coeffs_df, model=="ces", nestStr %in% c("iK+iL", "iK+iL+iQ", "iL+iQ+iK", "iQ+iK+iL"))
% standardScatterPlot(data=cesCoeffsData, mapping=aes(gamma_coef, lambda), facet_formula=country~nest, alpha=0.05) +
%     scale_x_continuous(breaks=c(0.9, 1.0, 1.1)) +
%     scale_y_continuous(breaks=c(0, 0.025, 0.05, 0.075)) +    
%     labs(x=expression(gamma), y=expression(lambda))
% @

<<CES_lambda_gamma_graph_kle, fig.pos="H", fig.align='center', fig.width=fourPanelScatterGraphWidth, fig.height=ninePanelScatterGraphHeight, fig.cap="$\\lambda$ and $\\gamma$ values for CES resample models with (kl)e nesting.">>=
cesCoeffsData_kle <- filter(coeffs_df, model=="ces", nestStr %in% c("iK+iL", "iK+iL+iQ", "iK+iL+iX", "iK+iL+iU"))
standardScatterPlot(data=cesCoeffsData_kle, mapping=aes(gamma_coef, lambda), facet_formula=country~nest, alpha=0.05) +
    scale_x_continuous(breaks=c(0.9, 1.0, 1.1)) +
    scale_y_continuous(breaks=c(0, 0.025, 0.05, 0.075)) +    
    labs(x=expression(gamma), y=expression(lambda))
@

<<CES_lambda_gamma_graph_lek, fig.pos="H", fig.align='center', fig.width=fourPanelScatterGraphWidth, fig.height=ninePanelScatterGraphHeight, fig.cap="$\\lambda$ and $\\gamma$ values for CES resample models with (le)k nesting.">>=
cesCoeffsData_lek <- filter(coeffs_df, model=="ces", nestStr %in% c("iK+iL", "iL+iQ+iK", "iL+iX+iK", "iL+iU+iK"))
standardScatterPlot(data=cesCoeffsData_lek, mapping=aes(gamma_coef, lambda), facet_formula=country~nest, alpha=0.05) +
    scale_x_continuous(breaks=c(0.9, 1.0, 1.1)) +
    scale_y_continuous(breaks=c(0, 0.025, 0.05, 0.075)) +    
    labs(x=expression(gamma), y=expression(lambda))
@

<<CES_lambda_gamma_graph_ekl, fig.pos="H", fig.align='center', fig.width=fourPanelScatterGraphWidth, fig.height=ninePanelScatterGraphHeight, fig.cap="$\\lambda$ and $\\gamma$ values for CES resample models with (ek)l nesting.">>=
cesCoeffsData_ekl <- filter(coeffs_df, model=="ces", nestStr %in% c("iK+iL", "iQ+iK+iL", "iX+iK+iL", "iU+iK+iL"))
standardScatterPlot(data=cesCoeffsData_ekl, mapping=aes(gamma_coef, lambda), facet_formula=country~nest, alpha=0.05) +
    scale_x_continuous(breaks=c(0.9, 1.0, 1.1)) +
    scale_y_continuous(breaks=c(0, 0.025, 0.05, 0.075)) +    
    labs(x=expression(gamma), y=expression(lambda))
@

% <<CES_sigma1_delta1_graph, fig.pos="H", fig.align='center', fig.width=fourPanelScatterGraphWidth, fig.height=ninePanelScatterGraphHeight, fig.cap="$\\sigma_1$ and $\\delta_1$ values for CES resample models.">>=
% standardScatterPlot(data=cesCoeffsData, mapping=aes(delta_1, sigma_1), facet_formula=country~nest, alpha=0.05) +
%     coord_trans(y="sigma") +
%     scale_x_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
%     scale_y_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
%     labs(x=expression(delta[1]), y=expression(sigma[1]))
% @

<<CES_sigma1_delta1_graph_kle, fig.pos="H", fig.align='center', fig.width=fourPanelScatterGraphWidth, fig.height=ninePanelScatterGraphHeight, fig.cap="$\\sigma_1$ and $\\delta_1$ values for CES resample models with (kl)e nesting.">>=
standardScatterPlot(data=cesCoeffsData_kle, mapping=aes(delta_1, sigma_1), facet_formula=country~nest, alpha=0.05) +
    coord_trans(y="sigma") +
    scale_x_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    scale_y_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    labs(x=expression(delta[1]), y=expression(sigma[1]))
@

<<CES_sigma1_delta1_graph_lek, fig.pos="H", fig.align='center', fig.width=fourPanelScatterGraphWidth, fig.height=ninePanelScatterGraphHeight, fig.cap="$\\sigma_1$ and $\\delta_1$ values for CES resample models with (le)k nesting.">>=
standardScatterPlot(data=cesCoeffsData_lek, mapping=aes(delta_1, sigma_1), facet_formula=country~nest, alpha=0.05) +
    coord_trans(y="sigma") +
    scale_x_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    scale_y_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    labs(x=expression(delta[1]), y=expression(sigma[1]))
@

<<CES_sigma1_delta1_graph_ekl, fig.pos="H", fig.align='center', fig.width=fourPanelScatterGraphWidth, fig.height=ninePanelScatterGraphHeight, fig.cap="$\\sigma_1$ and $\\delta_1$ values for CES resample models with (ek)l nesting.">>=
standardScatterPlot(data=cesCoeffsData_ekl, mapping=aes(delta_1, sigma_1), facet_formula=country~nest, alpha=0.05) +
    coord_trans(y="sigma") +
    scale_x_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    scale_y_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    labs(x=expression(delta[1]), y=expression(sigma[1]))
@

% # <<CES_sigma_delta_graph, fig.pos="H", fig.align='center', fig.width=4.5, fig.height=7, fig.cap="$\\sigma$ and $\\delta$ values for CES resample models.">>=
% #  standardScatterPlot(data=subset(cesCoeffsData, nestStr != "iK+iL"),
% #                      mapping=aes(delta, sigma), facet_formula=country~nest, alpha=0.05) +
% #      coord_trans(y="sigma") +
% #      scale_x_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
% #      scale_y_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
% #      labs(x=expression(delta), y=expression(sigma))
% #  @

<<CES_sigma_delta_graph_kle, fig.pos="H", fig.align='center', fig.width=4.5, fig.height=7, fig.cap="$\\sigma$ and $\\delta$ values for CES resample models with (kl)e nesting.">>=
standardScatterPlot(data=subset(cesCoeffsData_kle, nestStr != "iK+iL"),
                    mapping=aes(delta, sigma), facet_formula=country~nest, alpha=0.05) +
    coord_trans(y="sigma") +
    scale_x_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    scale_y_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    labs(x=expression(delta), y=expression(sigma))
@

<<CES_sigma_delta_graph_lek, fig.pos="H", fig.align='center', fig.width=4.5, fig.height=7, fig.cap="$\\sigma$ and $\\delta$ values for CES resample models with (le)k nesting.">>=
standardScatterPlot(data=subset(cesCoeffsData_lek, nestStr != "iK+iL"),
                    mapping=aes(delta, sigma), facet_formula=country~nest, alpha=0.05) +
    coord_trans(y="sigma") +
    scale_x_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    scale_y_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    labs(x=expression(delta), y=expression(sigma))
@

<<CES_sigma_delta_graph_ekl, fig.pos="H", fig.align='center', fig.width=4.5, fig.height=7, fig.cap="$\\sigma$ and $\\delta$ values for CES resample models with (ek)l nesting.">>=
standardScatterPlot(data=subset(cesCoeffsData_ekl, nestStr != "iK+iL"),
                    mapping=aes(delta, sigma), facet_formula=country~nest, alpha=0.05) +
    coord_trans(y="sigma") +
    scale_x_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    scale_y_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    labs(x=expression(delta), y=expression(sigma))
@

We can develop pseudo $\alpha$, $\beta$, and $\gamma$ parameters 
for the CES model, based on values of $\delta$ and $\delta_1$.
Table~\ref{tab:CES_abg_equations} shows the permuted equations for each CES nesting.

\begin{table} \caption{Equations for $\alpha$, $\beta$, and $\gamma$ for the various CES nestings.} 
\label{tab:CES_abg_equations} 
  \begin{center}
    \begin{tabular}{c c c c c} 
      Nesting    & Equation        & $\alpha$                & $\beta$                 & $\gamma$                \\
      \hline
      ($kl$)     & \ref{eq:CES}    & $\delta_1$              & $1 - \delta_1$          & 0                       \\
      ($kl$)$e$  & \ref{eq:CESkle} & $\delta \delta_1$       & $\delta (1 - \delta_1)$ & $1 - \delta$            \\   
      ($le$)$k$  & \ref{eq:CESlek} & $1 - \delta$            & $\delta \delta_1$       & $\delta (1 - \delta_1)$ \\
      ($ek$)$l$  & \ref{eq:CESekl} & $\delta (1 - \delta_1)$ & $1 - \delta$            & $\delta \delta_1$       \\
    \end{tabular}
  \end{center}
\end{table}

Figure~\ref{fig:CES_ternary_graph} shows 
$\alpha$, $\beta$, and $\gamma$ parameters for the CES model for all nests and countries.
Figure~\ref{fig:CES_ternary_graph} is comparable to
the ternary graph for the Cobb-Douglas models, 
Figure~\ref{fig:CD_Ternary_Scatter_Graph}.

% <<CES_ternary_graph, fig.pos="H", fig.align='center', fig.width=5.5, fig.height=7.5, fig.cap="$\\alpha$, $\\beta$, and $\\gamma$ values for CES resample models.">>=
% standardTriPlot(data=cesCoeffsData, facet_formula=country~nest)
% @

<<CES_ternary_graph_kle, fig.pos="H", fig.align='center', fig.width=5.5, fig.height=7.5, fig.cap="$\\alpha$, $\\beta$, and $\\gamma$ values for CES resample models with (kl)e nesting.">>=
standardTriPlot(data=cesCoeffsData_kle, facet_formula=country~nest)
@

<<CES_ternary_graph_lek, fig.pos="H", fig.align='center', fig.width=5.5, fig.height=7.5, fig.cap="$\\alpha$, $\\beta$, and $\\gamma$ values for CES resample models with (le)k nesting.">>=
standardTriPlot(data=cesCoeffsData_lek, facet_formula=country~nest)
@

<<CES_ternary_graph_ekl, fig.pos="H", fig.align='center', fig.width=5.5, fig.height=7.5, fig.cap="$\\alpha$, $\\beta$, and $\\gamma$ values for CES resample models with (ek)l nesting.">>=
standardTriPlot(data=cesCoeffsData_ekl, facet_formula=country~nest)
@

% <<CES_Spaghetti_Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=8.5, fig.cap="CES fitted models.">>=
% cesFittedData <- filter(fitted_df, model=="ces", nest %in% c("iK+iL", "iK+iL+iQ", "iL+iQ+iK", "iQ+iK+iL"))
% spaghettiPlot(data=cesFittedData, facet_formula=Country~nest, level=spaghettiGraphLevel) +
%     labs(y="Indexed GDP (1980=1 or 1991=1)")
% @

<<CES_Spaghetti_Graph_kle, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=8.5, fig.cap="CES fitted models with (kl)e nesting.">>=
cesFittedData <- filter(fitted_df, model=="ces", nest %in% c("iK+iL", "iK+iL+iQ", "iK+iL+iX", "iK+iL+iU"))
spaghettiPlot(data=cesFittedData, facet_formula=Country~nest, level=spaghettiGraphLevel) +
    labs(y="Indexed GDP (1980=1 or 1991=1)")
@

<<CES_Spaghetti_Graph_lek, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=8.5, fig.cap="CES fitted models with (le)k nesting.">>=
cesFittedData <- filter(fitted_df, model=="ces", nest %in% c("iK+iL", "iL+iQ+iK", "iL+iX+iK", "iL+iU+iK"))
spaghettiPlot(data=cesFittedData, facet_formula=Country~nest, level=spaghettiGraphLevel) +
    labs(y="Indexed GDP (1980=1 or 1991=1)")
@

<<CES_Spaghetti_Graph_ekl, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=8.5, fig.cap="CES fitted models with (ek)l nesting.">>=
cesFittedData <- filter(fitted_df, model=="ces", nest %in% c("iK+iL", "iQ+iK+iL", "iX+iK+iL", "iU+iK+iL"))
spaghettiPlot(data=cesFittedData, facet_formula=Country~nest, level=spaghettiGraphLevel) +
    labs(y="Indexed GDP (1980=1 or 1991=1)")
@

% <<CES_Spaghetti_Graph_CN, fig.pos="H", fig.align='center', fig.width=maxWidth, fig.height=maxHeight, fig.cap="CES fitted models for China with the (kl)e nesting.">>=
% spaghettiPlot(data=subset(cesFittedData, Country=="CN" & nest=="iK+iL+iQ"), level=spaghettiGraphLevel) +
%     labs(y="Indexed GDP (1980 = 1 or 1991 = 1")
% @

% <<LINEXq_Resample_Graph, fig.pos="H", fig.align='center', fig.width=2, fig.height=7, fig.cap="LINEX model (with energy) resampling results.">>=
% linexCoeffsData <- filter(coeffs_df, model=="linex", energy=="q")
% standardScatterPlot(data=linexCoeffsData, aes(a_0, c_t), facet_formula=country~energy, alpha=0.05) +
%     labs(x=expression(a[0]), y=expression(c[t])) +
%      scale_x_continuous(limits=c(-1, 2), breaks=c(-1, 0, 1.0, 2.0)) +
%      scale_y_continuous(limits=c(0, 10), breaks=c(0, 5, 10))
% @

<<LINEXq_Resample_Graph, fig.pos="H", fig.align='center', fig.width=4.5, fig.height=7, fig.cap="LINEX model (with energy) resampling results.">>=
linexCoeffsData <- filter(coeffs_df, model=="linex")
standardScatterPlot(data=linexCoeffsData, aes(a_0, c_t), facet_formula=country~energy, alpha=0.05) +
    labs(x=expression(a[0]), y=expression(c[t])) +
     scale_x_continuous(limits=c(-1, 2), breaks=c(-1, 0, 1.0, 2.0)) +
     scale_y_continuous(limits=c(0, 10), breaks=c(0, 5, 10))
@

% <<LINEX_Spaghetti_Graph, fig.pos="H", fig.align='center', fig.width=2, fig.height=7.1, fig.cap="LINEX fitted models.">>=
% linexFittedData <- filter(fitted_df, model=="linex", energy=="q")
% spaghettiPlot(data=linexFittedData, facet_formula=Country~energy, level=spaghettiGraphLevel) +
%     labs(y="Indexed GDP (1980=1 or 1991=1)")
% @

<<LINEX_Spaghetti_Graph, fig.pos="H", fig.align='center', fig.width=4.5, fig.height=7.1, fig.cap="LINEX fitted models.">>=
linexFittedData <- filter(fitted_df, model=="linex")
spaghettiPlot(data=linexFittedData, facet_formula=Country~energy, level=spaghettiGraphLevel) +
    labs(y="Indexed GDP (1980=1 or 1991=1)")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}
\label{sec:Acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Funding for Caleb Reese and Lucas Timmer was provided 
by the Jack and Lois Kuipers Applied Mathematics Endowment and 
the Calvin College Alumni Association. 
Partial support for Matthew Kuperus Heun was 
provided by a Calvin College Research Followship. 
Partial funding for Martin de Wit was provided 
by the South African National Research Foundation (NRF).\footnote{Any opinion, 
findings, and conclusions or recommendations expressed herein are those 
of the authors, and, therefore, NRF does not accept any liability in regard thereto.} 
The authors are grateful to Benjamin S. Warr (INSEAD) for providing 
the REXS database \citep{Warr:2012a}. 
The authors thank Loren L. Heun (Western Michigan University) 
for insightful critique of the statistical analyses presented herein.
The computations in this paper were accomplished using \url{http://dahl.calvin.edu}, 
a Beowulf cluster built with support from the National Science Foundation's 
Major Research Instrumentation and Cyber Infrastructure programs, 
grant number OCI-0722819.

%% References
%%
%% Following citation commands can be used in the body text:
%%
%%  \citet{key}  ==>>  Jones et al. (1990)
%%  \citep{key}  ==>>  (Jones et al., 1990)
%%
%% Multiple citations as normal:
%% \citep{key1,key2}         ==>> (Jones et al., 1990; Smith, 1989)
%%                            or  (Jones et al., 1990, 1991)
%%                            or  (Jones et al., 1990a,b)
%% \cite{key} is the equivalent of \citet{key} in author-year mode
%%
%% Full author lists may be forced with \citet* or \citep*, e.g.
%%   \citep*{key}            ==>> (Jones, Baker, and Williams, 1990)
%%
%% Optional notes as:
%%   \citep[chap. 2]{key}    ==>> (Jones et al., 1990, chap. 2)
%%   \citep[e.g.,][]{key}    ==>> (e.g., Jones et al., 1990)
%%   \citep[see][pg. 34]{key}==>> (see Jones et al., 1990, pg. 34)
%%  (Note: in standard LaTeX, only one note is allowed, after the ref.
%%   Here, one note is like the standard, two make pre- and post-notes.)
%%
%%   \citealt{key}          ==>> Jones et al. 1990
%%   \citealt*{key}         ==>> Jones, Baker, and Williams 1990
%%   \citealp{key}          ==>> Jones et al., 1990
%%   \citealp*{key}         ==>> Jones, Baker, and Williams, 1990
%%
%% Additional citation possibilities
%%   \citeauthor{key}       ==>> Jones et al.
%%   \citeauthor*{key}      ==>> Jones, Baker, and Williams
%%   \citeyear{key}         ==>> 1990
%%   \citeyearpar{key}      ==>> (1990)
%%   \citetext{priv. comm.} ==>> (priv. comm.)
%%   \citenum{key}          ==>> 11 [non-superscripted]
%% Note: full author lists depends on whether the bib style supports them;
%%       if not, the abbreviated list is printed even when full requested.
%%
%% For names like della Robbia at the start of a sentence, use
%%   \Citet{dRob98}         ==>> Della Robbia (1998)
%%   \Citep{dRob98}         ==>> (Della Robbia, 1998)
%%   \Citeauthor{dRob98}    ==>> Della Robbia

%% References with bibTeX database:

\bibliographystyle{model2-names}
%%\bibliography{<your-bib-database>}
\bibliography{EconGrowth.bib}

%% Authors are advised to submit their bibtex database files. They are
%% requested to list a bibtex style file in the manuscript if they do
%% not want to use model2-names.bst.

%% References without bibTeX database:

% \begin{thebibliography}{00}

%% \bibitem must have one of the following forms:
%%   \bibitem[Jones et al.(1990)]{key}...
%%   \bibitem[Jones et al.(1990)Jones, Baker, and Williams]{key}...
%%   \bibitem[Jones et al., 1990]{key}...
%%   \bibitem[\protect\citeauthoryear{Jones, Baker, and Williams}{Jones
%%       et al.}{1990}]{key}...
%%   \bibitem[\protect\citeauthoryear{Jones et al.}{1990}]{key}...
%%   \bibitem[\protect\astroncite{Jones et al.}{1990}]{key}...
%%   \bibitem[\protect\citename{Jones et al., }1990]{key}...
%%   \harvarditem[Jones et al.]{Jones, Baker, and Williams}{1990}{key}...
%%

% \bibitem[ ()]{}

% \end{thebibliography}

\end{document}

%%
%% End of file `elsarticle-template-2-harv.tex'.