### 2014-11-02 Articles/Article_JAE_2014/Education Investigation.Rmd
Beginning Education Investigation.Rmd.

---
title: "Education Investigation"
author: "MK Heun"
date: "November 2, 2014"
output: pdf_document
---

```{r Setup, echo=FALSE, message=FALSE, warning=FALSE}
require(EconData)
require(EconModels)
require(ggplot2)
```

This is a study of the effects of education on the estimation of parameters for South African growth.

Data
----------------------------------------------------------------
The following graph shows historical data, indexed to 1973.

```{r HistoricalData}
historicalData <- SUN
```

Formal Education Only
----------------------------------------------------------------
We begin by fitting GDP to the historical capital (K) and labor (L) data, using formal labor data only.

```{r}
summary(cars)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.