### 2014-11-03 Articles/Article_JAE_2014/Education Investigation.Rmd
Additional results on the education study.

---
title: "The Effect of Education Adjustment on Parameter Estimates 
for Cobb-Douglas Economic Growth in South Africa"
author: "Matthew Kuperus Heun"
date: "November 2, 2014"
output: pdf_document
---

```{r Setup, echo=FALSE, message=FALSE, warning=FALSE}
require(EconData)
require(EconModels)
require(ggplot2)
require(reshape2)
historicalData <- SUN
```

This is a study of the effects of education on the estimation of parameters for South African economic growth.

Historical Data
----------------------------------------------------------------
The following figure shows historical data for GDP, capital stock, and thermal energy data, indexed to 1973.

```{r HistoricalData, warning=FALSE, fig.cap="Historical GDP, capital stock, and thermal energy data."}
mdata <- melt(historicalData, 
              measure.vars = c("iGDP", "iK", "iQp"), 
              variable.name = "variable",
              value.name = "value")
mapping <- aes(x=Year, y=value, group=variable, linetype=variable)
ggplot(mdata, mapping=mapping) + 
  geom_line(mapping) +
  xy_theme() +
  xlab("") + 
  ylab("Indexed (1973=1)")
```

The following figure shows four different time series for labor.
$iLf$ is indexed formal labor.
$iLfe$ is indexed formal labor adjusted for education.
$iLfi$ is indexed formal plus informal labor.
$ilfie$ is indexed formal plus informal labor adjusted for education.
Adjustment for education is performed as described in du Plessis and Smit. (2007)

```{r HistoricalDataLabor, warning=FALSE, fig.cap="Historical labor time series."}
mdata_labor <- melt(historicalData, 
                    measure.vars = c("iLf", "iLfe", "iLfi", "iLfie"), 
                    variable.name = "variable",
                    value.name = "value")
mapping <- aes(x=Year, y=value, group=variable, linetype=variable)
ggplot(mdata_labor, mapping=mapping) + 
  geom_line(mapping) +
  xy_theme() +
  xlab("") + 
  ylab("Indexed (1973=1)") +
  scale_y_continuous(breaks=c(1, 1.5, 2.0))
```

Note that the effect of education is not observed until 1986, 
because the average education level is given as 4.8 years
from 1970--1985.

du Plessis and Smit (2007, p. 681) report that 
capital income share is 0.43
and wage income share is 0.57
for the time period 1985--1994.
For 1995--2004, 
capital income share is 0.46
and wage income share is 0.54

Cobb-Douglas Equation
----------------------------------------------------------------
We fit a Cobb-Douglas equation of the form

\begin{equation}
  GDP = \mathrm{e}^{\lambda t} k^\alpha l^\beta
\end{equation}

under the constraint $\alpha + \beta = 1$ to ensure constant returns to scale.
The GDP and capital stock time series are fixed.
However, we will try four different labor time series.

Without Education Adjustment
----------------------------------------------------------------
We begin by fitting GDP to the historical capital ($iK$) and formal sector labor ($iLf$) data.

```{r Formal}
naturalCoef(cdModel(formula = iGDP ~ iK + iLf + iYear, 
                    data = historicalData, constrained = TRUE))
```

Next, fit using formal and informal labor data ($iLfi$)
```{r FormalPlusInformal}
naturalCoef(cdModel(formula = iGDP ~ iK + iLfi + iYear, 
                    data = historicalData, constrained = TRUE))
```

Without adjusting for education, 
both formal and formal+informal labor time series produce 
$\alpha = 0$
and $\beta = 1$.
The Solow residual ($\lambda$) is on the order of 1.5 \%/year.

With Education Adjustment
----------------------------------------------------------------
Now we fit using education-adjusted formal sector labor data ($iLfe$).
```{r FormalPlusEducation}
naturalCoef(cdModel(formula = iGDP ~ iK + iLfe + iYear, 
                    data = historicalData, constrained = TRUE))
```

Finally, fit using education-adjusted formal and informal labor data ($iLfie$)
```{r FormalPlusInformalPlueEducation}
naturalCoef(cdModel(formula = iGDP ~ iK + iLfie + iYear, 
                    data = historicalData, constrained = TRUE))
```

When adjustment is made for education, we obtain significantly different 
coefficients for capital ($\alpha$) and labor ($\beta$), namely
$\alpha \approx 0.44$
and $\beta \approx 0.56$.
The Solow residual is $\lambda \approx 0.5$ \%/year.

Analysis
----------------------------------------------------------------
The education-adjusted coefficient estimates above 
($\alpha \approx 0.44$
and $\beta \approx 0.56$) 
align closely with income shares as reported in du Plessis and Smit. (2007, p. 681)

The fit to historical data is slightly better without education adjustment
than with education adjustment
($sse = 0.18$ without adjusting for education 
vs. $sse = 0.27$ with education adjustment).
However, the Solow residual is smaller when adjusting for education
($\lambda \approx 1.5$ \%/year without adjusting for education
vs. $\lambda \approx 0.5$ \%/year after adjusting for education).

Furthermore $sse$ improves when the informal labor sector is included,
regardless of whether labor is adjusted for education.

The improvement in Solow residual 
argues strongly in favor of including both education and the informal labor sector.

It appears that adjusting for education 
will be an important feature of these analyses moving forward.