### 2014-11-28 Articles/Article_EcolEcon_2014/Paper2.Rnw
Uncommented many lines for CES model.

% This article has been prepared for publication in Ecological Economics in RStudio with knitr.
% According to http://www.elsevier.com/author-schemas/the-elsarticle-latex-document-class, we should be using the
% elsarticle.cls file.
% According to http://cdn.elsevier.com/assets/pdf_file/0006/109392/journal_refstyles.pdf, we should be using
% elsarticle-template-2-harv.tex as the template for the text.
% Furthermore, we should be using model2-names.bst for the bibliographic references.
% The approach here is to load the frontmatter and backmatter from elsarticle-template-2-harv.tex
% both ahead of and behind the text for our paper.
% -- Matthew Kuperus Heun, 2014-08-01

%% This is file `elsarticle-template-2-harv.tex',
%%
%% Copyright 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle  Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for Elsevier's document class `elsarticle'
%% with harvard style bibliographic references
%%
%% $Id: elsarticle-template-2-harv.tex 155 2009-10-08 05:35:05Z rishi $
%% $URL: http://lenova.river-valley.com/svn/elsbst/trunk/elsarticle-template-2-harv.tex $
%%
\documentclass[preprint,authoryear,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,authoryear,1p,times]{elsarticle}
%% \documentclass[final,authoryear,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,authoryear,3p,times]{elsarticle}
%% \documentclass[final,authoryear,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,authoryear,5p,times]{elsarticle}
%% \documentclass[final,authoryear,5p,times,twocolumn]{elsarticle}

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%% or use the graphicx package for more complicated commands
%% \usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

\usepackage{wrapfig}    % Allows wrapping of text around figures
\usepackage{soul}       % Provides strikethrough text
\usepackage{float}      % Allows precise positioning of tables and figures within the text
\usepackage{amsmath}    % Allows \begin{equation*} \end{equation*} for unnumbered equations
\usepackage{multirow}   % To create multirow tables
\usepackage{hyperref}   % To create hyperlinks in the paper
\usepackage{microtype}  % produces hanging punctuation and beautiful type

% From http://economics.utoronto.ca/osborne/latex/BIBTEX.HTM
\newcommand{\citeapos}[1]{\citeauthor{#1}'s (\citeyear{#1})} % Posessive citations.

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon (default)
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   authoryear - selects author-year citations (default)
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%   longnamesfirst  -  makes first citation full author list
%%
%% \biboptions{longnamesfirst,comma}

% \biboptions{}

\journal{Ecological Economics}

\begin{document}

<<setup, echo=FALSE, message=FALSE, eval=TRUE>>=
require(EconModels)
require(EconData)
require(ggplot2)
require(dplyr) # Load in this order or the mutate and arrange functions will not work as desired.
# Provides access to data melting. 
# See http://cran.r-project.org/web/packages/reshape2/reshape2.pdf
require(reshape2) 
require(knitr)
require(mosaic)

opts_chunk$set(
  # ********************************************************
  eval=TRUE,
  # ********************************************************
  # tikz allows LaTeX code in graphical output. 
  # E.g., "$y$" for a variable name in the legend for a graph.  
  # dev='tikz',
  tidy=FALSE,
  comment=NA,
  # Tells whether to cache output from chunks are saved. Cacheing saves time. 
  # However, references to LaTeX tables DO NOT WORK if cacheing is turned on.
  cache=FALSE,
  warning=FALSE,  #Tells whether to show warnings in the output.
  # Turns off messages for all chunks. 
  # Set TRUE on an individual chunk to see it.
  message=FALSE, 
  # Tells whether to echo code for all chunks. 
  # Set TRUE on an individual chunk to see its code.
  echo=FALSE     
)
@

<<constants, include=FALSE, eval=TRUE>>=
spaghettiGraphLevel <- 1.0 # Spaghetti graph bands will encompass 95% of the data.
@

<<data, include=FALSE>>=
# Load historical data
historicalData <- plyr::rbind.fill(REXS1960, IST, Leeds) # From the EconData package

# Add some pre-econometric data
historicalData <- transform(historicalData, KPerGDP = iK/iGDP)
historicalData <- transform(historicalData, LPerGDP = iL/iGDP)
historicalData <- transform(historicalData, XpPerGDP = iXp/iGDP)
historicalData <- transform(historicalData, XfPerGDP = iXf/iGDP)
historicalData <- transform(historicalData, UPerGDP = iU/iGDP)

# The working directory is set to the directory that contains this file.
# To reach the post-processed data, we need to move 
# up two levels and into the data_postprocessed directory.
# So, in each of the loadPostProcessedData calls below, we use
# dir = "../../data_postprocessed".
pp_path <- file.path("..", "..", "data_postprocessed")

# Load coefficients from all models in one large data frame
REXS1960_Coeffs <- loadPostProcessedData(Source = "REXS1960", kind = "coeffs", dir = pp_path)
IST_Coeffs <- loadPostProcessedData(Source = "IST", kind = "coeffs", dir = pp_path)
Leeds_Coeffs <- loadPostProcessedData(Source = "Leeds", kind = "coeffs", dir = pp_path)
AllCoeffs <- rbind(REXS1960_Coeffs, IST_Coeffs, Leeds_Coeffs)
AllCoeffs$Source <- as.factor(replace(AllCoeffs$Source, which(AllCoeffs$Source=="REXS1960"), "REXS"))
AllCoeffs$Source <- relevelFactor(AllCoeffs$Source, levs = c("REXS", "IST", "Leeds"))
AllCoeffs$energy <- relevelFactor(AllCoeffs$energy, levs = energyLevels)
# Load fitted data from all models in one large data frame
REXS1960_Fitted <- loadPostProcessedData(Source = "REXS1960", kind = "fitted", dir = pp_path)
IST_Fitted <- loadPostProcessedData(Source = "IST", kind = "fitted", dir = pp_path)
Leeds_Fitted <- loadPostProcessedData(Source = "Leeds", kind = "fitted", dir = pp_path)
AllFitted <- rbind(REXS1960_Fitted, IST_Fitted, Leeds_Fitted)
AllFitted$Source <- replace(AllFitted$Source, which(AllFitted$Source=="REXS1960"), "REXS")
AllFitted$Source <- relevelFactor(as.factor(AllFitted$Source), levs = c("REXS", "IST", "Leeds"))
AllFitted$energy <- relevelFactor(AllFitted$energy, levs=energyLevels)

# Load models
REXS1960_Models <- loadPostProcessedData(Source = "REXS1960", kind = "models", dir=pp_path)
IST_Models <- loadPostProcessedData(Source = "IST", kind = "models", dir=pp_path)
Leeds_Models <- loadPostProcessedData(Source = "Leeds", kind = "models", dir=pp_path)
AllModels <- unlist(list(REXS1960_Models, IST_Models, Leeds_Models), recursive=FALSE)
# The function to extract data from the models
extract_data <- function(x, name) {
  # Names are in the form Source.Country.modelType.factor_string
  # Parse various pieces of name
  name_pieces <- strsplit(name, split=".", fixed=TRUE)[[1]]
  Source <- name_pieces[[1]]
  Country <- name_pieces[[2]]
  modelType <- name_pieces[[3]]
  factorString <- name_pieces[[4]]
  # Parse the factorString to get the energyType
  energyType <- parseFactorString(factorString)$energyType
  data.frame(year = (attr(x, which="data"))$Year, resid = resid(x), name=name, Source=Source, Country=Country,
             modelType=modelType, energyType=energyType)
}
residual_data <- leaf_apply(AllModels, extract_data, class=c("CDEmodel", "cesEst", "LINEXmodel"))
# Bind these all in a data frame. 
residuals_df <- do.call(rbind, residual_data)
# Re-order the factors as desired for our graphs.
residuals_df$Source <- as.factor(
  replace(as.character(residuals_df$Source), which(residuals_df$Source=="REXS1960"), "REXS"))
residuals_df$Source <- relevelFactor(residuals_df$Source, levs = c("REXS", "IST", "Leeds"))
residuals_df$energyType <- as.factor(
  replace(as.character(residuals_df$energyType), which(is.na(residuals_df$energyType)), "none"))
residuals_df$energyType <- relevelFactor(as.factor(residuals_df$energyType), levs = energyLevels)
@

% These equations are used in several locations in the paper.
% By defining commands for them, we ensure consistent appearance
% of the equations when they are used.

% Add an argument if you want a label. Example: "\label{eq:CES}"
\newcommand{\ceskl}[1]{\begin{equation} #1
  y = \gamma \: A \: \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{-1/\rho_1}; 
      A = \mathrm{e}^{\lambda (t-t_0)}
\end{equation}}

\newcommand{\ceskle}[1]{\begin{equation} #1
  y = \gamma \: A \: \left\{\delta \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) e^{-\rho} \right\}^{-1/\rho}; 
      A = \mathrm{e}^{\lambda (t-t_0)}
\end{equation}
}

\newcommand{\ceslek}[1]{\begin{equation} #1
  y = \gamma \: A \: \left\{\delta \left[\delta_1 l^{-\rho_1} 
      + (1-\delta_1) e^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) k^{-\rho} \right\}^{-1/\rho}; 
      A = \mathrm{e}^{\lambda (t-t_0)}
\end{equation}
}

\newcommand{\cesekl}[1]{\begin{equation} #1
  y = \gamma \: A \: \left\{\delta \left[\delta_1 e^{-\rho_1} 
      + (1-\delta_1) k^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) l^{-\rho} \right\}^{-1/\rho}; 
      A = \mathrm{e}^{\lambda (t-t_0)}
\end{equation}
}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{Paper 2 !DRAFT! Do not circulate!}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

% \author[BenWarr]{Benjamin Warr}
% \author[IST]{Tiago Domingos}
% \author[Leeds]{Paul Brockway}
% \author[Leeds]{Julia Steinberger}
% \author[CalvinEngr]{Matthew Kuperus Heun\corref{cor1}}
% \ead{mkh2@calvin.edu}

% \address[BenWarr]{Ben Warr's address, Zambia}
% \address[IST]{IST, Lisbon, Portugal}
% \address[Leeds]{Leeds University, Leeds, UK}
% \cortext[cor1]{Corresponding author}
% \address[CalvinEngr]{Engineering Department, Calvin College, Grand Rapids, MI 49546, USA}

\begin{abstract}
%% Text of abstract
**** Add abstract ****
\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
economic growth \sep exergy \sep energy \sep Cobb-Douglas \sep CES \sep LINEX
%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
\end{keyword}

\end{frontmatter}

% \linenumbers
%% main text

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} 
\label{sec:Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Coordinates of Analysis} 
\label{sec:Coordinates}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This section describes the coordinates of analysis and briefly reviews 
literature related to each.

%++++++++++++++++++++++++++++++
\subsection{Mathematical Forms of the Energy-augmented Production Function} 
\label{sec:Prod_Func_Forms}
%++++++++++++++++++++++++++++++

In this paper, we assess three prominent energy-augmented production functions 
that appear in the literature: 
Cobb-Douglas (CD), 
Constant Elasticity of Substitution (CES), and 
LINear EXponential (LINEX). 
The following subsections describe each.

%------------------------------
\subsubsection{Cobb-Douglas Production Function} 
\label{sec:CDe}
%------------------------------

The Cobb-Douglas production function can be expressed as
%
\begin{equation} \label{eq:CD}
  y = \theta \mathrm{e}^{\lambda(t-t_0)} k^\alpha l^\beta \;,
\end{equation}
%
\noindent where 
$y \equiv Y/Y_{0}$,
$\theta$ is a scale parameter,
e is the base of the natural logarithm, 
$\lambda$ is represents the pace of technological progress,
$t$ (time) is measured in years,
$k \equiv K/K_{0}$, 
$l \equiv L/L_{0}$, 
$Y$ (economic output) is represented by GDP, 
$K$ (capital) is expressed in currency units, 
$L$ (labor) is expressed in workers or work-hours/year, and
the 0 subscript indicates values at an initial year.%
  \footnote{Dimensionless, indexed quantities are represented by 
  lower-case symbols 
  ($y$, $k$, $l$, $e$, $q$, $x$, and $u$), and dimensional 
  quantities are represented by upper-case symbols 
  ($Y$, $K$, $L$, $E$, $Q$, $X$, and $U$). 
  Model parameters are represented by Greek letters
  ($\alpha$, $\beta$, $\lambda$, $\theta$).
  }
Constant returns to scale are represented by the constraint
$\alpha + \beta = 1$.

The capital-labor Cobb-Douglas production function shown in Equation~\ref{eq:CD}
can be augmented to include an energy term:
%
\begin{equation} \label{eq:CDe}
  y = \theta \mathrm{e}^{\lambda(t-t_0)}  k^\alpha l^\beta e^\gamma \;,
\end{equation}
%
where $e \equiv E/E_0$, and $E$ is in units of energy per time, typically TJ/year. 
The energy-augmented Cobb-Douglas production function 
is often assumed to have constant returns to scale for the three factors 
of production: $\alpha + \beta + \gamma = 1$.

The term $\mathrm{e}^{\lambda (t-t_0)}$ is known as total factor productivity~($A$),
and $\lambda$ is the Solow residual. 
Assuming that parameters $\theta$, $\alpha$, and $\beta$ are known from parameter estimation,
we can estimate the value of $\lambda$ as follows.
First, assume constant returns to scale such that $\gamma = 1 - \alpha - \beta$.
Then, take the natural logarithm ($\ln$) of Equation~\ref{eq:CD} to obtain
%
\begin{equation} \label{eq:lnCD}
  \ln y = \ln \theta + \ln A + \alpha \ln k + \beta \ln l + (1 - \alpha - \beta) \ln e \;.
\end{equation}
%
By taking the derivative of Equation~\ref{eq:lnCD} with respect to time ($t$) 
and noting that model parameters $\theta$, $\alpha$, and $\beta$ are 
constant with respect to time, we obtain
%
\begin{equation} \label{eq:ddtlnCD}
  \frac{1}{y}\frac{\mathrm{d}y}{\mathrm{d}t} 
      = \frac{1}{A}\frac{\mathrm{d}A}{\mathrm{d}t} 
      + \alpha \frac{1}{k}\frac{\mathrm{d}k}{\mathrm{d}t} 
      + \beta \frac{1}{l}\frac{\mathrm{d}l}{\mathrm{d}t} 
      + (1 - \alpha - \beta) \frac{1}{e}\frac{\mathrm{d}e}{\mathrm{d}t} \;.
\end{equation}
%
Solving for the Solow Residual term gives
%
\begin{equation} \label{eq:solow_resid}
  \frac{1}{A}\frac{\mathrm{d}A}{\mathrm{d}t} 
      = \frac{1}{y}\frac{\mathrm{d}y}{\mathrm{d}t} 
      - \left[
              \alpha \frac{1}{k}\frac{\mathrm{d}k}{\mathrm{d}t} 
              + \beta \frac{1}{l}\frac{\mathrm{d}l}{\mathrm{d}t} 
              + (1 - \alpha - \beta) \frac{1}{e}\frac{\mathrm{d}e}{\mathrm{d}t} 
        \right] \;.
\end{equation}
%
Recognizing that $A \equiv \mathrm{e}^{\lambda (t-t_0)}$ gives
%
\begin{equation}
  \frac{1}{A}\frac{\mathrm{d}A}{\mathrm{d}t} 
    = \frac{1}{\mathrm{e}^{\lambda (t-t_0)}} \; \lambda \mathrm{e}^{\lambda (t-t_0)}
    = \lambda \; ,
\end{equation}
%
which can be substituted into Equation~\ref{eq:solow_resid} to find
%
\begin{equation} \label{eq:lambda}
  \lambda 
    = \frac{1}{y}\frac{\mathrm{d}y}{\mathrm{d}t} 
    - \left[
              \alpha \frac{1}{k}\frac{\mathrm{d}k}{\mathrm{d}t} 
              + \beta \frac{1}{l}\frac{\mathrm{d}l}{\mathrm{d}t} 
              + (1 - \alpha - \beta) \frac{1}{e}\frac{\mathrm{d}e}{\mathrm{d}t} 
      \right] \;.
\end{equation}
%
Equation~\ref{eq:lambda} applies for any instant in time.

If we approximate derivatives in Equation~\ref{eq:lambda} with forward differences
between times $i$ and $j$, we find
%
\begin{equation} \label{eq:lambdaij}
  \lambda_{i,j}
    \approx \frac{1}{y_i}\frac{y_j - y_i}{t_j - t_i} 
    - \left[
            \alpha \frac{1}{k_i}\frac{k_j - k_i}{t_j - t_i} 
            + \beta \frac{1}{l_i}\frac{l_j - l_i}{t_j - t_i} 
            + (1 - \alpha - \beta) \frac{1}{e_i}\frac{e_j - e_i}{t_j - t_i} 
      \right] \;,
\end{equation}
% 
where $\lambda_{i,j}$ approximates the true, instantaneous value of $\lambda$
between times $i$ and $j$.

Interestingly, 
%
\begin{equation}
  \lambda_{1,n}
    \neq \sum_{i=2}^{n} \lambda_{i,i-1}  \;.
\end{equation}
% 
Rather,
%
\begin{equation} \label{eq:lambda1n}
  \lambda_{1,n}
    \approx \frac{1}{y_1}\frac{y_n - y_1}{t_n - t_1} 
      - \left[ 
            \alpha \frac{1}{k_1}\frac{k_n - k_1}{t_n - t_1} 
            + \beta \frac{1}{l_1}\frac{l_n - l_1}{t_n - t_1} 
            + (1 - \alpha - \beta) \frac{1}{e_1}\frac{e_n - e_1}{t_n - t_1} 
      \right] \; ,
\end{equation}
% 
which will become increasingly inaccurate over large time spans,
because $y_1$, $k_1$, $l_1$, and $e_1$ will be less representative
of the average value of $y$, $k$, $l$, and $e$ in the time span, respectively.
This suggests that an averaging approach such as
%
\begin{equation}
  \lambda_{1,n}
    \approx \frac{\sum_{i=2}^{n} \lambda_{i,i-1}}{n-1}
\end{equation}
%
is a better approximation of Equation~\ref{eq:lambda1n},
which is, itself, an approximation of the true value of $\lambda$ 
given by Equation~\ref{eq:lambda}.

%------------------------------
\subsubsection{Constant Elasticity of Substitution Production Function (CES)} 
\label{sec:CES}
%------------------------------

Other energy economists use an energy-augmented 
Constant Elasticity of Substitution (CES) production function to 
describe economic growth. The \texttt{R} \citep{R} package 
\texttt{micEconCES} \citep{Henningsen:2011td} estimates CES production 
functions of the following forms (among others):
%
\ceskl{\label{eq:CES}}
%
\ceskle{\label{eq:CESkle}}
%
\noindent Equation~\ref{eq:CES} is a CES production function with 
capital stock ($k$) and labor ($l$) factors of production. 
Equation~\ref{eq:CESkle} augments Equation~\ref{eq:CES} with energy 
using a ($kl$)($e$) nesting structure, as is typical in the literature. 
Equation~\ref{eq:CES} is a degenerate form of Equation~\ref{eq:CESkle} 
where $\delta \rightarrow 1$.

$\gamma$ is a fitting parameter that represents productivity. 
The fitting parameters $\rho_1$ and $\rho$ indicate elasticities 
of substitution ($\sigma_1$ and $\sigma$).
The elasticity of substitution 
between capital ($k$) and labor ($l$) is given by 
$\sigma_1 = \frac{1}{1+\rho_1}$, and
the elasticity of substitution between ($kl$) and ($e$) is given by 
$\sigma = \frac{1}{1+\rho}$. 
As $\rho_1 \rightarrow 0$, $\sigma_1 \rightarrow 1$,
and the embedded CES production function for $k$ and $l$ degenerates 
to the Cobb-Douglas production function.
Similarly, as $\rho \rightarrow 0$, $\sigma \rightarrow 1$,
and the CES production function for $(kl)$ and $(e)$ degenerates 
to the Cobb-Douglas production function.
As $\sigma \rightarrow \infty$ ($\rho \rightarrow -1$), 
($kl$) and ($e$) are perfect substitutes. 
As $\sigma \rightarrow 0$ ($\rho \rightarrow \infty$), 
($kl$) and ($e$) are perfect complements: 
no substitution is possible. Similarly, 
as $\sigma_1 \rightarrow 0$ ($\rho_1 \rightarrow \infty$),
$k$ and $l$ are perfect complements.
$\delta_1$ describes the relative importance of capital ($k$)
and labor ($l$), and
$\delta$ describes the importance of ($kl$) relative to ($e$).

Constraints on the fitting parameters include 
$\delta_1 \in [0,1]$,
$\delta \in [0,1]$,
$\rho_1 \in [-1,0) \cup (0,\infty)$, and
$\rho \in [-1,0) \cup (0,\infty)$.

Two other nestings of the factors of production ($k$, $l$, and $e$)
are possible with the CES model.
%
\ceslek{\label{eq:CESlek}}
%
\cesekl{\label{eq:CESekl}}
%
\noindent Note that the $\rho$ ($\sigma$) and $\delta$ parameters have different
meanings that depend upon the nesting of the factors of production.

%------------------------------
\subsubsection{LINEX Production Function} 
\label{sec:LINEX}
%------------------------------

A third production function, the LINear EXponential (LINEX) function
%
\begin{equation} \label{eq:LINEX}
  y = \theta \; \mathrm{e}^{a_0\left[2 \left(1 - \frac{1}{\rho_k} \right) 
    + c_t \left(\vphantom{\frac{1}{\rho_k}}\rho_l - 1 \right)\right]} \; e \;,
\end{equation}
%
was derived from thermodynamic considerations
\citep{Kummel:1980wx,Kummel:1982vy,Kummel:1985vz,Kummel:1989tf,Kummel:2002tx,Kummel:2010vz}. 
In contrast to both the Cobb-Douglas model (Equations~\ref{eq:CD} and 
\ref{eq:CDe}) and the CES model (Equations~\ref{eq:CES}--\ref{eq:CESekl}), 
the LINEX model (Equation~\ref{eq:LINEX}) does not include a generic, 
time-dependent augmentation term ($A$).
The LINEX model assumes energy ($e$) is the only factor of production.

The LINEX model does, however, include terms that represent
efficiencies among capital, labor, and energy, in the form of 
the ratios $\rho_k$ and $\rho_l$ which are defined by\footnote{The notation
for the LINEX model breaks from the conventions used in the Cobb-Douglas and CES
models:  $a_0$ and $c_t$ are parameters and $\rho_k$ and $\rho_l$ are not.}
%
\begin{equation} \label{eq:rho_k}
  \rho_k \equiv \frac{k}{(1/2)(l+e)}
\end{equation}
%
and
%
\begin{equation} \label{eq:rho_l}
  \rho_l \equiv \frac{l}{e}
\end{equation}
%
and represent 
%
(\ref{eq:rho_k}) capital stock increase relative to the average of labor and energy (capital deepening) and 
(\ref{eq:rho_l}) labor increase relative to energy.
The effect of energy ($e$) on output ($y$) is enhanced if either
%
\begin{itemize}
\item{capital stock ($k$) has increased more than the average 
      of labor ($l$) and energy ($e$), i.e. if $\rho_k >$ 1, or}
\item{labor ($l$) has increased more than energy ($e$), i.e. if $\rho_l >$ 1.}
\end{itemize}
%
An economy that increases 
capital stock ($k$) and labor ($l$) at the same rate as energy will have
capital and labor efficiency ratios ($\rho_k$ and $\rho_l$, respectively) of unity.
In that case, economic output ($y$) increases at the same rate as energy consumption ($e$).
According to the LINEX model, an economy that increases 
capital stock ($k$) without a commensurate increase in labor ($l$) 
and energy ($e$) will experience an increase in output ($y$) in excess 
of its increase of energy consumption ($e$), because
\(
\mathrm{e}^{a_0\left[2 \left(1 - \frac{1}{\rho_k} \right) 
    + c_t \left(\vphantom{\frac{1}{\rho_k}}\rho_l - 1 \right)\right]}
	> 1.
\) 
Similarly, an economy benefits by increasing labor ($l$) 
without a commensurate increase in energy ($e$). 
Thus, $\rho_k$ is an 
efficiency of using additional labor and energy to make 
additional capital stock available to the economy, 
and $\rho_l$ is an efficiency of using additional energy to make 
additional labor available to the economy. 
Increases in $\rho_k$ and $\rho_l$ provide upward pressure on economic output ($y$),
as the only factor of production ($e$) is used more efficiently.

The parameter $a_0$ represents the overall importance of efficiencies $\rho_k$ and $\rho_l$, 
and the parameter $c_t$ represents the relative importance 
of the labor efficiency term ($\rho_l$) compared to the capital stock efficiency term ($\rho_k$). 
Both $a_0$ and $c_t$ are assumed constant with respect to time.

%++++++++++++++++++++++++++++++
\subsection{Economies} 
\label{sec:Economies}
%++++++++++++++++++++++++++++++

Discuss economies here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Sources of Data}
\label{sec:Sources_of_Data}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Discuss data sources here.

%++++++++++++++++++++++++++++++
\subsection{Historical Data} 
\label{sec:Historical_Data}
%++++++++++++++++++++++++++++++

Figure~\ref{fig:Historical_Data_1} shows a graph of historical data.

<<Historical_Data_1, fig.pos="H", fig.align='center', fig.width=6.6, fig.height=7, fig.cap="Historical data.">>=
mdata <- melt(historicalData, 
              measure.vars = c("iGDP", "iK", "iL", "iXp", "iXf", "iU"), 
              variable.name = "variable",
              value.name = "value")
historicalPlot(mdata, 
               mapping=aes(x=Year, y=value, group=variable, linetype=variable), 
               facet_formula = Country ~ Source, 
               line_types = c(1,5,2,6,4,3)) + 
  xlab("") + 
  ylab("Indexed (1960=1 or 1970=1)")
@

Figure~\ref{fig:Historical_Data_2} shows the same data as Figure~\ref{fig:Historical_Data_1} 
in a different form.

<<Historical_Data_2, fig.pos="H", fig.align='center', fig.width=8, fig.height=7, fig.cap="Historical data.">>=
historicalPlot(mdata, 
               mapping=aes(x=Year, y=value, group=Source, linetype=Source), 
               facet_formula = Country ~ variable, 
               line_types = c(1,5,3)) + 
  xlab("") + 
  ylab("Indexed (1960=1 or 1970=1)")
@

Figure~\ref{fig:Preeconometric_1} shows a graph of factors of production relative to GDP.

<<Preeconometric_1, fig.pos="H", fig.align='center', fig.width=8, fig.height=7, fig.cap="Pre-econometric data.">>=
mdata2 <- melt(historicalData, 
               measure.vars = c("KPerGDP", "LPerGDP", "XpPerGDP", "XfPerGDP", "UPerGDP"), 
               variable.name = "variable",
               value.name = "value")

historicalPlot(mdata2, 
               mapping=aes(x=Year, y=value, group=variable, linetype=variable), 
               facet_formula = Country ~ Source, 
               line_types = c(1,5,2,6,4)) + 
  xlab("") + 
  ylab("Indexed Econometric Ratios (1960=1 or 1970=1)")
@

Figure~\ref{fig:Preeconometric_2} shows the same data as Figure~\ref{fig:Preeconometric_1} 
in a different form.

<<Preeconometric_2, fig.pos="H", fig.align='center', fig.width=8, fig.height=7, fig.cap="Pre-econometric data.">>=
historicalPlot(mdata2, 
               mapping=aes(x=Year, y=value, group=Source, linetype=Source), 
               facet_formula = Country ~ variable, 
               line_types = c(1,3,5)) + 
  xlab("") + 
  ylab("Indexed Econometric Ratios (1960=1 or 1970=1)")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Parameter Estimation}
\label{sec:Parameter_Estimation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%++++++++++++++++++++++++++++++
\subsection{Point Estimates}
%++++++++++++++++++++++++++++++

Each of the economic growth models discussed in 
Section~\ref{sec:Prod_Func_Forms} has parameters which must be estimated (fitted)
using data for each economy (see Sections~\ref{sec:Economies} and~\ref{sec:Sources_of_Data}).
We obtained parameter estimates by applying the method of least-squares to 
log-transformed data using the \texttt{R} \citep{R} functions
\texttt{lm} (for Cobb-Douglas and LINEX models) 
and \texttt{cesEst} \citep{Henningsen:2011td} (for CES models).

Although the details vary from model to model, each of our three models has the general form
%
\begin{equation}
  y = f(t,k,l,e; \theta) + \mathrm{error} \;,
\end{equation}
%
where $\theta$ is a vector of parameters for the function $f$.
Parameter estimates are chosen to minimize
%
\begin{equation} \label{eq:sse}
  sse = \sum_{i} (y_i - f(t,k,l,e; \theta) )^2
\end{equation}
%
within constraints.
For the optimal parameter values $\hat \theta$, we define the 
fitted response by
\[
\hat y_i = f(t_i, k_i, l_i, e_i; \hat \theta)
\]
and the residuals ($r_i$)
as the difference between the observed response and the fitted response
%
\begin{equation} \label{eq:residual}
  r_i 
  = y_i - \hat y_i
  = y_i - f(t,k,l,e; \hat\theta)
\end{equation}
%
The details for each of the models are discussed in 
Sections~\ref{sec:CD_Parameter_Estimation}--\ref{sec:LINEX_Parameter_Estimation} below.

%++++++++++++++++++++++++++++++
\subsection{Resampling Methods}
\label{sec:Resampling}
%++++++++++++++++++++++++++++++

Bootstrapping is a statistical resampling technique for assigning
measures of accuracy and precision to sample estimates by measuring
properties of an estimator when sampling from a resampling
distribution.  
Resampling distributions can be formed in a number of 
ways in accordance with the type of data, experimental design,
and modeling assumptions involved.  
In each case, many new resamples 
are created, each of which is a randomized version of the original sample data to 
which the desired analysis method can be applied.  
By investigating, for example,
the variability of a parameter estimate from one resample to another,
one can learn about the precision of the estimation method.

In the context of linear models (regression), resamples are generally created
by residual resampling.  
In our case, we formed resamples by adding to the fitted 
response ($\hat{y}$) the product of a residual from the original model fit and 
random sign ($-1$ or $1$, each with probability $0.5$).  
For the case of Cobb-Douglas and LINEX models, 
this residual resampling occurs on the 
log-transformed data.

Intuitively, this method assumes that the residuals are indicative of 
the variability (from many potential sources) inherent in the data such that
it would be unsurprising if the residual from 
any particular year had been observed in a different year. 
Thus, a resampled response $\tilde Y'$ can be computed as 
%
\begin{equation} \label{eq:Y_tilde}
  \tilde Y_i' = \hat y_i \pm r_j \; ,
\end{equation}
%
where 
%
\begin{equation}
  r_i = y_i - \hat y_i \; .
\end{equation}
%
In Equation~\ref{eq:Y_tilde}, both the sign ($\pm$) 
and the index ($j$,~typically different from $i$) 
of the residual ($r$) 
are chosen at random (with replacement). 
We repeated the resampling process 1000 times for each
combination of growth model, country, and data source, 
both with and without energy as appropriate.

The coefficients from the fit to a resampled time series
(the ``resample coefficients'') will be different from the coefficients 
obtained from the fit to historical data (the ``base coefficients'').
When these resample coefficients are highly variable, 
it is an indication
that the data do not determine the parameter estimates very precisely.
Lack of precision can stem from a number of factors, most obviously
a poor model fit, low model sensitivity to one or more parameters, 
or correlations between parameter estimates.

It is important to note that large residuals can arise from either
(a) poor quality historical time series data, or 
(b) a mathematical model that does not describe the underlying phenomena well.
It is also important to note that (c) even when the residuals are small and the 
model produces fitted values that track the observed data closely, it may
yet be difficult to estimate some or all of the model parameters precisely.
The resampling method employed herein reflects all three of these potential 
sources of uncertainty in parameter estimates.

%++++++++++++++++++++++++++++++
\subsection{Cobb-Douglas Models}
\label{sec:CD_Parameter_Estimation}
%++++++++++++++++++++++++++++++

The Cobb-Douglas model without energy is given by 
Equation~\ref{eq:CD}. Equation~\ref{eq:CD} was
reparameterized as
%
\begin{equation} \label{eq:CD_reparam}
  y = \theta \mathrm{e}^{\lambda (t-t_0)} k^{\alpha} l^{1-\alpha}
\end{equation}
%
\noindent to ensure $\alpha + \beta = 1$ for constant returns to scale. 
$\theta$, $\lambda$, and $\alpha$ were estimated by the 
\texttt{R} \citep{R} function \texttt{lm} in log-transform space. 
If the estimated value for $\alpha$ was found outside the interval $[0,1]$, 
we set $\alpha$ to its boundary value and re-estimated $\lambda$. 
The value of $\beta$ was found with $\beta = 1 - \alpha$.

To estimate the parameters $\theta$, $\lambda$, $\alpha$, $\beta$, and $\gamma$ 
in the energy-augmented Cobb-Douglas model, Equation~\ref{eq:CDe} 
was reparameterized as
%
\begin{equation} \label{eq:CDe_AB_reparam}
  y = \theta \mathrm{e}^{\lambda (t-t_0)} k^{\alpha} l^{\beta} e^{1 - \alpha - \beta}
\end{equation}
%
\noindent ensuring that $\alpha + \beta + \gamma = 1$,
thereby providing constant returns to scale.
The \texttt{R} \citep{R} function \texttt{lm} was used to estimate 
values of $\lambda$, $\theta$, $\alpha$, and $\beta$ in log-transform space.
If the fitted value for $\alpha$ or $\beta$ fell outside the interval $[0,1]$,
we fit along all boundaries 
($\alpha = 1$,
$\beta = 1$,
$\gamma = 1$,
$\alpha = 0$,
$\beta = 0$, and
$\gamma = 0$)
and chose the boundary fit with
minimum $sse$ (Equation~\ref{eq:sse}) as the winner.
$\gamma$ was recovered with $\gamma = 1 - \alpha - \beta$.

Figure~\ref{fig:CD_Ternary_Scatter_Graph} shows ternary plots 
for $\alpha$, $\beta$, and $\gamma$ parameter values 
for Cobb-Douglas (without and with energy) resample models.
Each reample is shown as a gray dot on the graph. 
The fit to historical data is shown as a gray crosshairs.

Figures~\ref{fig:CD_GDP_Spaghetti_Graphs_US}--\ref{fig:CD_GDP_Spaghetti_Graphs_CN} 
compare predictions 
from the Cobb-Douglas models 
(without and with energy) to historical data.
Historical data are shown as a black line. 
The fit to historical data is shown as a white line.
The gray band encompasses \Sexpr{spaghettiGraphLevel*100}\% of the fits to resampled data.

<<CD_lambda_theta_graph, fig.pos="H", fig.align='center', fig.width=6, fig.height=5.5, fig.cap="$\\lambda$ and $\\theta$ values for Cobb-Douglas resample models.">>=
cdCoeffsData <- subset(AllCoeffs, subset = model=="cd")
standardScatterPlot(data=cdCoeffsData, mapping=aes(x=scale, y=lambda, color=Source), 
                    facet_formula=country~energy, alpha=0.05) +
    scale_x_continuous(limits=c(0.9, 1.5), breaks=c(1.0, 1.2, 1.4)) +
    scale_y_continuous(limits=c(-0.02, 0.05), breaks=c(0.0, 0.025, 0.05)) +    
    labs(x=expression(theta), y=expression(lambda)) +
    guides(colour=guide_legend(override.aes = aes(alpha=1))) #+
#     geom_density2d()
@

<<CD_Ternary_Scatter_Graph, fig.pos="H", fig.align='center', fig.width=6, fig.height=5, fig.cap="Cobb-Douglas resample model parameters.">>=
standardTriPlot(data=cdCoeffsData, 
                mapping=aes(color=Source),
                facet_formula=country~energy) +
    guides(colour=guide_legend(override.aes = aes(alpha=1)))
@

Given the estimates for $\alpha$, $\beta$, and $\gamma$ for the Cobb-Douglas model,
we can estimate values for $\lambda$ on a year-by-year basis using Equation~\ref{eq:lambdaij}.
The results are shown in the Figure~\ref{fig:CD_lambda_time_series}.

<<CD_finite_diff_calcs, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=6, fig.cap='Solow residual time series.'>>=
# Add a bogus "none" column that we can use for the no-energy case
m_histData <- historicalData
m_histData$none <- 1.0
# Melt the the historical data to put energy in one column
m_histData <- melt(select(m_histData, Year, iGDP, iK, iL, iXp, iXf, iU, none, Country, Source), 
                   measure.vars = c("none", "iXp", "iXf", "iU"),
                   variable.name = "energyType",
                   value.name = "iE")
# Calculate finite difference approximations of time derivatives for output and all factors of production
fd <- m_histData %>% group_by(Source, Country, energyType) %>% arrange(Year) %>% 
  mutate(dydt = ediff(iGDP, pad="tail") / ediff(Year, pad="tail")) %>%
  mutate(dkdt = ediff(iK, pad="tail") / ediff(Year, pad="tail")) %>%
  mutate(dldt = ediff(iL, pad="tail") / ediff(Year, pad="tail")) %>%
  mutate(dedt = ediff(iE, pad="tail") / ediff(Year, pad="tail"))

origCDKLCoeffs <- subset(AllCoeffs, method=="orig" & model=="cd")
origCDKLCoeffs <- origCDKLCoeffs %>% select(method, country, Source, nestStr, energy, lambda, alpha, beta, gamma)
# Change name of Source REXS to REXS1960 and relevel to conform to information in fdlambda.
origCDKLCoeffs$Source <- as.character(origCDKLCoeffs$Source)
origCDKLCoeffs$Source <- replace(origCDKLCoeffs$Source, which(origCDKLCoeffs$Source=="REXS"), "REXS1960")
origCDKLCoeffs$Source <- as.factor(origCDKLCoeffs$Source)
origCDKLCoeffs$Source <- relevelFactor(origCDKLCoeffs$Source, levs = c("REXS1960", "IST", "Leeds"))

# Fill new columns in the fd data frame with 
# lambda, alpha, beta, and gamma values from the parameter estimation process.
# lambda, alpha, beta, and gamma values for the orig fits are in the origCDKLCoeffs data frame.
# Using the hint at 
# http://stackoverflow.com/questions/15059076/
# r-how-to-call-apply-like-function-on-each-row-of-dataframe-with-multiple-argum
labg <- apply(fd[ , c('Country', 'Source', 'energyType')], 1, function(x){
    row <- as.data.frame(subset(origCDKLCoeffs, 
                                country==x["Country"] & Source==x["Source"] & energy==x["energyType"]))
    abg <- subset(row, select=c(lambda, alpha, beta, gamma))
    if (nrow(abg) < 1){ abg[1, ] <- NA }
    return(abg)
  })
cols <- do.call(rbind, labg)
fd <- cbind(fd, cols)

# Now calculate lambda by finite difference for every year to make a time series
fdlambda <- mutate(fd, lambda_diff = dydt/iGDP - (alpha/iK*dkdt + beta/iL*dldt + gamma/iE*dedt))

# Calculate the fraction of GDP growth in each year that is explained by the solow residual
fdlambda <- mutate(fdlambda, solowFrac = lambda_diff / (dydt/iGDP))

# Calculate average values for lambda and solowFrac
avgs <- fdlambda %>% 
  group_by(Source, Country, energyType) %>% 
  summarise(
    avg.lambda_diff = mean(lambda_diff, na.rm=TRUE),
    avg.solowFrac = mean(solowFrac, na.rm=TRUE)
    )

@

<<CD_lambda_time_series, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=6, fig.cap='Solow residual time series.'>>=
# Create a time series graph of the lambda_diff values
ggplot(data=fdlambda, aes(x=Year, y=lambda_diff, color=Source)) + 
  geom_point(size = 1.0) +
  facet_grid(Country ~ energyType) +
  labs(x = "", y = expression(lambda)) + 
#   scale_y_continuous(limits=c(-0.1, 0.1)) +
  xy_theme()
@

<<CD_solowPerc_time_series, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=6, fig.cap='Solow residual time series.'>>=
# Create a time series graph of the solowPerc values
ggplot(data=fdlambda, aes(x=Year, y=solowFrac, color=Source)) + 
  geom_point(size = 1.0) +
  facet_grid(Country ~ energyType) +
  labs(x = "", y = "Fraction of GDP growth explained by Solow residual") + 
  scale_y_continuous(limits=c(-0.5, 2)) +
  xy_theme()
@

<<CD_GDP_Spaghetti_Graphs_US, fig.pos="H", fig.align='center', fig.width=4, fig.height=3.5, fig.cap="US Cobb-Douglas fitted models.">>=
cdFittedDataUS <- subset(AllFitted, subset = model=="cd" & Country=="US")
spaghettiPlot(data=cdFittedDataUS, facet_formula=Source~energy, level=spaghettiGraphLevel) +
    labs(y="Indexed GDP (1960=1)") +
  scale_x_continuous(breaks=c(1960, 1980, 2000))
@

<<CD_GDP_Spaghetti_Graphs_UK, fig.pos="H", fig.align='center', fig.width=4, fig.height=4.5, fig.cap="UK Cobb-Douglas fitted models.">>=
cdFittedDataUK <- subset(AllFitted, subset = model=="cd" & Country=="UK")
spaghettiPlot(data=cdFittedDataUK, facet_formula=Source~energy, level=spaghettiGraphLevel) +
    labs(y="Indexed GDP (1960=1)") +
  scale_x_continuous(breaks=c(1960, 1980, 2000))
@

<<CD_GDP_Spaghetti_Graphs_JP, fig.pos="H", fig.align='center', fig.width=4, fig.height=2.5, fig.cap="JP Cobb-Douglas fitted models.">>=
cdFittedDataJP <- subset(AllFitted, subset = model=="cd" & Country=="JP")
spaghettiPlot(data=cdFittedDataJP, facet_formula=Source~energy, level=spaghettiGraphLevel) +
    labs(y="Indexed GDP (1960=1)") +
  scale_x_continuous(breaks=c(1960, 1980, 2000))
@

<<CD_GDP_Spaghetti_Graphs_AT, fig.pos="H", fig.align='center', fig.width=5, fig.height=3.5, fig.cap="AT Cobb-Douglas fitted models.">>=
cdFittedDataAT <- subset(AllFitted, subset = model=="cd" & Country=="AT")
spaghettiPlot(data=cdFittedDataAT, facet_formula=Source~energy, level=spaghettiGraphLevel) +
    labs(y="Indexed GDP (1960=1)") +
  scale_x_continuous(breaks=c(1960, 1980, 2000))
@

<<CD_GDP_Spaghetti_Graphs_PT, fig.pos="H", fig.align='center', fig.width=5, fig.height=2.5, fig.cap="PT Cobb-Douglas fitted models.">>=
cdFittedDataPT <- subset(AllFitted, subset = model=="cd" & Country=="PT")
spaghettiPlot(data=cdFittedDataPT, facet_formula=Source~energy, level=spaghettiGraphLevel) +
    labs(y="Indexed GDP (1960=1)") +
  scale_x_continuous(breaks=c(1960, 1980, 2000))
@

<<CD_GDP_Spaghetti_Graphs_CN, fig.pos="H", fig.align='center', fig.width=4, fig.height=2.5, fig.cap="CN Cobb-Douglas fitted models.">>=
cdFittedDataCN <- subset(AllFitted, subset = model=="cd" & Country=="CN")
spaghettiPlot(data=cdFittedDataCN, facet_formula=Source~energy, level=spaghettiGraphLevel) +
    labs(y="Indexed GDP (1960=1)") +
  scale_x_continuous(breaks=c(1960, 1980, 2000))
@

<<CD_Residual_Plot, fig.pos="H", fig.align='center', fig.width=6, fig.height=6, fig.cap="Cobb-Douglas Residual Plots.">>=
# Use residuals_df to make a graph.
cd_residuals_df <- subset(residuals_df, subset = modelType=="cd")
ggplot(data=cd_residuals_df, aes(x=year, y=resid, color=Source)) + 
  geom_point(size = 1.0) +
  facet_grid(Country ~ energyType) +
  labs(x = "", y = "Residuals") + 
  xy_theme()
@

%++++++++++++++++++++++++++++++
\subsection{CES Models}
\label{sec:CES_Parameter_Estimation}
%++++++++++++++++++++++++++++++

Information on CES models goes here.

The CES model without and with energy is given by 
Equations~\ref{eq:CES}--\ref{eq:CESekl}. 
The \texttt{R} \citep{R} package \texttt{micEconCES} 
\citep{Henningsen:2011td} was used to estimate parameters 
$\lambda$, $\gamma$, $\delta$, $\delta_1$, $\rho$, and $\rho_1$. 
The \texttt{cesEst} function of \texttt{micEconCES} provides 
several algorithm options for parameter estimation.
The default algorithm (Levenberg-Marquardt) does not respect
parameter constraints (see Section~\ref{sec:CES}) and, 
in our testing, nearly always violated them, often returning
negative values for elasticity of substitution~($\sigma$) parameters.
Thus, we used the two fitting algorithms available in \texttt{cesEst} that
respect coefficient constraints: \texttt{PORT} and \texttt{L-BFGS-B}.

Our CES parameter estimation algorithm starts with an eleven-value grid search
in $\rho$ and $\rho_1$ (9, 2, 1, 0.43, 0.25, 0.1, -0.1, -0.5, -0.75, -0.9, -0.99),
which corresponds to $\sigma$ and $\sigma_1$ values of
0.1, 0.33, 0.5, 0.7, 0.8, 0.9, 1.11, 2, 4, 10, and 100, respectively.
During the grid search, values of $\rho$ and $\rho_1$ are fixed,
and values of $\lambda$, $\gamma$, $\delta$, and $\delta_1$ are estimated 
by gradient search with the \texttt{PORT} and \texttt{L-BFGS-B} algorithms.
In all, 121 gradient searches in $\lambda$, $\gamma$, $\delta$, and $\delta_1$
at grid points representing all combinations of $\rho$ and $\rho_1$ are attempted.
During the grid search portion of our algorithm, starting values for the free parameters
are $\lambda$ = 0.015/year, $\delta$ = 0.5, $\delta_1$ = 0.5, 
and $\gamma$ is set to a value such that the mean of the residuals is zero
by the \texttt{cesEst} function.

Next, a gradient search (using both \texttt{PORT} and \texttt{L-BFGS-B}) 
is attempted wherein all fitting parameters 
($\lambda$, $\gamma$, $\delta$, $\delta_1$, $\rho$, and $\rho_1$)
are allowed to float. 
The start values for fitting parameters are taken from 
the grid search fit that provided the lowest $sse$ (Equation~\ref{eq:sse}).

If resampled data are being fitted, a prior fit to historical data is available.
In the final step of our algorithm, 
a gradient search (using both \texttt{PORT} and \texttt{L-BFGS-B})
uses coefficients from the best 
fit to historical data as its starting point. 
In this final gradient search, all model parameters 
($\lambda$, $\gamma$, $\delta$, $\delta_1$, $\rho$, and $\rho_1$)
are considered free parameters.

The fit with lowest $sse$ of all above trials is deemed the winner,
and its fitting parameters are used as the model.

\citet{Henningsen:2011td}, 
in their detailed analysis of \citet{Kemfert:1998vd}, found that

\begin{quote}
\ldots the Levenberg-Marquardt and the \texttt{PORT} algorithms 
are---at least in this study---most likely to find the 
coefficients that give the best fit to the model, 
where the \texttt{PORT} algorithm can be used to restrict the estimates 
to the economically meaningful region.
\end{quote}

In our testing, we found that to be mostly true. 
\texttt{PORT} nearly always provided lower
$sse$ than \texttt{L-BFGS-B}, despite the fact that 
\texttt{L-BFGS-B} often reports convergence
and \texttt{PORT} does not for the same data
(i.e., for the same $y$, $k$, $l$, and $e$ data).

Coefficients for each resampled CES model for all countries and all nests
are shown in 
Figures~\ref{fig:CES_lambda_gamma_graph_X}--\ref{fig:CES_sigma_delta_graph_X}. 
The white crosshair on each graph shows the location 
of the fit to historical data. 
In Figures~\ref{fig:CES_sigma1_delta1_graph_X} and~\ref{fig:CES_sigma_delta_graph_X},
the bottom of the vertical axis represents $\sigma = 0$.
The top of the vertical axis represents $\sigma = \infty$.

Note that as $\delta$ or $\delta_1$ approaches 0 (1) 
in Equations~\ref{eq:CES}--\ref{eq:CESekl},
the corresponding value of $\sigma$ or $\sigma_1$ becomes irrelevant, and
resample points along the left (right) edge of
Figures \ref{fig:CES_sigma1_delta1_graph_X} and \ref{fig:CES_sigma_delta_graph_X}
are equivalent.
That is, all points along the left (right) edge of
Figures \ref{fig:CES_sigma1_delta1_graph_X} and \ref{fig:CES_sigma_delta_graph_X}
have the same $sse$. 
As a consequence, 
the \texttt{cesEst} fitting algorithm has no way 
to discriminate among points along the left (right) edge of
Figures \ref{fig:CES_sigma1_delta1_graph_X} and \ref{fig:CES_sigma_delta_graph_X}. 
Thus, resample point estimates where $\delta$ or $\delta_1$ approaches 0 (1)
tend to cluster at the locations of the initial grid search in 
$\sigma$ or $\sigma_1$
(0.1, 0.33, 0.5, 0.7, 0.8, 0.9, 1.11, 2, 4, 10, and 100).

<<CES_lambda_gamma_graph_X, fig.pos="H", fig.align='center', fig.width=5, fig.height=4.5, fig.cap="$\\lambda$ and $\\gamma$ values for CES resample models with primary exergy ($x_p$).", eval=FALSE>>=
cesCoeffsDataX <- filter(AllCoeffs, model=="ces", nestStrParen %in% c("(iK + iL) + ()", 
                                                                      "(iK + iL) + (iXp)", 
                                                                      "(iL + iXp) + (iK)", 
                                                                      "(iXp + iK) + (iL)"))
standardScatterPlot(data=cesCoeffsDataX, mapping=aes(x=gamma_coef, y=lambda, color=Source), 
                    facet_formula=country~nestStrParen, alpha=0.05) +
    scale_x_continuous(breaks=c(0.9, 1.0, 1.1)) +
    labs(x=expression(gamma), y=expression(lambda)) + 
    guides(colour=guide_legend(override.aes = aes(alpha=1)))
@

<<CES_sigma1_delta1_graph_X, fig.pos="H", fig.align='center', fig.width=5, fig.height=5, fig.cap="$\\sigma_1$ and $\\delta_1$ values for CES resample models with primary exergy ($x_p$).", eval=FALSE>>=
standardScatterPlot(data=cesCoeffsDataX, mapping=aes(x=delta_1, y=sigma_1, color=Source), 
                    facet_formula=country~nestStrParen, alpha=0.05) +
    coord_trans(y="sigma") +
    scale_x_continuous(breaks=c(0, 0.5, 1.0)) +
    scale_y_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    labs(x=expression(delta[1]), y=expression(sigma[1])) + 
    guides(colour=guide_legend(override.aes = aes(alpha=1)))
@

<<CES_sigma_delta_graph_X, fig.pos="H", fig.align='center', fig.width=4.4, fig.height=5, fig.cap="$\\sigma$ and $\\delta$ values for CES resample models using exergy ($x_p$).", eval=FALSE>>=
standardScatterPlot(data=subset(cesCoeffsDataX, nestStrParen != "(iK + iL) + ()"),
                    mapping=aes(x=delta, y=sigma, color=Source), 
                    facet_formula=country~nestStrParen, alpha=0.05) +
    coord_trans(y="sigma") +
    scale_x_continuous(breaks=c(0, 0.5, 1.0)) +
    scale_y_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    labs(x=expression(delta), y=expression(sigma)) +
    guides(colour=guide_legend(override.aes = aes(alpha=1)))
@

% We can develop pseudo $\alpha$, $\beta$, and $\gamma$ parameters 
% for the CES model, based on values of $\delta$ and $\delta_1$.
% Table~\ref{tab:CES_abg_equations} shows the permuted equations for each CES nesting.
% 
% \begin{table} \caption{Equations for $\alpha$, $\beta$, and $\gamma$ for the various CES nestings.} 
% \label{tab:CES_abg_equations} 
%   \begin{center}
%     \begin{tabular}{c c c c c} 
%       Nesting    & Equation        & $\alpha$                & $\beta$                 & $\gamma$                \\
%       \hline
%       ($kl$)     & \ref{eq:CES}    & $\delta_1$              & $1 - \delta_1$          & 0                       \\
%       ($kl$)$e$  & \ref{eq:CESkle} & $\delta \delta_1$       & $\delta (1 - \delta_1)$ & $1 - \delta$            \\   
%       ($le$)$k$  & \ref{eq:CESlek} & $1 - \delta$            & $\delta \delta_1$       & $\delta (1 - \delta_1)$ \\
%       ($ek$)$l$  & \ref{eq:CESekl} & $\delta (1 - \delta_1)$ & $1 - \delta$            & $\delta \delta_1$       \\
%     \end{tabular}
%   \end{center}
% \end{table}
% 
% Figure~\ref{fig:CES_ternary_graph} shows 
% $\alpha$, $\beta$, and $\gamma$ parameters for the CES model for all nests and countries.
% Figure~\ref{fig:CES_ternary_graph} is comparable to
% the ternary graph for the Cobb-Douglas models, 
% Figure~\ref{fig:CD_Ternary_Scatter_Graph}.
% 
% <<CES_ternary_graphX, fig.pos="H", fig.align='center', fig.width=5.5, fig.height=7.5, fig.cap="$\\alpha$, $\\beta$, and $\\gamma$ values for CES resample models with exergy ($x$).", eval=FALSE>>=
% standardTriPlot(data=cesCoeffsDataX, facet_formula=country~nestStrParen)
% @
% 
% Figure~\ref{fig:CES_Spaghetti_Graph} compares CES model predictions 
% to historical data for all nestings. 
% As with the Cobb-Douglas results, 
% historical data are shown as a black line, 
% the fit to historical data is shown as a white line, and 
% the gray band encompasses \Sexpr{spaghettiGraphLevel*100}\% of the fits to resampled data.
% 
% <<CES_Spaghetti_Graph_X, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=8.5, fig.cap="CES fitted models with exergy ($x$).", eval=FALSE>>=
% cesFittedDataX <- filter(fitted_df, model=="ces", nestStrParen %in% c("(iK + iL) + ()", 
%                                                                       "(iK + iL) + (iX)", 
%                                                                       "(iL + iX) + (iK)", 
%                                                                       "(iX + iK) + (iL)"))
% spaghettiPlot(data=cesFittedDataX, facet_formula=Country~nestStrParen, level=spaghettiGraphLevel) +
%     labs(y="Indexed GDP (1900=1 or 1950=1)")
% @
% 
% <<CES_lambda_gamma_graph_U, fig.pos="H", fig.align='center', fig.width=5, fig.height=4.5, fig.cap="$\\lambda$ and $\\gamma$ values for CES resample models with useful work ($u$).", eval=FALSE>>=
% cesCoeffsDataU <- filter(coeffs_df, model=="ces", nestStrParen %in% c("(iK + iL) + ()", 
%                                                                       "(iK + iL) + (iU)", 
%                                                                       "(iL + iU) + (iK)", 
%                                                                       "(iU + iK) + (iL)"))
% standardScatterPlot(data=cesCoeffsDataU, mapping=aes(gamma_coef, lambda), facet_formula=country~nestStrParen, alpha=0.05) +
%     scale_x_continuous(breaks=c(0.9, 1.0, 1.1)) +
%     labs(x=expression(gamma), y=expression(lambda))
% @
% 
% <<CES_sigma1_delta1_graph_U, fig.pos="H", fig.align='center', fig.width=5, fig.height=5, fig.cap="$\\sigma_1$ and $\\delta_1$ values for CES resample models with useful work ($u$).", eval=FALSE>>=
% standardScatterPlot(data=cesCoeffsDataU, mapping=aes(delta_1, sigma_1), facet_formula=country~nestStrParen, alpha=0.05) +
%     coord_trans(y="sigma") +
%     scale_x_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
%     scale_y_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
%     labs(x=expression(delta[1]), y=expression(sigma[1]))
% @
% 
% <<CES_sigma_delta_graph_U, fig.pos="H", fig.align='center', fig.width=4.4, fig.height=5, fig.cap="$\\sigma$ and $\\delta$ values for CES resample models with useful work ($u$).", eval=FALSE>>=
% standardScatterPlot(data=subset(cesCoeffsDataU, nestStrParen != "(iK + iL) + ()"),
%                     mapping=aes(delta, sigma), facet_formula=country~nestStrParen, alpha=0.05) +
%     coord_trans(y="sigma") +
%     scale_x_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
%     scale_y_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
%     labs(x=expression(delta), y=expression(sigma))
% @
% 
% <<CES_Spaghetti_Graph_U, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=8.5, fig.cap="CES fitted models with useful work ($u$).", eval=FALSE>>=
% cesFittedDataU <- filter(fitted_df, model=="ces", nestStrParen %in% c("(iK + iL) + ()", 
%                                                                       "(iK + iL) + (iU)", 
%                                                                       "(iL + iU) + (iK)", 
%                                                                       "(iU + iK) + (iL)"))
% spaghettiPlot(data=cesFittedDataU, facet_formula=Country~nestStrParen, level=spaghettiGraphLevel) +
%     labs(y="Indexed GDP (1900=1 or 1950=1)")
% @

%++++++++++++++++++++++++++++++
\subsection{LINEX Models}
\label{sec:LINEX_Parameter_Estimation}
%++++++++++++++++++++++++++++++

The LINEX model is given in Equation~\ref{eq:LINEX}. 
If the fitting algorithm drives $a_0 \rightarrow 0$, 
any changes to $c_t$ have no effect on $sse$. 
Therefore, we transformed the LINEX model into
%
\begin{equation} \label{eq:LINEX_transformed}
  y = \theta A e; 
  A = \mathrm{e}^{\left[2 a_0 \left(1 - \frac{1}{\rho_k} \right) 
    + a_1 \left(\vphantom{\frac{1}{\rho_k}}\rho_l - 1 \right)\right]},
\end{equation}
%
\noindent where $a_1 \equiv a_0 c_t$. 
We used the \texttt{lm} function in \texttt{R}
\citep{R} to estimate least-squares values for $a_0$ and $a_1$ in 
Equation~\ref{eq:LINEX_transformed} in log-transformed space. 
We recovered a value for $c_t$ with $c_t = a_1 / a_0$. 
Figure~\ref{fig:LINEXq_Resample_Graph}
shows the coefficients estimated from resampled data.

Figures~\ref{fig:LINEX_GDP_Spaghetti_Graphs_US}--\ref{fig:LINEX_GDP_Spaghetti_Graphs_CN} 
compare LINEX model predictions to historical data.
Again, historical data are shown as a black line, 
the fit to historical data is shown as a white line and 
the gray band encompasses \Sexpr{spaghettiGraphLevel*100}\% of the fits to resampled data.

<<LINEXq_Resample_Graph, fig.pos="H", fig.align='center', fig.width=6, fig.height=6, fig.cap="LINEX model resampling results.">>=
linexCoeffsData <- subset(AllCoeffs, subset = model=="linex")
standardScatterPlot(data=linexCoeffsData,
                    mapping=aes(x=a_0, y=c_t, color=Source),
                    facet_formula=country~energy, alpha=0.05) +
    scale_x_continuous(limits=c(0.0, 1.0), breaks=c(0.0, 0.5, 1.0)) +
    scale_y_continuous(limits=c(-10.0, 10.0), breaks=c(10.0, -5.0, 0.0, 5.0, 10.0)) +
    labs(x=expression(a[0]), y=expression(c[t])) +
    guides(colour=guide_legend(override.aes = aes(alpha=1)))
@

<<LINEXq_Resample_Graph_PT, fig.pos="H", fig.align='center', fig.width=6, fig.height=2.5, fig.cap="LINEX model resampling results for PT.">>=
linexCoeffsData <- subset(AllCoeffs, subset = model=="linex" & country=="PT")
standardScatterPlot(data=linexCoeffsData,
                    mapping=aes(x=a_0, y=c_t, color=Source),
                    facet_formula=country~energy, alpha=0.05) +
#     scale_x_continuous(limits=c(0.0, 1.0), breaks=c(0.0, 0.5, 1.0)) +
#     scale_y_continuous(limits=c(-10.0, 10.0), breaks=c(10.0, -5.0, 0.0, 5.0, 10.0)) +
    labs(x=expression(a[0]), y=expression(c[t])) +
    guides(colour=guide_legend(override.aes = aes(alpha=1)))
@

<<LINEX_GDP_Spaghetti_Graphs_US, fig.pos="H", fig.align='center', fig.width=3, fig.height=2.5, fig.cap="US LINEX fitted models.">>=
linexFittedDataUS <- subset(AllFitted, subset = model=="linex" & Country=="US")
spaghettiPlot(data=linexFittedDataUS, facet_formula=Source~energy, level=spaghettiGraphLevel) +
    labs(y="Indexed GDP (1960=1)") +
  scale_x_continuous(breaks=c(1960, 1980, 2000))
@

<<LINEX_GDP_Spaghetti_Graphs_UK, fig.pos="H", fig.align='center', fig.width=3, fig.height=4.5, fig.cap="UK LINEX fitted models.">>=
linexFittedDataUK <- subset(AllFitted, subset = model=="linex" & Country=="UK")
spaghettiPlot(data=linexFittedDataUK, facet_formula=Source~energy, level=spaghettiGraphLevel) +
    labs(y="Indexed GDP (1960=1)") +
  scale_x_continuous(breaks=c(1960, 1980, 2000))
@

<<LINEX_GDP_Spaghetti_Graphs_JP, fig.pos="H", fig.align='center', fig.width=3, fig.height=2.5, fig.cap="JP LINEX fitted models.">>=
linexFittedDataJP <- subset(AllFitted, subset = model=="linex" & Country=="JP")
spaghettiPlot(data=linexFittedDataJP, facet_formula=Source~energy, level=spaghettiGraphLevel) +
    labs(y="Indexed GDP (1960=1)") +
  scale_x_continuous(breaks=c(1960, 1980, 2000))
@

<<LINEX_GDP_Spaghetti_Graphs_AT, fig.pos="H", fig.align='center', fig.width=4, fig.height=3.5, fig.cap="AT LINEX fitted models.">>=
linexFittedDataAT <- subset(AllFitted, subset = model=="linex" & Country=="AT")
spaghettiPlot(data=linexFittedDataAT, facet_formula=Source~energy, level=spaghettiGraphLevel) +
    labs(y="Indexed GDP (1960=1)") +
  scale_x_continuous(breaks=c(1960, 1980, 2000))
@

<<LINEX_GDP_Spaghetti_Graphs_PT, fig.pos="H", fig.align='center', fig.width=4, fig.height=2.5, fig.cap="PT LINEX fitted models.">>=
linexFittedDataPT <- subset(AllFitted, subset = model=="linex" & Country=="PT")
spaghettiPlot(data=linexFittedDataPT, facet_formula=Source~energy, level=spaghettiGraphLevel) +
    labs(y="Indexed GDP (1960=1)") +
  scale_x_continuous(breaks=c(1960, 1980, 2000))
@

<<LINEX_GDP_Spaghetti_Graphs_CN, fig.pos="H", fig.align='center', fig.width=4, fig.height=2.5, fig.cap="CN LINEX fitted models.">>=
linexFittedDataCN <- subset(AllFitted, subset = model=="linex" & Country=="CN")
spaghettiPlot(data=linexFittedDataCN, facet_formula=Source~energy, level=spaghettiGraphLevel) +
    labs(y="Indexed GDP (1960=1)") +
  scale_x_continuous(breaks=c(1960, 1980, 2000))
@

<<LINEX_Residual_Plot, fig.pos="H", fig.align='center', fig.width=5, fig.height=6, fig.cap="LINEX Residual Plots.">>=
# Use residuals_df to make a graph.
linex_residuals_df <- subset(residuals_df, subset = modelType=="linex")
ggplot(data=linex_residuals_df, aes(x=year, y=resid, color=Source)) + 
  geom_point(size = 1.0) +
  facet_grid(Country ~ energyType) +
  labs(x = "", y = "Residuals") + 
  scale_x_continuous(breaks=c(1960, 1980, 2000)) + 
  scale_y_continuous(limits=c(-0.3, 0.2)) +
  xy_theme()
@

The LINEX model coefficients $a_0$ and $c_t$ imply time-dependent
values of $\alpha$, $\beta$, and $\gamma$ as given by the following equations~\citep{Warr:2012cg}:

\begin{equation} \label{eq:LINEX_alpha}
  \alpha \equiv \frac{k}{y}\frac{\partial y}{\partial k} = a_0 \left( \frac{l + e}{k} \right) \;,
\end{equation}

\begin{equation} \label{eq:LINEX_beta}
  \beta \equiv \frac{l}{y} \frac{\partial y}{\partial l} = a_0 \left[ c_t \left( \frac{l}{e}\right) - \frac{l}{k} \right] \;,
\end{equation}

\noindent{}and

\begin{equation} \label{eq:LINEX_beta_2}
  \gamma \equiv \frac{e}{y} \frac{\partial y}{\partial e} 
         = 1 - a_0 \left[ \frac{e}{k} + c_t \frac{l}{e} \right] 
         = 1 - \alpha - \beta \;.
\end{equation}

\noindent Thus, each fitted model implies a trajectory of $\alpha$, $\beta$, and $\gamma$ values as time proceeds.
If the trajectory stays within the triangle, 
the constraints on $\alpha$, $\beta$, and $\gamma$ are respected.
Conversely, if the trajectory goes outside the triangle, the constraints have been violated.

Figure~\ref{fig:LINEX_abg_traj_graph_REXS_UK_U}
shows trajectories for the UK and useful work ($u$) from the REXS data set.

<<LINEX_abg_traj_graph_REXS_UK_U, fig.pos="H", fig.align='center', fig.width=3, fig.height=3.5, fig.cap="UK LINEX trajectory for REXS with $u$.">>=
triPlot(data=fortify(REXS1960_Models$REXS1960$UK$linex$`iK+iL+iU`), grid_lines=5, x=alpha, y=beta, z=gamma, geom=geom_line, aes_string="colour = iYear + 1960")
@

Figure~\ref{fig:LINEX_abg_traj_graph_IST_UK_U}
shows trajectories for the UK and useful work ($u$) from the IST data set.

<<LINEX_abg_traj_graph_IST_UK_U, fig.pos="H", fig.align='center', fig.width=3, fig.height=3.5, fig.cap="UK LINEX trajectory for IST with $u$.">>=
triPlot(data=fortify(IST_Models$IST$UK$linex$`iK+iL+iU`), grid_lines=5, x=alpha, y=beta, z=gamma, geom=geom_line, aes_string="colour = iYear + 1960")
@

Figure~\ref{fig:LINEX_abg_traj_graph_Leeds_UK_U}
shows trajectories for the UK and useful work ($u$) from the Leeds data set.

<<LINEX_abg_traj_graph_Leeds_UK_U, fig.pos="H", fig.align='center', fig.width=3, fig.height=3.5, fig.cap="UK LINEX trajectory for Leeds with $u$.">>=
triPlot(data=fortify(Leeds_Models$Leeds$UK$linex$`iK+iL+iU`), grid_lines=5, x=alpha, y=beta, z=gamma, geom=geom_line, aes_string="colour = iYear + 1960")
@

% In our testing, Iran is the only country for which the trajectory stays within 
% the constraint triangle for the entire time period.
% Thus, in our fitted models, the LINEX model is nearly always
% violating constraints on $\alpha$, $\beta$, and $\gamma$.
% If the LINEX model were made to respect the constraints,
% it would fit worse than it already does for those countries where
% the constraints are violated (that is, for every country except Iran).
% We did nothing to correct this situation, because it is not clear what should be done.
% \citet{Warr:2012cg} constrained the values of $a_0$ and $c_t$, while still holding them constant
% for the entire time series, such that the constraints on $\alpha$, $\beta$, and $\gamma$
% were respected.
% But, Warr's approach necessarily leads to larger $sse$ values than unconstrained fitting.
% Because unconstrained LINEX models already fit worse than 
% constrained Cobb-Douglas and CES models,
% we chose to report the unconstrained LINEX model fits in this paper.
% Constraining $\alpha = 0$ in the case of China would imply that either
% $a_0$ or $c_t$ or both varied with time.
% At that point, $a_0$ and $c_t$ are no longer fitted parameters of the model, 
% and we have a different model: it is no longer LINEX.

%++++++++++++++++++++++++++++++
\subsection{Parameter Estimation Summary}
\label{sec:Parameter_Estimation_Summary}
%++++++++++++++++++++++++++++++

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion of Results}
\label{sec:Discussion_of_Results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:Conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Future Work}
\label{sec:FutureWork}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}
\label{sec:Acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section*{Reproducible Research}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% In the spirit of Reproducible Research \citep{Gandrud:2013vx}, 
% all data, spreadsheets, \texttt{R} code \citep{R}, and other materials 
% associated with this paper can be found at\\*
% \protect\url{https://github.com/MatthewHeun/Econ-Growth-R-Analysis}.

%% References
%%
%% Following citation commands can be used in the body text:
%%
%%  \citet{key}  ==>>  Jones et al. (1990)
%%  \citep{key}  ==>>  (Jones et al., 1990)
%%
%% Multiple citations as normal:
%% \citep{key1,key2}         ==>> (Jones et al., 1990; Smith, 1989)
%%                            or  (Jones et al., 1990, 1991)
%%                            or  (Jones et al., 1990a,b)
%% \cite{key} is the equivalent of \citet{key} in author-year mode
%%
%% Full author lists may be forced with \citet* or \citep*, e.g.
%%   \citep*{key}            ==>> (Jones, Baker, and Williams, 1990)
%%
%% Optional notes as:
%%   \citep[chap. 2]{key}    ==>> (Jones et al., 1990, chap. 2)
%%   \citep[e.g.,][]{key}    ==>> (e.g., Jones et al., 1990)
%%   \citep[see][pg. 34]{key}==>> (see Jones et al., 1990, pg. 34)
%%  (Note: in standard LaTeX, only one note is allowed, after the ref.
%%   Here, one note is like the standard, two make pre- and post-notes.)
%%
%%   \citealt{key}          ==>> Jones et al. 1990
%%   \citealt*{key}         ==>> Jones, Baker, and Williams 1990
%%   \citealp{key}          ==>> Jones et al., 1990
%%   \citealp*{key}         ==>> Jones, Baker, and Williams, 1990
%%
%% Additional citation possibilities
%%   \citeauthor{key}       ==>> Jones et al.
%%   \citeauthor*{key}      ==>> Jones, Baker, and Williams
%%   \citeyear{key}         ==>> 1990
%%   \citeyearpar{key}      ==>> (1990)
%%   \citetext{priv. comm.} ==>> (priv. comm.)
%%   \citenum{key}          ==>> 11 [non-superscripted]
%% Note: full author lists depends on whether the bib style supports them;
%%       if not, the abbreviated list is printed even when full requested.
%%
%% For names like della Robbia at the start of a sentence, use
%%   \Citet{dRob98}         ==>> Della Robbia (1998)
%%   \Citep{dRob98}         ==>> (Della Robbia, 1998)
%%   \Citeauthor{dRob98}    ==>> Della Robbia

%% References with bibTeX database:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{References}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{model2-names}
%%\bibliography{<your-bib-database>}
\bibliography{Paper2.bib}

%% Authors are advised to submit their bibtex database files. They are
%% requested to list a bibtex style file in the manuscript if they do
%% not want to use model2-names.bst.

%% References without bibTeX database:

% \begin{thebibliography}{00}

%% \bibitem must have one of the following forms:
%%   \bibitem[Jones et al.(1990)]{key}...
%%   \bibitem[Jones et al.(1990)Jones, Baker, and Williams]{key}...
%%   \bibitem[Jones et al., 1990]{key}...
%%   \bibitem[\protect\citeauthoryear{Jones, Baker, and Williams}{Jones
%%       et al.}{1990}]{key}...
%%   \bibitem[\protect\citeauthoryear{Jones et al.}{1990}]{key}...
%%   \bibitem[\protect\astroncite{Jones et al.}{1990}]{key}...
%%   \bibitem[\protect\citename{Jones et al., }1990]{key}...
%%   \harvarditem[Jones et al.]{Jones, Baker, and Williams}{1990}{key}...
%%

% \bibitem[ ()]{}

% \end{thebibliography}

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix
%% \section{}
%% \label{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Appendix}

Include data tables here?

\end{document}

%%
%% End of file `elsarticle-template-2-harv.tex'.