### 2014-11-30 Articles/Prod Func Primer/ProdFuncPrimer.tex
Edited the Primer.

%% This is file `elsarticle-template-2-harv.tex',
%%
%% Copyright 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle  Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for Elsevier's document class `elsarticle'
%% with harvard style bibliographic references
%%
%% $Id: elsarticle-template-2-harv.tex 155 2009-10-08 05:35:05Z rishi $
%% $URL: http://lenova.river-valley.com/svn/elsbst/trunk/elsarticle-template-2-harv.tex $
%%
\documentclass[preprint,authoryear,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,authoryear,1p,times]{elsarticle}
%% \documentclass[final,authoryear,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,authoryear,3p,times]{elsarticle}
%% \documentclass[final,authoryear,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,authoryear,5p,times]{elsarticle}
%% \documentclass[final,authoryear,5p,times,twocolumn]{elsarticle}

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%% or use the graphicx package for more complicated commands
%% \usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

\usepackage{wrapfig}    % Allows wrapping of text around figures
\usepackage{float}      % Allows precise positioning of tables and figures within the text
\usepackage{amsmath}    % Allows \begin{equation*} \end{equation*} for unnumbered equations
\usepackage{multirow}   % To create multirow tables
\usepackage{hyperref}   % To create hyperlinks in the paper
\usepackage{microtype}  % Produces hanging punctuation and beautiful type
\usepackage{booktabs}   % For nice tables
\usepackage{nicefrac}   % For beautiful fractions

% From http://economics.utoronto.ca/osborne/latex/BIBTEX.HTM
\newcommand{\citeapos}[1]{\citeauthor{#1}'s (\citeyear{#1})} % Posessive citations.

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon (default)
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   authoryear - selects author-year citations (default)
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%   longnamesfirst  -  makes first citation full author list
%%
%% \biboptions{longnamesfirst,comma}

% \biboptions{}

\journal{}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{Primer on Production Functions and Economic Growth}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

% \author[BenWarr]{Benjamin Warr}
% \author[IST]{Tiago Domingos}
% \author[Leeds]{Paul Brockway}
% \author[Leeds]{Julia Steinberger}
\author[CalvinEngr]{Matthew Kuperus Heun\corref{cor1}}
\ead{mkh2@calvin.edu}

% \address[BenWarr]{Ben Warr's address, Zambia}
% \address[IST]{IST, Lisbon, Portugal}
% \address[Leeds]{Leeds University, Leeds, UK}
\cortext[cor1]{Corresponding author}
\address[CalvinEngr]{Engineering Department, Calvin College, Grand Rapids, MI 49546, USA}

\begin{abstract}
%% Text of abstract
**** Add abstract ****
\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
production function \sep economic growth \sep energy \sep Cobb-Douglas \sep CES \sep LinEx
%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
\end{keyword}

\end{frontmatter}

% \linenumbers
%% main text

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} 
\label{sec:Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This primer on production functions 
provides background on several production functions
for the purpose of accounting for the role of energy
in economic growth.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Differential Economic Growth} 
\label{sec:diff_growth}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The differential equation for economic growth~($y$) with 
capital~($k$), labor~($l$), and energy~($e$) factors of production, 
i.e., $y = y(k, l, e; t)$ is

\begin{equation} \label{eq:diff_growth}
  \frac{1}{y}\frac{\mathrm{d}y}{\mathrm{d}t} 
      = \lambda \frac{1}{t - t_0}
      + \alpha \frac{1}{k}\frac{\mathrm{d}k}{\mathrm{d}t} 
      + \beta \frac{1}{l}\frac{\mathrm{d}l}{\mathrm{d}t} 
      + \gamma \frac{1}{e}\frac{\mathrm{d}e}{\mathrm{d}t} \; ,
\end{equation}
%
where 
$y \equiv Y/Y_{0}$,
$t$ (time) is often measured in years,
$\lambda$ represents the time dependence of the change of $y$,
$k \equiv K/K_0$, 
$l \equiv L/L_0$, 
$e \equiv E/E_0$.
$Y$ (economic output) is measured by GDP in currency units, 
$K$ (capital) is expressed in currency units, 
$L$ (labor) is expressed in workers or work-hours/year, and
the subscript 0 indicates values at an initial year.
Constant returns to scale are represented by the constraint
$\alpha + \beta + \gamma = 1$.

The output elasticities $\alpha$, $\beta$, and $\gamma$ are defined as

\begin{equation} \label{eq:alaph_def}
  \alpha \equiv \frac{k}{y} \frac{\partial k / \partial t}{\partial y / \partial t} \; , \;
  \beta  \equiv \frac{l}{y} \frac{\partial k / \partial t}{\partial y / \partial t} \; , \;
  \gamma \equiv \frac{e}{y} \frac{\partial e / \partial t}{\partial y / \partial t} \; .
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Leontief Production Function} 
\label{sec:Leontief}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The Leontief production function is given by 
%
\begin{equation} \label{eq:Leontief}
  y = \min \left[ a k, b l, c e\right]
\end{equation}
%
where $a$, $b$, and $c$ are technologically-determined constants. 
The Leontief function is a recipe for production.

An example of a Leontief system is the production of automobiles.
If $Y$ is the production rate of cars (cars/year),
$W$ is wheels (wheels/year), 
$S$ is steering wheels (steering wheels/year), and
$R$ is rear-view mirrors (rear-view mirrors/year), 
a Leontief production function is
%
\begin{equation} \label{eq:cars}
  Y = \min \left[ a W, \; b S, \; c R\right] \; ,
\end{equation}
%
where $a$ = 1 car/4 wheels, $b$ = 1 car/steering wheel, and $c$ = 1 car/rear-view mirror.

The Leontief production function indicates that 
oversupplying a factor of production~($W$, $S$, or $R$) will not increase output~($Y$).
Put another way, wheels~($W$), steering wheels~($S$), and rear-view mirrors~($R$)
are not substitutable.

In the Leontief world, there is always at least one limiting factor of production.
With 8 wheels, 10 steering wheels, and 10 rear-view mirrors,
an automaker can produce only 2 cars.
Under these conditions, wheels are the limiting factor of production, 
and there will be 8 unused steering wheels and 8 unused rear-view mirrors, 
a non-optimal production system indeed.
The optimal (profit-maximizing) production strategy to make 2 cars would be to supply
8 wheels, 2 steering wheels, and 2 rear-view mirrors.

In modern economies, 
constraints exist in terms of capital~($k$), labor~($l$), and energy~($e$).~\citep{Kummel:2010vz}
%
\begin{itemize}

  \item Given a fully-utilized machine, throwing more energy at it will \emph{not} increase output~($y$).%
  \footnote{
  In fact, adding additional energy to the machine may cause damage.
  }
  Capital~($k$) constrains energy~($e$).

  \item Given a state of automation of the economy at an instant in time, 
  and assuming full utilization of the capital~($k$)
  in terms of both labor~($l$) and energy~($e$) for operation,
  adding labor to the economy doesn't increase output~($y$). 
  Capital~($k$) and energy~($e$) constrain labor~($l$).

\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Linear Production Function} 
\label{sec:linear}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The linear production function is given by
%
\begin{equation} \label{eq:linear}
  y = a k + b l + c e \; .
\end{equation}
%
In Equation~\ref{eq:linear}, $a$, $b$, and $c$ are positive constants.
In the linear world, factors of production are perfect substitutes.

Continuing the cooking analogy, consider producing toast ($Y$) 
with a fatty substance~(butter, $B$, or margarine, $M$).
%
\begin{equation} \label{eq:toast}
  Y = a B + b M
\end{equation}
%
The linear model (Equation~\ref{eq:toast}) indicates that butter~($B$) and margarine~($M$) 
can be used interchangably as the fatty substance.
Butter~($B$) and margarine~($M$) are perfect substitutes.%
  \footnote{
  Butter and margarine are perfect substitutes for all but the most discerning toast consumers.
  }%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Cobb-Douglas Production Function} 
\label{sec:CD}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The Cobb-Douglas production function is given by
%
\begin{equation} \label{eq:CD}
  y = A k^\alpha l^\beta e^\gamma \; ; \; A \equiv \mathrm{e}^{\lambda(t-t_0)} \; .
\end{equation}
%
$A$ is known as total factor productivity,
and $\lambda$ is the Solow residual.
Constant returns to scale is represented by $\alpha + \beta + \gamma = 1$.

%++++++++++++++++++++++++++++++
\subsection{Cobb-Douglas and substutitability} 
\label{sec:CD-substitution}
%++++++++++++++++++++++++++++++

In the Cobb-Douglas world, factors of production are substitutable, 
eg., a reduction of labor~($l$) can be compensated by an increase of capital~($k$)
without affecting output~($y$).%
  \footnote{
  In fact, the capital-for-labor substitution has a name: automation.
  }
Furthermore, an increase in one factor of production (say, $k$)
without a concomitant increase in the others ($l$ or $e$) 
will always result in an increase of output ($y$), provided that the output elasticity 
($\alpha$) is non-zero.

%++++++++++++++++++++++++++++++
\subsection{The cost share theorem} 
\label{sec:CD-cost-share}
%++++++++++++++++++++++++++++++

The so-called ``cost share theorem'' says that coefficients ($\alpha$, $\beta$, and $\gamma$) 
in the Cobb-Douglas model are equal to cost shares in the economy.~\citep[p.~150]{Kummel:2010vz}
\citet[Eqn.~12]{Kummel:2010vz} shows (through the use of Lagrange multipliers) 
that the ``cost share theorem'' is strictly true for a profit maximizing economy
if and only if there are \emph{no technological constraints} on economic production.

There are many technological constraints that affect economic production, such as
the Second Law of Thermodynamics and
the productive capability and energy efficiency 
of capital~($k$) at any instant in time.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Constant Elasticity of Substitution Production Function (CES)} 
\label{sec:CES}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The Constant Elasticity of Substitution production function
provides smooth transition between perfectly complementary (Leontief) 
and perfectly substitutable (linear) factors of production.

%++++++++++++++++++++++++++++++
\subsection{CES with two factors of production} 
\label{sec:CES-2}
%++++++++++++++++++++++++++++++

With two factors of production (capital, $k$ and labor, $l$), 
the CES production function is given by
%
\begin{equation} \label{eq:CES}
  y = A \: \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{-1/\rho_1} \; ; \;
      A = \mathrm{e}^{\lambda (t-t_0)}   \; .
\end{equation}
%
In Equation~\ref{eq:CES}, capital~($k$) and labor~($l$) are said to be ``nested''
together in the production function.

$\delta_1$ describes the importance of capital ($k$)
relative to labor ($l$).
As $\delta_1 \rightarrow 1$, only capital~($k$) is necessary for production.
As $\delta_1 \rightarrow 0$, only labor~($l$) is necessary for production.

$\rho_1$ indicates the elasticity 
of substitution~($\sigma_1$) between capital~($k$) and labor~($l$).
The elasticity of substitution~($\sigma_1$)
between capital~($k$) and labor~($l$) is related to $\rho_1$ and given by 
%
\begin{equation} \label{eq:sigma_1}
 \sigma_1 = \frac{1}{1+\rho_1} \; .
\end{equation}
%
Solving for $\rho_1$ gives
%
\begin{equation} \label{eq:rho_1}
  \rho_1 = \frac{1 - \sigma_1}{\sigma_1} \; .
\end{equation}

Constraints on the parameters include 
$\delta_1 \in [0,1]$ and
$\rho_1 \in (-1,0) \cup (0,\infty)$.

%++++++++++++++++++++++++++++++
\subsection{CES with three factors of production} 
\label{sec:CES-3}
%++++++++++++++++++++++++++++++

A doubly-nested form of the CES production function is required
if a third factor of production~($e$) is to be included.
%
\begin{equation} \label{eq:CES-kle}
  y = A \: \left\{\delta \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) e^{-\rho} \right\}^{-1/\rho} \; ; \; 
      A = \mathrm{e}^{\lambda (t-t_0)}
\end{equation}
%
In Equation~\ref{eq:CES-kle}, $\delta$ 
describes the importance of the capital-labor ($kl$) nesting 
relative to energy~($e$).

Important questions arise: why nest capital~($k$) and labor~($l$) together 
and separate energy~($e$)
in the~($kl$)($e$) nest structure?
Why not nest, e.g., labor~($l$) and energy~($e$) in a~($le$)($k$) structure?%
  \footnote{
  Note that the $\rho$ (and, therefore, $\sigma$) and $\delta$ parameters have different
  meanings depending upon how the factors of production are nested.
  }%
%
\begin{equation} \label{eq:CES-lek}
  y = A \: \left\{\delta \left[\delta_1 l^{-\rho_1} 
      + (1-\delta_1) e^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) k^{-\rho} \right\}^{-1/\rho} \; ; \; 
      A = \mathrm{e}^{\lambda (t-t_0)}
\end{equation}
%
Or nest energy and capital in a~($ek$)($l$) structure?
%
\begin{equation} \label{eq:CES-ekl}
  y = A \: \left\{\delta \left[\delta_1 e^{-\rho_1} 
      + (1-\delta_1) k^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) l^{-\rho} \right\}^{-1/\rho} \; ; \; 
      A = \mathrm{e}^{\lambda (t-t_0)}
\end{equation}

%++++++++++++++++++++++++++++++
\subsection{CES assumptions} 
\label{sec:CES_assumptions}
%++++++++++++++++++++++++++++++

The CES production function was derived by assuming profit maximization
and perfect competition in the markets for both output and 
factors of production.~\citep[p.~228]{Arrow:1961aa}
In particular the assumption that 
``the real wage equals the marginal product of labor''~\citep[p.~6]{Klump:2011aa}
is made.

As I have written before, these assumptions become problematic
when considering energy.
Many countries have parastatal or full government control 
over electricity generation and/or fuel markets.  
Many countries subsidize energy markets.  
And, there are always health and environmental externalities 
that are not priced into fossil fuel energy markets.  
So, the conditions for perfect competition are not present for energy markets.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Relating CES to Other Production Functions} 
\label{sec:CES-others}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Equation~\ref{eq:CES} simplifies to the other production functions 
with different values of $\sigma$.
For the two-factor CES production function, 
Table~\ref{tab:ces_simplifications} describes the options.

\begin{table}
\caption[CES Simplifications]{CES Simplifications.}
\begin{center}
 \begin{tabular}{ c c c c }
\toprule 
$\sigma$ & $\rho$ & CES simplifies to \dots & Factors of production are \dots \\
\midrule
0        & $\infty$ & Leontief              & perfect complements             \\
% \midrule
1        & 0        & Cobb-Douglas          & substitutable                   \\
% \midrule
$\infty$ & $-1$     & linear                & perfect substitutes             \\
\bottomrule
\end{tabular}
\end{center}
\label{tab:ces_simplifications}
\end{table}

Figure~\ref{fig:CES_isoquants} shows the simplifications from Table~\ref{tab:ces_simplifications}
graphically.
%
\begin{figure}[!ht]
\centering\
\includegraphics[width=0.8\linewidth]{figure/CESIsoquants.png}
\caption[CES isoquants]{CES isoquants (lines of constant output, $Y$). 
$K$ is capital and $N$ is labor. From \citet{Klump:2011aa}.}
\label{fig:CES_isoquants}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{LinEx Production Function} 
\label{sec:linex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

An additional production function, the Linear Exponential (LinEx%
    \footnote{
    The name ``LinEx'' comes from the form of the equation:
    production is \emph{linear} in energy~($e$) and \emph{exponential}
    in ratios of the factors of production~($k$, $l$, and $e$).
    }%
) function is given by K\"{u}mmel in several places~\citep{Kummel:1980wx,Kummel:1982vy,
Kummel:1985vz,Kummel:1989tf,Kummel:2002tx,Kummel:2010vz} as
%
\begin{equation} \label{eq:LinEx-kummel}
  y = e \exp{\left[ a \left( 2 - \frac{l + e}{k} \right)  + a c \left( \frac{l}{e} - 1 \right)\right]} \; .
\end{equation}
%
An alternative presentation that affords easier interpretation is
%
\begin{equation} \label{eq:LinEx}
  y = A e \; ; \; A \equiv \mathrm{e}^{a\left[2 \left(1 - \frac{1}{\rho_k} \right) 
    + c \left(\vphantom{\frac{1}{\rho_k}}\rho_l - 1 \right)\right]} \; .
\end{equation}
%
The LinEx function was derived from thermodynamic considerations. 
A key feature of the derivation involves
constraints among capital~($k$), labor~($l$), and energy~($e$).
These constraints were discussed in Section~\ref{sec:Leontief} above.%
  \footnote{
  As an aside, I question whether the constraints 
  described in Section~\ref{sec:Leontief} are the right ones.
  }%

In contrast to both the Cobb-Douglas model (Equation~\ref{eq:CD}) 
and the CES model~(Section~\ref{sec:CES}), 
the LiNex model (Equation~\ref{eq:LinEx}) does not include a generic, 
time-dependent augmentation term~($A$).
The LinEx model assumes energy~($e$) is the only factor of production.

The LinEx model does, however, include terms that represent
efficiencies among capital, labor, and energy, in the form of 
the ratios $\rho_k$ and $\rho_l$ which are defined by
%
\begin{equation} \label{eq:rho_k}
  \rho_k \equiv \frac{k}{(\nicefrac{1}{2})(l+e)}
\end{equation}
%
and
%
\begin{equation} \label{eq:rho_l}
  \rho_l \equiv \frac{l}{e}
\end{equation}
%
and represent 
%
(\ref{eq:rho_k}) capital stock increase relative to the average of labor and energy (capital deepening) and 
(\ref{eq:rho_l}) labor increase relative to energy.
The effect of energy ($e$) on output ($y$) is enhanced when $A > 1$,
which occurs if either
%
\begin{itemize}
\item{capital stock ($k$) has increased more than the average 
      of labor ($l$) and energy ($e$), i.e. if $\rho_k >$ 1 or}
\item{labor ($l$) has increased more than energy ($e$), i.e. if $\rho_l >$ 1.}
\end{itemize}
%
An economy that increases 
capital stock ($k$) and labor ($l$) at the same rate as energy ($e$) will have
capital and labor efficiency ratios ($\rho_k$ and $\rho_l$, respectively) of unity.
In that case, economic output ($y$) increases at the same rate as energy consumption ($e$).
According to the LiNex model, an economy that increases 
capital stock ($k$) without a commensurate increase in the average of labor
and energy [$(\nicefrac{1}{2})(l+e)$] will experience an increase in output ($y$) in excess 
of its increase of energy consumption ($e$), because $\rho_k > 1$ and
%
\begin{equation}
    \mathrm{e}^{a \left[2 \left(1 - \frac{1}{\rho_k} \right) 
    + c \left(\vphantom{\frac{1}{\rho_k}}\rho_l - 1 \right)\right]}
	> 1 \; .
\end{equation}
%
Similarly, an economy benefits by increasing labor ($l$) 
without a commensurate increase in energy ($e$), because $\rho_l > 1$. 
Thus, $\rho_k$ is an 
efficiency of using additional labor~($l$) and energy~($e$) to make 
additional capital stock~($k$) available to the economy, 
and $\rho_l$ is an efficiency of using additional energy~($e$) to make 
additional labor~($l$) available to the economy. 
Increases in $\rho_k$ and $\rho_l$ provide upward pressure on economic output~($y$),
as the only factor of production~($e$) is used more efficiently.

The parameter $a$ represents the overall importance of efficiencies $\rho_k$ and $\rho_l$, 
and the parameter $c$ represents the relative importance 
of labor efficiency~($\rho_l$) compared to capital stock efficiency~($\rho_k$).

\citet{Kummel:2010vz} use primary exergy for $e$ and find that 
both $a$ and $c$ vary with time.
\citet{Warr:2012cg} use useful work for $e$ and find that 
constant values of $a$ and $c$ are sufficient to describe economic growth 
over the long run.

For the Leontief, linear, Cobb-Douglas, and CES production functions, 
output elasticities are constant with respect to time.
However, for the LinEx function, 
the values of coefficients $a$ and $c$ imply time-dependent
values of $\alpha$, $\beta$, and $\gamma$ as given by the following equations~\citep{Warr:2012cg}:
%
\begin{equation} \label{eq:LINEX_alpha}
  \alpha \equiv \frac{k}{y}\frac{\partial y}{\partial k} = a \left( \frac{l + e}{k} \right) \;,
\end{equation}

\begin{equation} \label{eq:LINEX_beta}
  \beta \equiv \frac{l}{y} \frac{\partial y}{\partial l} 
  = a \left[ c \left( \frac{l}{e}\right) - \frac{l}{k} \right] \;,
\end{equation}
%
and
%
\begin{equation} \label{eq:LINEX_beta_2}
  \gamma \equiv \frac{e}{y} \frac{\partial y}{\partial e} 
         = 1 - a \left[ \frac{e}{k} + c \frac{l}{e} \right] 
         = 1 - \alpha - \beta \;.
\end{equation}
%
The LinEx model assumes constant returns to scale.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:Conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Common forms of macroeconomic production functions have been reviewed.
Relationships among production functions have been examined.
Key assumptions for each have been discussed.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Future Work}
\label{sec:FutureWork}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

After reviewing this literature, I think there is a need
for a book entitled \emph{Energy in the Production Function}
which would start with Equation~\ref{eq:diff_growth}
and carefully derive each of the production functions.
In each derivation, assumptions should be clearly exposed.
The constraints associated with the LiNex equation should be critically
examined.
New constraints should be introduced to round out the picture of 
economic growth in an energy- and materials-constrained world.
If possible, the price elasticity of supply, especially for energy, should be 
included in the production function.
I'm unsure how this should be done, 
although a profitable route might be found by
re-thinking the marginal product of energy.
Finally, a new form of the production function should be derived
and applied.

This new piece should be an entry point for people interested 
in the role of energy in macroeconomic growth
who have a differential equations background.
\emph{Lecture Notes in Energy} might be a good home for it.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section*{Acknowledgements}
% \label{sec:Acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% References
%%
%% Following citation commands can be used in the body text:
%%
%%  \citet{key}  ==>>  Jones et al. (1990)
%%  \citep{key}  ==>>  (Jones et al., 1990)
%%
%% Multiple citations as normal:
%% \citep{key1,key2}         ==>> (Jones et al., 1990; Smith, 1989)
%%                            or  (Jones et al., 1990, 1991)
%%                            or  (Jones et al., 1990a,b)
%% \cite{key} is the equivalent of \citet{key} in author-year mode
%%
%% Full author lists may be forced with \citet* or \citep*, e.g.
%%   \citep*{key}            ==>> (Jones, Baker, and Williams, 1990)
%%
%% Optional notes as:
%%   \citep[chap. 2]{key}    ==>> (Jones et al., 1990, chap. 2)
%%   \citep[e.g.,][]{key}    ==>> (e.g., Jones et al., 1990)
%%   \citep[see][pg. 34]{key}==>> (see Jones et al., 1990, pg. 34)
%%  (Note: in standard LaTeX, only one note is allowed, after the ref.
%%   Here, one note is like the standard, two make pre- and post-notes.)
%%
%%   \citealt{key}          ==>> Jones et al. 1990
%%   \citealt*{key}         ==>> Jones, Baker, and Williams 1990
%%   \citealp{key}          ==>> Jones et al., 1990
%%   \citealp*{key}         ==>> Jones, Baker, and Williams, 1990
%%
%% Additional citation possibilities
%%   \citeauthor{key}       ==>> Jones et al.
%%   \citeauthor*{key}      ==>> Jones, Baker, and Williams
%%   \citeyear{key}         ==>> 1990
%%   \citeyearpar{key}      ==>> (1990)
%%   \citetext{priv. comm.} ==>> (priv. comm.)
%%   \citenum{key}          ==>> 11 [non-superscripted]
%% Note: full author lists depends on whether the bib style supports them;
%%       if not, the abbreviated list is printed even when full requested.
%%
%% For names like della Robbia at the start of a sentence, use
%%   \Citet{dRob98}         ==>> Della Robbia (1998)
%%   \Citep{dRob98}         ==>> (Della Robbia, 1998)
%%   \Citeauthor{dRob98}    ==>> Della Robbia

%% References with bibTeX database:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{References}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{model2-names}
%%\bibliography{<your-bib-database>}
\bibliography{ProdFuncPrimer.bib}

%% Authors are advised to submit their bibtex database files. They are
%% requested to list a bibtex style file in the manuscript if they do
%% not want to use model2-names.bst.

%% References without bibTeX database:

% \begin{thebibliography}{00}

%% \bibitem must have one of the following forms:
%%   \bibitem[Jones et al.(1990)]{key}...
%%   \bibitem[Jones et al.(1990)Jones, Baker, and Williams]{key}...
%%   \bibitem[Jones et al., 1990]{key}...
%%   \bibitem[\protect\citeauthoryear{Jones, Baker, and Williams}{Jones
%%       et al.}{1990}]{key}...
%%   \bibitem[\protect\citeauthoryear{Jones et al.}{1990}]{key}...
%%   \bibitem[\protect\astroncite{Jones et al.}{1990}]{key}...
%%   \bibitem[\protect\citename{Jones et al., }1990]{key}...
%%   \harvarditem[Jones et al.]{Jones, Baker, and Williams}{1990}{key}...
%%

% \bibitem[ ()]{}

% \end{thebibliography}

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix
%% \section{}
%% \label{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \section{Appendix}
%
% Include data tables here?

\end{document}

%%
%% End of file `elsarticle-template-2-harv.tex'.