### 2014-12-01 Workflow/Workflow.Rmd
DownloadPostProcessed.R now exists. It accepts a comma-separated list of <Source>s.

---
title: "Workflow"
author: "Matthew Kuperus Heun"
date: "6 August 2014"
output: html_document
---

This file contains information about the workflow for resampling analyses of a data source.

### Add new data
* Add a file that contains the historical data from `<Source>` (name should be `<Source>.txt`) 
in the `data` directory at the top level of the repository.   
* Add any Excel workbooks or other metadata in `data/Excel Workbooks/<Source>`.
* Add any new countries to the `countryAbbrevs` variable in `Packages/EconData/R/Constants.R`.

### Create historical data frame
* Run `Scripts/PreProcess.R -S <Source>` to create the historical data frame for `<Source>`.
The file will be stored at `Packages/EconData/data/<Source>.rda`.
This command is fast; it can be run on a local machine.    
* Build and reload the `EconData` package on a local machine.    
* Push the changes to GitHub.   
* On `acc`, pull from GitHub.     
* Install the `EconData` package from GitHub to dahl
    + `ssh` to node-00 on dahl. (For some reason, this doesn't work from `fs1` or any other node.)   
    + Open an R shell by saying `R` at the command prompt.    
    + Say `require(devtools)`.
    + Say `install_github("MatthewHeun/Econ-Growth-R-Analysis/Packages/EconData")`.

`install_github` makes the historical data available to the nodes of `dahl`.
This command doesn't work from `fs1` or any of the other nodes of `dahl` (`node-01` through `node-42`).

### Create the batch script
* Run the script `r-versions` on `dahl` to see which nodes are operational and which nodes have a working version of `R`. (This script is located at `/usr/local/bin/r-versions`.)
* Create the file `Scripts/<Source>_batch.bash` file.
    + It is probably best to modify an existing `<Source>_batch.bash` file.      
    + Be sure to edit the `$SRC` variable.
    + Distribute the work over the nodes in a way that makes sense.
    + Avoid non-working nodes or nodes that have an out-of-date version of `R` as revealed by `r-versions`.
* Push the `<Source>_batch.bash` file to GitHub.
* On `acc`, pull from GitHub.

### Submit the job to dahl
* `cd` to the `Econ-Growth-R-Analysis` directory from `fs1.calvin.edu`.
* Run `Scripts/<Source>_batch.bash "-n 2 -C -d"` followed by `more data_resample/<SOURCE>/node*.txt` to obtain a printout of what will be executed.    
* Run `Scripts/<Source>_batch.bash "-n 2 -C"` to test that things work as expected.
* Then do `Scripts/<Source>_batch.bash "-n 1000 -C"` to do a complete run.
* Check `dahl`'s status at (http://fs1.calvin.edu/ganglia/?m=load_one&r=hour&s=by%2520name&c=dahl&h=&sh=1&hc=4&z=small).
* Check progress with `more data_resample/<Source>/node*.txt`.

### Post-process the results
* Run `Scripts/PostProcess.R -S <Source>` on a node of `dahl` to create:       
    + `data_postprocessed/<Source>_Coeffs.rda` and
    + `data_postprocessed/<Source>_Fitted.rda`.  
* `PostProcess.R` will accept multiple `<Source>`s. E.g., `Scripts/PostProcess.R -S IST,Leeds,SUN,REXS1960`
* Note that this command must be issued from a node of the cluster (e.g., `node-00`), not from `fs1.calvin.edu`.
`fs1` does not have a full install of `R`.

### Download post-processed results to local machine
* From the `Econ-Growth-R-Analysis` directory of a local machine, 
say `Scripts/DownloadPostProcessed.R -u <user> -S <Source> -d`, where `<user>` is 
a username on `dahl`, `<Source>` is a comma-separated list of data sources of interest, and `-d` (debug) will 
show the files to be transferred to the local machine, without actually transferring them.
* Verify that the list of files to be transferred looks right.
* From a local computer, say `Scripts/DownloadPostProcessed.R -u <user> -S <Source>` to download the files.
* The files are downloaded to the `data_postprocessed` directory on the local machine.
* Use the function `loadPostProcessedData` in the `EconData` package to load the downloaded data into
the environment. 
* The `DownloadPostProcessed.R` script connects to `dahl.calvin.edu` (`node-00`). 
The choice to connect to `node-00` was made to minimize the risk that a file won't propagate across the cluster's 
file system prior to the download request.

The data files are saved with `saveRDS` by `PostProcess.R`, so `loadPostProcessedData` uses `readRDS` to 
load the data into memory. 
As a consequence, `loadPostProcessedData` returns an object which must be assigned to a variable in `R`.