### 2015-06-03 Articles/Presentation_ESEE_2015/ESEE_2015.Rnw
Rearranged much of the analysis.

% This article has been prepared for publication in RStudio with knitr.
% According to http://www.elsevier.com/author-schemas/the-elsarticle-latex-document-class, we should be using the
% elsarticle.cls file.
% According to http://cdn.elsevier.com/assets/pdf_file/0006/109392/journal_refstyles.pdf, we should be using
% elsarticle-template-2-harv.tex as the template for the text.
% Furthermore, we should be using model2-names.bst for the bibliographic references.
% The approach here is to load the frontmatter and backmatter from elsarticle-template-2-harv.tex
% both ahead of and behind the text for our paper.
% -- Matthew Kuperus Heun, 2014-08-01

%% This is file `elsarticle-template-2-harv.tex',
%%
%% Copyright 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle  Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for Elsevier's document class `elsarticle'
%% with harvard style bibliographic references
%%
%% $Id: elsarticle-template-2-harv.tex 155 2009-10-08 05:35:05Z rishi $
%% $URL: http://lenova.river-valley.com/svn/elsbst/trunk/elsarticle-template-2-harv.tex $
%%
\documentclass[preprint,authoryear,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,authoryear,1p,times]{elsarticle}
%% \documentclass[final,authoryear,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,authoryear,3p,times]{elsarticle}
%% \documentclass[final,authoryear,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,authoryear,5p,times]{elsarticle}
%% \documentclass[final,authoryear,5p,times,twocolumn]{elsarticle}

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%% or use the graphicx package for more complicated commands
%% \usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

\usepackage{wrapfig}    % Allows wrapping of text around figures
\usepackage{soul}       % Provides strikethrough text
\usepackage{float}      % Allows precise positioning of tables and figures within the text
\usepackage{amsmath}    % Allows \begin{equation*} \end{equation*} for unnumbered equations
\usepackage{multirow}   % To create multirow tables
\usepackage{hyperref}   % To create hyperlinks in the paper
\usepackage{microtype}  % produces hanging punctuation and beautiful type
\usepackage{booktabs}   % for beautiful tables

% From http://economics.utoronto.ca/osborne/latex/BIBTEX.HTM
\newcommand{\citeapos}[1]{\citeauthor{#1}'s (\citeyear{#1})} % Posessive citations.

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon (default)
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   authoryear - selects author-year citations (default)
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%   longnamesfirst  -  makes first citation full author list
%%
%% \biboptions{longnamesfirst,comma}

% \biboptions{}

\journal{Ecological Economics}

\begin{document}

<<setup, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE>>=
require(EconData)
require(EconModels)
require(dplyr)
require(reshape2)
require(ggplot2)
require(mosaic)
require(xtable)

# Tell whether to re-run the analysis to create the models.
# This file will use previously-saved models to generate
# the paper if set to FALSE, greatly reducing time.
# If something changes in the data or in the code that
# performs the analysis, set 
# rerunModels = TRUE
# once to re-generate the information needed to 
# create the paper.
# ********************************************************
rerunModels=FALSE
# ********************************************************

opts_chunk$set(
  # ********************************************************
  eval=TRUE,
  # ********************************************************
  # tikz allows LaTeX code in graphical output. 
  # E.g., "$y$" for a variable name in the legend for a graph.  
  # dev='tikz',
  tidy=FALSE,
  comment=NA,
  # Tells whether to cache output from chunks are saved. Cacheing saves time. 
  # However, references to LaTeX tables DO NOT WORK if cacheing is turned on.
  cache=FALSE,
  warning=FALSE,  #Tells whether to show warnings in the output.
  # Turns off messages for all chunks. 
  # Set TRUE on an individual chunk to see it.
  message=FALSE, 
  # Tells whether to echo code for all chunks. 
  # Set TRUE on an individual chunk to see its code.
  echo=FALSE,
  # Sets all table captions to the top.
  xtable.caption.placement="top"
)
@

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{Effect of quality-adjusted production function inputs}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

\author[IST]{Jo\~{a}o dos Santos}
\author[Leeds]{Paul Brockway}
\author[CalvinEngr]{Matthew Kuperus Heun\corref{cor1}}
\ead{mkh2@calvin.edu}
\author[Leeds]{Marco Sekai}
\author[IST]{Tiago Domingos}
\author[Leeds]{Julia Steinberger}

\address[IST]{IST, Lisbon, Portugal}
\address[Leeds]{Leeds University, Leeds, UK}
\address[CalvinEngr]{Engineering Department, Calvin College, Grand Rapids, MI 49546, USA}
\cortext[cor1]{Corresponding author}

\begin{abstract}
%% Text of abstract
**** Add abstract ****
\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
economic growth \sep exergy \sep energy \sep Cobb-Douglas \sep CES
%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
\end{keyword}

\end{frontmatter}

% \linenumbers
%% main text

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} 
\label{sec:Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This is a study of the effects of quality-adjusted production function inputs
on the importance of energy in the production function.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Coordinates of Analysis} 
\label{sec:Coordinates}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This section describes the coordinates of analysis and briefly reviews 
literature related to each.

%++++++++++++++++++++++++++++++
\subsection{Mathematical Forms of the Energy-augmented Production Function} 
\label{sec:Prod_Func_Forms}
%++++++++++++++++++++++++++++++

In this paper, we assess two prominent energy-augmented production functions 
that appear in the literature: 
Cobb-Douglas (CD) and
Constant Elasticity of Substitution (CES).
The following subsections describe each.

%------------------------------
\subsubsection{Cobb-Douglas Production Function} 
\label{sec:CDe}
%------------------------------

The Cobb-Douglas production function can be expressed as
%
\begin{equation} \label{eq:CD}
  y = \theta A k^\alpha l^\beta \; ; \; A \equiv \mathrm{e}^{\lambda(t-t_0)} \; ,
\end{equation}
%
where 
$y \equiv Y/Y_{0}$,
$\theta$ is a scale parameter,
e is the base of the natural logarithm, 
$\lambda$ is represents the pace of technological progress,
$t$ (time) is measured in years,
$k \equiv K/K_{0}$, 
$l \equiv L/L_{0}$, 
$Y$ (economic output) is represented by GDP, 
$K$ (capital) is expressed in currency units, 
$L$ (labor) is expressed in workers or work-hours/year, and
the 0 subscript indicates values at an initial year.%
  \footnote{Dimensionless, indexed quantities are represented by 
  lower-case symbols 
  ($y$, $k$, $l$, $e$, $q$, $x$, and $u$), and dimensional 
  quantities are represented by upper-case symbols 
  ($Y$, $K$, $L$, $E$, $Q$, $X$, and $U$). 
  Model parameters are represented by Greek letters
  ($\alpha$, $\beta$, $\lambda$, $\theta$).
  }
Constant returns to scale are represented by the constraint
$\alpha + \beta = 1$.

The capital-labor Cobb-Douglas production function shown in Equation~\ref{eq:CD}
can be augmented to include an energy term:
%
\begin{equation} \label{eq:CDe}
  y = \theta A k^\alpha l^\beta e^\gamma \; ; \; A \equiv \mathrm{e}^{\lambda(t-t_0)} \; ,
\end{equation}
%
where $e \equiv E/E_0$, and $E$ is in units of energy per time, typically TJ/year.
The energy-augmented Cobb-Douglas production function 
is often assumed to have constant returns to scale for the three factors 
of production: $\alpha + \beta + \gamma = 1$.
The term $A$ is known as total factor productivity,
and $\lambda$ is the Solow residual.

%------------------------------
\subsubsection{Constant Elasticity of Substitution Production Function (CES)} 
\label{sec:CES}
%------------------------------

Other energy economists use an energy-augmented 
Constant Elasticity of Substitution (CES) production function to 
describe economic growth. The \texttt{R} 
% \citep{R} 
package \texttt{micEconCES} 
% \citep{Henningsen:2011td} 
estimates CES production 
functions of the following form
%
\begin{equation} \label{eq:CESgeneric}
  y = \gamma \: A \: \left\{\delta \left[\delta_1 x_1^{-\rho_1} 
      + (1-\delta_1)x_2^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) x_3^{-\rho} \right\}^{-1/\rho}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)},
\end{equation}
%
where $x_1$, $x_2$, and $x_3$ are factors of production
and permutations of capital~($k$), labor~($l$), and energy~($e$).

The CES model without energy is given in Equation~\ref{eq:CESkl}.
%
\begin{equation} \label{eq:CESkl}
  y = \gamma \: A \: \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{-1/\rho_1}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)} . 
\end{equation}
%
Equation~\ref{eq:CESkle} augments Equation~\ref{eq:CESkl} with energy 
using a ($kl$)($e$) nesting structure, as is typical in the literature. 
Equation~\ref{eq:CESkl} is a degenerate form of Equation~\ref{eq:CESkle} 
where $\delta \rightarrow 1$. 
%
\begin{equation} \label{eq:CESkle}
  y = \gamma \: A \: \left\{\delta \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) e^{-\rho} \right\}^{-1/\rho}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)}.
\end{equation}
%

In the CES production function, 
$\gamma$ is a fitting parameter that accounts for an atypical first year. 
The fitting parameters $\rho_1$ and $\rho$ indicate elasticities 
of substitution ($\sigma_1$ and $\sigma$).
The elasticity of substitution 
between capital ($k$) and labor ($l$) is given by 
$\sigma_1 = \frac{1}{1+\rho_1}$, and
the elasticity of substitution between ($kl$) and ($e$) is given by 
$\sigma = \frac{1}{1+\rho}$. 
As $\rho_1 \rightarrow 0$, $\sigma_1 \rightarrow 1$,
and the embedded CES production function for $k$ and $l$ degenerates 
to the Cobb-Douglas production function.
Similarly, as $\rho \rightarrow 0$, $\sigma \rightarrow 1$,
and the CES production function for $(kl)$ and $(e)$ degenerates 
to the Cobb-Douglas production function.
As $\sigma \rightarrow \infty$ ($\rho \rightarrow -1$), 
($kl$) and ($e$) are perfect substitutes. 
As $\sigma \rightarrow 0$ ($\rho \rightarrow \infty$), 
($kl$) and ($e$) are perfect complements: 
no substitution is possible. Similarly, 
as $\sigma_1 \rightarrow 0$ ($\rho_1 \rightarrow \infty$),
$k$ and $l$ are perfect complements.
$\delta_1$ describes the relative importance of capital ($k$)
and labor ($l$), and
$\delta$ describes the importance of ($kl$) relative to ($e$).

Constraints on the fitting parameters include 
$\delta_1 \in [0,1]$,
$\delta \in [0,1]$,
$\rho_1 \in [-1,0) \cup (0,\infty)$, and
$\rho \in [-1,0) \cup (0,\infty)$.

Two other nestings of the factors of production ($k$, $l$, and $e$)
are possible with the CES model.
%
\begin{equation} \label{eq:CESlek}
  y = \gamma \: A \: \left\{\delta \left[\delta_1 l^{-\rho_1} 
      + (1-\delta_1) e^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) k^{-\rho} \right\}^{-1/\rho}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)}
\end{equation}
%
\begin{equation} \label{eq:CESekl}
  y = \gamma \: A \: \left\{\delta \left[\delta_1 e^{-\rho_1} 
      + (1-\delta_1) k^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) l^{-\rho} \right\}^{-1/\rho}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)}
\end{equation}
% 
Note that $\rho$ ($\sigma$), $\rho_1$ ($\sigma_1$), $\delta$, and $\delta_1$ have different
meanings depending upon the nesting of the factors of production.

%++++++++++++++++++++++++++++++
\subsection{Economies} 
\label{sec:Economies}
%++++++++++++++++++++++++++++++

Discuss economies here.

UK and Portugal.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Sources of Data}
\label{sec:Sources_of_Data}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Discuss data sources here.

%++++++++++++++++++++++++++++++
\subsection{Historical Data} 
\label{sec:Historical_Data}
%++++++++++++++++++++++++++++++

Historical data are stored in the \texttt{IST} and \texttt{Leeds} data sets.

<<Formulas>>=
# Set up the formulas for each type of analysis.
# These formulas contain energy. 
# We can use keep_summands to drop the energy term (3rd summand on the rhs)
# later if needed.
formulas <- 
  list( 
    PT = c(unadjusted = build_formula("iGDP", list("iKstkS.L", "iL", "iXpMP", "iYear")),
             adjusted = build_formula("iGDP", list("iKservS.L", "ihLest", "iUMP", "iYear"))),
    UK = c(unadjusted = build_formula("iGDP", list("iKstkO.WwithRD", "iL", "iXp", "iYear")),
             adjusted = build_formula("iGDP", list("iKservO.WwithRD", "ihLest", "iU", "iYear")))
  )
countries <- names(formulas)
flavors <- names(formulas$PT)
@

<<Data>>=
# Create the historicalData data frames
historicalData <- list()
commonVars <- c("Year", "Country", "Source")
historicalData$PT <- filter(IST, Country == "PT") 
historicalData$PT <- historicalData$PT[ , unique(c(commonVars,
                                                   all.vars(formulas$PT$unadjusted), 
                                                   all.vars(formulas$PT$adjusted)))]
historicalData$UK <- filter(Leeds, Country == "UK")
historicalData$UK <- historicalData$UK[ , unique(c(commonVars,
                                                   all.vars(formulas$UK$unadjusted), 
                                                   all.vars(formulas$UK$adjusted)))]
@

For unadjusted variables, we use the time series shown in Table~\ref{tab:unadjusted_variables}.

\begin{table} \caption{Unadjusted time-series variables.} 
\label{tab:unadjusted_variables} 
  \begin{center}
    \begin{tabular}{r l l} 
      \toprule
      Variable    & Portugal           & UK                 \\
      \midrule
      Output      & \Sexpr{summands(formulas$PT$unadjusted, n=1, right=FALSE)}   
                                                             & \Sexpr{summands(formulas$UK$unadjusted, n=1, right=FALSE)}\\
      Capital     & \Sexpr{summands(formulas$PT$unadjusted, 1)} & \Sexpr{summands(formulas$UK$unadjusted, 1)} \\
      Labor       & \Sexpr{summands(formulas$PT$unadjusted, 2)} & \Sexpr{summands(formulas$UK$unadjusted, 2)} \\
      Energy      & \Sexpr{summands(formulas$PT$unadjusted, 3)} & \Sexpr{summands(formulas$UK$unadjusted, 3)} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

For quality-adjusted variables, we use the time series shown in Table~\ref{tab:quality_adjusted_variables}.

\begin{table} \caption{Quality-adjusted time series variables.} 
\label{tab:quality_adjusted_variables} 
  \begin{center}
    \begin{tabular}{r l l} 
      \toprule
      Variable    & Portugal         & UK               \\
      \midrule
      Output      & \Sexpr{summands(formulas$PT$adjusted, n=1, right=FALSE)}   
                                                           & \Sexpr{summands(formulas$UK$adjusted, n=1, right=FALSE)} \\
      Capital     & \Sexpr{summands(formulas$PT$adjusted, 1)} & \Sexpr{summands(formulas$UK$adjusted, 1)} \\
      Labor       & \Sexpr{summands(formulas$PT$adjusted, 2)} & \Sexpr{summands(formulas$UK$adjusted, 2)} \\
      Energy      & \Sexpr{summands(formulas$PT$adjusted, 3)} & \Sexpr{summands(formulas$UK$adjusted, 3)} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

Unadjusted and quality-adjusted data for both Portugal and the United Kingdom 
are shown in Figure~\ref{fig:Historical Data Graph}.

<<Historical Data Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Historical data.", warning=FALSE>>=
# Melt the data into a form that is useful in a couple places
histDataForGraph <- leaf_apply(formulas, class = "formula", f=function(form, name){
  meta <- strsplit(name, split="[.]") %>% unlist
  Country <- meta[1]
  flavor <- meta[2]
  measure.vars <- c(summands(form, right=FALSE), summands(form, 1:3))
  trimmed_historical_data <- historicalData[[Country]][c("Year", measure.vars)]
  mtrimmed_historical_data <- melt(trimmed_historical_data, measure.vars = measure.vars, 
                                   value.name = "value", variable.name = "variable")
  mtrimmed_historical_data$Country <- Country
  mtrimmed_historical_data$flavor <- if (flavor == "adjusted") "Quality-adjusted" else "Unadjusted"
  return(mtrimmed_historical_data)
}) %>% do.call(rbind, .)

# Relevel factors before plotting
histDataForGraph$Country <- relevelFactor(as.factor(histDataForGraph$Country), c("PT", "UK"))
histDataForGraph$flavor <- relevelFactor(as.factor(histDataForGraph$flavor), c("Unadjusted", "Quality-adjusted"))
histDataForGraph$variable <- relevelFactor(as.factor(histDataForGraph$variable), unique(histDataForGraph$variable))

historicalPlot(histDataForGraph, 
               mapping=aes(x=Year, y=value, group=variable, linetype=variable), 
               facet_formula = Country ~ flavor, 
               line_types = c(1,2,4,3,2,4,3,2,3,2,3)) + 
  scale_y_continuous(breaks=c(1, 2, 4, 6, 8, 10, 12)) +
  coord_cartesian(ylim=c(0.5, 13)) +
  xlab("") + 
  ylab("Indexed (1960=1)")
@

<<Pre-econometric graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Pre-econometric data.", warning=FALSE>>=
pre_econometric_data <- leaf_apply(formulas, class = "formula", f=function(form, name){
  meta <- strsplit(name, split="[.]") %>% unlist
  Country <- meta[1]
  flavor <- meta[2]
  nums <- summands(form, n=1:3)
  denom <- summands(form, n=1, right=FALSE)
  ratios <- lapply(nums, function(num) {paste(num, denom, sep="/")}) %>% unlist
  coltemplate <- build_formula(NULL, c("Year", ratios))
  trimmed_historical_data <- historicalData[[Country]][all.vars(coltemplate)]
  ratio_data <- evalFormula(coltemplate, data=trimmed_historical_data)$right
  mratio_data <- melt(ratio_data, measure.vars = ratios, value.name = "ratio", variable.name = "variable")
  mratio_data$Country <- Country
  mratio_data$flavor <- if (flavor == "adjusted") "Quality-adjusted" else "Unadjusted"
  return(mratio_data)
}) %>% do.call(rbind, .)

# Fix levels
pre_econometric_data$flavor <- relevelFactor(as.factor(as.character(pre_econometric_data$flavor)), 
                                               c("Unadjusted", "Quality-adjusted"))
pre_econometric_data$Country <- relevelFactor(as.factor(pre_econometric_data$Country), 
                                              c("PT", "UK"))
pre_econometric_data$variable <- relevelFactor(as.factor(pre_econometric_data$variable),
                                               unique(pre_econometric_data$variable))
# Now make the plot of pre-econometric data
historicalPlot(pre_econometric_data, 
               mapping=aes(x=Year, y=ratio, group=variable, linetype=variable), 
               facet_formula = Country ~ flavor, 
               line_types = c(2,4,3, 2,4,3, 2,3,2, 3)) + 
  xlab("") + 
  ylab("Indexed (1960=1)")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Parameter Estimation}
\label{sec:parameter_estimation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Parameter estimation is accomplished with \texttt{formula}s of the form
%
\begin{equation} \label{eq:formulas}
  y \sim x_1 + x_2 + x_3 + time,
\end{equation}
%
where $y$ is economic output and 
$x_1$, $x_2$, and $x_3$ are factors of production
and permutations of
capital ($k$), labor ($l$), and energy ($e$).
(The factors of production may be quality-adjusted.)
$time$ is the time variable, usually \texttt{iYear}. 
The \texttt{formula}s are used in a fitting function such as 
\texttt{cdModel} to create a \texttt{model} object.
The function \texttt{naturalCoef} extracts reasonable
coefficients from \texttt{model} objects.

Note that in the fitting process for the Cobb-Douglas equation,
$\texttt{scale} = \theta$, 
$\alpha_1 = \alpha$,
$\alpha_2 = \beta$, and
$\alpha_3 = \gamma$,
when the \texttt{formula} is specified as 
%
\begin{equation} \label{eq:formula_kle}
  \texttt{y} \sim \texttt{k + l + e + time}.
\end{equation}
%
Note that for the CES model, $\texttt{scale} = \gamma$ and is expected to be close to unity.

For the CES model, values of $\alpha_1$, $\alpha_2$, and $\alpha_3$ are calculated by
$\alpha_1 = \delta_1$, $\alpha_2 = 1 - \delta_1$, and $\alpha_3 = 0$ 
for the CES model with two factors of production
and by 
$\alpha_1 = \delta \delta_1$, $\alpha_2 = \delta(1 - \delta_1)$, and $\alpha_3 = 1 - \delta$
for the CES model with three factors of production.
The values of $\alpha_1$, $\alpha_2$, and $\alpha_3$ are interpreted as
$\alpha$, $\beta$, and $\gamma$ (coefficients for capital, labor, and energy, respectively)
as shown in Table \ref{tab:CES_abg_equations}, 
assuming factors of production are specified in the fitting \texttt{formula} as
\texttt{y~$\sim$~k~+~l~+~e~+~time}.

\begin{table} \caption{Equations for $\alpha$, $\beta$, and $\gamma$ 
                        (importance of capital, labor, and energy, respectively)
                        for the various CES nestings, 
                        provided that the \texttt{formula} is specified as \texttt{y~$\sim$~k~+~l~+~e~+~time}.} 
\label{tab:CES_abg_equations} 
  \begin{center}
    \begin{tabular}{c c c c c} 
      \toprule
      Nesting         & \texttt{nest}     & $\alpha$                & $\beta$                 & $\gamma$       \\
      \midrule
      ($kl$) + ()     & \texttt{c(1,2)}   & $\delta_1$              & $1 - \delta_1$          & 0          \\
      ($kl$) + ($e$)  & \texttt{c(1,2,3)} & $\delta \delta_1$       & $\delta (1 - \delta_1)$ & $1 - \delta$ \\   
      ($le$) + ($k$)  & \texttt{c(2,3,1)} & $1 - \delta$            & $\delta \delta_1$       & $\delta (1 - \delta_1)$ \\
      ($ek$) + ($l$)  & \texttt{c(3,1,2)} & $\delta (1 - \delta_1)$ & $1 - \delta$            & $\delta \delta_1$ \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

<<Models>>=
fileName <- "ESEE2015Models.Rdata"
if (!rerunModels){
  # load from the file on disk
  load(fileName)
} else {
  # re-run the models and save to disk.
  # In RStudio, be sure to set the working directory to the source file directory.

  models <- list()
  for(country in countries){
    for (flavor in flavors){
      fnoE <- keep_summands(formulas[[country]][[flavor]], -3) # Drop the energy term
      fwithE <- formulas[[country]][[flavor]]
      # CD models
      models[[country]][[flavor]][["noE"]][["CD"]]   <- cdModel(fnoE, data = historicalData[[country]])
      models[[country]][[flavor]][["withE"]][["CD"]] <- cdModel(fwithE, data = historicalData[[country]])
      # CES models
      models[[country]][[flavor]][["noE"]][["CES"]][["kl"]] <- cesModel(fnoE, 
                                                                        data = historicalData[[country]], 
                                                                        nest=c(1,2))
      models[[country]][[flavor]][["withE"]][["CES"]] <- 
        lapply(nests, function(nest){
          cesModel(formula=fwithE, data=historicalData[[country]], nest=nest)
        })
    }      
  }
  # Save models to disk.
  save(models, file=fileName)
}
# Put all the coefficients into a data frame
AllCoeffs <- leaf_apply(models, class = c("CDEmodel", "cesModel", "LINEXmodel"), f = function(mod, id){
  # Parse the id to find out whether this is quality-adjusted data, etc.
  id_list <- strsplit(id, split = "[.]")[[1]]
  Country <- id_list[[1]]
  flavor <- id_list[[2]]
  energy <- id_list[[3]]
  model <- id_list[[4]]
  nest <- if(length(id_list) > 4) id_list[[5]] else NA
  res <- rbind(list(Country = Country, 
                    flavor = if (flavor=="unadjusted") "Unadjusted" else "Quality-adjusted",
                    energy = if (energy=="noE") "Without energy" else "With energy",
                    model=model, 
                    nest=nest))
  coeffs <- naturalCoef(mod)
  res <- cbind(res, coeffs)
  # Now calculate the mean squared error (mse) for each model.
  sqResids <- resid(mod)^2
  mse <- sum(sqResids)/(length(sqResids))
  res <- cbind(res, list(mse=mse))
  return(res)
}) %>% do.call(plyr::rbind.fill, .)
# Get the unadjusted, no-energy values for mse and lambda and place them on each appropriate row
AllCoeffs_unadj_noE <- filter(AllCoeffs, flavor=="Unadjusted", energy=="Without energy") %>% 
                              select(Country, flavor, energy, model, nest, mse, lambda)
colnames(AllCoeffs_unadj_noE)[colnames(AllCoeffs_unadj_noE)=="mse"] <- "mse_unadj_noE"
colnames(AllCoeffs_unadj_noE)[colnames(AllCoeffs_unadj_noE)=="lambda"] <- "lambda_unadj_noE"
AllCoeffs <- merge(AllCoeffs, AllCoeffs_unadj_noE[c("Country", "model", "mse_unadj_noE", "lambda_unadj_noE")])
# Calculate differences relative to the unadjusted, no-energy condition
AllCoeffs <- mutate(AllCoeffs, Dlambda = abs(lambda) - abs(lambda_unadj_noE), Dmse = mse - mse_unadj_noE)

# Relevel factors
AllCoeffs$Country <- relevelFactor(as.factor(unlist(AllCoeffs$Country)), c("PT", "UK"))
AllCoeffs$flavor <- relevelFactor(as.factor(unlist(AllCoeffs$flavor)), c("Unadjusted", "Quality-adjusted"))
AllCoeffs$energy <- relevelFactor(as.factor(unlist(AllCoeffs$energy)), c("Without energy", "With energy"))
AllCoeffs$model <- relevelFactor(as.factor(unlist(AllCoeffs$model)), c("CD", "CES"))
AllCoeffs$nest <- relevelFactor(as.factor(unlist(AllCoeffs$nest)), c("kl", "kle", "lek", "ekl"))
# Rearrange for easier viewing
AllCoeffs <- arrange(AllCoeffs, Country, model, flavor, energy, nest)
@

<<Fitted_and_Residuals>>=
fittedAndResidData <- function(model){
  df <- data.frame(select(model$data, Year))
  # Trim fitted to be same length as df to eliminate problems with possible NAs at end of fitted
  df$fitted <- yhat(model)[c(1:length(df$Year))]
  df$Residuals <- resid(model)[c(1:length(df$Year))]
  return(df)  
}

temp_data <- leaf_apply(models, class=c("CDEmodel", "cesModel", "LINEXmodel"), f = function(model, id){
  df <- fittedAndResidData(model)
  f <- model$formula
  response <- summands(f, right = FALSE)
  df[[response]] <- model$data[[response]] # Add iGDP to the data frame
  id_list <- strsplit(id, split = "[.]")[[1]]
  df$Country <- id_list[[1]]
  df$flavor <- if (id_list[[2]] == "unadjusted") "Unadjusted" else "Quality-adjusted"
  df$energy <- if (id_list[[3]] == "noE") "Without energy" else "With energy"
  df$model <- id_list[[4]]
  df$nest <- if (length(id_list) > 4) id_list[[5]] else NA
  return(df)  
}) %>% do.call(rbind, .)

# Get the unadjusted, without-energy residuals and store them in a data frame
r_unadj_noE <- filter(temp_data, flavor=="Unadjusted" & energy=="Without energy") %>% 
               select(Year, Country, Residuals, model)
colnames(r_unadj_noE)[colnames(r_unadj_noE)=="Residuals"] <- "r_unadj_noE"
# Merge to get the unadjusted, without-energy values next to each country, year, and model
fitted_and_resid_data <- merge(temp_data, r_unadj_noE)
# Create a column containing ∆r, the difference between a residual value
# and the residual value in the same year for the unadjusted, no-energy case
fitted_and_resid_data[["Dr"]] <- abs(fitted_and_resid_data[["Residuals"]]) - abs(fitted_and_resid_data[["r_unadj_noE"]])

# Relevel factors before plotting
fitted_and_resid_data$Country <- relevelFactor(as.factor(fitted_and_resid_data$Country), c("PT", "UK"))
fitted_and_resid_data$flavor <- relevelFactor(as.factor(fitted_and_resid_data$flavor), c("Unadjusted", "Quality-adjusted"))
fitted_and_resid_data$energy <- relevelFactor(as.factor(fitted_and_resid_data$energy), c("Without energy", "With energy"))
fitted_and_resid_data$nest <- relevelFactor(as.factor(fitted_and_resid_data$nest), c("kl", "kle", "lek", "ekl"))

# Make data viewing easier
fitted_and_resid_data <- arrange(fitted_and_resid_data, Country, model, flavor, energy, nest)
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\label{sec:results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%++++++++++++++++++++++++++++++
\subsection{Fits to historical data} 
\label{sec:fits_to_historical_data}
%++++++++++++++++++++++++++++++

<<Fitted and Residual Graph Functions>>=
fittedGraph <- function(data){
  ggplot() +
    geom_line(data=filter(data, energy=="Without energy"), mapping=aes(x=Year, y=iGDP)) +
    geom_line(data=data, mapping=aes(x=Year, y=fitted, linetype=energy)) +
    scale_linetype_manual(values=c(2,3), name="") +
    facet_grid(Country ~ flavor) +
    scale_x_continuous(breaks=c(1960, 1980, 2000)) +
    scale_y_continuous(breaks=c(1, 2, 4, 6, 8, 10, 12)) +
    xlab("") +
    ylab("Indexed GDP (1960=1)") +
    xy_theme()
}
@

Both historical GDP and fitted GDP are shown in 
Figures~\ref{fig:Graphs CD-fitted}--\ref{fig:Graphs CESekl-fitted}.
%
<<Graphs CD-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Cobb-Douglas models. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CD"))
@
%
<<Graphs CESkle-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($kl$)$e$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CES" & (nest=="kle" | nest=="kl")))
@
%
<<Graphs CESlek-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($le$)$k$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CES" & (nest=="lek" | nest=="kl")))
@
%
<<Graphs CESekl-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($ek$)$l$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CES" & (nest=="ekl" | nest=="kl")))
@

%++++++++++++++++++++++++++++++
\subsection{Fitting residuals} 
\label{sec:fitting_residuals}
%++++++++++++++++++++++++++++++

<<ResidualPlotFunction>>=
residualPlot <- function(data){
  ggplot() +
    geom_line(data=data, mapping=aes(x=Year, y=Residuals, linetype=energy)) +
    scale_linetype_manual(values=c(2,3), name="") +
    facet_grid(Country ~ flavor) +
    scale_x_continuous(breaks=c(1960, 1980, 2000)) +
    xlab("") +
    ylab("r") + 
    xy_theme()
}
@

Because we fit in log-space, fitting residuals ($r_i$) are defined as
%
\begin{equation} \label{eq:log-residuals}
  r_i \equiv \ln(y_i) - \ln(\hat{y}_i) =\ln\left( \frac{y_i}{\hat{y}_i} \right), 
\end{equation}
%
where $r_i$ will be zero when there is agreement between 
historical~($y_i$) and fitted~($\hat{y}_i$) economic output.

The mean squared error~($mse$) for any fitted model can be calculated by
%
\begin{equation} \label{eq:mse}
  mse \equiv \frac{1}{N} \sum_{i=1}^N{r_i^2} .
\end{equation}

Figures~\ref{fig:Graphs CD-resid}--\ref{fig:Graphs CESekl-resid}
show fitting residuals for all models.
%
<<Graphs CD-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Fitting residuals for Cobb-Douglas models.">>=
residualPlot(filter(fitted_and_resid_data, model=="CD"))
@
%
<<Graphs CESkle-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Fitting residuals for CES models with $(kl$)$e$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CES" & (nest=="kle" | nest=="kl")))
@
%
<<Graphs CESlek-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Fitting residuals for CES models with ($le$)$k$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CES" & (nest == "lek" | nest=="kl")))
@
%
<<Graphs CESekl-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Fitting residuals for CES models with $(ek$)$l$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CES" & (nest=="ekl" | nest=="kl")))
@

%++++++++++++++++++++++++++++++
\subsection{Dynamic Solow residuals} 
\label{sec:dynamic_solow_residuals}
%++++++++++++++++++++++++++++++

<<Dynamic Solow Residuals>>=
# Calculate finite difference approximations of time derivatives for output and all factors of production
dynamicSR <- function(model, id, lag=1, pad="tail"){
  id_list <- strsplit(id, split = "[.]")[[1]]
  energy <- id_list[[3]]
  modelStr <- id_list[[4]]

  f <- model$formula
  denom <- summands(f, n=length(summands(f))) # time variable
  nums <- all.vars(f) [1:(length(all.vars(f))-1)] # y, k, l, and e (if present)
  df <- model$data[c(denom, nums)] # columns are: iYear, y, k, l, and e (if present)

  # Calculate names for the columns of finite difference derivatives
  derivNames <- lapply(1:length(nums), function(i){paste0("d", nums[[i]], ".d", denom)}) %>% unlist
  derivCols <- lapply(1:length(nums), function(i){
    # Calculate finite difference derivative columns
    ediff(df[[i+1]], pad=pad, lag=lag) / ediff(df[[1]], pad=pad, lag=lag)
  }) %>% do.call(cbind, .) %>% as.data.frame
  names(derivCols) <- derivNames  
  df <- cbind(derivCols, df[ , c(nums, denom)]) # puts derivatives first, then raw numbers, then time
  df[["dydtOvery"]] <- 1 / df[[ nums[[1]] ]] * df[[ derivNames[[1]] ]]
  coeffs <- naturalCoef(model)
  df <- cbind(df, coeffs)

  if (modelStr == "CD"){
    lambda_fd <- 
      df[["dydtOvery"]] - 
      df[["alpha_1"]] / df[[nums[[2]]]] * df[[derivNames[[2]]]] - 
      df[["alpha_2"]] / df[[nums[[3]]]] * df[[derivNames[[3]]]]
    if (energy == "withE"){
      # Subtract the energy term also
      lambda_fd <- lambda_fd - 
        df[["alpha_3"]] / df[[nums[[4]]]] * df[[derivNames[[4]]]]
    }
    # Fraction of growth explained by Solow Residual
    solowFrac <- lambda_fd / df[["dydtOvery"]]
    # Add to data frame
    df$lambda_fd <- lambda_fd
    df$solowFrac <- solowFrac
    return(df)
  } else if (modelStr == "CES"){
    # Assign nice variable names so that the equations below are easier to read (and debug, and maintain)
    # This code is nest-aware.
    n <- model$nest
    for (i in 1:length(model$nest)){
      df[[paste0("x", i)]] <- df[[ nums[ model$nest[i]+1 ] ]] # GDP is 1st col, add 1 to get to the factors of production
      df[[paste0("dx", i, "dt")]] <- df[[ derivNames[ model$nest[i]+1 ] ]]
    }
    if (energy == "noE"){
      # Calculate terms for the finite-difference Solow residual (lambda_fd)
      df <- df %>% 
        transform(b = delta_1 * x1^(-rho_1) + (1 - delta_1) * x2^(-rho_1)) %>%
        transform(dbdtOverrho1 = -delta_1 * x1^(-(rho_1+1))*dx1dt - (1 - delta_1) * x2^(-(rho_1+1))*dx2dt) %>%
        transform(lambda_fd = dydtOvery + 1/b * dbdtOverrho1) %>%
        transform(solowFrac = lambda_fd / dydtOvery)
      return(df)
    } else if (energy == "withE"){
      df <- df %>% 
        transform(b = delta_1 * x1^(-rho_1) + (1 - delta_1) * x2^(-rho_1)) %>%
        transform(a = delta * b^(rho/rho_1) + (1 - delta) * x3^(-rho)) %>%
        transform(dbdtOverrho1 = -delta_1 * x1^(-(rho_1+1))*dx1dt - (1 - delta_1) * x2^(-(rho_1+1))*dx2dt) %>%
        transform(dadtOverrho = delta * b^(rho/rho_1 - 1) * dbdtOverrho1 - (1 - delta) * x3^(-(rho+1))*dx3dt) %>%
        transform(lambda_fd = dydtOvery + 1/a * dadtOverrho) %>%
        transform(solowFrac = lambda_fd / dydtOvery)
      return(df)
    } else {
      stop (paste("Unknown energy:", energy))
    }
  } else {
    stop(paste("Unknown modelStr", modelStr))
  }
}

lag.to.use <- 1 # years
temp_data <- leaf_apply(models, class=c("CDEmodel", "cesModel", "LINEXmodel"), 
                        f = function(model, id, lag=lag.to.use, pad="tail"){
  df <- dynamicSR(model, id=id, lag=lag, pad=pad)
  id_list <- strsplit(id, split = "[.]")[[1]]
  df$Country <- id_list[[1]]
  df$flavor <- if (id_list[[2]] == "unadjusted") "Unadjusted" else "Quality-adjusted"
  df$energy <- if (id_list[[3]] == "noE") "Without energy" else "With energy"
  df$model <- id_list[[4]]
  df$nest <- if (length(id_list) > 4) id_list[[5]] else NA
  df$Year <- model$data[["Year"]]
  return(df)  
}) %>% do.call(plyr::rbind.fill, .)

# Get the unadjusted, without-energy Solow residuals and store them in a data frame
sr_unadj_noE <- filter(temp_data, flavor=="Unadjusted" & energy=="Without energy") %>% 
                select(Year, Country, lambda_fd, model)
colnames(sr_unadj_noE)[colnames(sr_unadj_noE)=="lambda_fd"] <- "sr_unadj_noE"
# Merge to get the unadjusted, without-energy values next to each country, year, and model
solowResiduals <- merge(temp_data, sr_unadj_noE)

# Create a column containing ∆lambda_df, the difference between a lambda_df value
# and the lambda_df value in the same year for the unadjusted, no-energy case
solowResiduals[["Dlambda_fd"]] <- abs(solowResiduals[["lambda_fd"]]) - abs(solowResiduals[["sr_unadj_noE"]])

# Relevel factors before plotting
solowResiduals$Country <- relevelFactor(as.factor(solowResiduals$Country), c("PT", "UK"))
solowResiduals$flavor <- relevelFactor(as.factor(solowResiduals$flavor), c("Unadjusted", "Quality-adjusted"))
solowResiduals$energy <- relevelFactor(as.factor(solowResiduals$energy), c("Without energy", "With energy"))
solowResiduals$nest <- relevelFactor(as.factor(solowResiduals$nest), c("kl", "kle", "lek", "ekl"))

# Make data viewing easier
solowResiduals <- arrange(solowResiduals, Country, model, flavor, energy, nest)
@

<<DynamicSR_Graph>>=
dynamicSRGraph <- function(data){
ggplot(data=data) +
  geom_line(data=filter(data, energy=="Without energy"), mapping=aes(x=Year, y=dydtOvery)) +
  geom_line(mapping=aes(x=Year, y=lambda_fd, linetype=energy)) +
  scale_linetype_manual(values=c(2,3), name="") +
  facet_grid(Country ~ flavor) +
  scale_x_continuous(breaks=c(1960, 1980, 2000)) +
  xlab("") + 
  ylab(expression(lambda[ij])) +
  xy_theme()  
}
@

``Dynamic'' Solow residuals~($\lambda_{i,j}$) can be calculated 
for the Cobb-Douglas and CES models
as shown in Appendices~\ref{sec:dynamic_sr_CD} and~\ref{sec:dynamic_sr_CES},
respectively.

Figure~\ref{fig:DynamicSR_CD_Graph} shows the evolution 
of the Solow residual~($\lambda_{i,j}$) over time
as calculated by Equation~\ref{eq:lambdaij} 
with the timespan for finite differences ($t_j - t_i$) set to 
\Sexpr{lag.to.use}~\Sexpr{if (lag.to.use == 1) "year" else "years"}.
%
<<DynamicSR_CD_Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residuals for the Cobb-Douglas model. Dashed lines are $\\lambda$ values. Solid lines are $\\frac{1}{y}\\frac{\\mathrm{d} y}{\\mathrm{d} t}$.">>=
dynamicSRGraph(filter(solowResiduals, model=="CD"))
@
%
<<DynamicSR_CESkle_Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residuals for the CES model with ($kl$)$e$ nesting. Dashed lines are $\\lambda$ values. Solid lines are $\\frac{1}{y}\\frac{\\mathrm{d} y}{\\mathrm{d} t}$.">>=
dynamicSRGraph(filter(solowResiduals, model=="CES" & (nest=="kle" | nest=="kl")))
@
%
<<DynamicSR_CESlek_Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residuals for the CES model with ($le$)$k$ nesting. Dashed lines are $\\lambda$ values. Solid lines are $\\frac{1}{y}\\frac{\\mathrm{d} y}{\\mathrm{d} t}$.">>=
dynamicSRGraph(filter(solowResiduals, model=="CES" & (nest=="lek" | nest=="kl")))
@
%
<<DynamicSR_CESekl_Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residuals for the CES model with ($ek$)$l$ nesting. Dashed lines are $\\lambda$ values. Solid lines are $\\frac{1}{y}\\frac{\\mathrm{d} y}{\\mathrm{d} t}$.">>=
dynamicSRGraph(filter(solowResiduals, model=="CES" & (nest=="ekl" | nest=="kl")))
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Analysis}
\label{sec:analysis}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For now, I've organized the analysis around questions that we would like to answer.

%++++++++++++++++++++++++++++++
\subsection{Does quality-adjusting the factors of production 
            decrease the mean squared error and/or the Solow residual?} 
\label{sec:quality_adj_energy_and_solow_residual}
%++++++++++++++++++++++++++++++

We can test our hypothesis that quality-adjusting the factors of production
and including energy will both reduce the Solow residuals~($\lambda$)
and improve the fit to historical data
(thereby, reducing the fitting residuals, $r_i$)
by calculating 
$\Delta \lambda$ and $\Delta mse$, where
%
\begin{equation} \label{eq:delta_lambda}
  \Delta \lambda \equiv \left| \lambda \right| - \left| \lambda_{Unadjusted, \, Without \, energy} \right|
\end{equation}
%
and
%
\begin{equation} \label{eq:delta_mse}
  \Delta mse \equiv mse - mse_{Unadjusted, \, Without \, energy}.
\end{equation}
%
where~$N$ is the number of years of data.
When $\Delta \lambda$ and $\Delta mse$ are negative, 
we observe reduction in the Solow residual ($\lambda$) and
the fitting residuals ($mse$)
relative to the unadjusted, no-energy condition.

Table~\ref{tab:DSRmse_CD} shows the results for the Cobb-Douglas models.
%
<<MyXTableFunction>>=
myXTable <- function(data, caption, label, digits=2){
  table <- xtable(data, caption=caption, label=label, digits=digits)
  print(table, 
        caption.placement="top", 
        size="\\tiny", 
        include.rownames=FALSE, 
        math.style.negative=TRUE)
}
@
%
<<DSRmse_CD, results="asis">>=
myXTable(filter(AllCoeffs, model=="CD") %>% select(Country, model, flavor, energy, Dlambda, Dmse),
         caption="$\\Delta \\lambda$ and $\\Delta mse$ for Cobb-Douglas models.",
         label=paste0("tab:", opts_current$get("label")), 
         digits=6)
@
% 
Tables~\ref{tab:DSRmse_CESkle}--\ref{tab:DSRmse_CESekl} show results 
for CES models with various nestings.
%
<<DSRmse_CESkle, results="asis">>=
myXTable(filter(AllCoeffs, model=="CES" & (nest=="kl" | nest=="kle")) %>% 
           select(Country, model, flavor, energy, nest, Dlambda, Dmse),
         caption="$\\Delta \\lambda$ and $\\Delta mse$ for CES models with ($kl$)$e$ nesting.",
         label=paste0("tab:", opts_current$get("label")), 
         digits=6)
@
%
<<DSRmse_CESlek, results="asis">>=
myXTable(filter(AllCoeffs, model=="CES" & (nest=="kl" | nest=="lek")) %>% 
           select(Country, model, flavor, energy, nest, Dlambda, Dmse),
         caption="$\\Delta \\lambda$ and $\\Delta mse$ for CES models with ($le$)$k$ nesting.",
         label=paste0("tab:", opts_current$get("label")), 
         digits=6)
@
%
<<DSRmse_CESekl, results="asis">>=
myXTable(filter(AllCoeffs, model=="CES" & (nest=="kl" | nest=="ekl")) %>% 
           select(Country, model, flavor, energy, nest, Dlambda, Dmse),
         caption="$\\Delta \\lambda$ and $\\Delta mse$ for CES models with ($ek$)$l$ nesting.",
         label=paste0("tab:", opts_current$get("label")), 
         digits=6)
@
%
Figure~\ref{fig:DSRmse_CD_graph} summarizes the data for the Cobb-Douglas model.
%
<<DSRmse_CD_graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Change in Solow Residuals~($\\Delta\\lambda$) and mean squared error~($\\Delta mse$) for the Cobb-Douglas model relative to the unadjusted, no-energy condition.">>=
ggplot(data=filter(AllCoeffs, model=="CD")) +
  geom_point(mapping=aes(x=Dmse, y=Dlambda, color = energy)) +
  facet_grid(Country ~ flavor) +
  xlab(expression(Delta * mse)) + 
  ylab(expression(Delta * lambda)) +
  xy_theme() + 
  theme(legend.title=element_blank())
@
%
Figure~\ref{fig:DSRmse_CES_graph} summarizes the data for the CES models.
%
<<DSRmse_CES_graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Change in Solow Residuals~($\\Delta\\lambda$) and mean squared error~($\\Delta mse$) for the Cobb-Douglas model relative to the unadjusted, no-energy condition.">>=
ggplot(data=filter(AllCoeffs, model=="CES")) +
  geom_point(mapping=aes(x=Dmse, y=Dlambda, color = energy, shape=nest)) +
  facet_grid(Country ~ flavor) +
  xlab(expression(Delta * mse)) + 
  ylab(expression(Delta * lambda)) +
  xy_theme() +
  theme(legend.title=element_blank())
@

%++++++++++++++++++++++++++++++
\subsection{Does quality-adjusting the factors of production 
            decrease fitting residuals and/or the dynamic Solow residual?} 
\label{sec:quality_adj_energy_and_dynamic_solow_residual}
%++++++++++++++++++++++++++++++

<<SR and R trajectory data>>=
traj.data <- merge(solowResiduals[c("Country", "Year", "model", "flavor", "energy", "nest", "Dlambda_fd")], 
                   fitted_and_resid_data[c("Country", "Year", "model", "flavor", "energy", "nest", "Dr")]) %>%
                   arrange(Country, model, flavor, energy, nest)
traj.data.stats <- traj.data %>% 
  group_by(model, Country, flavor, energy, nest) %>%
  summarise(Dlambda_fd.mean = mean(Dlambda_fd, na.rm=TRUE), Dr.mean = mean(Dr))

############################################
# Dlambda_fd.p <- traj.data %>% 
#   group_by(model, Country, flavor, energy, nest) %>% 
#   select(Dlambda_fd) %>%
#   t.test(., alternative="less")[["p.value"]]
############################################  
@

<<SR and R trajectory plot function>>=
residualsTrajGraph <- function(data){
  plot <- ggplot(data=data) + 
    geom_point(mapping=aes(x=Dr, y=Dlambda_fd, color=energy)) + 
    facet_grid(Country ~ flavor) +
    xlab(expression(Delta*r[i])) +
    ylab(expression(Delta*lambda[ij])) +
    xy_theme() + 
    theme(legend.title=element_blank())
  return(plot)
}
@

We can test our hypothesis that quality-adjusting the factors of production
and including energy will reduce the Solow residuals~($\lambda$)
and the fitting residuals~($r$) by plotting
$\Delta \lambda_i$ vs. $\Delta r_i$, where
%
\begin{equation} \label{eq:delta_lambda_i}
  \Delta \lambda_i \equiv \left| \lambda_i \right| - \left| \lambda_{i, \, Unadjusted, \, Without \, energy} \right|
\end{equation}
%
and
%
\begin{equation} \label{eq:delta_r}
  \Delta r_i \equiv \left| r_i \right| - \left| r_{i, \,Unadjusted, \, Without \, energy} \right|.
\end{equation}
%
If most of the points fall in the lower-left quadrant of a graph, 
we can conclude that both Solow residual~($\lambda$) and
fitting residual~($r$) have decreased due to either 
quality-adjustment of the factors of production or 
including energy.

Figure~\ref{fig:Solow_residual_trajectories_CD} indicates that
for the Cobb-Douglas model, 
neither 
quality-adjusting the factors of production
nor including energy 
consistently reduces
the Solow residual or the fitting residual.

<<Solow_residual_trajectories_CD, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residual and fitting residual trajectories relative to the Unadjusted, Without-energy condition for the Cobb-Douglas model.">>=
residualsTrajGraph(filter(traj.data, model=="CD"))
@

We can also calculate the average values for $\Delta \lambda_{i,j}$ and $\Delta r_{i}$ 
as shown in Table~\ref{tab:Trajectory_Means_CD}.
%
<<Trajectory_Means_CD, results="asis">>=
myXTable(data = filter(traj.data.stats, model=="CD") %>% select(-nest), 
         caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for Cobb-Douglas models.",
         label=paste0("tab:", opts_current$get("label")),
         digits=6)
@
%
The following tables and figures show the same analysis for the CES model.
%
<<Solow_residual_trajectories_CESkle, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residual and fitting residual trajectories relative to the Unadjusted, Without-energy condition for the CES model with ($kl$)$e$ nesting.">>=
residualsTrajGraph(filter(traj.data, model=="CES" & (nest=="kl" | nest=="kle")))
@
%
<<Trajectory_Means_CESkle, results="asis">>=
myXTable(data = filter(traj.data.stats, model=="CES" & (nest=="kl" | nest=="kle")), 
         caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for CES models with ($kl$)$e$ nesting.",
         label=paste0("tab:", opts_current$get("label")),
         digits=6)
@

<<Solow_residual_trajectories_CESlek, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residual and fitting residual trajectories relative to the Unadjusted, Without-energy condition for the CES model with ($le$)$k$ nesting.">>=
residualsTrajGraph(filter(traj.data, model=="CES" & (nest=="kl" | nest=="lek")))
@

<<Trajectory_Means_CESlek, results="asis">>=
myXTable(data = filter(traj.data.stats, model=="CES" & (nest=="kl" | nest=="lek")), 
         caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for CES models with ($le$)$k$ nesting.",
         label=paste0("tab:", opts_current$get("label")),
         digits=6)
@

<<Solow_residual_trajectories_CESekl, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residual and fitting residual trajectories relative to the Unadjusted, Without-energy condition for the CES model with ($ek$)$l$ nesting.">>=
residualsTrajGraph(filter(traj.data, model=="CES" & (nest=="kl" | nest=="ekl")))
@

<<Trajectory_Means_CESekl, results="asis">>=
myXTable(data = filter(traj.data.stats, model=="CES" & (nest=="kl" | nest=="ekl")), 
         caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for CES models with ($ek$)$l$ nesting.",
         label=paste0("tab:", opts_current$get("label")),
         digits=6)
@

% %++++++++++++++++++++++++++++++
% \subsection{What effect does including energy and quality-adjustment have 
%             on the Solow residual in the Cobb-Douglas model?} 
% \label{sec:quality_adj_energy_and_solow_residual}
% %++++++++++++++++++++++++++++++
% 
% We'll first look at Portugal.
% 
% %------------------------------
% \subsubsection{Portugal} 
% %------------------------------
% 
% First, use non-quality-adjusted data for Portugal and the Cobb-Douglas model.
% 
% <<Portugal C-D undadjusted without energy>>=
% naturalCoef(models$PT$unadjusted$noE$CD)
% @
% 
% <<Portugal C-D undadjusted with energy>>=
% naturalCoef(models$PT$unadjusted$withE$CD)
% @
% 
% The Solow residual (\texttt{lambda}) is always negative. 
% However, the absolute value of \texttt{lambda} decreases from 
% \Sexpr{naturalCoef(models$PT$unadjusted$noE$CD)$lambda} to 
% \Sexpr{naturalCoef(models$PT$unadjusted$withE$CD)$lambda}
% when adding energy.
% So it appears that for Portugal, 
% using unadjusted inputs and including energy as a factor of production
% reduces the absolute value of the Solow residual somewhat.
% 
% What about for quality-adjusted data and Portugal?
% 
% <<Portugal C-D adjusted without energy>>=
% naturalCoef(models$PT$adjusted$noE$CD)
% @
% 
% <<Portugal C-D adjusted with energy>>=
% naturalCoef(models$PT$adjusted$withE$CD)
% @
% 
% Using quality-adjusted inputs, the Solow residual (\texttt{lambda}) moves from 
% \Sexpr{naturalCoef(models$PT$adjusted$noE$CD)$lambda} to
% \Sexpr{naturalCoef(models$PT$adjusted$withE$CD)$lambda} when including energy.
% The absolute value of the Solow residual again decreases when adding energy.
% But, with quality-adjusted inputs, the effect is much stronger.
% 
% Table~\ref{tab:lambda_CD_PT} shows
% the effect of both quality-corrected factors of production and energy.
% We find that including energy as a factor of production 
% reduces the absolute value of the Solow residual (\texttt{lambda})
% whether or not the factors of production are quality-adjusted. 
% Quality-adjustment gets part of the way toward eliminating the Solow residual.
% 
% \begin{table} \caption{Portugal Solow residual results with Cobb-Douglas.} 
% \label{tab:lambda_CD_PT} 
%   \begin{center}
%     \begin{tabular}{r c c} 
%       \toprule
%                       & Unadjusted     & Quality-adjusted   \\
%       \midrule
%       Without energy  & \Sexpr{naturalCoef(models$PT$unadjusted$noE$CD)$lambda}    & \Sexpr{naturalCoef(models$PT$adjusted$noE$CD)$lambda} \\
%       With energy     & \Sexpr{naturalCoef(models$PT$unadjusted$withE$CD)$lambda}  & \Sexpr{naturalCoef(models$PT$adjusted$withE$CD)$lambda} \\
%       \bottomrule
%     \end{tabular}
%   \end{center}
% \end{table}
% 
% Note that for Portugal the contribution of capital (\texttt{alpha\_1})
% always shifts to energy (\texttt{alpha\_3}) when energy is included 
% as a factor of production.
% 
% What about for the UK?
% 
% %------------------------------
% \subsubsection{United Kingdom} 
% %------------------------------
% 
% Table~\ref{tab:lambda_CD_UK} shows fitted Cobb-Douglas coefficients for the UK. 
% 
% \begin{table} \caption{UK Solow residual results with Cobb-Douglas.} 
% \label{tab:lambda_CD_UK} 
%   \begin{center}
%     \begin{tabular}{r c c} 
%       \toprule
%                       & Unadjusted     & Quality-adjusted   \\
%       \midrule
%       Without energy  & \Sexpr{naturalCoef(models$UK$unadjusted$noE$CD)$lambda}   
%                                                 & \Sexpr{naturalCoef(models$UK$adjusted$noE$CD)$lambda} \\
%       With energy     & \Sexpr{naturalCoef(models$UK$unadjusted$withE$CD)$lambda} 
%                                                 & \Sexpr{naturalCoef(models$UK$adjusted$withE$CD)$lambda} \\
%       \bottomrule
%     \end{tabular}
%   \end{center}
% \end{table}
% 
% For the UK, the quality-adjusted, with energy Solow residual is 
% \Sexpr{naturalCoef(models$UK$adjusted$withE$CD)$lambda * 100}\%,
% a very small number indeed.
% For the UK, it appears that \emph{both} quality-adjustment and addition of energy as a factor of production
% are required to effectively eliminate the Solow residual,
% reducing it to \Sexpr{naturalCoef(models$UK$adjusted$withE$CD)$lambda}.
% 
% Adding energy alone actually \emph{in}creases the Solow residual from 
% \Sexpr{naturalCoef(models$UK$unadjusted$noE$CD)$lambda} to
% \Sexpr{naturalCoef(models$UK$unadjusted$withE$CD)$lambda}.
% The details for each model with unadjusted factors of production are shown below.
% 
% <<UK CD Unadjusted Coefficients>>=
% naturalCoef(models$UK$unadjusted$noE$CD)
% naturalCoef(models$UK$unadjusted$withE$CD)
% @
% 
% When energy is included, much of the effect of capital (\texttt{alpha\_1}) is 
% attributed to energy (\texttt{alpha\_3}), 
% as \texttt{alpha\_1} moves from 
% \Sexpr{naturalCoef(models$UK$unadjusted$noE$CD)$alpha_1} to
% \Sexpr{naturalCoef(models$UK$unadjusted$withE$CD)$alpha_1}.
% 
% When we look at quality-adjusted inputs, we find similar results.
% 
% <<UK CD Quality-adjusted Coefficients>>=
% naturalCoef(models$UK$adjusted$noE$CD)
% naturalCoef(models$UK$adjusted$withE$CD)
% @
% 
% \texttt{alpha\_1} moves from 
% \Sexpr{naturalCoef(models$UK$adjusted$noE$CD)$alpha_1} to
% \Sexpr{naturalCoef(models$UK$adjusted$withE$CD)$alpha_1}.
% 
% Now, let's compare the fits to historical data.
% 
% 
% 
% %++++++++++++++++++++++++++++++
% \subsection{Do we obtain a negative Solow residual for Portugal with the CES model?} 
% \label{sec:Solow_negative_for_PT_in_CES?}
% %++++++++++++++++++++++++++++++
% 
% First, look at the CES model without energy.
% Nesting is unabmiguous, so this fit does not need to be repeated
% for each nesting option. 
% (This means that the top rows of Tables~\ref{tab:lambda_CES-kle_PT}--\ref{tab:lambda_CES-ekl_PT}
% are identical.)
% 
% <<PT CES-kl, warning=FALSE>>=
% naturalCoef(models$PT$unadjusted$noE$CES$kl)
% naturalCoef(models$PT$adjusted$noE$CES$kl)
% @
% 
% Both unadjusted and quality-adjusted variables give positive Solow residuals,
% \Sexpr{naturalCoef(models$PT$unadjusted$noE$CES$kl)$lambda} for unadjusted inputs and
% \Sexpr{naturalCoef(models$PT$adjusted$noE$CES$kl)$lambda} for quality-adjusted inputs.
% Furthermore, quality-adjustment reduces the magnitude of the Solow residual.
% 
% %------------------------------
% \subsubsection{($kl$)$e$ nesting} 
% %------------------------------
% 
% Now, look at unadjusted inputs using the ($kl$)$e$ nesting.
% 
% <<PT CES-kle unadjusted, warning=FALSE>>=
% naturalCoef(models$PT$unadjusted$withE$CES$kle)
% @
% 
% Now look at quality-adjusted inputs using the ($kl$)$e$ nesting.
% 
% <<PT CES-kle quality-adjusted, warning=FALSE>>=
% naturalCoef(models$PT$adjusted$withE$CES$kle)
% @
% 
% Table~\ref{tab:lambda_CES-kle_PT} summarizes the Portugal Solow residual results with CES and the ($kl$)$e$ nesting.
% 
% \begin{table} \caption{Portugal Solow residual results with CES and ($kl$)$e$ nesting.} 
% \label{tab:lambda_CES-kle_PT} 
%   \begin{center}
%     \begin{tabular}{r c c} 
%       \toprule
%                       & Unadjusted     & Quality-adjusted   \\
%       \midrule
%       Without energy  & \Sexpr{naturalCoef(models$PT$unadjusted$noE$CES$kl)$lambda} 
%                                                             & \Sexpr{naturalCoef(models$PT$adjusted$noE$CES$kl)$lambda} \\
%       With energy     & \Sexpr{naturalCoef(models$PT$unadjusted$withE$CES$kle)$lambda} 
%                                                             & \Sexpr{naturalCoef(models$PT$adjusted$withE$CES$kle)$lambda} \\
%       \bottomrule
%     \end{tabular}
%   \end{center}
% \end{table}
% 
% Table~\ref{tab:sse_CES-kle_PT} shows the $sse$ values for each of these models.
% 
% \begin{table} \caption{PT $sse$ with CES and ($kl$)$e$ nesting.} 
% \label{tab:sse_CES-kle_PT} 
%   \begin{center}
%     \begin{tabular}{r c c} 
%       \toprule
%                       & Unadjusted     & Quality-adjusted   \\
%       \midrule
%       Without energy  & \Sexpr{EconModels:::sse(models$PT$unadjusted$noE$CES$kl)$sse} 
%                                                             & \Sexpr{EconModels:::sse(models$PT$adjusted$noE$CES$kl)$sse} \\
%       With energy     & \Sexpr{EconModels:::sse(models$PT$unadjusted$withE$CES$kle)$sse} 
%                                                             & \Sexpr{EconModels:::sse(models$PT$adjusted$withE$CES$kle)$sse} \\
%       \bottomrule
%     \end{tabular}
%   \end{center}
% \end{table}
% 
% In contrast to the Cobb-Douglas model,
% the CES model with ($kl$)$e$ nesting
% produces positive Solow residuals in all cases. 
% We also see that both 
% quality-adjusting the inputs and
% including energy
% reduce the Solow residual (\texttt{lambda}).
% 
% %------------------------------
% \subsubsection{($le$)$k$ nesting} 
% %------------------------------
% 
% Next, look at unadjusted inputs using the ($le$)$k$ nesting.
% 
% <<PT CES-lek unadjusted, warning=FALSE>>=
% naturalCoef(models$PT$unadjusted$withE$CES$lek)
% @
% 
% Now look at quality-adjusted inputs using the ($le$)$k$ nesting.
% 
% <<PT CES-lek quality-adjusted, warning=FALSE>>=
% naturalCoef(models$PT$adjusted$withE$CES$lek)
% @
% 
% Table~\ref{tab:lambda_CES-lek_PT} summarizes the Portugal Solow residual results with CES and the ($le$)$k$ nesting.
% 
% \begin{table} \caption{Portugal Solow residual results with CES and ($le$)$k$ nesting.} 
% \label{tab:lambda_CES-lek_PT} 
%   \begin{center}
%     \begin{tabular}{r c c} 
%       \toprule
%                       & Unadjusted     & Quality-adjusted   \\
%       \midrule
%       Without energy  & \Sexpr{naturalCoef(models$PT$unadjusted$noE$CES$kl)$lambda} 
%                                         & \Sexpr{naturalCoef(models$PT$adjusted$noE$CES$kl)$lambda} \\
%       With energy     & \Sexpr{naturalCoef(models$PT$unadjusted$withE$CES$lek)$lambda} 
%                                         & \Sexpr{naturalCoef(models$PT$adjusted$withE$CES$lek)$lambda} \\
%       \bottomrule
%     \end{tabular}
%   \end{center}
% \end{table}
% 
% Table~\ref{tab:sse_CES-lek_PT} shows the $sse$ values for each of these models.
% 
% \begin{table} \caption{PT $sse$ with CES and ($le$)$k$ nesting.} 
% \label{tab:sse_CES-lek_PT} 
%   \begin{center}
%     \begin{tabular}{r c c} 
%       \toprule
%                       & Unadjusted     & Quality-adjusted   \\
%       \midrule
%       Without energy  & \Sexpr{EconModels:::sse(models$PT$unadjusted$noE$CES$kl)$sse} 
%                                         & \Sexpr{EconModels:::sse(models$PT$adjusted$noE$CES$kl)$sse} \\
%       With energy     & \Sexpr{EconModels:::sse(models$PT$unadjusted$withE$CES$lek)$sse} 
%                                         & \Sexpr{EconModels:::sse(models$PT$adjusted$withE$CES$lek)$sse} \\
%       \bottomrule
%     \end{tabular}
%   \end{center}
% \end{table}
% 
% Similar to the CES model with ($kl$)$e$ nesting,
% the CES model with ($le$)$k$ nesting
% produces positive Solow residuals in all cases. 
% After energy has been added, quality-adjustment 
% has very little effect on the Solow residual,
% creating a slight increase from 
% \Sexpr{naturalCoef(models$PT$unadjusted$withE$CES$lek)$lambda} to
% \Sexpr{naturalCoef(models$PT$adjusted$withE$CES$lek)$lambda}.
% 
% 
% %------------------------------
% \subsubsection{($ek$)$l$ nesting} 
% %------------------------------
% 
% Next, look at unadjusted inputs using the ($ek$)$l$ nesting.
% 
% <<PT CES-ekl unadjusted, warning=FALSE>>=
% naturalCoef(models$PT$unadjusted$withE$CES$ekl)
% @
% 
% Now look at quality-adjusted inputs using the ($ek$)$l$ nesting.
% 
% <<PT CES-ekl quality-adjusted, warning=FALSE>>=
% naturalCoef(models$PT$adjusted$withE$CES$ekl)
% @
% 
% Table~\ref{tab:lambda_CES-ekl_PT} summarizes the Portugal Solow residual results with CES and the ($ek$)$l$ nesting.
% 
% \begin{table} \caption{Portugal Solow residual results with CES and ($ek$)$l$ nesting.} 
% \label{tab:lambda_CES-ekl_PT} 
%   \begin{center}
%     \begin{tabular}{r c c} 
%       \toprule
%                       & Unadjusted     & Quality-adjusted   \\
%       \midrule
%       Without energy  & \Sexpr{naturalCoef(models$PT$unadjusted$noE$CES$kl)$lambda} 
%                                         & \Sexpr{naturalCoef(models$PT$adjusted$noE$CES$kl)$lambda} \\
%       With energy     & \Sexpr{naturalCoef(models$PT$unadjusted$withE$CES$ekl)$lambda} 
%                                         & \Sexpr{naturalCoef(models$PT$adjusted$withE$CES$ekl)$lambda} \\
%       \bottomrule
%     \end{tabular}
%   \end{center}
% \end{table}
% 
% Table~\ref{tab:sse_CES-ekl_PT} shows the $sse$ values for each of these models.
% 
% \begin{table} \caption{PT $sse$ with CES and ($ek$)$l$ nesting.} 
% \label{tab:sse_CES-ekl_PT} 
%   \begin{center}
%     \begin{tabular}{r c c} 
%       \toprule
%                       & Unadjusted     & Quality-adjusted   \\
%       \midrule
%       Without energy  & \Sexpr{EconModels:::sse(models$PT$unadjusted$noE$CES$kl)$sse} 
%                                         & \Sexpr{EconModels:::sse(models$PT$adjusted$noE$CES$kl)$sse} \\
%       With energy     & \Sexpr{EconModels:::sse(models$PT$unadjusted$withE$CES$ekl)$sse} 
%                                         & \Sexpr{EconModels:::sse(models$PT$adjusted$withE$CES$ekl)$sse} \\
%       \bottomrule
%     \end{tabular}
%   \end{center}
% \end{table}
% 
% For the ($ek$)$l$ nesting, 
% the Solow residual is positive, and 
% quality-adjusting the inputs 
% reduces the Solow residual (\texttt{lambda}).
% Adding energy decreases the Solow residual 
% (from \Sexpr{naturalCoef(models$PT$unadjusted$withE$CES$ekl)$lambda} to 
% \Sexpr{naturalCoef(models$PT$adjusted$withE$CES$ekl)$lambda})
% when enegy is included in the model.
% But, adding energy slightly increases the Solow residual
% (from \Sexpr{naturalCoef(models$PT$adjusted$noE$CES$kl)$lambda} to \Sexpr{naturalCoef(models$PT$adjusted$withE$CES$ekl)$lambda})
% when quality-adjusted inputs are used.
% 
% 
% %++++++++++++++++++++++++++++++
% \subsection{Do we obtain a positive Solow residual for the UK with the CES model?} 
% \label{sec:Solow_positive_for_UK_in_CES?}
% %++++++++++++++++++++++++++++++
% 
% First, look at the CES model without energy.
% Nesting is unabmiguous, so this fit does not need to be repeated
% for each nesting option. 
% (This means that the top rows of Tables~\ref{tab:lambda_CES-kle_UK}--\ref{tab:lambda_CES-ekl_UK}
% are identical.)
% 
% <<UK CES-kl, warning=FALSE>>=
% naturalCoef(models$UK$unadjusted$noE$CES$kl)
% naturalCoef(models$UK$adjusted$noE$CES$kl)
% @
% 
% Both unadjusted and quality-adjusted variables give positive Solow residuals for the UK,
% \Sexpr{naturalCoef(models$UK$unadjusted$noE$CES$kl)$lambda} for unadjusted inputs and
% \Sexpr{naturalCoef(models$UK$adjusted$noE$CES$kl)$lambda} for quality-adjusted inputs.
% Furthermore, quality-adjustment reduces the magnitude of the UK Solow residual.
% 
% %------------------------------
% \subsubsection{($kl$)$e$ nesting} 
% %------------------------------
% 
% Now, look at unadjusted inputs using the ($kl$)$e$ nesting.
% 
% <<UK CES-kle unadjusted>>=
% naturalCoef(models$UK$unadjusted$withE$CES$kle)
% @
% 
% Now look at quality-adjusted inputs using the ($kl$)$e$ nesting.
% 
% <<UK CES-kle quality-adjusted>>=
% naturalCoef(models$UK$adjusted$withE$CES$kle)
% @
% 
% Table~\ref{tab:lambda_CES-kle_UK} summarizes the UK Solow residual results with CES and the ($kl$)$e$ nesting.
% 
% \begin{table} \caption{UK Solow residual results with CES and ($kl$)$e$ nesting.} 
% \label{tab:lambda_CES-kle_UK} 
%   \begin{center}
%     \begin{tabular}{r c c} 
%       \toprule
%                       & Unadjusted     & Quality-adjusted   \\
%       \midrule
%       Without energy  & \Sexpr{naturalCoef(models$UK$unadjusted$noE$CES$kl)$lambda} & \Sexpr{naturalCoef(models$UK$adjusted$noE$CES$kl)$lambda} \\
%       With energy     & \Sexpr{naturalCoef(models$UK$unadjusted$withE$CES$kle)$lambda} 
%                                                                 & \Sexpr{naturalCoef(models$UK$adjusted$withE$CES$kle)$lambda} \\
%       \bottomrule
%     \end{tabular}
%   \end{center}
% \end{table}
% 
% Table~\ref{tab:sse_CES-kle_UK} shows the $sse$ values for each of these models.
% 
% \begin{table} \caption{UK $sse$ with CES and ($kl$)$e$ nesting.} 
% \label{tab:sse_CES-kle_UK} 
%   \begin{center}
%     \begin{tabular}{r c c} 
%       \toprule
%                       & Unadjusted     & Quality-adjusted   \\
%       \midrule
%       Without energy  & \Sexpr{EconModels:::sse(models$UK$unadjusted$noE$CES$kl)$sse} & \Sexpr{EconModels:::sse(models$UK$adjusted$noE$CES$kl)$sse} \\
%       With energy     & \Sexpr{EconModels:::sse(models$UK$unadjusted$withE$CES$kle)$sse} 
%                                                             & \Sexpr{EconModels:::sse(models$UK$adjusted$withE$CES$kle)$sse} \\
%       \bottomrule
%     \end{tabular}
%   \end{center}
% \end{table}
% 
% Like the Cobb-Douglas model, the CES model produces positive Solow residuals
% with the ($kl$)$e$ nesting.
% We also see that both 
% quality-adjusting the inputs and
% including energy
% reduce the Solow residual (\texttt{lambda}).
% 
% %------------------------------
% \subsubsection{($le$)$k$ nesting}
% %------------------------------
% 
% Next, look at unadjusted inputs using the ($le$)$k$ nesting.
% 
% <<UK CES-lek unadjusted>>=
% naturalCoef(models$UK$unadjusted$withE$CES$lek)
% @
% 
% Now look at quality-adjusted inputs using the ($le$)$k$ nesting.
% 
% <<UK CES-lek quality-adjusted>>=
% naturalCoef(models$UK$adjusted$withE$CES$lek)
% @
% 
% Table~\ref{tab:lambda_CES-lek_UK} summarizes the UK Solow residual results with CES and the ($le$)$k$ nesting.
% 
% \begin{table} \caption{UK Solow residual results with CES and ($le$)$k$ nesting.} 
% \label{tab:lambda_CES-lek_UK} 
%   \begin{center}
%     \begin{tabular}{r c c} 
%       \toprule
%                       & Unadjusted     & Quality-adjusted   \\
%       \midrule
%       Without energy  & \Sexpr{naturalCoef(models$UK$unadjusted$noE$CES$kl)$lambda} & \Sexpr{naturalCoef(models$UK$adjusted$noE$CES$kl)$lambda} \\
%       With energy     & \Sexpr{naturalCoef(models$UK$unadjusted$withE$CES$lek)$lambda} 
%                                                             & \Sexpr{naturalCoef(models$UK$adjusted$withE$CES$lek)$lambda} \\
%       \bottomrule
%     \end{tabular}
%   \end{center}
% \end{table}
% 
% Table~\ref{tab:sse_CES-lek_UK} shows the $sse$ values for each of these models.
% 
% \begin{table} \caption{UK $sse$ with CES and ($le$)$k$ nesting.} 
% \label{tab:sse_CES-lek_UK} 
%   \begin{center}
%     \begin{tabular}{r c c} 
%       \toprule
%                       & Unadjusted     & Quality-adjusted   \\
%       \midrule
%       Without energy  & \Sexpr{EconModels:::sse(models$UK$unadjusted$noE$CES$kl)$sse} & \Sexpr{EconModels:::sse(models$UK$adjusted$noE$CES$kl)$sse} \\
%       With energy     & \Sexpr{EconModels:::sse(models$UK$unadjusted$withE$CES$lek)$sse} 
%                                                             & \Sexpr{EconModels:::sse(models$UK$adjusted$withE$CES$lek)$sse} \\
%       \bottomrule
%     \end{tabular}
%   \end{center}
% \end{table}
% 
% 
% Similar to the CES model with ($kl$)$e$ nesting,
% the CES model with ($le$)$k$ nesting
% produces positive Solow residuals in all cases. 
% Both quality-adjustment and inclusion of energy 
% decrease the magnitude of the Solow residual.
% 
% %------------------------------
% \subsubsection{($ek$)$l$ nesting}
% %------------------------------
% 
% Next, look at unadjusted inputs using the ($el$)$l$ nesting.
% 
% <<UK CES-ekl unadjusted>>=
% naturalCoef(models$UK$unadjusted$withE$CES$ekl)
% @
% 
% Now look at quality-adjusted inputs using the ($ek$)$l$ nesting.
% 
% <<UK CES-ekl quality-adjusted>>=
% naturalCoef(models$UK$adjusted$withE$CES$ekl)
% @
% 
% Table~\ref{tab:lambda_CES-ekl_UK} summarizes the UK Solow residual results with CES and the ($ek$)$l$ nesting.
% 
% \begin{table} \caption{UK Solow residual results with CES and ($ek$)$l$ nesting.} 
% \label{tab:lambda_CES-ekl_UK} 
%   \begin{center}
%     \begin{tabular}{r c c} 
%       \toprule
%                       & Unadjusted     & Quality-adjusted   \\
%       \midrule
%       Without energy  & \Sexpr{naturalCoef(models$UK$unadjusted$noE$CES$kl)$lambda} & \Sexpr{naturalCoef(models$UK$adjusted$noE$CES$kl)$lambda} \\
%       With energy     & \Sexpr{naturalCoef(models$UK$unadjusted$withE$CES$ekl)$lambda} 
%                                                             & \Sexpr{naturalCoef(models$UK$adjusted$withE$CES$ekl)$lambda} \\
%       \bottomrule
%     \end{tabular}
%   \end{center}
% \end{table}
% 
% Table~\ref{tab:sse_CES-ekl_UK} shows the $sse$ values for each of these models.
% 
% \begin{table} \caption{UK $sse$ with CES and ($ek$)$l$ nesting.} 
% \label{tab:sse_CES-ekl_UK} 
%   \begin{center}
%     \begin{tabular}{r c c} 
%       \toprule
%                       & Unadjusted     & Quality-adjusted   \\
%       \midrule
%       Without energy  & \Sexpr{EconModels:::sse(models$UK$unadjusted$noE$CES$kl)$sse} & \Sexpr{EconModels:::sse(models$UK$adjusted$noE$CES$kl)$sse} \\
%       With energy     & \Sexpr{EconModels:::sse(models$UK$unadjusted$withE$CES$ekl)$sse} 
%                                                             & \Sexpr{EconModels:::sse(models$UK$adjusted$withE$CES$ekl)$sse} \\
%       \bottomrule
%     \end{tabular}
%   \end{center}
% \end{table}
% 
% 
% Similar to the CES model with ($ek$)$l$ nesting,
% the CES model with ($ek$)$l$ nesting
% produces positive Solow residuals in all cases. 
% Both quality-adjustment and inclusion of energy 
% decrease the magnitude of the Solow residual.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:Conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Future Work}
\label{sec:FutureWork}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}
\label{sec:Acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section*{Reproducible Research}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% In the spirit of Reproducible Research \citep{Gandrud:2013vx}, 
% all data, spreadsheets, \texttt{R} code \citep{R}, and other materials 
% associated with this paper can be found at\\*
% \protect\url{https://github.com/MatthewHeun/Econ-Growth-R-Analysis}.

%% References
%%
%% Following citation commands can be used in the body text:
%%
%%  \citet{key}  ==>>  Jones et al. (1990)
%%  \citep{key}  ==>>  (Jones et al., 1990)
%%
%% Multiple citations as normal:
%% \citep{key1,key2}         ==>> (Jones et al., 1990; Smith, 1989)
%%                            or  (Jones et al., 1990, 1991)
%%                            or  (Jones et al., 1990a,b)
%% \cite{key} is the equivalent of \citet{key} in author-year mode
%%
%% Full author lists may be forced with \citet* or \citep*, e.g.
%%   \citep*{key}            ==>> (Jones, Baker, and Williams, 1990)
%%
%% Optional notes as:
%%   \citep[chap. 2]{key}    ==>> (Jones et al., 1990, chap. 2)
%%   \citep[e.g.,][]{key}    ==>> (e.g., Jones et al., 1990)
%%   \citep[see][pg. 34]{key}==>> (see Jones et al., 1990, pg. 34)
%%  (Note: in standard LaTeX, only one note is allowed, after the ref.
%%   Here, one note is like the standard, two make pre- and post-notes.)
%%
%%   \citealt{key}          ==>> Jones et al. 1990
%%   \citealt*{key}         ==>> Jones, Baker, and Williams 1990
%%   \citealp{key}          ==>> Jones et al., 1990
%%   \citealp*{key}         ==>> Jones, Baker, and Williams, 1990
%%
%% Additional citation possibilities
%%   \citeauthor{key}       ==>> Jones et al.
%%   \citeauthor*{key}      ==>> Jones, Baker, and Williams
%%   \citeyear{key}         ==>> 1990
%%   \citeyearpar{key}      ==>> (1990)
%%   \citetext{priv. comm.} ==>> (priv. comm.)
%%   \citenum{key}          ==>> 11 [non-superscripted]
%% Note: full author lists depends on whether the bib style supports them;
%%       if not, the abbreviated list is printed even when full requested.
%%
%% For names like della Robbia at the start of a sentence, use
%%   \Citet{dRob98}         ==>> Della Robbia (1998)
%%   \Citep{dRob98}         ==>> (Della Robbia, 1998)
%%   \Citeauthor{dRob98}    ==>> Della Robbia

%% References with bibTeX database:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{References}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \bibliographystyle{model2-names}
%%\bibliography{<your-bib-database>}
% \bibliography{Paper2.bib}

%% Authors are advised to submit their bibtex database files. They are
%% requested to list a bibtex style file in the manuscript if they do
%% not want to use model2-names.bst.

%% References without bibTeX database:

% \begin{thebibliography}{00}

%% \bibitem must have one of the following forms:
%%   \bibitem[Jones et al.(1990)]{key}...
%%   \bibitem[Jones et al.(1990)Jones, Baker, and Williams]{key}...
%%   \bibitem[Jones et al., 1990]{key}...
%%   \bibitem[\protect\citeauthoryear{Jones, Baker, and Williams}{Jones
%%       et al.}{1990}]{key}...
%%   \bibitem[\protect\citeauthoryear{Jones et al.}{1990}]{key}...
%%   \bibitem[\protect\astroncite{Jones et al.}{1990}]{key}...
%%   \bibitem[\protect\citename{Jones et al., }1990]{key}...
%%   \harvarditem[Jones et al.]{Jones, Baker, and Williams}{1990}{key}...
%%

% \bibitem[ ()]{}

% \end{thebibliography}

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix
%% \section{}
%% \label{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Derivation of dynamic Solow residual for the CD equation}
\label{sec:dynamic_sr_CD}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Assuming that parameters $\theta$, $\alpha$, and $\beta$
in Equation~\ref{eq:CDe} are known from parameter estimation,
we can estimate the value of $\lambda$ as follows.
First, assume constant returns to scale such that 
$\alpha + \beta + \gamma = 1$, 
and calculate $\gamma = 1 - \alpha - \beta$.
Then, take the natural logarithm ($\ln$) of Equation~\ref{eq:CDe} to obtain
%
\begin{equation} \label{eq:lnCD}
  \ln y = \ln \theta + \ln A + \alpha \ln k + \beta \ln l + (1 - \alpha - \beta) \ln e \;.
\end{equation}
%
By taking the derivative of Equation~\ref{eq:lnCD} with respect to time ($t$) 
and noting that model parameters $\theta$, $\alpha$, and $\beta$ are 
constant with respect to time, we obtain
%
\begin{equation} \label{eq:ddtlnCD}
  \frac{1}{y}\frac{\mathrm{d}y}{\mathrm{d}t} 
      = \frac{1}{A}\frac{\mathrm{d}A}{\mathrm{d}t} 
      + \alpha \frac{1}{k}\frac{\mathrm{d}k}{\mathrm{d}t} 
      + \beta \frac{1}{l}\frac{\mathrm{d}l}{\mathrm{d}t} 
      + (1 - \alpha - \beta) \frac{1}{e}\frac{\mathrm{d}e}{\mathrm{d}t} \;.
\end{equation}
%
Solving for the total factor productivity term gives
%
\begin{equation} \label{eq:solow_resid}
  \frac{1}{A}\frac{\mathrm{d}A}{\mathrm{d}t} 
      = \frac{1}{y}\frac{\mathrm{d}y}{\mathrm{d}t} 
      - \left[
              \alpha \frac{1}{k}\frac{\mathrm{d}k}{\mathrm{d}t} 
              + \beta \frac{1}{l}\frac{\mathrm{d}l}{\mathrm{d}t} 
              + (1 - \alpha - \beta) \frac{1}{e}\frac{\mathrm{d}e}{\mathrm{d}t} 
        \right] \;.
\end{equation}
%
Recognizing that $A \equiv \mathrm{e}^{\lambda (t-t_0)}$ gives
%
\begin{equation} \label{eq:lambda_calc}
  \frac{1}{A}\frac{\mathrm{d}A}{\mathrm{d}t} 
    = \frac{1}{\mathrm{e}^{\lambda (t-t_0)}} \; \lambda \mathrm{e}^{\lambda (t-t_0)}
    = \lambda \; ,
\end{equation}
%
which can be substituted into Equation~\ref{eq:solow_resid} to find
%
\begin{equation} \label{eq:lambda}
  \lambda 
    = \frac{1}{y}\frac{\mathrm{d}y}{\mathrm{d}t} 
    - \left[
              \alpha \frac{1}{k}\frac{\mathrm{d}k}{\mathrm{d}t} 
              + \beta \frac{1}{l}\frac{\mathrm{d}l}{\mathrm{d}t} 
              + (1 - \alpha - \beta) \frac{1}{e}\frac{\mathrm{d}e}{\mathrm{d}t} 
      \right] \;.
\end{equation}
%
Equation~\ref{eq:lambda} applies for any instant in time.

If we approximate derivatives in Equation~\ref{eq:lambda} with forward differences
between times $i$ and $j$, we find
%
\begin{equation} \label{eq:lambdaij}
  \lambda_{i,j}
    = \frac{1}{y_i}\frac{y_j - y_i}{t_j - t_i} 
    - \left[
            \alpha \frac{1}{k_i}\frac{k_j - k_i}{t_j - t_i} 
            + \beta \frac{1}{l_i}\frac{l_j - l_i}{t_j - t_i} 
            + (1 - \alpha - \beta) \frac{1}{e_i}\frac{e_j - e_i}{t_j - t_i} 
      \right] \;,
\end{equation}
% 
where $\lambda_{i,j}$ approximates the true, instantaneous value of $\lambda$
from Equation~\ref{eq:lambda} between times $i$ and $j$.

Interestingly, 
%
\begin{equation}
  \lambda_{1,n}
    \neq \sum_{i=2}^{n} \lambda_{i,i-1}  \;.
\end{equation}
% 
Rather,
%
\begin{equation} \label{eq:lambda1n}
  \lambda_{1,n}
    = \frac{1}{y_1}\frac{y_n - y_1}{t_n - t_1} 
      - \left[ 
            \alpha \frac{1}{k_1}\frac{k_n - k_1}{t_n - t_1} 
            + \beta \frac{1}{l_1}\frac{l_n - l_1}{t_n - t_1} 
            + (1 - \alpha - \beta) \frac{1}{e_1}\frac{e_n - e_1}{t_n - t_1} 
      \right] \; ,
\end{equation}
% 
which will become increasingly inaccurate over large time spans,
because $y_1$, $k_1$, $l_1$, and $e_1$ will be less representative
of the average value of $y$, $k$, $l$, and $e$ in the time span, respectively.
This suggests that an averaging approach such as
%
\begin{equation}
  \lambda_{1,n}
    = \frac{\sum_{i=2}^{n} \lambda_{i,i-1}}{n-1}
\end{equation}
%
is a better approximation of Equation~\ref{eq:lambda1n},
which is, itself, an approximation of the true value of $\lambda$ 
given by Equation~\ref{eq:lambda}.

The fraction~($f$) of GDP growth explained by the Solow residual
for the CD model
can be given as
%
\begin{equation} \label{eq:frac_growth_explained_by_solow}
  f \equiv \frac{\lambda}{\frac{1}{y}\frac{\mathrm{d}y}{\mathrm{d}t}} 
            = 1 - \frac{\alpha \frac{1}{k}\frac{\mathrm{d}k}{\mathrm{d}t} 
                      + \beta \frac{1}{l}\frac{\mathrm{d}l}{\mathrm{d}t} 
                      + (1 - \alpha - \beta) \frac{1}{e}\frac{\mathrm{d}e}{\mathrm{d}t}}
                  {\frac{1}{y}\frac{\mathrm{d}y}{\mathrm{d}t}} \; .
\end{equation}
%
Using forward differences to estimate $f$ yields
%
\begin{equation} \label{eq:f_finite_diff_1}
  f_{i,j} = \frac{\lambda_{i,j}}{\frac{1}{y_i}\frac{y_j - y_i}{t_j - t_i}}
       = 1 - \frac{\alpha \frac{1}{k_i}\frac{k_j - k_i}{t_j - t_i} 
            + \beta \frac{1}{l_i}\frac{l_j - l_i}{t_j - t_i} 
            + (1 - \alpha - \beta) \frac{1}{e_i}\frac{e_j - e_i}{t_j - t_i} }
            {\frac{1}{y_i}\frac{y_j - y_i}{t_j - t_i}} \; .
\end{equation}
%
Cancelling $t_j - t_i$ terms and simplifying gives
%
\begin{equation} \label{eq:f_finite_diff_2}
  f_{i,j} = 1 - \frac{\alpha \left( \frac{k_j}{k_i} - 1 \right) 
                      + \beta \left( \frac{l_j}{l_i} - 1 \right)
                      + (1 - \alpha - \beta) \left( \frac{e_j}{e_i} - 1 \right)}
            {\left( \frac{y_j}{y_i} - 1 \right)} \; ,
\end{equation}
%
which is an estimate of the instantaneous value of $f$ 
given by Equation~\ref{eq:frac_growth_explained_by_solow}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Derivation of dynamic Solow residual for the CES equation}
\label{sec:dynamic_sr_CES}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this section, we derive the dynamic Solow residual ($\lambda$) 
for the CES equation with three generic factors of production, namely
$x_1$, $x_2$, and $x_3$.

We can define
%
\begin{equation} \label{eq:a_def}
  a \equiv \delta b^{\rho/\rho_1} + (1-\delta) x_3^{-\rho}
\end{equation}
%
and
%
\begin{equation} \label{eq:b_def}
  b \equiv \delta_1 x_1^{-\rho_1} + (1 - \delta_1) x_2^{-\rho_1} \; ,
\end{equation}
%
such that Equation~\ref{eq:CESgeneric} can be restated as
%
\begin{equation} \label{eq:simple_y} 
  y = \gamma A a^{-1/\rho} \; .
\end{equation}
%
Taking the natural logarithm of Equation~\ref{eq:simple_y}
and realizing that $\gamma$ is not a function of time, we find
%
\begin{equation}
  \frac{1}{y} \frac{\mathrm{d} y}{\mathrm{d} t} = \frac{1}{A} \frac{\mathrm{d} A}{\mathrm{d} t} 
                      + \left( - \frac{1}{\rho} \right) \frac{1}{a} \frac{\mathrm{d} a}{\mathrm{d} t} \; .
\end{equation}
%
By Equation~\ref{eq:lambda_calc}, and after rearranging, 
the dynamic Solow residual for the CES equation can be stated as
%
\begin{equation} \label{eq:lambda_for_CES}
  \lambda = \frac{1}{y} \frac{\mathrm{d} y}{\mathrm{d} t} + \frac{1}{a} \frac{1}{\rho} \frac{\mathrm{d} a}{\mathrm{d} t} \; .
\end{equation}
%
To find $\frac{\mathrm{d} a}{\mathrm{d} t}$ and $\frac{\mathrm{d} b}{\mathrm{d} t}$, 
we take the time derivatives 
of Equations~\ref{eq:a_def} and~\ref{eq:b_def}
and rearrange slightly to obtain
%
\begin{equation} \label{eq:da_dt}
  \frac{1}{\rho} \frac{\mathrm{d} a}{\mathrm{d} t} = \delta b^{(\rho/\rho_1 - 1)} \frac{1}{\rho_1} \frac{\mathrm{d} b}{\mathrm{d} t}
                                      - (1-\delta) x_3^{-(\rho + 1)} \frac{\mathrm{d} x_3}{\mathrm{d} t}
\end{equation}
%
and
%
\begin{equation} \label{eq:db_dt}
  \frac{1}{\rho_1} \frac{\mathrm{d} b}{\mathrm{d} t} = - \delta_1 x_1^{-(\rho_1 + 1)} \frac{\mathrm{d} x_1}{\mathrm{d} t}
                                      - (1 - \delta_1) x_2^{-(\rho_1 + 1)} \frac{\mathrm{d} x_2}{\mathrm{d} t} \; .
\end{equation}
%
Equations~\ref{eq:lambda_for_CES}--\ref{eq:db_dt} can be used to 
calculate a dynamic time series for $\lambda$ given 
fitted parameters from the CES model
($\rho, \rho_1, \delta, \, \mathrm{and} \; \delta_1$)
and time series for economic output~($y$) and
factors of production 
($x_1, x_2, \, \mathrm{and} \; x_3$).
The time derivatives $\left( 
\frac{\mathrm{d} y}{\mathrm{d} t}, 
\frac{\mathrm{d} x_1}{\mathrm{d} t},
\frac{\mathrm{d} x_2}{\mathrm{d} t}, \, \mathrm{and} \;
\frac{\mathrm{d} x_3}{\mathrm{d} t} \right)$
can be approximated from historical data
by forward finite differences in a manner similar to Equation~\ref{eq:lambdaij}.

\end{document}

%%
%% End of file `elsarticle-template-2-harv.tex'.