### 2015-06-26 Articles/Article_JAE_2014/Results_Comparisons/Results_Comparison.Rmd
Comparison between Stan and Matt.

---
title: "Comparison of CD Fits"
author: "MK Heun"
date: "June 26, 2015"
output: pdf_document
---

This document compares fits performed by Matt and Stan using
ZA data with `iLfi` for labor.

```{r Setup, echo=FALSE, include=FALSE}
require(EconData)
require(EconModels)
require(dplyr)
```

First, set up the data and the formula.

```{r Data and formula}
form <- iGDP ~ iK + iLfi + iYear
zaData <- filter(SUN, Country=="ZA")
# Keeps only the variabls that are in the formula
zaData <- subset(zaData, select=c(all.vars(form), "Year"))
# Removes any years that do not have complete information 
zaData <- zaData[complete.cases(zaData), ]
# Tells how many rows are available.
nrow(zaData)
head(zaData)
tail(zaData)
```

Here are Stan's results as reported in his email to me on 26 June 2015:

```{r Stan\'s results}
alpha_1 <- 0.0533081
alpha_2 <- 0.9466919
alpha_3 <- 0
logscale <- -0.0753839
lambda <- 0.0137483
# Calculate sse in log space using all available data
zaData <- within(zaData, {
  residual <- log(iGDP) - (logscale + lambda*iYear + alpha_1*log(iK) + alpha_2*log(iLfi)) 
})
sse_stan <- sum(zaData$residual^2)
sse_stan
```

The above result does not match the number reported by Stan
during the telecon on 26 June 2015 (0.168).

One question arises immediately:
which years are being fitted?
There are `r nrow(zaData)` rows of data
where complete cases exist.
Yet, Stan's email indicated that he is fitting 41 observations, 1973--2013.
I suspect that Stan is trimming the rows
where `iYear < 0`.
So, let's make another data set where the
rows where `iYear < 0` are trimmed
and re-calculate `sse`.

```{r Stan\'s results, trimmed data}
zaData_trimmed <- zaData[zaData$iYear >= 0, ]
sse_stan_trimmed <- sum(zaData_trimmed$residual^2)
sse_stan_trimmed
```

For a correct comparison, I'll also fit using the 1973--2013 data only.

```{r Matt\'s results trimmed}
mod_trimmed <- cdModel(form, data = zaData_trimmed, constrained = TRUE)
summary(mod_trimmed)
naturalCoef(mod_trimmed)
sum(resid(mod_trimmed)^2)
```

If we use the \emph{exact same data}, specifically 1973--2013, we obtain exactly the same results.

For completeness, here are Matt's results when all available years are fitted:

```{r Matt\'s results}
mod <- cdModel(form, data = zaData, constrained = TRUE)
summary(mod)
sum(resid(mod)^2)
```

This number (`r sum(resid(mod)^2)`) is what Matt reported
during the telecon on 26 June 2015.

It is unsurprising that Stan (`r sse_stan_trimmed`)
reported a lower `sse` value than Matt (`r sum(resid(mod)^2)`)
during the telecon of 26 June 2015,
because Stan was using fewer observations!