### 2016-08-08 NSF-ClusterGrant/EconOnePager.Rmd
ieee formatting, 3rd person

---
title: "Econ Modeling One-pager"
author: "R Pruim"
date: "August 1, 2016"
output:
  pdf_document: 
    keep_tex: yes
  html_document: default
  word_document: default
bibliography: OnePager.bib
csl: ieee.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Development of energy and economic policy is often informed 
by energy-economic modeling via aggregate production functions (APFs) fitted
to historical data in a theory-to-policy process utilized 
in the literature [@Kummel:1985vz, @Turner:2012], 
in government budget offices [@Congressional-Budget-Office:2001], and
at the World Bank [@Bussolo:2010].
Despite more than 60 years of criticism
[@robinson1953production, @shaikh1974laws, @felipe2003aggregation], 
the use of APFs continues and appears to be expanding, 
with APFs moving from academic study
[@solow1957technical, @easterly1995soviet] 
to macroeconomic modeling that informs energy and economic policy
[@szeto2001econometric, @pauw2003functional, @sancho2009calibration].
Empirical applications of APFs
are numerous, including analysis of 
(a) technical change [@jorgenson1967explanation], 
(b) substitution elasticities [@kemfert2000energy],
(c) inter-country economic output 
	and structure [@easterly1995soviet, @toh1996differential], and 
(d) CO$_2$ emissions policy [@van2008production].
One reason for the enduring allure of APFs is
their promise to describe the intricate, complex 
workings of entire economies by simple, aggregated measures
of inputs (capital, labor, and sometimes energy) 
and output (typically, Gross Domestic Product, GDP).

The theory-to-policy process is not merely academic: 
the policies it generates affect economic development and 
quality of life for people worldwide.
Development of energy policy is often informed by economic considerations
via aggregate production functions (APFs) as part of a theory-to-policy
process comprised of six steps:

 1. selecting a theoretical energy-economy framework,
 2. formulating modeling approaches to be considered,
 3. econometrically fitting an APF to historical data,
 4. comparing and evaluating modeling approaches considered,
 5. interpreting the economy, and
 6. formulating economic policy.

Choices made in Steps 1--4
can lead to very different interpretations of the economy in Step 5
and subequently to the justification for different policies in Step 6.
In a recently submitted paper [@Heun:2016], Drs. Heun and Pruim and co-authors 
investigated these effects
by analyzing data from two countries (Portugal and the United Kingdom) 
using the Constant Elasticity of Substitution (CES) APF (which subsumes
the familiar Cobb-Douglas models as a special case)
to evaluate four modeling choices:
(a) whether or not to simplify models by employing a certain cost-share principle 
that is often assumed in economic modeling,
(b) whether or not to include energy as a factor of production along with capital and labor,
(c) whether or not to make certain quality adjustments to factors of production in an
attempt to better quantify their impact on the economy, and 
(d) which CES nesting structure to use for models with three factors of production.

This paper includes a discussion of two revealing examples for which different
upstream modeling choices lead to very different policies. In the first example,
the $(kl)e$ nesting structure implies significant investment in energy, while
other nesting structures suggest otherwise. In the second example, unadjusted
factors of production suggest balanced investment in labor and energy, while 
quality-adjusting suggests significant investment in labor over energy. 
Divergent outcomes provide cautionary tales for policymakers and demand
greater understanding of upstream modeling choices and 
their downstream implications.

One challenge in this work is the computational difficulty involved in fitting CES models 
and quantifying the precision of the estimated parameters that result.  Because of 
numerical challenges, we fit each model with multiple non-linear optimization algorithms,
and because of boundary issues, we explicity refit models along boundaries.  Then, because
parameter precision is challenging to estimate using standard methods, we employ 
computationally intensive resampling and/or Bayesian models.  Our current exploration
is a sort of pilot study considering a limited number of countries, a limited number of 
modeling choices and reducing the number of resamples to an amount that can be readily handled 
on our current hardware in a few hours.  Further work in this area, especially any simulations
to explore the quality of the bootstrap resampling or Bayesian estimation, would require days 
to run currently.
Futhermore, we already have data from several additional countries that we would like to
anaylze by similar methods.

**Additional paragraph here** (when Matt returns to radio contact) about future related 
integrating the paper described above with cointegration work that Tiago has done.  [Tiago is
a possible letter writer, so this makes a nice connection between the grant proposal and his
letter, and also demonstrates its usefulness for things being done beyond Calvin.]

# References