### 2017-01-19 NSF-ClusterGrant/EconOnePager.Rmd

---
title: "Econ Modeling One-pager"
author: "R Pruim"
date: "August 1, 2016"
output:
  html_document: default
  pdf_document:
    keep_tex: yes
  word_document: default
csl: ieee.csl
bibliography: OnePager.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Development of energy and economic policy is often informed 
by energy-economic modeling via aggregate production functions (APFs) fitted
to historical data in a theory-to-policy process utilized 
in the literature [@Kummel:1985vz; @Turner:2012], 
in government budget offices [@Congressional-Budget-Office:2001], and
at the World Bank [@Bussolo:2010].
Applications of APFs are numerous, including analysis of 
(a) technical change [@jorgenson1967explanation], 
(b) substitution elasticities [@kemfert2000energy],
(c) inter-country economic output and structure [@easterly1995soviet; @toh1996differential], and 
(d) CO$_2$ emissions policy [@van2008production].
One reason for the enduring allure of APFs is
their promise to describe the intricate, complex 
workings of entire economies by simple, aggregated measures
of inputs (capital, labor, and sometimes energy) 
and output (typically, Gross Domestic Product, GDP).

Development of energy policy is often informed by economic considerations
via aggregate production functions (APFs) as part of a theory-to-policy
process comprised of six steps:
1. selecting a theoretical energy-economy framework;
2. formulating modeling approaches to be considered;
3. econometrically fitting an APF to historical data;
4. comparing and evaluating modeling approaches considered;
5. interpreting the economy; and
6. formulating economic policy.

Choices made in Steps 1--4
can lead to very different interpretations of the economy in Step 5
and subequently to the justification for different policies in Step 6.
In a recently submitted paper [@Heun:2016], Drs.\ Heun and Pruim and co-authors 
investigated these effects
by analyzing data from two countries (Portugal and the United Kingdom) 
using the Constant Elasticity of Substitution (CES) APF (which subsumes
the familiar Cobb-Douglas models as a special case)
to evaluate four modeling choices:
(a) whether or not to simplify models by employing a certain cost-share principle 
that is often assumed in economic modeling,
(b) whether or not to include energy as a factor of production along with capital and labor,
(c) whether or not to make certain quality adjustments to factors of production in an
attempt to better quantify their impact on the economy, and 
(d) which CES nesting structure to use for models with three factors of production.
[@Heun:2016] includes a discussion of two revealing examples for which different
upstream modeling choices lead to very different policies. 
*Divergent outcomes provide cautionary tales for policymakers and demand
greater understanding of upstream modeling choices and 
their downstream implications.*

One challenge in this work is the computational difficulty involved in fitting CES models 
and quantifying the precision of the estimated parameters that result.  Because of 
numerical challenges, we fit each model with multiple non-linear optimization algorithms,
and because of boundary issues, we explicity refit models along boundaries.  Then, because
estimating parameter precision is challenging using standard methods, we employ 
computationally intensive resampling and/or Bayesian models.  *Our current exploration
forms a pilot study considering a limited number of countries, a limited number of 
modeling choices and reducing the number of resamples to an amount that can be readily 
handled on our current hardware in a few hours.*

Plans for future work will be even more computationally intensive than the work
done to date.
*An enhanced cluster computing capability at Calvin College will 
enable expansion of this research into several new areas
and enable collaboration with international colleagues.*

 1. Drs. Heun and Pruim have in hand data from several additional countries that they 
 would like to anaylze by similar methods as discussed above.

 2. Collaborators in Portugal have completed a paper that employs the 
Cobb-Douglas APF in combination with statistical co-integration techniques 
to understand the relationship between energy and the economy in 
Portugal [@jsantos2016].  
At the Second Exergy Economics Workshop (Sussex, England, July 2016), 
plans were made to collaborate with them to generalize this to the CES APF
using techniques from [@Heun:2016].

 3. The use of bootstrapping and Bayesian methods to esimate the precision 
 of CES parameters is new, and a computationally intensive simulation is planned
 to investigate the quality and performance of the these methods in this setting.

# References