### 2017-01-19 Articles/Brockway_Landscape_2017/Indexing_Demo.Rnw
Minor changes.

\documentclass{article}

\begin{document}

\title{Demonstration of Effect of Indexing on\\
Cobb-Douglas and CES\\
Aggregate Production Functions}
\author{Matthew Kuperus Heun}
\maketitle

<<setup, include=FALSE>>=
library(EconModels)
library(dplyr)
library(tidyr)
library(ggplot2)

opts_chunk$set(
  # ********************************************************
  eval = TRUE,
  include = TRUE,
  echo = TRUE
  # ********************************************************
)
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Models} 
\label{sec:models}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For demonstration purposes, 
we focus on models that assume capital and labor factors of production only.

%++++++++++++++++++++++++++++++
\subsection{Raw historical data} 
\label{sec:raw}
%++++++++++++++++++++++++++++++

The Cobb-Douglas model for using raw historical data is given by 
%
\begin{equation} \label{eq:CD_model_nonindexed}
  Y = \theta \: \mathrm{e}^{\lambda t} \: K^{\alpha_k} \, L^{\alpha_l} \; ,
\end{equation}
%
where $Y$, $K$, and $L$ are the raw (non-indexed) time series.
Time ($t$) is indexed by difference to the intial year (1960).

The CES model with raw historical data is given by
%
\begin{equation} \label{eq:CESkl_nonindexed}
  Y = \gamma \: \mathrm{e}^{\lambda t} \left[\delta K^{-\rho}
      + (1-\delta)L^{-\rho} \right]^{-1/\rho} \; ,
\end{equation}
%
where $\rho$ is related to the elasticity of substitution~($\sigma$) by 
%
\begin{equation} \label{eq:sigma}
  \sigma = \frac{1}{1 + \rho} \; .
\end{equation}

%++++++++++++++++++++++++++++++
\subsection{Indexed factors of production} 
\label{sec:indexed}
%++++++++++++++++++++++++++++++

The Cobb-Douglas model with indexed historical data is given by
%
\begin{equation} \label{eq:CD_model}
  y = \theta \: \mathrm{e}^{\lambda t} \: k^{\alpha_k} \, l^{\alpha_l} \; ,
\end{equation}
%
where $y$, $k$, and $l$ are indexed by ratio to the initial year (1960). 
For example,
%
\begin{equation} \label{eq:indexing}
  y = Y/Y_{1960} \; .
\end{equation}

The CES model with indexed historical data is given by 
%
\begin{equation} \label{eq:CESkl}
  y = \gamma \: \mathrm{e}^{\lambda t} \left[\delta k^{-\rho}
      + (1-\delta)l^{-\rho} \right]^{-1/\rho} \; .
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data} 
\label{sec:data}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

All historical data are from the Penn World Table.
Figure~\ref{fig:GDP-graph} below shows historical data as dots.

<<load-data>>=
UKdata <- read.csv(file.path("data", "IndexingDemoRawData.csv"))
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Fitting} 
\label{sec:fitting}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We fit both the Cobb-Douglas and CES aggregate production functions 
to both raw and indexed data.
Figure~\ref{fig:GDP-graph} below shows fitted GDP as lines.

<<fitting, warning=FALSE>>=
cd_nonindexed <- cdModel(GDP ~ K + L + iYear, data = UKdata)
cd_indexed <- cdModel(iGDP ~ iK + iL + iYear, data = UKdata)
ces_nonindexed <- cesModel(GDP ~ K + L + iYear, data = UKdata)
ces_indexed <- cesModel(iGDP ~ iK + iL + iYear, data = UKdata)
@

<<quadplot-data-frame, echo = FALSE>>=
quadplotdata <- UKdata %>% 
  mutate(
    yhat_cd_nonindexed = yhat(cd_nonindexed),
    yhat_cd_indexed = yhat(cd_indexed),
    yhat_ces_nonindexed = yhat(ces_nonindexed),
    yhat_ces_indexed = yhat(ces_indexed),
    Country = NULL,
    K = NULL,
    iK = NULL,
    L = NULL,
    iL = NULL
  )
# Eliminate attributes to avoid warning.
attr(quadplotdata$yhat_cd_nonindexed, "label") <- NULL
attr(quadplotdata$yhat_cd_indexed, "label") <- NULL
# Prepare format for quad plot
quadplotdata <- quadplotdata %>% 
  gather(key = variable, value = y_hat, 
         yhat_cd_nonindexed, yhat_cd_indexed, 
         yhat_ces_nonindexed, yhat_ces_indexed) %>% 
  mutate(
    y_historical = 
      case_when(
        .$variable %in% c("yhat_cd_nonindexed", "yhat_ces_nonindexed") ~ .$GDP,
        .$variable %in% c("yhat_cd_indexed", "yhat_ces_indexed")       ~ .$iGDP
      ),
    GDP = NULL, 
    iGDP = NULL,
    indexed = case_when(
        .$variable %in% c("yhat_cd_nonindexed", "yhat_ces_nonindexed") ~ "raw",
        .$variable %in% c("yhat_cd_indexed", "yhat_ces_indexed")       ~ "indexed"
    ),
    indexed = factor(indexed, c("raw", "indexed")),
    apf = case_when(
        .$variable %in% c("yhat_cd_nonindexed", "yhat_cd_indexed")     ~ "CD",
        .$variable %in% c("yhat_ces_nonindexed", "yhat_ces_indexed")   ~ "CES"
    )
  )
@

Figure~\ref{fig:GDP-graph} compares each type of fitted model (lines) with historical data (dots).
Note that all models provide excellent fits.

<<GDP-graph, fig.align="center", fig.width=6.5, fig.height=5, fig.cap="Fits to historical data. Dots are historical data. Lines are fits.">>=
ggplot(data = quadplotdata, mapping = aes_string(x = "Year")) + 
  geom_point(mapping = aes_string(y = "y_historical"), 
             size = 0.2) + 
  geom_line(mapping = aes_string(y = "y_hat")) + 
  facet_grid(indexed ~ apf, scales = "free") + 
  xlab(NULL) +
  ylab("GDP [raw: millions 2005$, indexed: 1960 = 1]") +
  xy_theme()
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Coefficients} 
\label{sec:coefficients}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%++++++++++++++++++++++++++++++
\subsection{Cobb-Douglas} 
\label{sec:cd_coeffs}
%++++++++++++++++++++++++++++++

The coefficients for Cobb-Douglas models are shown below.
<<cd-coeffs>>=
cd_nonindexed
cd_indexed
@

Note that output elasticity
\verb+alpha_1+ ($\alpha_k$) 
and 
Solow residual growth rate
\verb+lambda+
are identical,
regardless of whether the historical data are indexed or raw.

The pre-multiplier (in this case \texttt{logscale})
accounts for the fact that historical data have been indexed.
We can convert \texttt{logscale} to $\theta$ as follows:

<<cd-logscale-to-theta>>=
exp(cd_nonindexed$coefficients[["logscale"]])
exp(cd_indexed$coefficients[["logscale"]])
@

Note that $\theta$ for the model fit to indexed data is near 1.0, as expected.

%++++++++++++++++++++++++++++++
\subsection{CES}
\label{sec:ces_coeffs}
%++++++++++++++++++++++++++++++

The following shows parameters for CES models
fitted to raw and indexed historical data.

<<ces-coeffs>>=
ces_nonindexed
ces_indexed
@

Similar to the Cobb-Douglas model,
the pre-multiplier term
(in this case \verb+gamma+)
is different depending upon whether the historical data were indexed or raw.
As expected, and similar to the Cobb-Douglas model,
the growth rate of the Solow residual
(\verb+gamma+)
is identical, regardless of whether the historical data are indexed or raw.
We note that the value of \verb+gamma+ is close to 1.0,
when fitting to indexed data,
as expected.
Furthermore, the elasticity of substitution exponents
(\verb+rho+)
are identical regardless of whether the historical data are indexed or not.

However, the share parameter
(\verb+delta+)
varies significantly, depending upon whether 
the historical data are indexed.
Any inferences about the economy that depend upon the share parameter (\verb+delta+)
will be significantly different depending upon whether the 
historical data are indexed prior to fitting.

\end{document}