### 2017-01-25 Articles/Brockway_Landscape_2017/Indexing_Demo.Rnw
Updates to indexing demo.

\documentclass{article}

\begin{document}

\title{Demonstration of Effect of Indexing on\\
Cobb-Douglas (C-D) and 
Constant Elasticity of Substitution (CES)\\
Aggregate Production Functions (APFs)}
\author{Matthew Kuperus Heun}
\maketitle

<<setup, include=FALSE>>=
library(EconModels)
library(dplyr)
library(tidyr)
library(ggplot2)

opts_chunk$set(
  # ********************************************************
  eval = TRUE,
  include = TRUE,
  echo = TRUE
  # ********************************************************
)
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Models} 
\label{sec:models}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For demonstration purposes, 
we focus on models that assume capital and labor factors of production only.

The C-D model for using raw historical data is given by 
%
\begin{equation} \label{eq:CD_model_nonindexed}
  Y = \theta \: \mathrm{e}^{\lambda t} \: K^{\alpha_k} \, L^{\alpha_l} \; ,
\end{equation}
%
where $Y$, $K$, and $L$ are the raw (non-indexed) time series.
Time ($t$) is indexed by difference to the intial year (1960).

The Constant Elasticity of Substitution~(CES) model with raw historical data is given by
%
\begin{equation} \label{eq:CESkl_nonindexed}
  Y = \gamma \: \mathrm{e}^{\lambda t} \left[\delta K^{-\rho}
      + (1-\delta)L^{-\rho} \right]^{-1/\rho} \; ,
\end{equation}
%
where $\rho$ is related to the elasticity of substitution~($\sigma$) by 
%
\begin{equation} \label{eq:sigma}
  \sigma = \frac{1}{1 + \rho} \; .
\end{equation}

The C-D model with indexed historical data is given by
%
\begin{equation} \label{eq:CD_model}
  y = \theta \: \mathrm{e}^{\lambda t} \: k^{\alpha_k} \, l^{\alpha_l} \; ,
\end{equation}
%
where $y$, $k$, and $l$ are indexed by ratio to the initial year (1960). 
For example,
%
\begin{equation} \label{eq:indexing}
  y = Y/Y_{1960} \; .
\end{equation}

The CES model with indexed historical data is given by 
%
\begin{equation} \label{eq:CESkl}
  y = \gamma \: \mathrm{e}^{\lambda t} \left[\delta k^{-\rho}
      + (1-\delta)l^{-\rho} \right]^{-1/\rho} \; .
\end{equation}

When a third factor of production (in this case energy)
is added to the C-D APF with raw historical data,
we obtain
%
\begin{equation} \label{eq:CDe_model_raw}
  Y = \theta \: \mathrm{e}^{\lambda t} \: K^{\alpha_k} \, L^{\alpha_l} \, E^{\alpha_e} \; ,
\end{equation}
%
where $E$ is total primary energy supply.
For the CES APF with raw historical data, we obtain
%
\begin{equation} \label{eq:CES3_raw}
  Y = \theta \: \mathrm{e}^{\lambda t} \, \left\{\delta \left[\delta_1 K^{-\rho_1}
      + (1-\delta_1) L^{-\rho_1} \right]^{\rho/\rho_1}
      + (1-\delta) E^{-\rho} \right\}^{-1/\rho} \; .
\end{equation}
%
For indexed historical data, we define
$e \equiv E / E_{1960}$ to obtain
%
\begin{equation} \label{eq:CDe_model_indexed}
  y = \theta \: \mathrm{e}^{\lambda t} \: k^{\alpha_k} \, l^{\alpha_l} \, e^{\alpha_e} \; ,
\end{equation}
%
and
%
\begin{equation} \label{eq:CES3_indexed}
  y = \theta \: \mathrm{e}^{\lambda t} \, \left\{\delta \left[\delta_1 k^{-\rho_1}
      + (1-\delta_1) l^{-\rho_1} \right]^{\rho/\rho_1}
      + (1-\delta) e^{-\rho} \right\}^{-1/\rho} \; .
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data} 
\label{sec:data}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Historical data for $Y$, $K$, and $L$ are from the Penn World Table.
Historical data for $E$ are from the International Energy Agency
time series for total primary energy supply.
Figure~\ref{fig:GDP-graph} below shows historical data for $Y$ and $y$ as dots.

<<load-data>>=
UKdata <- read.csv(file.path("data", "IndexingDemoRawData.csv"))
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Fitting} 
\label{sec:fitting}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We fit both the C-D and CES aggregate production functions (APFs)
both with and without energy to both raw and indexed data.
Figure~\ref{fig:GDP-graph} below shows fitted GDP as lines.

<<fitting, warning=FALSE>>=
cd_nonindexed <- cdModel(GDP ~ K + L + iYear, data = UKdata)
cd_indexed <- cdModel(iGDP ~ iK + iL + iYear, data = UKdata)
ces_nonindexed <- cesModel(GDP ~ K + L + iYear, data = UKdata)
ces_indexed <- cesModel(iGDP ~ iK + iL + iYear, data = UKdata)
cd_nonindexed_e <- cdModel(GDP ~ K + L + E + iYear, data = UKdata)
cd_indexed_e <- cdModel(iGDP ~ iK + iL + iE + iYear, data = UKdata)
ces_nonindexed_e <- cesModel(GDP ~ K + L + E + iYear, data = UKdata)
ces_indexed_e <- cesModel(iGDP ~ iK + iL + iE + iYear, data = UKdata)
@

<<quadplot-data-frame, echo = FALSE>>=
quadplotdata <- UKdata %>% 
  mutate(
    yhat_cd_nonindexed = yhat(cd_nonindexed),
    yhat_cd_indexed = yhat(cd_indexed),
    yhat_ces_nonindexed = yhat(ces_nonindexed),
    yhat_ces_indexed = yhat(ces_indexed),
    yhat_cd_nonindexed_e = yhat(cd_nonindexed_e),
    yhat_cd_indexed_e = yhat(cd_indexed_e),
    yhat_ces_nonindexed_e = yhat(ces_nonindexed_e),
    yhat_ces_indexed_e = yhat(ces_indexed_e),
    Country = NULL,
    K = NULL, iK = NULL, L = NULL, iL = NULL, E = NULL, iE = NULL
  )
# Eliminate attributes to avoid warning.
attr(quadplotdata$yhat_cd_nonindexed, "label") <- NULL
attr(quadplotdata$yhat_cd_indexed, "label") <- NULL
attr(quadplotdata$yhat_cd_nonindexed_e, "label") <- NULL
attr(quadplotdata$yhat_cd_indexed_e, "label") <- NULL
# Prepare format for quad plot
quadplotdata <- quadplotdata %>% 
  gather(key = variable, value = y_hat, 
         yhat_cd_nonindexed, yhat_cd_indexed, 
         yhat_ces_nonindexed, yhat_ces_indexed,
         yhat_cd_nonindexed_e, yhat_cd_indexed_e,
         yhat_ces_nonindexed_e, yhat_ces_indexed_e
  ) %>% 
  mutate(
    y_historical = 
      case_when(
        .$variable %in% c("yhat_cd_nonindexed", "yhat_ces_nonindexed",
                          "yhat_cd_nonindexed_e", "yhat_ces_nonindexed_e") ~ .$GDP,
        .$variable %in% c("yhat_cd_indexed", "yhat_ces_indexed",
                          "yhat_cd_indexed_e", "yhat_ces_indexed_e")       ~ .$iGDP
      ),
    GDP = NULL, 
    iGDP = NULL,
    indexed = case_when(
      .$variable %in% c("yhat_cd_nonindexed", "yhat_ces_nonindexed",
                        "yhat_cd_nonindexed_e", "yhat_ces_nonindexed_e") ~ "raw",
      .$variable %in% c("yhat_cd_indexed", "yhat_ces_indexed",
                        "yhat_cd_indexed_e", "yhat_ces_indexed_e")       ~ "indexed"
    ),
    indexed = factor(indexed, c("raw", "indexed")),
    apf = case_when(
      .$variable %in% c("yhat_cd_nonindexed", "yhat_cd_indexed",
                        "yhat_cd_nonindexed_e", "yhat_cd_indexed_e")     ~ "CD",
      .$variable %in% c("yhat_ces_nonindexed", "yhat_ces_indexed",
                        "yhat_ces_nonindexed_e", "yhat_ces_indexed_e")   ~ "CES"
    ),
    energy = case_when(
      .$variable %in% c("yhat_cd_nonindexed", "yhat_cd_indexed",
                        "yhat_ces_nonindexed", "yhat_ces_indexed")       ~ "w/o energy",
      .$variable %in% c("yhat_cd_nonindexed_e", "yhat_cd_indexed_e",
                        "yhat_ces_nonindexed_e", "yhat_ces_indexed_e")   ~ "w/ energy"
    ),
    energy = factor(energy, c("w/o energy", "w/ energy"))
  )
@

Figures~\ref{fig:GDP-graph} and~\ref{fig:GDP-graph-e} 
compare each type of fitted model (lines) with historical data (dots).
Note that all models provide excellent fits.

<<GDP-graph, fig.align="center", fig.width=6.5, fig.height=5, fig.cap="Fits to historical data for models that exclude energy. Dots are historical data. Lines are fits.">>=
ggplot(data = quadplotdata %>% filter(energy == "w/o energy"), 
       mapping = aes_string(x = "Year")) + 
  geom_point(mapping = aes_string(y = "y_historical"), 
             size = 0.2) + 
  geom_line(mapping = aes_string(y = "y_hat")) + 
  facet_grid(indexed ~ apf, scales = "free") + 
  xlab(NULL) +
  ylab("GDP [raw: millions 2005$, indexed: 1960 = 1]") +
  xy_theme()
@

<<GDP-graph-e, fig.align="center", fig.width=6.5, fig.height=5, fig.cap="Fits to historical data for models that include energy. Dots are historical data. Lines are fits.">>=
ggplot(data = quadplotdata %>% filter(energy == "w/ energy"), 
       mapping = aes_string(x = "Year")) + 
  geom_point(mapping = aes_string(y = "y_historical"), 
             size = 0.2) + 
  geom_line(mapping = aes_string(y = "y_hat")) + 
  facet_grid(indexed ~ apf, scales = "free") + 
  xlab(NULL) +
  ylab("GDP [raw: millions 2005$, indexed: 1960 = 1]") +
  xy_theme()
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Coefficients} 
\label{sec:coefficients}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%++++++++++++++++++++++++++++++
\subsection{Without energy} 
\label{sec:coefficients_without_energy}
%++++++++++++++++++++++++++++++

%------------------------------
\subsubsection{Cobb-Douglas} 
\label{sec:cd_coeffs}
%------------------------------

The coefficients for C-D models are shown below.
<<cd-coeffs>>=
summary(cd_nonindexed)
summary(cd_indexed)
@

Note that output elasticity
\verb+alpha_1+ ($\alpha_k$) 
and 
Solow residual growth rate
\verb+lambda+
are identical,
regardless of whether the historical data are indexed or raw.

The pre-multiplier (in this case \texttt{logscale})
accounts for the fact that historical data have been indexed.
We can convert \texttt{logscale} to $\theta$ as follows:

<<cd-logscale-to-theta>>=
exp(coef(cd_nonindexed)[["logscale"]])
exp(coef(cd_indexed)[["logscale"]])
@

Note that $\theta$ for the model fit to indexed data is near 1.0, as expected.

%------------------------------
\subsubsection{CES}
\label{sec:ces_coeffs}
%------------------------------

The following shows parameters for CES models
fitted to raw and indexed historical data.

<<ces-coeffs>>=
summary(ces_nonindexed)
summary(ces_indexed)
@

Similar to the C-D model,
the pre-multiplier term
(in this case \verb+gamma+)
is different depending upon whether the historical data were indexed or raw.
We note that the value of \verb+gamma+ is close to 1.0,
when fitting to indexed data,
as expected.
Also as expected, and similar to the C-D model,
the growth rate of the Solow residual
(\verb+lambda+)
is identical, regardless of whether the historical data are indexed or raw.
Furthermore, the elasticity of substitution exponents
(\verb+rho+)
are identical regardless of whether the historical data are indexed or not.

However, the share parameter
(\verb+delta+)
varies significantly, depending upon whether 
the historical data are indexed.
Any interpretation of the economy that relies upon the share parameter (\verb+delta+)
will be significantly different depending upon whether the 
CES function is fitted to raw or indexed data.

%++++++++++++++++++++++++++++++
\subsection{With energy} 
\label{sec:coefficients_with_energy}
%++++++++++++++++++++++++++++++

%------------------------------
\subsubsection{Cobb-Douglas with energy} 
\label{sec:cd_coeffs_e}
%------------------------------

The coefficients for C-D models with energy are shown below.
<<cd-coeffs-e>>=
summary(cd_nonindexed_e)
summary(cd_indexed_e)
@

Again, output elasticities
\verb+alpha_1+ ($\alpha_k$) 
\verb+alpha_2+ ($\alpha_l$) 
and 
Solow residual growth rate
\verb+lambda+
are identical,
regardless of whether the historical data are indexed or raw.

The \texttt{logscale} pre-multiplier
is converted to $\theta$ below:

<<cd-logscale-to-theta-e>>=
exp(coef(cd_nonindexed_e)[["logscale"]])
exp(coef(cd_indexed_e)[["logscale"]])
@

Note that $\theta$ for the model fit to indexed data is near 1.0, as expected.

%------------------------------
\subsubsection{CES}
\label{sec:ces_coeffs_e}
%------------------------------

The following shows parameters for CES models with energy
fitted to raw and indexed historical data.

<<ces-coeffs-e>>=
summary(ces_nonindexed_e)
summary(ces_indexed_e)
@

Note that the value of \verb+delta+ is shown as 1.0 for both models.
However, the value is not exactly 1.0:

<<ces-e-delta>>=
1 - coef(ces_nonindexed_e)[["delta"]]
1 - coef(ces_indexed_e)[["delta"]]
@

These values are close enough to exactly 1 to make 
energy ($E$ and $e$) the only important factor of production in the fitting process.
As such, the values of \verb+rho_1+ and \verb+rho+
make little difference to the computed value of SSE, regardless of the values of the other parameters.
Thus, the value of \verb+rho+ doesn't move from its initial guess (1).
These effects are reflected in the non-convergence of the indexed model.
It is surprising that the non-indexed model reports convergence, because
the indexed model (which reported non-convergence) has a better fit than 
the indexed model (which reported convergence).
All in all, I think 
this is not a very good example 
on which to demonstrate the effect of indexing historical data
with three factors of production.

That said, and similar to the CES model without energy,
the estimate of pre-multiplier term
(in this case \verb+gamma+)
is different depending upon whether the historical data were indexed or raw.
We note that the value of \verb+gamma+ is close to 1.0,
when fitting to indexed data,
as expected.
The growth rate of the Solow residual
(\verb+lambda+)
is slightly different, depending upon whether the historical data are indexed or raw,
but that difference could be caused by the fact that the parameters are not determined 
very precisely.
The elasticity of substitution exponents
(\verb+rho_1+ and \verb+rho+)
are similar, but not identical, when we move from raw to indexed historical data.
Again, this could be a reflection of the fact that these parameters are not determined very precisely.

Similar to the CES model with two factors of production, 
the share parameter (\verb+delta+)
varies significantly, depending upon whether 
the historical data are indexed.
And again, 
any interpretation of the economy that relies upon the share parameter (\verb+delta+)
will be significantly different depending upon whether the 
CES function is fitted to raw or indexed data.

%++++++++++++++++++++++++++++++
\subsection{With generated data} 
\label{sec:generated_data}
%++++++++++++++++++++++++++++++

We can try again with generated data, 
hoping to avoid the messiness that accompanies 
raw data that is many orders of magnitude different from 
indexed data.

<<gen-data>>=
# First, assume some nested 3-factor CES coefficients
sigma_1 <- 0.4
sigma <- 1.2
coef_indexed_gen <- c(gamma = 1, lambda = 0.021, 
                      delta_1 = 0.4, delta = 0.6, 
                      rho_1 = 1/sigma_1 - 1, rho = 1/sigma - 1,
                      nu = 1)
# Calculate what output would have been with those coefficients
y_gen <- cesCalc(xNames = c("iK", "iL", "iE"), data = UKdata, 
                 coef = coef_indexed_gen, tName = "iYear", 
                 nested = TRUE)
# Put this generated data into the data frame 
# that we'll use for fitting.
UKwithGenData <- UKdata %>% 
  mutate(
    # Add the generated GDP values to our data frame
    y_gen = y_gen,
    # But also add some generated historical data.
    # Perturb by rather small values to make fitting easier.
    Y_gen = y_gen * 1.01,
    K_gen = iK * 1.02,
    L_gen = iL * 1.03,
    E_gen = iE * 1.04
  )

# Fit to the generated GDP time series using indexed data. 
ces_indexed_e_gen <- cesModel(y_gen ~ iK + iL + iE + iYear, 
                              data = UKwithGenData)
summary(ces_indexed_e_gen)
# Compare with our assumed coefficients.  
# We should get exact same coefficients back, and we do!
coef_indexed_gen
coef(ces_indexed_e_gen)
# Fit to generated raw data.
ces_raw_e_gen <- cesModel(Y_gen ~ K_gen + L_gen + E_gen + iYear, 
                          data = UKwithGenData)
summary(ces_raw_e_gen)
# Compare.
coef_indexed_gen
coef(ces_raw_e_gen)
@

We find that the values for \verb+lambda+
are identical when comparing fits to indexed and raw historical data (in this case, generated).
Values for \verb+gamma+ are different, as expected.
Values for \verb+delta_1+ and \verb+delta+ are different, as expected.
The values for \verb+rho_1+ (in the inner next)
and \verb+rho_1+ are identical
across models fitted to raw and indexed historical data.

To summarize, 
when moving from indexed historical data to raw historical data,
CES parameters have the following behavio(u)r:
%
\begin{itemize}

  \item $\gamma$: changes

  \item $\lambda$: stays same

  \item $\delta_1$, $\delta$: change

  \item $\rho_1$, $\rho$: stay same

\end{itemize}

\end{document}