### 2018-07-10 Articles/Court_Analysis_2018/FittingForCourt2018.Rmd
Fitting information for Victor Court.

---
title: "Analysis for Victor Court"
author: "Matthew Kuperus Heun"
date: "10 July 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analysis for Victor Court

Victor has asked me to use my CES fitting functions on his US data
as a point of comparison.

```{r, include=FALSE}
library(dplyr)
library(EconModels)
library(ggplot2)
library(magrittr)
library(tidyr)
```

## Load and sanitize data

First, load the data.

```{r}
USData_raw <- read.csv(file = "CES_USA.csv", sep = ";")
head(USData_raw, n = 3)
```

Next, sanitize the data a bit
by removing non-indexed columns and
renaming the columns.
Lower-case variable names indicate indexing to 1 in 1960.

```{r}
USData <- USData_raw %>% 
  select(-K.L...US.2010.per.hour., 
         -Capital...US.2010., 
         -Labor..millions.of.hours., 
         -Primary.exergy..TJ.,
         -GDP....US.2010.) %>% 
  rename(
    k = K..1960...1.,
    l = L..1960...1.,
    e = E..1960...1.,
    y = Y..1960...1.
  ) %>% 
  mutate(
    Country = "US"
  ) %>% 
  select(Country, Year, y, k, l, e)
head(USData, n = 3)
```

Now review the data.

```{r}
USData %>% 
  gather(key = variable, value = value, k, l, e, y) %>% 
  mutate(
    # Put variables in a reasonable order for the graph legend
    variable = factor(variable, levels = c("y", "k", "e", "l"))
  ) %>% 
  ggplot(mapping = aes(x = Year, y = value, linetype = variable)) +
  geom_line() +
  geom_hline(yintercept = 1, colour = "gray50", size = 0.1) +
  labs(y = NULL, x = NULL, linetype = NULL) +

  xy_theme()
```

## Perform the fit

We'll fit with the $(kl)e$ nesting.

```{r}
kleModel <- cesModel(y ~ k + l + e, data = USData)
kleModel
nckle <- naturalCoef(kleModel)
nckle
```

Note that $\delta = 1$, so this is a boundary model.
We also have $\rho_1 = -1$ (therefore, $\sigma_1 = \infty$). 
We can query which boundary model we have: 
```{r}
kleModel$bname
```
So this is boundary model 7 of Table 2 in the ``Cautionary Tales'' paper.

With $\theta = \mathrm{e}^{\mathrm{logscale}} = \mathrm{e}^
{`r naturalCoef(kleModel)[["logscale"]]`} =
`r naturalCoef(kleModel)[["scale"]]`$,
these results mean that boundary model 7

$$y = \theta A \left[ \delta_1 x_1 + (1 - \delta_1) x_2 \right]$$
is filled with coefficients like this:

$$y = `r exp(coef(kleModel)[["logscale"]])` 
      \mathrm{e}^{-`r coef(kleModel)[["lambda"]]` (t-1960)} 
      \left[ `r coef(kleModel)[["delta_1"]]` \, k 
          + `r 1 - coef(kleModel)[["delta_1"]]` \, l \right] \; .$$

## Compare with historical data

We can compare the fitted model against the historical data
as shown in the following graph.

```{r}
USData %>% 
  mutate(
    `y predicted` = yhat(kleModel) %>% as.numeric() # avoids an attributes warning
  ) %>% 
  select(Year, y, `y predicted`) %>% 
  gather(key = var, value = value, y, `y predicted`) %>% 
  ggplot(mapping = aes(x = Year, y = value, linetype = var)) +
  geom_line() + 
  geom_hline(yintercept = 1, colour = "gray50", size = 0.1) +
  labs(x = NULL, y = NULL, linetype = NULL) +
  xy_theme()
```

## Comments

1. The fit is rather good.
2. It is likely that the fitted model excludes energy ($e$), 
   because the energy dip in 1980 is too large. 
   Instead, by picking up the smaller dip in the $l$ data series, the fit is better.
3. I find it interesting, but don't quite know what to make of it, that
   the coefficients on $k$ and $l$ are approximately the neoclassical numbers 
   (0.7 and 0.3, respectively).