### 2018-07-11 Articles/Court_Analysis_2018/FittingForCourt2018.Rmd
Edits to Victor Court analysis.

---
title: "US CES Analysis for Victor Court"
author: "Matthew Kuperus Heun"
date: "10 July 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(dplyr)
library(EconModels)
library(ggplot2)
library(magrittr)
library(tidyr)
```

## Analysis for Victor Court

Victor Court has asked me to fit US energy and economic data
using the CES aggregate production function. 
The energy and economic data have been supplied by Victor.

## Load and sanitize data

First, load the data.

```{r}
USData_raw <- read.csv(file = "CES_USA.csv", sep = ";")
head(USData_raw, n = 3) # Show the first 3 rows
```

Next, sanitize the data a bit
by removing non-indexed columns and
renaming the columns.
Lower-case variable names indicate indexing to 1 in 1960.

```{r}
USData <- USData_raw %>% 
  select(-K.L...US.2010.per.hour., 
         -Capital...US.2010., 
         -Labor..millions.of.hours., 
         -Primary.exergy..TJ.,
         -GDP....US.2010.) %>% 
  rename(
    k = K..1960...1.,
    l = L..1960...1.,
    e = E..1960...1.,
    y = Y..1960...1.
  ) %>% 
  mutate(
    Country = "US"
  ) %>% 
  select(Country, Year, y, k, l, e) # Re-order the columns
head(USData, n = 3)
```

Now review the data in graphical form.

```{r}
USData %>% 
  gather(key = variable, value = value, k, l, e, y) %>% 
  mutate(
    # Put variables in a reasonable order for the graph legend
    variable = factor(variable, levels = c("y", "k", "e", "l"))
  ) %>% 
  ggplot(mapping = aes(x = Year, y = value, linetype = variable)) +
  geom_line() +
  geom_hline(yintercept = 1, colour = "gray50", size = 0.1) +
  labs(y = NULL, x = NULL, linetype = NULL) +

  xy_theme()
```

## Perform the fit

We'll fit with the $(kl)e$ nesting.

```{r}
kleModel <- cesModel(y ~ k + l + e, data = USData)
kleModel
```
Note that $\delta = 1$, so this is a boundary model.
We also have $\rho_1 = -1$ (therefore, $\sigma_1 = \infty$).

We can query which boundary model we have obtained: 
```{r}
kleModel$bname
```
So this is boundary model 7 of Table 2 in the ``Cautionary Tales'' paper.

We can see the ``natural'' coefficients for the model.
```{r}
nckle <- naturalCoef(kleModel)
nckle
```

With $\theta = \mathrm{e}^{\mathrm{logscale}} = \mathrm{e}^
{`r nckle[["logscale"]]`} = `r nckle[["scale"]]`$,
these results mean that boundary model 7

$$y = \theta A \left[ \delta_1 x_1 + (1 - \delta_1) x_2 \right]$$
is filled with coefficients like this:

$$y = `r nckle[["scale"]]` 
      \mathrm{e}^{`r coef(kleModel)[["lambda"]]` (t-1960)} 
      \left[ `r coef(kleModel)[["delta_1"]]` \, k 
          + `r 1 - coef(kleModel)[["delta_1"]]` \, l \right] \; .$$

At the moment, I'm a bit unclear about the $\theta$ coefficient.
Inserting values for 1960 ($k = 1$, $l = 1$)
gives $y = \theta \mathrm{e}^0 [1] = \theta = `r nckle[["scale"]]` \ne 1$,
but it should be about 1.

I probably have something wrong in my thinking about this.  
Any ideas?

## Compare with historical data

We can compare the fitted model against the historical data
as shown in the following graph.

```{r}
USData %>% 
  mutate(
    `y fitted` = yhat(kleModel) %>% as.numeric() # avoids an attributes warning
  ) %>% 
  select(Year, y, `y fitted`) %>% 
  gather(key = var, value = value, y, `y fitted`) %>% 
  ggplot(mapping = aes(x = Year, y = value, linetype = var)) +
  geom_line() + 
  geom_hline(yintercept = 1, colour = "gray50", size = 0.1) +
  labs(x = NULL, y = NULL, linetype = NULL) +
  xy_theme()
```

## Comments

1. Visually, the fit is quite good.
2. It is likely that the fitted model excludes energy ($e$), 
   because the energy dip in 1980 puts the energy trajectory at a very different
   multiple of the $y$ curve before and after 1980. 
   Mixing $e$ into the fit makes the fitted model worse!
   In contrast, the $l$ curve stays at about the same multiple of $y$
   throughout the entire time series, yet retains the dip in 1980. 
   The $l$ curve provides a better approximation to the shape of $y$.
   So, the fitting algorithm chooses to use $l$ instead of $e$ to fit $y$.
3. I find it interesting, but don't quite know what to make of it, that
   the coefficients on $k$ and $l$ are approximately the neoclassical numbers 
   (0.7 and 0.3, respectively).
   Maybe it's a result of the fact that the $k$ and $l$ lines are
   approximately a ratio of 2 apart from each other.