### 2018-08-13 README.Rmd
remove lattice and latticeExtra; add README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# MacroGrowth

The MacroGrowth package provides sophisticated tools for fitting macroeconomic 
growth models to data. MacroGrowth has several desirable features for economic modelers:

* Functions that fit several different macroeconomic growth models.
* Functions fit along 
boundaries of the economically-meaningful region, 
thereby ensuring that the best possible fit is found.
* Options to perform residual resampling, 
thereby providing a way to assess the variability of estimated parameters.
* A consistent formula-based interface to fitting functions.

The functions in this package were developed for and used first in the paper
Heun, M.K., J. Santos, P.E. Brockway, R. Pruim, T. Domingos, and M. Sakai.
"From Theory to Econometrics to Energy Policy: 
Cautionary Tales for Policymaking Using Aggregate Production Functions".
*Energies*, pp. 1-44, 10 (203), 2017.
<http://doi.org/10.3390/en10020203>.

## Installation

You can install MacroGrowth from github with:

```{r gh-installation, eval = FALSE}
# install devtools if not already installed
# install.packages("devtools")
devtools::install_github("EconModels/MacroGrowth")
# To get vignettes built locally, use
devtools::install_github("EconModels/MacroGrowth", build_vignettes = TRUE)
```